---
title: "Introduction"
author: "Connor Quiroz, Lana Huang, Christopher D. Golden, Sarah Glaser, and Jessica A. Gephart"
date: "2026-08-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
inland <- consumption_full %>%
  group_by(year, consumer_iso3c, habitat) %>%
  summarize(consumption_live_t = sum(consumption_live_t)) %>%
  group_by(year, consumer_iso3c) %>%
  mutate(prop_consumption = consumption_live_t / sum(consumption_live_t)) %>% 
  filter(year == 2023) %>%
  filter(habitat == "inland") %>%
  left_join(
    ecoregion %>%
      select(consumer_iso3c = iso3c, realm)
  ) %>%
  group_by(realm) %>%
  mutate(realm_prop = mean(prop_consumption > 0.6),
         n_countries = n()) %>%
  distinct(realm, realm_prop, n_countries) %>%
  arrange(-realm_prop) %>%
  drop_na()

marine <- consumption_full %>%
  group_by(year, consumer_iso3c, habitat) %>%
  summarize(consumption_live_t = sum(consumption_live_t)) %>%
  group_by(year, consumer_iso3c) %>%
  mutate(prop_consumption = consumption_live_t / sum(consumption_live_t)) %>% 
  filter(year == 2023) %>%
  filter(habitat == "marine") %>%
  left_join(
    ecoregion %>%
      select(consumer_iso3c = iso3c, realm)
  ) %>%
  group_by(realm) %>%
  mutate(realm_prop = mean(prop_consumption > 0.8),
         n_countries = n()) %>%
  distinct(realm, realm_prop, n_countries) %>%
  arrange(-realm_prop) %>%
  drop_na()

inland
marine
```

For example, nations in the Central Indo-Pacific, Tropical Atlantic, and Temperate Northern Atlantic majorly consume marine species, whereas landlocked nations in Asia, Africa, and South America consume more inland species.

```{r}
 consumption_full %>%
  group_by(year, consumer_iso3c, consumption_source) %>%
  summarize(consumption_live_t = sum(consumption_live_t)) %>%
  group_by(year, consumer_iso3c) %>%
  mutate(prop_consumption = consumption_live_t / sum(consumption_live_t)) %>% 
  filter(year == 2023) %>%
  filter(consumption_source == "foreign") %>%
  add_region("consumer_iso3c", "region") %>%
  filter(region %in% c("North America", "Europe", "Africa")) %>%
  View()
```

North America, Europe, and Africa have particularly increased their foreign consumption over the years, exceeding domestic consumption in some countries like the United States, Germany, and Botswana