---
title: "Introduction"
author: "Connor Quiroz, Lana Huang, Christopher D. Golden, Sarah Glaser, and Jessica A. Gephart"
date: "2026-08-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Read in data

```{r}
# Read in aquatic reliance data
aquatic_reliance <- read.csv("../output/fao_aquatic_reliance_source.csv")

# Read in consumption data
consumption_full <- read_parquet("../data/consumption_full.parquet")

# Read in ecoregion data (all shapes associated with soverign)
ecoregion_full <- 
  read.csv("../data/Spatially joined ecoregion to EEZ/ecoregion.csv") %>%
  rename(
    "iso3c" = "ISO_SOV1",
    ecoregion = "ECOREGION",
    province = "PROVINCE",
    realm = "REALM",
    shape_area = "Shape_Area"
    )

# Read in ecoregions for largest sovereign shape
ecoregion_unique <- read.csv("../data/Spatially joined ecoregion to EEZ/ecoregion_clean.csv") %>%
  select(-FREQUENCY) %>%
  rename(
    iso3c = "ISO_SOV1",
    shape_area = "MAX_Shape_Area"
    )

# Join largest shape for the respective sovereign country
# to their ecoregion names
ecoregion <- left_join(
  ecoregion_unique,
  ecoregion_full,
  by = c("iso3c", "shape_area")
  ) %>%
  filter(realm != "")

# Assign countries not associated with a marine ecoregion to "landlocked" code
landlocked_countries <- aquatic_reliance %>%
  distinct(iso3c) %>%
  drop_na() %>%
  add_region(col = "iso3c", region.col.name = "region") %>%
  filter(!iso3c %in% ecoregion$iso3c) %>%
  mutate(
    ecoregion = paste0("Landlocked, ", region),
    province = paste0("Landlocked, ", region),
    realm = paste0("Landlocked, ", region)) %>%
  select(-region)

# Clean ecoregion data
ecoregion <- bind_rows(
  ecoregion,
  landlocked_countries
  ) %>%
  arrange(iso3c) %>%
  select(-shape_area)

# Clean up environment
rm(list = c("ecoregion_full", "ecoregion_unique"))
```

# Analysis

## Consumption hotspots for inland- and marine-source aquatic animal foods

```{r}
# Here, we calculate the proportion of countries in each realm and the
# total number of countries in each realm to find: 

# What proportion of countries in each realm source more than 60% of their
# aquatic food consumption from inland sources? Use 2023
inland <- consumption_full %>%
  group_by(year, consumer_iso3c, habitat) %>%
  summarize(consumption_live_t = sum(consumption_live_t)) %>%
  group_by(year, consumer_iso3c) %>%
  mutate(prop_consumption = consumption_live_t / sum(consumption_live_t)) %>% 
  filter(year == 2023) %>%
  filter(habitat == "inland") %>%
  left_join(
    ecoregion %>%
      select(consumer_iso3c = iso3c, realm)
  ) %>%
  group_by(realm) %>%
  mutate(realm_prop = mean(prop_consumption > 0.6),
         n_countries = n()) %>%
  distinct(realm, realm_prop, n_countries) %>%
  arrange(-realm_prop) %>%
  drop_na()

# What proportion of countries in each realm source more than 60% of their
# aquatic food consumption from inland sources? Use 2023
marine <- consumption_full %>%
  group_by(year, consumer_iso3c, habitat) %>%
  summarize(consumption_live_t = sum(consumption_live_t)) %>%
  group_by(year, consumer_iso3c) %>%
  mutate(prop_consumption = consumption_live_t / sum(consumption_live_t)) %>% 
  filter(year == 2023) %>%
  filter(habitat == "marine") %>%
  left_join(
    ecoregion %>%
      select(consumer_iso3c = iso3c, realm)
  ) %>%
  group_by(realm) %>%
  mutate(realm_prop = mean(prop_consumption > 0.8),
         n_countries = n()) %>%
  distinct(realm, realm_prop, n_countries) %>%
  arrange(-realm_prop) %>%
  drop_na()

inland
marine
```

For example, nations in the Central Indo-Pacific, Tropical Atlantic, and Temperate Northern Atlantic majorly consume marine species, whereas landlocked nations in Asia, Africa, and South America consume more inland species.

## Which countries source a majority of their consumed aquatic foods from foreign sources?

```{r}
 consumption_full %>%
  group_by(year, consumer_iso3c, consumption_source) %>%
  summarize(consumption_live_t = sum(consumption_live_t)) %>%
  group_by(year, consumer_iso3c) %>%
  mutate(prop_consumption = consumption_live_t / sum(consumption_live_t)) %>% 
  filter(year == 2023) %>%
  filter(consumption_source == "foreign") %>%
  add_region("consumer_iso3c", "region") %>%
  filter(region %in% c("North America", "Europe", "Africa")) %>%
  View()
```

North America, Europe, and Africa have particularly increased their foreign consumption over the years, exceeding domestic consumption in some countries like the United States, Germany, and Botswana.