---
title: "01_Analysis"
author: "Connor Quiroz"
date: "2025-03-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(arrow)
library(countrycode)
library(exploreARTIS)
```

```{r read in data created from 00_Preprocess_Data}
# Read in consumption data
consumption_full <- read_parquet("../output/consumption_full.parquet")

# Read in FAO food data
fao_food <- read_parquet("../output/fao_food.parquet")

# Read in historical FAO food data
fao_historical <- read_parquet("../output/fao_historical.parquet")

# Get years for each FAO dataset
fao_years <- paste0("Y", as.character(2010:2019))
historical_years <- paste0("Y", as.character(1996:2013))
```

```{r function for preprocessing FAO data}
fao_preprocess <- function(data = data, colnames = "", element = "") {
  # Convert FAO data to long format
fao_food_long <- data %>%
  pivot_longer(cols = colnames, names_to = "year") %>%
  mutate(year = str_replace(year, "Y", ""))


# Extract FAO animal protein totals
aquatic_animal_proteins <- fao_food_long %>%
  filter(Element == element,
         Item %in% c("Aquatic Animals, Others", "Cephalopods",
                     "Crustaceans", "Demersal Fish", "Fish, Seafood",
                     "Freshwater Fish", "Marine Fish, Other", "Meat, Aquatic Mammals",
                     "Molluscs, Other"),) %>%
  mutate(iso3c = countrycode(Area, origin = 'country.name', destination = 'iso3c')) %>%
  relocate(iso3c, .before = "Area") %>%
  group_by(iso3c, year) %>%
  summarize(a_a_protein = sum(value, na.rm = TRUE)) %>%
  ungroup()

# Calculate the total animal protein in FAO FBS
total_protein <- fao_food_long %>%
  filter(Element == element,
         Item == "Animal Products") %>%
  mutate(iso3c = countrycode(Area, origin = 'country.name', destination = 'iso3c')) %>%
  relocate(iso3c, .before = "Area") %>%
  group_by(iso3c, year) %>%
  summarize(t_p = sum(value, na.rm = TRUE)) %>%
  ungroup()

# Produce proportions aquatic animal protein of total conusmed protein
fao_prop_imports <- left_join(aquatic_animal_proteins, total_protein) %>%
  group_by(iso3c, year) %>%
  summarize(prop_aquatic_animal = a_a_protein / t_p) %>%
  mutate(year = as.numeric(year))

return(fao_prop_imports)
}
```

# Ensure FAO Data can be stiched together

Old FAO and New FAO Food Balance Sheets data overlap in years 2010-2013. Since both datasets use different methods between calculating protein supply quantity, we want to make sure these methods do not produce significantly different results - we will perform t-tests on overlapping years to compare the values. If they are not significantly different across years, then we will stitch the years together

```{r compare old vs new FAO FBS data}
# Get 1996-2019 FAO data
historical_fao_props <- fao_preprocess(data = fao_historical, 
                                         colnames = historical_years,
                                         element = "Protein supply quantity (g/capita/day)")

# Get 2010-2019 FAO data
fao_prop_imports <- fao_preprocess(data = fao_food, 
                                   colnames = fao_years,
                                   element = "Protein supply quantity (t)")

# Check 2010: p value = 0.37
t.test(historical_fao_props %>% filter(year == 2010) %>% pull(prop_aquatic_animal),
       fao_prop_imports %>% filter(year == 2010) %>% pull(prop_aquatic_animal))

# Check 2011: p-value = 0.37
t.test(historical_fao_props %>% filter(year == 2011) %>% pull(prop_aquatic_animal),
       fao_prop_imports %>% filter(year == 2011) %>% pull(prop_aquatic_animal))

# Check 2012: p-value = 0.37
t.test(historical_fao_props %>% filter(year == 2012) %>% pull(prop_aquatic_animal),
       fao_prop_imports %>% filter(year == 2012) %>% pull(prop_aquatic_animal))

# Check 2013: p-value = 0.43
t.test(historical_fao_props %>% filter(year == 2013) %>% pull(prop_aquatic_animal),
       fao_prop_imports %>% filter(year == 2013) %>% pull(prop_aquatic_animal))


# All overlapping years were not significantly different. We can merge safely merge datasets
fao_prop_imports <- bind_rows(historical_fao_props %>% filter(year < 2010), fao_prop_imports)
```


```{r calculate importance on protein consumption (ARTIS + FAO)}
# Calculate proportion of consumed materials from different habitats and methods
habitat_method_props <- consumption_full %>%
  group_by(year, consumer_iso3c, habitat, method, consumption_source) %>%
  summarize(total_group_catch = sum(consumption_live_t)) %>%
  filter(!habitat == "unknown" | 
         !method == "unknown") %>%
  group_by(year, consumer_iso3c) %>%
  mutate(total_catch = sum(total_group_catch)) %>%
  group_by(year, consumer_iso3c, habitat, method, consumption_source) %>%
  summarize(prop_consumption = total_group_catch / total_catch) %>%
  ungroup()

# Join artis and fao food balance sheet data (combines habitat/method props to prop animal protein imports byt fao)
artis_fao_fbb <- left_join(habitat_method_props, fao_prop_imports, by = c("consumer_iso3c" = "iso3c", "year"))

# How important is marine capture on countries' animal sourced protein supply?
supply_importance <- artis_fao_fbb %>%
  group_by(year, consumer_iso3c, habitat, method, consumption_source) %>%
  mutate(importance_on_protein_cons = prop_consumption * prop_aquatic_animal) %>%
  add_region("consumer_iso3c", region.col.name = "region") %>%
  filter(!is.na(region)) %>%
  ungroup()

# Remove NA prop values from supply importance
supply_importance <- supply_importance %>%
  filter(!is.na(prop_aquatic_animal)) %>%
  relocate(region, .after = "consumer_iso3c")

# Clean up some names (i.e., capitalize data entries)
supply_importance <- supply_importance %>%
  mutate(consumption_source = str_to_title(consumption_source),
         habitat = str_to_title(habitat),
         method = str_to_title(method))
```

```{r tests to ensure dataset works as is}
# Make sure all items add to 1
supply_importance %>%
  group_by(consumer_iso3c, year) %>%
  summarize(total = sum(prop_consumption))

# Make sure everything sums to only 1 within year / consuming country
supply_importance %>%
  group_by(consumer_iso3c, year) %>%
  summarize(total = sum(prop_consumption)) %>%
  ungroup() %>%
  filter(total > 1.001)
```

```{r save new dataframe}
# Everything works as it it's supposed to
# Save to parquet
write_parquet(supply_importance, "../output/supply_importance.parquet")
```

