---
title: "02.1_Supplementary_Information"
authors: "Connor Quiroz, Lana Huang, Christopher D. Golden, Sarah Glaser, and Jessica A. Gephart"
date: "2026-03-04"
output: html_document
---

# Chapter 1 Supplementary information

# Load in packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(exploreARTIS)
library(cowplot)
library(ggrepel)
library(countrycode)
library(colorspace)
library(ggpubr)
library(arrow)
library(rstan)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggridges)
library(bayesplot)
library(purrr)
library(patchwork)
library(reshape2)
```

# Read in data

### Read in aquatic reliance data & FAO population data

```{r read in FAO derived data}
# Read in population
population_fao_wb <- read.csv("../output/population_fao_wb.csv")

# Read in aquatic reliance data
aquatic_reliance <- read.csv("../output/fao_aquatic_reliance_source.csv") %>%
  left_join(population_fao_wb,by=c("year","iso3c"))%>%
  mutate(protein_consumed_kg_percap=protein_consumed_t*1000/population_fao)

# Read in FAO sourcing data
aquatic_fbs <- read.csv("../output/fao_fbs_aquatic_source.csv")
```

## Get list of response variables (models)

This code chunk is also in the main results. This code chunk extracts the name of each model name used in the Bayesian model. In code lines further down, we call the "response_vars" object to generate statistics and figures. This code chunk will only run if the response_vars object has not been created.

```{r acquire animal consumption data to be fed through analysis}
if ( !file.exists("../output/response_vars.rds") ) {
  
  # Calculate animal consumption
terr_animal_cons_cap <- aquatic_reliance %>%
  filter(year == 2019) %>%
  group_by(iso3c) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))

animal_cons_cap <- aquatic_reliance %>%
  filter(year == 2019) %>%
  group_by(iso3c) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))

# Calculate aquatic animal consumption
aquatic_animal_cons_cap <- aquatic_reliance %>%
  filter(year == 2019, food_group == "aquatic") %>%
  group_by(iso3c) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))

# Calculate animal reliance consumption
aquatic_rel <- aquatic_animal_cons_cap %>%
  left_join(animal_cons_cap, by = "iso3c") %>%
  group_by(iso3c) %>%
  summarize(prop_animal_protein = protein_consumed_kg_percap.x / protein_consumed_kg_percap.y)

# Add items not created in the loop to the list
response_vars <- c("terr_animal_cons_cap","animal_cons_cap", "aquatic_animal_cons_cap", "aquatic_rel", "aquatic_protein_roc", "terrestrial_protein_roc")

for (i in c("capture", "aquaculture", "domestic", "foreign", "marine", "inland")) {
  
  # Calculate consumption per capita by method
  if (i %in% c("capture", "aquaculture")) {
    assign(
      paste0(i, "_cap"),
      aquatic_reliance %>%
        filter(year == 2019, food_group == "aquatic", method == i) %>%
        group_by(iso3c) %>%
        summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))
    )
  }
  
  # Calculate consumption per capita by consumption source
  if (i %in% c("domestic", "foreign")) {
    assign(
      paste0(i, "_cap"),
      aquatic_reliance %>%
        filter(year == 2019, food_group == "aquatic", consumption_source == i) %>%
        group_by(iso3c) %>%
        summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))
    )
  }
  
  # Calculate consumption per capita by habitat
  if (i %in% c("marine", "inland")) {
    assign(
      paste0(i, "_cap"),
      aquatic_reliance %>%
        filter(year == 2019, food_group == "aquatic", habitat == i) %>%
        group_by(iso3c) %>%
        summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))
    )
  }
  
  response_vars <- c(response_vars, paste0(i, "_cap"))
  
  # Calculate reliance by method, production source, or habitat
  assign(
    paste0(i, "_rel"),
    get(paste0(i, "_cap")) %>%
      left_join(animal_cons_cap, by = "iso3c") %>%
      group_by(iso3c) %>%
      summarize(prop_animal_protein = protein_consumed_kg_percap.x / protein_consumed_kg_percap.y)
  )
  
  response_vars <- c(response_vars, paste0(i, "_rel"))
  
  # Calculate production sources' proportion of total aquatic animal consumption per capita
  assign(
    paste0(i, "_prop_of_aquatic_cons"),
    get(paste0(i, "_cap")) %>%
      left_join(aquatic_animal_cons_cap, by = "iso3c") %>%
      group_by(iso3c) %>%
      summarize(aquatic_source_prop = protein_consumed_kg_percap.x / protein_consumed_kg_percap.y)
  )
  
  response_vars <- c(response_vars, paste0(i, "_prop_of_aquatic_cons"))
}

#calculate rate of change in aquatic protein quantity
aquatic_protein_roc <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, year, method) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap, na.rm = TRUE)) %>%
  distinct() %>%
  filter(method == "capture") %>%
  split(.$iso3c) %>%
  map(~lm(protein_consumed_kg_percap~year, data = .x)) %>%
  map_df(broom::tidy, .id = 'iso3c') %>%
  filter(term == 'year') %>%
  select(c("iso3c", "aquatic_estimate" = "estimate"))

#calculate rate of change in terrestrial protein quantity
terrestrial_protein_roc <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "terrestrial") %>%
  group_by(iso3c, year) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap, na.rm = TRUE)) %>%
  distinct() %>%
  split(.$iso3c) %>%
  map(~lm(protein_consumed_kg_percap~year, data = .x)) %>%
  map_df(broom::tidy, .id = 'iso3c') %>%
  filter(term == 'year') %>%
  select(c("iso3c", "terrestrial_estimate" = "estimate"))

# Save response_vars to RDS format
saveRDS(response_vars, "../data/response_vars.rds")
}
```

## Read in list of response variables (models)

```{r read in response variables}
# Read in list of response variables used in Bayesian models
response_vars <- readRDS("../data/response_vars.rds")
```

# RQ 1a: How do aquatic food reliance trends disaggregated by source differ from aggregate aquatic food reliance?

## SI figure: Multi metric plot with 4 panels created for each individual region

The first row of the figure has 2 panels, and the following two rows have 1 panel, making a 4 panel plot. The regions include Africa, Asia, South America, North America, Europe, and Oceania.

The top left panel highlights the number of countries with increasing / decreasing *aggregate* aquatic animal reliance

The top right panel highlights the number of countries with increasing / decreasing *disaggregate* aquatic animal reliance. Where the y-axis represents the disaggregated production sources for aquatic foods.

The middle panel measures the 2020 aggregate aquatic animal reliance for each country.

The bottom panel visualizes for each country in each region the average annual change in their aggregate aquatic animal reliance, with separate colors as stacked bars representing the disaggregate reliance.

```{r individual continent production contribution}
# shared color palettes

# Production-source palette (matches main figure)
pal_consource <- c(
  "inland aquaculture" = "#E8B000",
  "inland capture"     = "#D4601A",
  "marine aquaculture" = "#9B6DC5",
  "marine capture"     = "#2B6AA8",
  "unknown"            = "#A8A8A8"
)

# Diverging palette for direction-of-change panels
pal_direction <- c(
  "decreasing" = "#B2182B",
  "increasing" = "#2166AC"
)

# Neutral fill for reliance bar chart
col_reliance_bar <- "#606060"


# Shared base theme

theme_si <- function(base_size = 8,
                     show_x_title = TRUE,
                     show_y_title = TRUE,
                     show_legend  = TRUE) {

  base <- theme_minimal(base_size = base_size) +
    theme(
      text                  = element_text(family = "sans", color = "black"),
      plot.background       = element_rect(fill = "transparent", color = NA),
      panel.background      = element_rect(fill = "transparent", color = NA),
      panel.grid.minor      = element_blank(),
      panel.grid.major.y    = element_line(
        color = scales::alpha("black", 0.06), linewidth = 0.3
      ),
      panel.grid.major.x    = element_line(
        color = scales::alpha("black", 0.25), linewidth = 0.3
      ),
      axis.text             = element_text(size = base_size - 2.5, color = "black"),
      axis.title.x          = element_text(
        size = base_size - 1, face = "bold",
        margin = margin(t = 3)
      ),
      axis.title.y          = element_text(
        size = base_size - 1, face = "bold",
        margin = margin(r = 3)
      ),
      strip.text            = element_text(size = base_size - 1, face = "bold"),
      legend.position       = if (show_legend) "bottom" else "none",
      legend.title          = element_text(size = base_size - 1.5, face = "bold"),
      legend.text           = element_text(size = base_size - 2.5),
      legend.key.size       = unit(0.25, "cm"),
      legend.box            = "vertical",
      legend.spacing        = unit(0.05, "cm"),
      legend.margin         = margin(t = 0, b = 0),
      legend.box.margin     = margin(t = -2, b = -2),
      legend.background     = element_rect(fill = "transparent", color = NA),
      legend.box.background = element_rect(fill = "transparent", color = NA),
      plot.margin           = margin(3, 4, 3, 4)
    )

  if (!show_x_title) base <- base + theme(axis.title.x = element_blank())
  if (!show_y_title) base <- base + theme(axis.title.y = element_blank())
  base
}

# suppress y-axis grid (used for horizontal count bars
theme_si_hbar <- function(base_size = 8, show_legend = TRUE) {
  theme_si(base_size = base_size, show_legend = show_legend) +
    theme(
      panel.grid.major.y = element_blank(),
      axis.text.y        = element_blank(),
      axis.ticks.y       = element_blank(),
      axis.title.y       = element_blank()
    )
}


# Shared geom helpers

agg_point <- function() {
  geom_point(
    aes(x = agg_estimate, y = iso3c),
    shape  = 21,
    fill   = "grey20",
    color  = "white",
    size   = 1.2,
    stroke = 0.3
  )
}


# Fit linear models to estimate the average annual change in disaggregated
# aquatic animal reliance (2010-2020) by production source (habitat x method)
# for each country
disagg <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "aquatic") %>%
  group_by(year, iso3c, habitat, method) %>%
  summarize(prop_animal_protein = sum(prop_animal_protein), .groups = "drop") %>%
  group_by(iso3c, habitat, method) %>%
  nest() %>%
  mutate(
    model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
    tidy  = map(model, broom::tidy)
  ) %>%
  unnest(tidy) %>%
  filter(term == "year") %>%
  mutate(
    group_label = paste(habitat, method),
    group_label = if_else(group_label == "unknown unknown", "unknown", group_label)
  ) %>%
  drop_na(estimate) %>%
  group_by(group_label) %>%
  mutate(pct_above = mean(estimate > 0) * 100) %>%
  ungroup()

# Fit linear models to estimate the average annual change in aggregate
# aquatic animal reliance (2010-2020)
agg <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, year) %>%
  summarise(prop_animal_protein = sum(prop_animal_protein), .groups = "drop") %>%
  group_by(iso3c) %>%
  nest() %>%
  mutate(
    model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
    tidy  = map(model, broom::tidy)
  ) %>%
  unnest(tidy) %>%
  filter(term == "year") %>%
  mutate(group_label = "aggregate") %>%
  drop_na(estimate) %>%
  group_by(group_label) %>%
  mutate(pct_above = mean(estimate > 0) * 100) %>%
  ungroup()

# Clean aggregate estimate data
agg_estimate <- bind_rows(agg, disagg) %>%
  group_by(iso3c) %>%
  mutate(agg_est = estimate[group_label == "aggregate"][1]) %>%
  add_region("iso3c", "region") %>%
  ungroup() %>%
  filter(group_label == "aggregate") %>%
  select(iso3c, agg_estimate = estimate)

# Extract rate of change in aggregate aquatic animal reliance from 2010-2020 
roc_aquatic_total <- aquatic_reliance %>%
  filter(
    !(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
    data_source == "new FBS",
    food_group  == "aquatic"
  ) %>%
  group_by(iso3c, year) %>%
  summarize(prop_animal_protein = sum(prop_animal_protein, na.rm = TRUE),
            .groups = "drop") %>%
  distinct() %>%
  split(list(.$iso3c), drop = TRUE) %>%
  map(~ lm(prop_animal_protein ~ year, data = .x)) %>%
  map_df(broom::tidy, .id = "combined_id") %>%
  filter(term == "year") %>%
  select(iso3c = combined_id, roc = estimate) %>%
  mutate(
    direction      = if_else(roc > 0, "increasing", "decreasing"),
    method_habitat = "total"
  ) %>%
  add_region(col = "iso3c", region.col.name = "continent")

counts_by_continent <- roc_aquatic_total %>%
  group_by(continent, method_habitat) %>%
  count(direction) %>%
  group_by(continent) %>%
  mutate(prop = n / sum(n))


# Extract legend

extract_legend <- function(p) {
  legs <- cowplot::get_plot_component(p, "guide-box", return_all = TRUE)
  good <- Filter(function(x) !inherits(x, "zeroGrob"), legs)
  if (length(good) == 0) return(NULL)
  good[[1]]
}


# Loop across regions

regions <- c("Asia", "Africa", "Europe", "North America", "South America", "Oceania")

for (i in regions) {

  # Panel 1: aggregate + disaggregated rate-of-change bars
  p_bars <- disagg %>%
    add_region("iso3c", "region") %>%
    filter(region == i) %>%
    left_join(agg_estimate, by = "iso3c") %>%
    mutate(iso3c = fct_reorder(iso3c, -agg_estimate)) %>%
    ggplot(aes(x = estimate, y = iso3c, fill = group_label)) +
    geom_col() +
    agg_point() +
    scale_fill_manual(
      values = pal_consource,
      name   = "Production source",
      drop   = FALSE
    ) +
    guides(fill = guide_legend(ncol = 2)) +
    labs(x = "Average annual change in aquatic animal reliance (2010–2020)") +
    theme_si(show_y_title = FALSE, show_legend = TRUE)

  # Panel 2: aggregate direction counts (horizontal stacked bar)
  p_agg_counts <- counts_by_continent %>%
    filter(method_habitat == "total", continent == i) %>%
    ggplot(aes(y = continent, x = n, fill = direction)) +
    geom_col(width = 0.55) +
    scale_fill_manual(
      values = pal_direction,
      labels = c(
        "decreasing" = "Declining reliance",
        "increasing" = "Increasing reliance"
      ),
      name = "Direction\nof change"
    ) +
    guides(fill = guide_legend(nrow = 2)) +
    labs(x = "No. countries (aggregate)") +
    theme_si_hbar(show_legend = TRUE)

  # Panel 3: disaggregated direction counts
  num_regional_countries <- disagg %>%
    add_region("iso3c", "region") %>%
    filter(region == i) %>%
    distinct(iso3c) %>%
    nrow()

  p_disagg_counts <- disagg %>%
    add_region("iso3c", "region") %>%
    filter(region == i) %>%
    group_by(region, group_label) %>%
    count(increasing = estimate > 0) %>%
    ungroup() %>%
    mutate(direction = if_else(increasing, "increasing", "decreasing")) %>%
    ggplot(aes(y = group_label, x = n, fill = direction)) +
    geom_col(width = 0.55) +
    scale_fill_manual(values = pal_direction, guide = "none") +
    labs(
      x = paste0(
        "No. countries (disaggregate, n = ", num_regional_countries, ")"
      )
    ) +
    theme_si(show_y_title = FALSE, show_legend = FALSE) +
    theme(panel.grid.major.y = element_blank())

  # Panel 4: 2020 reliance, ordered by rate of change
  p_reliance <- aquatic_reliance %>%
    filter(year == 2020, food_group == "aquatic") %>%
    group_by(iso3c) %>%
    summarize(prop_animal_protein = sum(prop_animal_protein), .groups = "drop") %>%
    left_join(agg_estimate, by = "iso3c") %>%
    add_region("iso3c", "region") %>%
    filter(region == i) %>%
    ggplot(aes(
      x = fct_reorder(iso3c, agg_estimate),
      y = prop_animal_protein * 100
    )) +
    geom_col(fill = col_reliance_bar, width = 0.7) +
    labs(y = "Aquatic animal\nreliance, 2020 (%)") +
    theme_si(show_x_title = FALSE, show_legend = FALSE) +
    theme(
      axis.text.x        = element_text(size = 4.5, angle = 45, hjust = 1),
      panel.grid.major.x = element_blank(),
      panel.grid.major.y = element_line(
        color = scales::alpha("black", 0.25), linewidth = 0.3
      )
    )

  # Assemble legends
  legend_bars       <- extract_legend(p_bars)
  legend_agg_counts <- extract_legend(p_agg_counts)

  p_bars_nl       <- p_bars       + theme(legend.position = "none")
  p_agg_counts_nl <- p_agg_counts + theme(legend.position = "none")

  legend_row <- cowplot::plot_grid(
    legend_agg_counts,
    legend_bars,
    ncol       = 2,
    rel_widths = c(1, 1.45)
  )

  # Title and divider
  title_panel <- ggplot() +
    labs(title = i) +
    theme_void() +
    theme(
      plot.title = element_text(
        hjust = 0.5, face = "bold", size = 9
      )
    )

  divider_panel <- ggplot() +
    theme_void() +
    geom_hline(yintercept = 0.5, linewidth = 0.4, color = "grey60")

  # Combine top count panels 
  top_counts <- cowplot::plot_grid(
    p_agg_counts_nl,
    p_disagg_counts,
    ncol       = 2,
    rel_widths = c(1, 1)
  )

  # Final assembly
  final_plot <- cowplot::plot_grid(
    title_panel,
    divider_panel,
    top_counts,
    p_reliance,
    p_bars_nl,
    legend_row,
    ncol        = 1,
    rel_heights = c(0.12, 0.04, 0.6, 0.8, 2.5, 0.32)
  )

  # Save plot
  ggsave(
    filename = paste0("../images/SI/multi_metrics_", i, "_SI.pdf"),
    plot     = final_plot,
    device   = cairo_pdf,
    width    = 5,
    height   = 7,
    units    = "in"
  )

}
```


# RQ 1b: How does aquatic animal reliance change with the relative contribution of aquatic versus terrestrial animal consumption

## SI Figure: Proportion of countries in each quadrant by region

This plot identifies the proportion of countries that reside in each quadrant by region. The top right quadrant indicates an increasing terrestrial animal & aquatic animal consumption. The top left quadrant indicates an increasing terrestrial animal & decreasing aquatic animal consumption. The bottom left indicates both decreasing terrestrial & aquatic animal consumption. The bottom right quadrant indicates a decreasing terrestrial & increasing aquatic animal consumption.


```{r proportion of countries in each quadrant by region}
# Calculate rate of change in aquatic protein quantity
aquatic_protein_quant <- aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, year) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap, na.rm = TRUE),
            .groups = "drop") %>%
  distinct() %>%
  split(.$iso3c) %>%
  map(~ lm(protein_consumed_kg_percap ~ year, data = .x)) %>%
  map_df(broom::tidy, .id = "iso3c") %>%
  filter(term == "year") %>%
  select(iso3c, aquatic_estimate = estimate)

# Calculate rate of change in terrestrial protein quantity
terrestrial_protein_quant <- aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         data_source == "new FBS", food_group == "terrestrial") %>%
  group_by(iso3c, year) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap, na.rm = TRUE),
            .groups = "drop") %>%
  distinct() %>%
  split(.$iso3c) %>%
  map(~ lm(protein_consumed_kg_percap ~ year, data = .x)) %>%
  map_df(broom::tidy, .id = "iso3c") %>%
  filter(term == "year") %>%
  select(iso3c, terrestrial_estimate = estimate)

# Calculate rate of change in aquatic reliance
aquatic_protein_reliance <- aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, year) %>%
  summarize(prop_animal_protein = sum(prop_animal_protein, na.rm = TRUE),
            .groups = "drop") %>%
  distinct() %>%
  split(.$iso3c) %>%
  map(~ lm(prop_animal_protein ~ year, data = .x)) %>%
  map_df(broom::tidy, .id = "iso3c") %>%
  filter(term == "year") %>%
  select(iso3c, aquatic_reliance_estimate = estimate)

# Join and compute quadrant proportions per region
quad_df <- aquatic_protein_quant %>%
  full_join(terrestrial_protein_quant, by = "iso3c") %>%
  full_join(aquatic_protein_reliance,  by = "iso3c") %>%
  add_region(col = "iso3c", region.col.name = "region") %>%
  mutate(
    quad_raw = case_when(
      aquatic_estimate > 0 & terrestrial_estimate > 0 ~ "Q1",
      aquatic_estimate < 0 & terrestrial_estimate > 0 ~ "Q2",
      aquatic_estimate < 0 & terrestrial_estimate < 0 ~ "Q3",
      aquatic_estimate > 0 & terrestrial_estimate < 0 ~ "Q4"
    )
  ) %>%
  group_by(region, quad_raw) %>%
  summarise(n = n_distinct(iso3c), .groups = "drop") %>%
  group_by(region) %>%
  mutate(
    prop     = n / sum(n),
    quadrant = factor(
      quad_raw,
      levels = c("Q1", "Q2", "Q3", "Q4"),
      labels = c(
        "Aquatic \u2191 / Terrestrial \u2191",
        "Aquatic \u2193 / Terrestrial \u2191",
        "Aquatic \u2193 / Terrestrial \u2193",
        "Aquatic \u2191 / Terrestrial \u2193"
      )
    )
  ) %>%
  ungroup()

# Plot parameters
max_val    <- 0.75   # axis limit — corresponds to 75% proportion
half_width <- 0.20   # angular half-width of each spike in radians

# Quadrant colours — harmonized with project palette
# Q1 (both up) = teal: growing together
# Q2 (aquatic down, terr up) = dark red: aquatic declining
# Q3 (both down) = dark wine: both declining
# Q4 (aquatic up, terr down) = blue: aquatic gaining share
quad_colors <- c(
  "Aquatic \u2191 / Terrestrial \u2191" = "#7BAEC4",   # mid blue — ambiguous
  "Aquatic \u2193 / Terrestrial \u2191" = "#B2182B",   # strong red — reliance clearly ↓
  "Aquatic \u2193 / Terrestrial \u2193" = "#D4956A",   # warm sand — ambiguous
  "Aquatic \u2191 / Terrestrial \u2193" = "#2166AC"    # strong blue — reliance clearly ↑
)
# Build spike polygons — each spike is a triangle from the origin
# r = prop * max_val * sqrt(2) scales so that prop = 1 fills to the axis limit
# Number of arc points — more = smoother curve, 20 is plenty at this scale
n_arc <- 20

spike_df <- quad_df %>%
  mutate(
    center_angle = case_when(
      quad_raw == "Q1" ~  pi / 4,
      quad_raw == "Q2" ~ 3*pi / 4,
      quad_raw == "Q3" ~ 5*pi / 4,
      quad_raw == "Q4" ~ 7*pi / 4
    ),
    r     = prop * max_val * sqrt(2),
    group = paste(region, quad_raw)
  ) %>%
  rowwise() %>%
  reframe(
    region   = region,
    quadrant = quadrant,
    prop     = prop,
    group    = group,
    # Fan of arc angles from left edge to right edge
    angle = c(NA, seq(center_angle - half_width,
                      center_angle + half_width,
                      length.out = n_arc)),
    # Origin point first (NA angle → 0,0), then arc points at radius r
    x = c(0, r * cos(seq(center_angle - half_width,
                         center_angle + half_width,
                         length.out = n_arc))),
    y = c(0, r * sin(seq(center_angle - half_width,
                         center_angle + half_width,
                         length.out = n_arc)))
  )

# Reference circle data — drawn at 50% and 100% of max_val
# Gives readers two anchors for judging spike length without cluttering axis labels
circle_df <- expand.grid(
  region = unique(spike_df$region),
  theta  = seq(0, 2 * pi, length.out = 200)
) %>%
  as_tibble() %>%
  crossing(r_frac = c(0.5, 1.0)) %>%
  mutate(
    x = r_frac * max_val * cos(theta),
    y = r_frac * max_val * sin(theta)
  )

# Axis quadrant label positions — placed just outside max_val in each quadrant
quad_axis_labels <- tibble(
  x     = c( 0.58, -0.58, -0.58,  0.58),
  y     = c( 0.58,  0.58, -0.58, -0.58),
  label = c("A\u2191 T\u2191", "A\u2193 T\u2191", "A\u2193 T\u2193", "A\u2191 T\u2193")
)

# Single reference label — annotates the outer circle in top-right of each facet
ref_label_df <- tibble(
  region = unique(spike_df$region)
) %>%
  bind_rows(
    tibble(region = unique(spike_df$region)) %>%
      mutate(x = -max_val * 0.5 * cos(pi/4) - 0.3,
             y =  max_val * 0.5 * sin(pi/4) + 0.03,
             label = "37.5%"),
    tibble(region = unique(spike_df$region)) %>%
      mutate(x =  max_val * cos(pi/4) + 0.04,
             y =  max_val * sin(pi/4) + 0.04,
             label = "75%")
  )

# Assemble plot
plot <- ggplot() +

  # Reference circles — dashed, at 50% and 100% of axis range
  geom_path(
    data    = circle_df %>% filter(r_frac == 1.0),
    aes(x = x, y = y, group = interaction(region, r_frac)),
    colour    = "grey70",
    linewidth = 0.25,
    linetype  = "dashed"
  ) +
  geom_path(
    data    = circle_df %>% filter(r_frac == 0.5),
    aes(x = x, y = y, group = interaction(region, r_frac)),
    colour    = "grey80",
    linewidth = 0.18,
    linetype  = "dashed"
  ) +

  # Axis lines
  geom_hline(yintercept = 0, linewidth = 0.35, colour = "grey30") +
  geom_vline(xintercept = 0, linewidth = 0.35, colour = "grey30") +

  # Spikes
  geom_polygon(
    data = spike_df,
    aes(x = x, y = y, fill = quadrant, group = group),
    alpha = 0.88
  ) +

  # 75% reference label — top-right corner only, one per facet
  geom_text(
    data  = ref_label_df,
    aes(x = x, y = y, label = label),
    size  = 1.8,
    color = "grey55",
    hjust = 0,
    vjust = 0
  ) +

  facet_wrap(~ region, nrow = 2) +
  coord_equal(
    xlim   = c(-max_val - 0.12, max_val + 0.12),
    ylim   = c(-max_val - 0.12, max_val + 0.12)
  ) +
  scale_x_continuous(breaks = NULL) +
  scale_y_continuous(breaks = NULL) +
  scale_fill_manual(
    values = quad_colors,
    name   = NULL
  ) +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE,
                              override.aes = list(alpha = 1))) +
  labs(
    x = "Aquatic protein consumption \u2192",
    y = "Terrestrial protein consumption \u2192"
  ) +
  theme_minimal(base_size = 8) +
  theme(
    panel.grid            = element_blank(),
    panel.background      = element_rect(fill = "transparent", color = NA),
    plot.background       = element_rect(fill = "transparent", color = NA),
    strip.background      = element_blank(),
    strip.text            = element_text(size = 6.5, face = "bold"),
    axis.text             = element_blank(),
    axis.title.x          = element_text(size = 6, face = "bold",
                                          margin = margin(t = 2)),
    axis.title.y          = element_text(size = 6, face = "bold",
                                          margin = margin(r = 2)),
    legend.position       = "bottom",
    legend.box            = "vertical",
    legend.title          = element_blank(),
    legend.text           = element_text(size = 5.5),
    legend.key.size       = unit(0.3, "cm"),
    legend.spacing        = unit(0.1, "cm"),
    legend.margin         = margin(t = 2, b = 0),
    legend.background     = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    panel.spacing         = unit(0.8, "lines")
  )

ggsave(
  filename = "../images/SI/quadrant_spike_region_SI.pdf",
  plot     = plot,
  device   = cairo_pdf,
  width    = 5,
  height   = 3.5,
  units    = "in"
)
```


# RQ 2: What biophysical and socioeconomic factors are associated with reliance on farmed and wild aquatic animals for protein and overall consumption?

## SI Figure: Directed Acylcic Diagram (DAG). The DAG was not created in R, but in LucidChart.

## SI Figure: Extra posterior distributions that show the partial slopes for models that use disaggregated reliance as the response.

Posterior distributions are displayed with 90%, 80%, and 50% credibility intervals, represented by the lightest, darker, and darkest colors respectively. The point represents the posterior median.

```{r Extra posteriors}
model_type = "realm"

# Variable list with coastline_km area instead of EEZ area
predictors <- c(
  "coastline_km",
  "log_gdp",
  "gdp_trade",
  "inundation_area_log1000km2",
  "supermarkets",
  "pop_10k_pct_coast"
)

# Get which variable is selected (for unique file path)
var_used = predictors[1]

# Create file path to save files to
base_path <- file.path(
  "..", "output", "bayesian_output",
  paste0("random_effect_", model_type),
  paste0("var_", var_used)
)

# Create data frame to hold all posteriors across variables
slopes_df <- data.frame()

for (i in seq_along(response_vars)) {

  # Extract MCMC samples
  x <- rstan::extract(
    readRDS(file.path(base_path, paste0("fit_", response_vars[i], ".rds"))),
    permuted = TRUE
  )
  beta_global <- x$beta
  X           <- readRDS(file.path(base_path, paste0("X_", response_vars[i], ".rds")))

  n_iter <- nrow(beta_global)
  K      <- ncol(beta_global)
  vars   <- colnames(X)

  # Extract global slopes and calculate credibility intervals
  ci_df <- expand_grid(iter = 1:n_iter, k = 1:K) %>%
    mutate(
      value = map2_dbl(iter, k, ~ beta_global[.x, .y]),
      var   = vars[k]
    ) %>%
    group_by(var) %>%
    summarise(
      lo95 = quantile(value, 0.025),
      lo80 = quantile(value, 0.10),
      lo50 = quantile(value, 0.25),
      med  = quantile(value, 0.50),
      hi50 = quantile(value, 0.75),
      hi80 = quantile(value, 0.90),
      hi95 = quantile(value, 0.975),
      .groups = "drop"
    ) %>%
    mutate(model = response_vars[i])

  slopes_df <- bind_rows(slopes_df, ci_df)
}

# Load script to convert model name strings to readable labels
source("../R/clean_model_name.R")

# Clean variable names for axis labels
var_labels <- c(
  coastline_km               = "Coastline length (km)",
  gdp_trade                  = "GDP from trade (%)",
  inundation_area_log1000km2 = "Inundated area (log 1000 km²)",
  log_gdp                    = "GDP (log scale)",
  pop_10k_pct_coast          = "Pop. within 10 km of coast (%)",
  supermarkets               = "Supermarkets per 100,000"
)

# Model colours — harmonized palette, reliance models only
model_colors <- c(
  "Aquatic reliance"      = "#1A7A6E",
  "Aquaculture reliance"  = "#E8B000",
  "Capture reliance"      = "#D4601A",
  "Domestic reliance"     = "#7BAEC4",
  "Foreign reliance"      = "#C45C3A",
  "Marine reliance"       = "#9B6DC5",
  "Inland reliance"       = "#D4956A"
)

# Interval colours — three lightness steps per model colour
col_95 <- colorspace::lighten(model_colors, 0.65)
col_80 <- colorspace::lighten(model_colors, 0.35)
col_50 <- colorspace::lighten(model_colors, 0.05)

# Build plotting data
df_plot <- slopes_df %>%
  mutate(
    clean_name = clean_model_name(model),
    nonzero    = (lo95 < 0 & hi95 < 0) | (lo95 > 0 & hi95 > 0),
    suffix     = sub(".*_", "", model),
    suffix_num = suppressWarnings(as.numeric(suffix))
  ) %>%
  filter(
    str_detect(clean_name, "reliance"),
    model != "animal_cons_cap"
  ) %>%
  arrange(suffix_num, suffix) %>%
  mutate(
    clean_name = factor(clean_name, levels = unique(clean_name)),
    col95      = coalesce(col_95[as.character(clean_name)], "grey80"),
    col80      = coalesce(col_80[as.character(clean_name)], "grey60"),
    col50      = coalesce(col_50[as.character(clean_name)], "grey40"),
    panel_id   = as.numeric(factor(clean_name))
  )

# Shared symmetric x-axis limits so zero lines align across all panels
x_limit <- max(abs(c(df_plot$lo95, df_plot$hi95)), na.rm = TRUE) * 1.05
x_limit <- ceiling(x_limit * 10) / 10

panel_names <- levels(df_plot$clean_name)

# Function to draw a single credible interval panel
# left_col = TRUE shifts the tag inward to clear the y-axis label text
make_panel <- function(df, show_yaxis = FALSE, title = "", tag = "",
                       left_col = FALSE) {

  df_sig  <- df %>% filter(nonzero)
  df_nsig <- df %>% filter(!nonzero)

  # Left-column panels have wide y-axis text; nudge the tag right so it
  # doesn't sit directly on top of the longest label
  tag_pos <- if (left_col) c(0.52, 1) else c(0.01, 1)

  ggplot(df, aes(y = var)) +

    # Non-significant rows — reduced opacity, drawn behind
    geom_segment(data = df_nsig,
                 aes(x = lo95, xend = hi95, color = col95),
                 linewidth = 2.2, lineend = "butt", alpha = 0.3) +
    geom_segment(data = df_nsig,
                 aes(x = lo80, xend = hi80, color = col80),
                 linewidth = 2.2, lineend = "butt", alpha = 0.3) +
    geom_segment(data = df_nsig,
                 aes(x = lo50, xend = hi50, color = col50),
                 linewidth = 2.2, lineend = "butt", alpha = 0.3) +
    geom_point(data = df_nsig,
               aes(x = med), shape = 21, size = 1.6,
               fill = "white", color = "grey60", stroke = 0.3, alpha = 0.5) +

    # Significant rows — full opacity, drawn on top
    geom_segment(data = df_sig,
                 aes(x = lo95, xend = hi95, color = col95),
                 linewidth = 2.2, lineend = "butt", alpha = 1) +
    geom_segment(data = df_sig,
                 aes(x = lo80, xend = hi80, color = col80),
                 linewidth = 2.2, lineend = "butt", alpha = 1) +
    geom_segment(data = df_sig,
                 aes(x = lo50, xend = hi50, color = col50),
                 linewidth = 2.2, lineend = "butt", alpha = 1) +
    geom_point(data = df_sig,
               aes(x = med), shape = 21, size = 1.6,
               fill = "white", color = "grey20", stroke = 0.4) +

    geom_vline(xintercept = 0, linetype = "dashed",
               linewidth = 0.35, color = "grey40") +
    scale_color_identity() +
    scale_x_continuous(limits = c(-x_limit, x_limit)) +
    scale_y_discrete(labels = var_labels) +
    labs(title = title, tag = tag) +
    theme_minimal(base_size = 8) +
    theme(
      plot.title         = element_text(size = 5.5, face = "bold", hjust = 0.5,
                                        margin = margin(b = 2)),
      plot.tag           = element_text(size = 7, face = "bold"),
      plot.tag.position  = tag_pos,
      panel.grid.major.y = element_blank(),
      panel.grid.minor   = element_blank(),
      panel.grid.major.x = element_line(
        color = scales::alpha("black", 0.12), linewidth = 0.3
      ),
      axis.text.x        = element_text(size = 5.5, color = "black"),
      axis.title         = element_blank(),
      plot.margin        = margin(4, 6, 4, 6),
      axis.text.y        = if (show_yaxis) element_text(size = 5.5, color = "black",
                                                         hjust = 1) else element_blank(),
      axis.ticks.y       = element_blank()
    )
}

# Build all seven panels
# Panels 1 and 4 are left-column (show y-axis + shifted tag)
# Panel 7 is the lone bottom panel — also left-column for y-axis and tag
panels <- lapply(seq_along(panel_names), function(i) {
  name     <- panel_names[i]
  df_sub   <- df_plot %>% filter(clean_name == name)
  show_y   <- i %in% c(1, 4, 7)
  left_col <- i %in% c(1, 4, 7)
  make_panel(df_sub,
             show_yaxis = show_y,
             left_col   = left_col,
             title      = name,
             tag        = letters[i])
})

# Centered x-axis label placed below the grid
xlab_plot <- ggplot() +
  theme_void() +
  annotate("text", x = 0.5, y = 0.5,
           label    = "Posterior slope estimate",
           size     = 2.4,
           fontface = "bold",
           hjust    = 0.5,
           vjust    = 0.5)

empty <- ggplot() + theme_void()

# Rows 1 and 2 are full 3-panel; row 3 has one centered panel flanked by
# empty spacers — widths kept equal so the single panel aligns to column 2
row1 <- panels[[1]] | panels[[2]] | panels[[3]]
row2 <- panels[[4]] | panels[[5]] | panels[[6]]
row3 <- (empty | panels[[7]] | empty) + plot_layout(widths = c(.0002, 1, 1.118))

grid_plot <- wrap_plots(row1, row2, row3, ncol = 1) +
  plot_layout(heights = c(1, 1, 1)) &
  theme(plot.margin = margin(3, 4, 3, 4))

# Stack grid above centered x-axis label
final_plot <- cowplot::plot_grid(
  grid_plot,
  xlab_plot,
  ncol        = 1,
  rel_heights = c(1, 0.07)
)

ggsave(
  filename = paste0("../images/SI/posteriors_partial_slopes_SI.pdf"),
  plot     = final_plot,
  device   = cairo_pdf,
  width    = 5,
  height   = 3,
  units    = "in"
)
```


## SI Figure: Multicollinearity check

Create a correlation matrix for the explanatory variables to test for collinearity. On the plot, the upper triangle represents the shade of correlation on the color scale, while the bottom triangle displays the correlation between variables.
 
```{r multicollinearity check}
# Clean variable names
var_labels <- c(
  coastline_km               = "Coastline distance (km)",
  gdp_trade                  = "GDP from trade (%)",
  inundation_area_log1000km2 = "Inundated area (log 1000 km²)",
  log_gdp                    = "GDP (log scale)",
  pop_10k_pct_coast          = "Pop. within 10 km of coast (%)",
  supermarkets               = "Supermarkets per 100,000"
)

# var_labels <- c(
#   eez_total                   = "EEZ area",
#   gdp_trade                  = "GDP from trade (%)",
#   inundation_area_log1000km2 = "Inundated area (log 1000 km²)",
#   log_gdp                    = "GDP (log scale)",
#   pop_10k_pct_coast          = "Pop. within 10 km of coast (%)",
#   supermarkets               = "Supermarkets per 100,000"
# )

# Use the realm model setup
model_type = "realm"

# Read in explanatory variables to assess collinearity
X <- readRDS(file.path(base_path, paste0("X_", response_vars[i], ".rds")))

# Rename columns to clean labels before computing correlation
# so axis-adjacent labels are publication-ready throughout
colnames(X) <- var_labels[colnames(X)]

# Create correlation matrix
cor_mat <- cor(X, use = "pairwise.complete.obs")

# Helper functions to split matrix into triangles and diagonal
get_lower_tri <- function(m) { m[upper.tri(m)] <- NA; m }
get_upper_tri <- function(m) { m[lower.tri(m)] <- NA; m }
get_diag      <- function(m) { m[!diag(nrow(m))] <- NA; m }

# Melt each part
lower_melt <- melt(get_lower_tri(cor_mat), na.rm = TRUE)
upper_melt <- melt(get_upper_tri(cor_mat), na.rm = TRUE) %>%
  filter(Var1 != Var2) %>%
  mutate(label     = sprintf("%.2f", value))
diag_melt  <- melt(get_diag(cor_mat), na.rm = TRUE) %>%
  mutate(label = str_wrap(as.character(Var1), width = 14))

# Collinearity plot
plot <- ggplot() +

  # Lower triangle — filled heatmap tiles
  geom_tile(
    data  = lower_melt,
    aes(x = Var2, y = Var1, fill = value),
    color = "white",
    linewidth = 0.4
  ) +

  # Diagonal — white tiles with border, holds variable name labels
  geom_tile(
    data      = diag_melt,
    aes(x = Var2, y = Var1),
    fill      = "white",
    color     = "grey40",
    linewidth = 0.4
  ) +

  # Upper triangle — correlation values, coloured to match heatmap
  geom_text(
    data  = upper_melt,
    aes(x = Var2, y = Var1, label = label, color = value),
    size  = 2.2,
    fontface = "plain"
  ) +

  # Diagonal variable name labels
  geom_text(
    data     = diag_melt,
    aes(x = Var2, y = Var1, label = label),
    size     = 1.9,
    fontface = "bold",
    color    = "grey15",
    lineheight = 0.9
  ) +

  # Fill scale — aligned with project diverging palette
  scale_fill_gradient2(
    low      = "#2B6AA8",   # blue — strong negative correlation
    mid      = "#F7F7F7",   # near-white midpoint
    high     = "#B2182B",   # red — strong positive correlation
    midpoint = 0,
    limits   = c(-1, 1),
    name     = "Pearson\ncorrelation",
    guide    = guide_colorbar(
      title.position = "top",
      title.hjust    = 0.5,
      barheight      = unit(0.88, "npc"),
      barwidth       = unit(0.35, "cm"),
      frame.colour   = "grey40",
      ticks.colour   = "grey40"
    )
  ) +

  # Text colour scale — matches fill so values read clearly against tiles
  scale_color_gradient2(
    low      = "#2B6AA8",
    mid      = "#888888",   # mid-grey so near-zero values are still legible
    high     = "#B2182B",
    midpoint = 0,
    limits   = c(-1, 1),
    guide    = "none"
  ) +

  coord_fixed() +
  scale_x_discrete(limits = rev) +
  scale_y_discrete(limits = levels(factor(lower_melt$Var1))) +
  theme_minimal(base_size = 8) +
  theme(
    axis.title        = element_blank(),
    axis.text         = element_blank(),
    axis.ticks        = element_blank(),
    panel.grid        = element_blank(),
    legend.title      = element_text(size = 6.5, face = "bold"),
    legend.text       = element_text(size = 6),
    plot.background   = element_rect(fill = "transparent", color = NA),
    panel.background  = element_rect(fill = "transparent", color = NA),
    plot.margin       = margin(4, 4, 4, 4)
  )

# Save
ggsave(
  filename = "../images/SI/collinearity_plot_SI.pdf",
  plot     = plot,
  device   = cairo_pdf,
  width    = 5,
  height   = 4,
  units    = "in"
)
```

# Bayesian model assessment

## Trace Plot

```{r trace plot}
# Create trace plots for the hierarchcical Bayesian model

# Create separate datasets for each chain
model_type = "realm"
trace_df <- data.frame()

for (i in seq_along(response_vars))  {
  
  # Load in Stan data
  fit <- rstan::extract(readRDS(paste0("../output/bayesian_output/", model_type, "/fit_", response_vars[i],".stan")), permuted = T)
  
  chain_1 <- data.frame(beta = fit$beta[1:3000,1],
           beta_0 = fit$beta_0[1:3000],
           sigma = fit$sigma[1:3000],
           beta_0i = fit$beta_0i[1:3000,1],
           z_beta_0 = fit$z_beta_0[1:3000,1],
           model = response_vars[i],
           iteration = 1:3000,
           chain = "1")
  
  chain_2 <- data.frame(beta = fit$beta[3001:6000,1],
           beta_0 = fit$beta_0[3001:6000],
           sigma = fit$sigma[3001:6000],
           beta_0i = fit$beta_0i[3001:6000,1],
           z_beta_0 = fit$z_beta_0[3001:6000,1],
           model = response_vars[i],
           iteration = 1:3000,
           chain = "2")
  
  chain_3 <- data.frame(beta = fit$beta[6001:9000,1],
           beta_0 = fit$beta_0[6001:9000],
           sigma = fit$sigma[6001:9000],
           beta_0i = fit$beta_0i[6001:9000,1],
           z_beta_0 = fit$z_beta_0[6001:9000,1],
           model = response_vars[i],
           iteration = 1:3000,
           chain = "3")
  
  chains <- chain_1 %>% bind_rows(chain_2) %>% bind_rows(chain_3)
  
  trace_df <- bind_rows(trace_df, chains)
    
}

(plot <- trace_df %>%
  pivot_longer(cols = -c(model, iteration, chain), names_to = "parameter") %>%
  ggplot(aes(x = iteration, y = value, color = chain)) +
  geom_line() +
  scale_color_viridis_d(option = "plasma", end = 0.95) +
  facet_grid(model ~ parameter) +
  guides(color = "none") +
  labs(x = "Parameter", y = "Parameter value") +
  theme_pubclean())

# ggsave(paste0("../images/bayesian plots/", model_type, "/trace.jpg"), plot = plot, device = "jpg", width = 18, height = 30, units = "in", dpi = 300)
```

## Hierarchical Bayesian model R^2

```{r calculate and visualize R^2}
# Calcualte the hierarchical R^2 across models where we disaggregate the total R^2 to the fixed and random effects. We will also include a combined effect. 
R2_df <- data.frame(
  fixed_r2    = numeric(0),
  random_r2   = numeric(0),
  combined_r2 = numeric(0),
  model       = character(0)
)

# use the realm model setup
model_type = "realm"

# Iterate through each model to extract R^2
for (i in seq_along(response_vars))  {
  
fit <- rstan::extract(readRDS(paste0("../output/bayesian_output/", model_type, "/fit_", response_vars[i],".stan")), permuted = T)
  
  # Z: N x 16 realm indicator matrix (1 for realm membership)
b_s <- fit$beta_0i[1:3000,]   # iterations x 16

realm_ids <- readRDS(paste0("../output/bayesian_output/", model_type, "_ids.rds"))

X <- readRDS(paste0("../output/bayesian_output/", model_type, "/X_",response_vars[i],".rds"))
realm_ids <- readRDS(paste0("../output/bayesian_output/", model_type, "/ids_",response_vars[i],".rds"))

Z <- model.matrix(~ 0 + factor(realm_ids))

# random-only linear predictor: iterations x N
mu_random <- b_s %*% t(Z)


beta_s <- fit$beta[1:3000, ]   # 8000 × K
mu_fixed <- beta_s %*% t(X)    # 8000 × N

var_fixed_s  <- apply(mu_fixed,  1, var)   # length = iterations
var_random_s <- apply(mu_random, 1, var)   # length = iterations
var_resid_s  <- fit$sigma^2               # or fit$sigma^2 depending on your model

mu_total     <- mu_fixed + mu_random
var_total_s  <- apply(mu_total, 1, var)

R2_fixed_s   <- var_fixed_s  / (var_total_s + var_resid_s)
R2_random_s  <- var_random_s / (var_total_s + var_resid_s)
R2_cond_s    <- var_total_s  / (var_total_s + var_resid_s)

# Create new data frame with Current iteration R2 values
new_data <- data.frame(fixed_r2 = R2_fixed_s, random_r2 = R2_random_s, combined_r2 = R2_cond_s, model = response_vars[i])

# Add on R2 values to R2 data frame
R2_df <- bind_rows(R2_df, new_data)
  
}

# Plot all model R^2
(plot <- R2_df %>%
  pivot_longer(cols = c("fixed_r2", "random_r2", "combined_r2")) %>%
  mutate(model = factor(model, levels = ordered_names),
         name = case_when(name == "combined_r2" ~ "combined",
                          name == "fixed_r2" ~ "fixed",
                          name == "random_r2" ~ "random")) %>%
  ggplot(aes(x = value, fill = name)) +
  geom_histogram(position = "identity", alpha = 0.7) +
  facet_wrap(~ model) +
  theme_pubclean() +
  scale_fill_viridis_d(option = "plasma", end = 0.95) +
  labs(x = bquote("Posterior" ~ R^2), y = "# iterations", fill = "Effect type") +
  theme(legend.position = "bottom",
        strip.text.x = element_text(margin = margin(0.05,0,0.05,0, "cm"), size = 6),
        axis.text = element_text(size = 6)))

ggsave(paste0("../images/bayesian plots/", model_type, "/r2.jpg"), plot = plot, device = "jpg", width = 9, height = 5, units = "in", dpi = 300)
```

## R hat

```{r r hat}
model_type = "continental"
rhat_df <- data.frame()
for (i in seq_along(response_vars)) {
  fit <- readRDS(paste0("../output/bayesian_output/", model_type,
                     "/fit_", response_vars[i], ".stan"))
  

  if (model_type == "simple") {
    
    vals <- rstan::summary(fit, pars = c("beta", "beta_0", "sigma", "lp__"))$summary[, "Rhat"]
    
  } else if (model_type != "simple") {
    
    vals <- rstan::summary(fit, pars = c("beta_0", "beta", "sigma", "beta_0i", "sigma_beta_0", "z_beta_0", "lp__"))$summary[, "Rhat"]
    
  }
  # Add iteration i rhat values to total dataframe for plotting
  new_data <- data.frame(rhat = vals, model = response_vars[i])
  rhat_df <- bind_rows(rhat_df, new_data)
}

# Create rhat plot
(plot <- rhat_df %>%
  ggplot(aes(x = rhat)) +
  geom_histogram() +
  facet_wrap(~ model) +
  labs(x = "R-hat", y = "# parameters") +
  theme_pubclean() +
    theme(axis.text = element_text(size = 7)))

# Save plot
ggsave(paste0("../images/bayesian plots/", model_type, "/rhat.jpg"), plot = plot, device = "jpg", width = 7, height = 5, units = "in", dpi = 300)
```

## Divergent transitions

```{r divergent transitions}
# Change to either continental / realm
model_type = "simple"

# Rename the divergent transition column names to model names
divergent_transitions <- matrix(nrow = 3, ncol = 23)
colnames(divergent_transitions) <- response_vars

for (i in seq_along(response_vars)) {
  
  # Extract number of divergent samples
  sampler_params <- get_sampler_params(readRDS(paste0("../output/bayesian_output/", model_type, "/fit_", response_vars[i],".stan")), inc_warmup = FALSE)
  num_transitions <- sapply(sampler_params, function(x) sum(x[, "divergent__"]))
  
  divergent_transitions[,i] <- num_transitions
  
}

# Create dataframe for plotting divergent transitions
plot_df <- divergent_transitions %>%
  as.data.frame() %>%
  mutate(chain = as.character(1:3)) %>%
  pivot_longer(!"chain", names_to = "model", values_to = "divergent") %>%
  mutate(
    suffix = sub(".*_", "", model)
  )

# compute the order of model names by suffix
ordered_names <- plot_df %>%
  distinct(model, suffix) %>%
  arrange(suffix) %>%
  pull(model)

# Plot # divergent transitions
(plot <- plot_df %>%
  mutate(model = factor(model, levels = ordered_names)) %>%
  ggplot(aes(y = model, x = divergent, fill = chain)) +
  geom_col() +
  scale_fill_viridis_d(end = 0.8) +
  labs(y = "Model", x = "# divergent transitions") +
  theme_light() +
    theme(legend.position = "bottom"))

ggsave(paste0("../images/bayesian plots/", model_type, "/divergent_samples.jpg"), plot = plot, device = "jpg", width = 7, height = 5, units = "in", dpi = 300)
``` 

## Generate Posterior Predictive distribuitons

This code chunk will only run if not run before as it takes awhile to run.

```{r ppd distributions}
if ( !file.exists(paste0("../output/bayesian_output/", model_type, "/posterior_predictive_all_models.parquet")) ) {
  
  model_type = "realm"
# Load in function to clean the model output names from snake case to a publication-friendly read
source("../R/clean_model_name.R")

# Function to extract posterior predictive draws for a model.
extract_ppd <- function(model_name, model_type) {

  # Load fitted Stan model
  fit <- readRDS(
    paste0("../output/bayesian_output/", model_type, "/fit_", model_name, ".stan")
  )

  x <- rstan::extract(fit, permuted = TRUE)

  # Posterior predictive draws (S × N)
  y_pred <- x$y_pred

  # Load X and y to get N
  X <- readRDS(
    paste0("../output/bayesian_output/", model_type, "/X_", model_name, ".rds")
  )
  y <- readRDS(
    paste0("../output/bayesian_output/", model_type, "/y_", model_name, ".rds")
  )

  N <- length(y)

  # Load realm names (vector of length N)
  realm_levels <- readRDS(
    paste0("../output/bayesian_output/", model_type, "/realms_", model_name, ".rds")
  )

  # Safety check
  if (length(realm_levels) != N) {
    stop(paste0("realm_levels length mismatch for model ", model_name))
  }

  # Format long
  ppd_long <- as.data.frame(y_pred) %>%
    mutate(draw = row_number()) %>%
    pivot_longer(-draw, names_to = "obs", values_to = "ppd") %>%
    mutate(
      obs = as.integer(gsub("V", "", obs)),
      region = realm_levels[obs],
      y_obs = y[obs],
      model = model_name
    )

  return(ppd_long)
}


# Loop over all models and bind results
ppd_all_models <- map_dfr(response_vars, ~ extract_ppd(.x, model_type)) %>%
  mutate(clean_name = clean_model_name(model)) # Add column for clean model names


# Calculate posterior predictive means
ppd_means <- ppd_all_models %>%
  group_by(model, obs, region) %>%
  summarise(
    ppd_mean = mean(ppd),
    y_obs    = first(y_obs),
    .groups  = "drop"
  )

# Save combined posterior predictive dataset
write_parquet(
  ppd_all_models,
  paste0("../output/bayesian_output/", model_type, "/posterior_predictive_all_models.parquet")
)
  
}
```

## Plot posterior predictive distributions

```{r create ppd plot}
# Read in posterior predictive distrbituion data
ppd_all_models <- read_parquet(
  paste0("../output/bayesian_output/", model_type, "/posterior_predictive_all_models.parquet")
)

# Get letters for panel numbers
panel_tags <- letters[1:15] 

# Function to draw a single panel ---
make_ppd_panel <- function(df, title, tag = "") {
  ggplot(df, aes(x = ppd, group = draw)) +
    geom_density(alpha = 0.08, color = "grey60") +
    geom_density(aes(x = y_obs), color = "black", linewidth = 1.1) +
    labs(y = "Density") +
    annotate(
      "text",
      x = -Inf, y = Inf,
      label = tag,
      hjust = -0.3, vjust = 1.3,
      size = 3, fontface = "bold"
    ) +
    theme_minimal(base_size = 8) +
    theme(
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.y = element_line(color = scales::alpha("black", 0.25), linewidth = 0.3),
      panel.grid.minor.y = element_blank(),
      panel.background = element_rect(fill = "transparent", color = NA),
      plot.background = element_rect(fill = "transparent", color = NA),
      strip.background = element_blank(),
      strip.text = element_blank(),
      axis.text = element_text(size = 6, color = "black"),
      axis.title.y = element_text(size = 7, margin = margin(r = 5), face = "bold"),
      axis.text.x = element_text(size = 5.5, angle = 30, hjust = 1),
      axis.title.x = element_blank()
    )
}

#  Split df_plot into 9 panels
xlab_plot <- ggplot() +
  theme_void() +
  annotate("text", x = 0.5, y = 0.5,
           label = "Response value",
           size = 2.5, fontface = "bold")

draw_sample <- sample(unique(ppd_all_models$draw), size = 30)

df_plot <- ppd_all_models %>%
  filter(
    model != "animal_cons_cap",
    str_detect(clean_name, "capita") |
      str_detect(clean_name, "Aquatic reliance") |
      str_detect(clean_name, "reliance")
  ) %>%
  filter(draw %in% draw_sample) %>%
  mutate(clean_name = factor(clean_name))

panel_names <- levels(df_plot$clean_name)

layout_order <- c(15, 3, 5,
                  1, 7, 9,
                  13, 11, 4,
                  2, 6, 8,
                  10, 14, 12)

panel_tags <- letters[1:15]

# assign tags in layout order
tags_by_panel_index <- rep("", length(panel_names))
tags_by_panel_index[layout_order] <- panel_tags

# Create panel plots
panels <- lapply(seq_along(panel_names), function(i) {
  nm <- panel_names[i]
  df_sub <- df_plot %>% filter(clean_name == nm)
  make_ppd_panel(df_sub, nm, tag = tags_by_panel_index[i])
})


# Create empty plots for bottom row in the combined plot (so final plot that sits alone at the bottom can be centered)
empty <- ggplot() + theme_void()

# Create panels that will display in the ppd plot
row1 <- panels[[15]]  | panels[[3]]  | panels[[5]]
row2 <- panels[[1]]  | panels[[7]]  | panels[[9]]
row3 <- panels[[13]]  | panels[[11]]  | panels[[4]]
row4 <- panels[[2]] | panels[[6]] | panels[[8]]
row5 <- panels[[10]] | panels[[14]] | panels[[12]]
row6 <- empty | xlab_plot | empty

(final_plot <- wrap_plots(
  row1, row2, row3, row4, row5, row6,
  ncol = 1
) + plot_layout(heights = c(1,1,1,1,1,0.15)))

ggsave(paste0("../images/SI/ppd_SI.jpg"), plot = final_plot, device = "jpg", width = 4, height = 5, units = "in", dpi = 300)
```
