---
title: "02.1_Supplementary_Information"
author: "Connor Quiroz and Jessica Gephart"
date: "2026-03-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(exploreARTIS)
library(cowplot)
library(ggrepel)
library(countrycode)
library(colorspace)
library(ggpubr)
library(arrow)
library(rstan)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggridges)
library(bayesplot)
library(purrr)
library(patchwork)
library(reshape2)
```

# Chapter 1 Supplementary information

## Read in data

### Read in aquatic reliance data & FAO population data

```{r read in FAO derived data}
# Read in population
population_fao_wb <- read.csv("../output/population_fao_wb.csv")

# Read in aquatic reliance data
aquatic_reliance <- read.csv("../output/fao_aquatic_reliance_source.csv") %>%
  left_join(population_fao_wb,by=c("year","iso3c"))%>%
  mutate(protein_consumed_kg_percap=protein_consumed_t*1000/population_fao)

# Read in FAO sourcing data
aquatic_fbs <- read.csv("../output/fao_fbs_aquatic_source.csv")

# Read in list of response variables used in Bayesian models
response_vars <- readRDS("../output/response_vars.rds")
```

### Get list of response variables

```{r acquire animal consumption data to be fed through analysis}
# Calculate animal consumption
terr_animal_cons_cap <- aquatic_reliance %>%
  filter(year == 2019) %>%
  group_by(iso3c) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))

animal_cons_cap <- aquatic_reliance %>%
  filter(year == 2019) %>%
  group_by(iso3c) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))

# Calculate aquatic animal consumption
aquatic_animal_cons_cap <- aquatic_reliance %>%
  filter(year == 2019, food_group == "aquatic") %>%
  group_by(iso3c) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))

# Calculate animal reliance consumption
aquatic_rel <- aquatic_animal_cons_cap %>%
  left_join(animal_cons_cap, by = "iso3c") %>%
  group_by(iso3c) %>%
  summarize(prop_animal_protein = protein_consumed_kg_percap.x / protein_consumed_kg_percap.y)

# Add items not created in the loop to the list
response_vars <- c("terr_animal_cons_cap","animal_cons_cap", "aquatic_animal_cons_cap", "aquatic_rel", "aquatic_protein_roc", "terrestrial_protein_roc")

for (i in c("capture", "aquaculture", "domestic", "foreign", "marine", "inland")) {
  
  # Calculate consumption per capita by method
  if (i %in% c("capture", "aquaculture")) {
    assign(
      paste0(i, "_cap"),
      aquatic_reliance %>%
        filter(year == 2019, food_group == "aquatic", method == i) %>%
        group_by(iso3c) %>%
        summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))
    )
  }
  
  # Calculate consumption per capita by consumption source
  if (i %in% c("domestic", "foreign")) {
    assign(
      paste0(i, "_cap"),
      aquatic_reliance %>%
        filter(year == 2019, food_group == "aquatic", consumption_source == i) %>%
        group_by(iso3c) %>%
        summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))
    )
  }
  
  # Calculate consumption per capita by habitat
  if (i %in% c("marine", "inland")) {
    assign(
      paste0(i, "_cap"),
      aquatic_reliance %>%
        filter(year == 2019, food_group == "aquatic", habitat == i) %>%
        group_by(iso3c) %>%
        summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))
    )
  }
  
  response_vars <- c(response_vars, paste0(i, "_cap"))
  
  # Calculate reliance by method, production source, or habitat
  assign(
    paste0(i, "_rel"),
    get(paste0(i, "_cap")) %>%
      left_join(animal_cons_cap, by = "iso3c") %>%
      group_by(iso3c) %>%
      summarize(prop_animal_protein = protein_consumed_kg_percap.x / protein_consumed_kg_percap.y)
  )
  
  response_vars <- c(response_vars, paste0(i, "_rel"))
  
  # Calculate production sources' proportion of total aquatic animal consumption per capita
  assign(
    paste0(i, "_prop_of_aquatic_cons"),
    get(paste0(i, "_cap")) %>%
      left_join(aquatic_animal_cons_cap, by = "iso3c") %>%
      group_by(iso3c) %>%
      summarize(aquatic_source_prop = protein_consumed_kg_percap.x / protein_consumed_kg_percap.y)
  )
  
  response_vars <- c(response_vars, paste0(i, "_prop_of_aquatic_cons"))
}

#calculate rate of change in aquatic protein quantity
aquatic_protein_roc <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, year, method) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap, na.rm = TRUE)) %>%
  distinct() %>%
  filter(method == "capture") %>%
  split(.$iso3c) %>%
  map(~lm(protein_consumed_kg_percap~year, data = .x)) %>%
  map_df(broom::tidy, .id = 'iso3c') %>%
  filter(term == 'year') %>%
  select(c("iso3c", "aquatic_estimate" = "estimate"))

#calculate rate of change in terrestrial protein quantity
terrestrial_protein_roc <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "terrestrial") %>%
  group_by(iso3c, year) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap, na.rm = TRUE)) %>%
  distinct() %>%
  split(.$iso3c) %>%
  map(~lm(protein_consumed_kg_percap~year, data = .x)) %>%
  map_df(broom::tidy, .id = 'iso3c') %>%
  filter(term == 'year') %>%
  select(c("iso3c", "terrestrial_estimate" = "estimate"))
```

# RQ 1

## 1A

### Individual continent production contribution (i.e., disaggregate reliance) to rate of change in aggregate reliance + other exploratory metrics.

```{r individual continent production contribution}
# Fit linear models to estimate the average annual change in disaggregated
# aquatic animal reliance (2010-2020) by production source (habitat x method)
# for each country
disagg <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "aquatic") %>%
  # group_by(iso3c) %>%
  # mutate(n = n_distinct(year)) %>%
  # filter(n >= 3) %>%
  group_by(year, iso3c, habitat, method) %>%
  mutate(prop_animal_protein = sum(prop_animal_protein)) %>%
  group_by(iso3c, habitat, method) %>%
  # filter(n() >= 3) %>%
  nest() %>%
  mutate(
    model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
    tidy  = map(model, broom::tidy)
  ) %>%
  unnest(tidy) %>%
  filter(term == "year") %>%
  mutate(
    consource_habitat_method = paste(habitat, method),
    group_label = consource_habitat_method
  ) %>%
  drop_na(estimate) %>%
  filter(!str_detect(consource_habitat_method, "unknown")) %>%
  group_by(group_label) %>%
  mutate(pct_above = mean(estimate > 0) * 100) %>%
  # mutate(estimate = round(estimate, 2)) %>%
  ungroup()

# Fit linear models to estimate the average annual change in aggregate
# aquatic animal reliance (2010-2020)
agg <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "aquatic") %>%
  # group_by(iso3c) %>%
  # mutate(n = n_distinct(year)) %>%
  # filter(n >= 3) %>%
  group_by(iso3c, year) %>%
  summarise(prop_animal_protein = sum(prop_animal_protein), .groups = "drop") %>%
  group_by(iso3c) %>%
  nest() %>%
  mutate(
    model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
    tidy  = map(model, broom::tidy)
  ) %>%
  unnest(tidy) %>%
  filter(term == "year") %>%
  mutate(group_label = "aggregate") %>%
  drop_na(estimate) %>%
  group_by(group_label) %>%
  mutate(pct_above = mean(estimate > 0) * 100) %>%
  ungroup()

# Clean aggregate estimate data
agg_estimate <- bind_rows(agg, disagg) %>%
  group_by(iso3c) %>%
  mutate(agg_est = estimate[group_label == "aggregate"][1]) %>%
  add_region("iso3c", "region") %>%
  ungroup() %>%
  mutate(
    iso3c  = fct_reorder(iso3c, -agg_est),
    is_agg = group_label == "aggregate"
  ) %>%
  filter(group_label == "aggregate") %>%
  select(iso3c, agg_estimate = estimate)

# Set plot title size
axis.title.size <- 8
for (i in c("Asia", "Africa", "Europe", "North America", "South America", "Oceania")) {

# Plot disaggregated aquatic animal reliance by production source (2010-2020)
# for each country in Asia, ordered by aggregate rate of change 
plot <- disagg %>%
  group_by(iso3c) %>%
  add_region("iso3c", "region") %>%
  ungroup() %>%
  left_join(agg_estimate, by = "iso3c") %>%
  mutate(iso3c = fct_reorder(iso3c, -agg_estimate)) %>%
  filter(region == i) %>%
  ggplot() +
  geom_col(aes(x = estimate, y = iso3c, fill = group_label)) +
  geom_point(aes(x = agg_estimate, y = iso3c), shape = 16) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.4) +
  scale_alpha_identity() +
  scale_size_identity() +
  scale_fill_manual(values = c("#D8C84A","#E68A2E", "#7A5FA3", "#3A77B5")) +
  guides(fill = guide_legend(ncol = 2)) +
  theme(axis.text.y = element_text(size = 4.75),
        axis.title.x = element_text(size = axis.title.size, face = "bold"),
    panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(color = scales::alpha("black", 0.05), linewidth = 0.3),
    panel.grid.major.x = element_line(color = scales::alpha("black", 0.25), linewidth = 0.3),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),
    axis.text = element_text(size = 6, color = "black"),
    axis.title.y = element_blank(),
    legend.position = "bottom",
    legend.title = element_text(size = 7, face = "bold"),
    legend.text = element_text(size = 6),
    legend.key.size = unit(0.3, "cm"),
    legend.box = "vertical",
    legend.spacing = unit(0.1, "cm"),
    legend.margin = margin(t = 0, b = 0),
    legend.box.margin = margin(t = -2, b = -2),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA)) +
      labs(fill = "Production\nsource",
           x = "Average annual change in aquatic animal reliance (2010–2020)")

# Lana's analysis calcuating the # of countries with increasing reliance and decreasing relaince from 2010-2020.

# Extract rate of change in aggregate aquatic animal reliance from 2010-2020 
roc_aquatic_total <- aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, year) %>%
  summarize(prop_animal_protein = sum(prop_animal_protein, na.rm = TRUE), .groups = "drop") %>%
  distinct() %>%
  split(list(.$iso3c), drop = TRUE) %>%
  map(~lm(prop_animal_protein~year, data = .x)) %>%
  map_df(broom::tidy, .id = 'combined_id') %>%
  filter(term == 'year') %>%
  select(c("iso3c"="combined_id", "roc" = "estimate"))%>%
  mutate(direction = case_when(roc > 0 ~ "inc", roc < 0 ~ "dec"))%>%
  mutate(method_habitat = "total")%>%
  add_region(col = "iso3c", region.col.name = "continent")

# Get the counts by continent
Counts_by_continent<-roc_aquatic_total%>%
  group_by(continent, method_habitat)%>%
  count(direction)%>%
  pivot_wider(names_from = direction, values_from = n) %>%
  rename(decreasing = dec, increasing = inc) %>%
  pivot_longer(cols = c(decreasing, increasing), names_to = "direction",values_to = "count") %>%
  group_by(continent) %>%
  mutate(prop = count / sum(count))

# Plot counts by target continent
region_counts <- Counts_by_continent %>%
  filter(method_habitat == "total", continent == i) %>%
  ggplot(aes(y = continent, x = prop, fill = direction)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("#b2182b", "#2166ac"),
                    labels  = c("decreasing" = "Declining reliance",
              "increasing" = "Increasing reliance")) +
  guides(fill = guide_legend(nrow = 2)) +
  labs(x = "Proportion of countries with increasing\nor decreasing reliance (aggregate)", fill = "Direction of change") +
  # scale_y_continuous(labels = scales::label_number(accuracy = 1)) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text = element_text(size = 4.75),
        axis.title.y = element_blank(),
        axis.title.x = element_text(size = axis.title.size-2, face = "bold"),
    panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_line(color = scales::alpha("black", 0.25), linewidth = 0.3),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),
    legend.position = "bottom",
    legend.title = element_text(size = 7, face = "bold"),
    legend.text = element_text(size = 6),
    legend.key.size = unit(0.3, "cm"),
    legend.box = "vertical",
    legend.spacing = unit(0.1, "cm"),
    legend.margin = margin(t = 0, b = 0),
    legend.box.margin = margin(t = -2, b = -2),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    aspect.ratio = 0.15)

# Proportion of countries with increasing reliance by production source for each region
(disagg_props <- disagg %>%
  add_region("iso3c", "region") %>%
  filter(region == i) %>%
  group_by(region, group_label) %>%
  summarize(prop_greater_0 = mean(estimate > 0)) %>%
  mutate(other_prop = 1 - prop_greater_0) %>%
  pivot_longer(cols = c("other_prop", "prop_greater_0")) %>%
  ggplot(aes(y = group_label, x = value, fill = name)) +
  geom_col() +
  # geom_vline(xintercept = 0.5, linetype = "dashed") +
  scale_fill_manual(values = c("#b2182b", "#2166ac")) +
  guides(fill = "none")  +
  labs(x = "Proportion of countries with increasing\nor decreasing reliance (disaggregate)") +
  theme(axis.title.y = element_blank(),
      axis.text = element_text(size = 4.75),
      axis.title.x = element_text(size = axis.title.size-2, face = "bold"),
      panel.grid.minor.y = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.y = element_blank(),
      panel.grid.major.x = element_line(color = scales::alpha("black", 0.25), linewidth = 0.3),
      panel.background = element_rect(fill = "transparent", color = NA),
      plot.background = element_rect(fill = "transparent", color = NA)))

# Plot Aggregate aquatic animal reliance for target continent's region,
# ordered by the rate of change in aggregate aquatic animal reliance 
# from 2010-2020
regional_reliance <- aquatic_reliance %>%
    filter(year == 2020, food_group == "aquatic") %>%
    group_by(iso3c) %>%
    summarize(prop_animal_protein = sum(prop_animal_protein)) %>%
    left_join(agg_estimate, by = "iso3c") %>%
    add_region("iso3c", "region") %>%
    filter(region == i) %>%
    ggplot(aes(
      x = fct_reorder(iso3c, agg_estimate), y = prop_animal_protein * 100
    )) +
    geom_col(fill = "#555555") +
    theme(
      axis.text.x = element_text(size = 4.5, angle = 45),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 4.5),
      axis.title.y = element_text(size = axis.title.size, face = "bold"),
      panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(color = scales::alpha("black", 0.25), linewidth = 0.3),
      panel.background = element_rect(fill = "transparent", color = NA),
      plot.background = element_rect(fill = "transparent", color = NA),
      axis.text = element_text(size = 6, color = "black")
    ) +
    labs(y = "Aquatic animal\nreliance, 2020 (%)")


# Function to extract legends
extract_legend <- function(p) {
  legs <- cowplot::get_plot_component(p, "guide-box", return_all = TRUE)
  # Filter out empty (zeroGrob) legends
  good <- Filter(function(x) !inherits(x, "zeroGrob"), legs)
  if (length(good) == 0) return(NULL)
  good[[1]]
}

# Extract legends
legend_plot          <- extract_legend(plot)
legend_region_counts <- extract_legend(region_counts)

# Remove legends from panels
plot_nolegend          <- plot          + theme(legend.position = "none")
region_counts_nolegend <- region_counts + theme(legend.position = "none")

# Combinelegends
legend_row <- cowplot::plot_grid(
  legend_region_counts,
  legend_plot,
  ncol       = 2,
  rel_widths = c(1, 1)
)

title <- ggplot() +
  labs(title = i) +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

divider <- ggplot() +
  theme_void() +
  theme(panel.border = element_blank()) +
  geom_hline(yintercept = 0.5, linewidth = 0.5, color = "black")

# Row for global counts and disaggregated counts
top_plot_panel <- cowplot::plot_grid(region_counts_nolegend, disagg_props)
    

# Final plot with legends at the bottom
(final_plot <- cowplot::plot_grid(
  title,
  divider,                    # <-- line between title and panels
  top_plot_panel,
  regional_reliance,
  plot_nolegend,
  legend_row,
  ncol        = 1,
  rel_heights = c(0.15, 0.05, 0.65, 0.85, 2.5, 0.25)  # very small height for divider
))

# Save plot
ggsave(paste0("../images/SI/multi_metrics_",i,"_SI.jpg"), plot = final_plot, device = "jpg", units = "in", width = 5, height = 7, dpi = 500)

}
```

### Country level trends in delta 2010-2020 aquatic animal reliance

```{r Country level trends in delta 2010-2020 aquatic animal reliance}
# Fit linear models to estimate the average annual change in disaggregated
# aquatic animal reliance (2010-2020) by production source (habitat x method)
# for each country
disagg <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c) %>%
  mutate(n = n_distinct(year)) %>%
  filter(n >= 3) %>%
  group_by(iso3c, habitat, method) %>%
  filter(n() >= 3) %>%
  nest() %>%
  mutate(
    model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
    tidy  = map(model, broom::tidy)
  ) %>%
  unnest(tidy) %>%
  filter(term == "year") %>%
  mutate(
    consource_habitat_method = paste(habitat, method),
    group_label = consource_habitat_method
  ) %>%
  drop_na(estimate) %>%
  filter(!str_detect(consource_habitat_method, "unknown")) %>%
  group_by(group_label) %>%
  mutate(pct_above = mean(estimate > 0) * 100) %>%
  ungroup()

# Fit linear models to estimate the average annual change in aggregate
# aquatic animal reliance (2010-2020)
agg <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c) %>%
  mutate(n = n_distinct(year)) %>%
  filter(n >= 3) %>% # Get countries with only 3 years or more of data (so lm can properly fit points)
  group_by(iso3c, year, data_source, food_group) %>%
  summarise(prop_animal_protein = sum(prop_animal_protein)) %>%
  group_by(iso3c) %>%
  nest() %>%
  mutate(
    model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
    tidy  = map(model, broom::tidy)
  ) %>%
  unnest(tidy) %>%
  filter(term == "year") %>%
  mutate(
    group_label = "aggregate"
  ) %>%
  drop_na(estimate) %>%
  group_by(group_label) %>%
  mutate(pct_above = mean(estimate > 0) * 100) %>%
  ungroup()

# Combine aggregate and disaggregate slopes
all_slopes <- bind_rows(agg, disagg)

# Calculate the percent of countries that have slopes above and below 0.
all_slopes <- all_slopes %>%
  mutate(
    pct_below = 100 - pct_above,
    label_majority = if_else(
      pct_above >= 50,
      sprintf("%.0f%%", pct_above),
      sprintf("%.0f%%", pct_below)
    ),
    label_x = if_else(
      pct_above >= 50,
      0.002,   # right side, close to center
      -0.002   # left side, close to center
    ),
    label_hjust = if_else(pct_above >= 50, 0, 1),
    color = case_when(pct_above >= 50 ~ "#2166ac",
                      TRUE ~ "#b2182b")
  )

# Plot data
(plot <- ggplot(all_slopes, aes(
  x = estimate,
  y = fct_reorder(group_label, pct_above)
)) +
  geom_density_ridges_gradient(
    aes(fill = after_stat(x > 0)),
    scale = 1.4,
    rel_min_height = 0.01,
    color = "black"
  ) +
  scale_fill_manual(
    values = c(`TRUE` = "#2166ac", `FALSE` = "#b2182b")
  ) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_text(
    aes(x = label_x, label = label_majority, hjust = label_hjust),
    color = all_slopes$color,
    vjust = -2,
    size = 2.5
  ) +
  guides(fill = "none") +
  labs(x = expression(bold(Delta[2010-2020] ~ "Aquatic animal protein reliance %")), y = "Production source") +
  # xlim(c(-0.0105, 0.0105)) +

  # --- Matching theme from aa_reliance_multiple_trends ---
  theme_minimal(base_size = 8) +
  theme(
    # Grid and background
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(color = scales::alpha("black", 0.25), linewidth = 0.3),
    panel.grid.minor.y = element_blank(),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),

    # Axis styling
    axis.text = element_text(size = 6, color = "black"),
    axis.title.x = element_text(size = 7, face = "bold", margin = margin(t = 5)),
    axis.title.y = element_text(size = 7, face = "bold"),

    # Ridge spacing
    axis.text.y = element_text(size = 6),

    # No legend
    legend.position = "none"
  ))



# ggsave("../images/pct_ichange_reliance.jpg", plot = plot, device = "jpg", units = "in", width = 5, height = 4)

cor(joined_df$eez_total,joined_df$coastline_km)
```

# RQ 2

## What biophysical and socioeconomic factors are associated with reliance on farmed and wild aquatic animals for protein and overall consumption?

```{r}
model_type = "realm"

# Create data frame to hold all posteriors across variables
slopes_df <- data.frame()

for (i in seq_along(response_vars)) {
  
  # Extract MCMC samples
  x <- rstan::extract(readRDS(paste0("../output/bayesian_output/", model_type, "/fit_", response_vars[i],".stan")), permuted = T)
  beta_global <- x$beta       # iter × K (slopes)
  
  X <- readRDS(paste0("../output/bayesian_output/", model_type, "/X_", response_vars[i],".rds"))
  
  n_iter <- nrow(beta_global)
  K <- ncol(beta_global)
  vars <- colnames(X)

# Extract global slopes
post_long_slopes <- expand_grid(
  iter = 1:n_iter,
  k = 1:K
) %>%
  mutate(
    value = map2_dbl(iter, k, ~beta_global[.x, .y]),
    var = vars[k]
  )

# Calculate credibility intervals
ci_df <- post_long_slopes %>%
  group_by(var) %>%
  summarise(
    lo95 = quantile(value, 0.025),
    lo80 = quantile(value,0.1),
    lo50 = quantile(value, 0.25),
    med = quantile(value, 0.5),
    hi50 = quantile(value, 0.75),
    hi80 = quantile(value,0.9),
    hi95 = quantile(value, 0.975),
    .groups = "drop"
  ) %>%
  mutate(model =  response_vars[i])

slopes_df <- slopes_df %>%
  bind_rows(ci_df)

}

clean_model_name <- function(x) {

  # Suffix dictionary
  suffix_map <- c(
    cap = "consumption per capita",
    rel = "reliance",
    roc = "rate of change",
    prop_of_aquatic_cons = "share of aquatic foods"
  )

  # Domain dictionary
  domain_map <- c(
    terr_animal = "Terrestrial animal",
    terrestrial = "Terrestrial",
    animal = "Animal",
    aquatic_animal = "Aquatic animal",
    aquatic = "Aquatic",
    capture = "Capture",
    aquaculture = "Aquaculture",
    domestic = "Domestic",
    foreign = "Foreign",
    marine = "Marine",
    inland = "Inland"
  )

  # Metric dictionary
  metric_map <- c(
    cons = "consumption",
    protein = "protein"
  )

  # Split name
  parts <- strsplit(x, "_")[[1]]

  # Detect suffix safely
  if (endsWith(x, "prop_of_aquatic_cons")) {
    suffix <- "prop_of_aquatic_cons"
    core <- parts[1:(length(parts)-4)]
  } else {
    suffix <- parts[length(parts)]
    core <- parts[1:(length(parts)-1)]
  }

  suffix_clean <- suffix_map[[suffix]]

  # Detect domain (first one or two parts)
  if (length(core) >= 2 && paste(core[1:2], collapse = "_") %in% names(domain_map)) {
    domain <- paste(core[1:2], collapse = "_")
    rest <- core[-(1:2)]
  } else {
    domain <- core[1]
    rest <- core[-1]
  }

  domain_clean <- domain_map[[domain]]

  # Detect metric (optional)
  if (length(rest) >= 1 && rest[1] %in% names(metric_map)) {
    metric_clean <- metric_map[[rest[1]]]
  } else {
    metric_clean <- NULL
  }

  # Build label
  if (!is.null(metric_clean) && suffix != "prop_of_aquatic_cons") {
    paste(domain_clean, metric_clean, suffix_clean)
  } else {
    paste(domain_clean, suffix_clean)
  }
}

var_labels <- c(
  coastline_km = "Coastline length (km)",
  gdp_trade = "GDP from trade (%)",
  inundation_area_log1000km2 = "Inundated area (log 1000 km²)",
  log_gdp = "GDP (log scale)",
  prop_pop_coast = "Population within 10 km of coast (%)",
  supermarkets = "Supermarkets per 100,000 people"
)

theme_tile_pub <- function() {
  theme_minimal(base_size = 8) +
    theme(
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.y = element_line(color = scales::alpha("black", 0.25), linewidth = 0.3),
      panel.grid.minor.y = element_blank(),
      panel.background = element_rect(fill = "transparent", color = NA),
      plot.background = element_rect(fill = "transparent", color = NA),
      axis.text.x = element_text(size = 6,angle = 45, vjust = 0.5, color = "black"),
      axis.text.y = element_text(size = 6, color = "black"),
      axis.title = element_text(size = 7, face = "bold"),
      legend.title = element_text(size = 7, face = "bold"),
      legend.text = element_text(size = 6),
      legend.key.size = unit(0.3, "cm"),
      legend.box = "vertical",
      legend.spacing = unit(0.1, "cm"),
      legend.margin = margin(t = 0, b = 0),
      legend.box.margin = margin(t = -2, b = -2),
      legend.background = element_rect(fill = "transparent", color = NA),
      legend.box.background = element_rect(fill = "transparent", color = NA),
      plot.title = element_text(size = 8, face = "bold", hjust = 0.5)
    )
}

library(colorspace)
library(dplyr)
library(ggplot2)
library(stringr)
library(colorspace)


# 1. Base colors per model
# 1. Base colors per model
model_colors <- c(
  "Aquatic reliance" = "#1f78b4",
  "Aquaculture reliance"                                      = "#e31a1c",
  "Capture reliance"                    = "#6a3d9a",
  "Domestic reliance"                       = "#b15928",
  "Foreign reliance"                        = "#a6cee3",
  "Marine reliance"                         = "#b2df8a",
  "Inland reliance"                         = "#fb9a99"
)

col_95 <- lighten(model_colors, 0.65)
col_80 <- lighten(model_colors, 0.4)
col_50 <- lighten(model_colors, 0.10)

# 2. Add clean_name once
slopes_df <- slopes_df %>%
  mutate(clean_name = sapply(model, clean_model_name))

# 3. Build plotting data
df_plot <- slopes_df %>%
  mutate(
    clean_name = sapply(model, clean_model_name),
    nonzero = (lo95 < 0 & hi95 < 0) | (lo95 > 0 & hi95 > 0),
    suffix = sub(".*_", "", model),
    suffix_num = suppressWarnings(as.numeric(suffix))
  )  %>%
  filter(model != "animal_cons_cap") %>%
  mutate(
    nonzero    = (lo95 < 0 & hi95 < 0) | (lo95 > 0 & hi95 > 0),
    suffix     = sub(".*_", "", model),
    suffix_num = suppressWarnings(as.numeric(suffix))
  ) %>%
  filter(
    str_detect(clean_name, "reliance")
  ) %>%
  arrange(suffix_num, suffix) %>%
  mutate(
    clean_name = factor(clean_name, levels = unique(clean_name)),
    col95 = col_95[as.character(clean_name)],
    col80 = col_80[as.character(clean_name)],
    col50 = col_50[as.character(clean_name)],
    panel_id = as.numeric(factor(clean_name)),
    left_col = panel_id %% 3 == 1
  )

library(dplyr)
library(ggplot2)
library(patchwork)

panel_tags <- letters[1:9] 


# --- 1. Function to draw a single panel ---
make_panel <- function(df, show_yaxis = FALSE, tag = "") {
  ggplot(df, aes(y = var)) +
    geom_segment(aes(x = lo95, xend = hi95, color = col95),
                 linewidth = 2, lineend = "square") +
    geom_segment(aes(x = lo80, xend = hi80, color = col80),
                 linewidth = 2, lineend = "square") +
    geom_segment(aes(x = lo50, xend = hi50, color = col50),
                 linewidth = 2, lineend = "square") +
    geom_point(aes(x = med), size = 1.8, color = "black") +
    geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.4) +
    scale_color_identity() +
    scale_y_discrete(labels = var_labels) +

    # ← add tag in upper-left corner
    annotate("text", x = -Inf, y = Inf, label = tag,
             hjust = -0.3, vjust = 1.3, size = 3.2, fontface = "bold") +

    theme_minimal(base_size = 8) +
    theme(
      plot.title = element_blank(),
      panel.grid.major.y = element_blank(),
      panel.grid.minor   = element_blank(),
      strip.text         = element_text(size = 5, face = "bold"),
      axis.text.x        = element_text(size = 6, color = "black"),
      axis.title         = element_blank(),
      panel.spacing      = unit(1.2, "lines"),
      plot.margin        = margin(5, 5, 5, 5),
      axis.text.y        = if (show_yaxis) element_text(size = 6, color = "black") else element_blank(),
      axis.ticks.y       = if (show_yaxis) element_line() else element_blank()
    )
}

# --- 2. Split df_plot into 9 panels ---
panel_names <- levels(df_plot$clean_name)

panels <- lapply(seq_along(panel_names), function(i) {
  name <- panel_names[i]
  df_sub <- df_plot %>% filter(clean_name == name)

  show_y <- i %in% c(1, 4, 7)
  wrapped_name <- stringr::str_wrap(name, width = 20)

  make_panel(df_sub, show_yaxis = show_y, tag = panel_tags[i]) +
    ggtitle(wrapped_name)
})

library(ggplot2)
library(patchwork)

# --- x-axis label plot ---
xlab_plot <- ggplot() +
  theme_void() +
  annotate("text", x = 0.5, y = 0.5,
           label = "Posterior slope estimate",
           size = 2, fontface = "bold")

# --- empty spacer ---
empty <- ggplot() + theme_void()

# --- assemble ---
row1 <- (panels[[1]] | panels[[2]] | panels[[3]])
row2 <- (panels[[4]] | panels[[5]] | panels[[6]])
row3 <- (empty | panels[[7]] | empty) + plot_layout(widths = c(0.135, 1, 1))
row4 <- (empty | xlab_plot | empty) + plot_layout(widths = c(1, 1, 1))

(final_plot_with_xlab <- wrap_plots(row1, row2, row3, row4, ncol = 1) +
  plot_layout(heights = c(1, 1, 1, 0.12)))

# --- add x-axis label below ---
# final_plot_with_xlab <- final_plot_with_xlab /
#   (empty | xlab_plot | empty) +
#   plot_layout(heights = c(3, 0.12))


# ggsave(paste0("../images/SI/",model_type,"_posteriors_SI.jpg"), plot = final_plot_with_xlab, device = "jpg", width = 5, height = 3, units = "in", dpi = 300)
```

## SI Figure 2: Proportion of countries in each quadrant by region

```{r proportion of countries in each quadrant by region}
# Calculate rate of change in aquatic protein quantity
aquatic_protein_quant <- aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, year) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap, na.rm = TRUE)) %>%
  distinct() %>%
  split(.$iso3c) %>%
  map(~lm(protein_consumed_kg_percap ~ year, data = .x)) %>%
  map_df(broom::tidy, .id = 'iso3c') %>%
  filter(term == 'year') %>%
  select(iso3c, aquatic_estimate = estimate)

# Calculate rate of change in terrestrial protein quantity
terrestrial_protein_quant <- aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         data_source == "new FBS", food_group == "terrestrial") %>%
  group_by(iso3c, year) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap, na.rm = TRUE)) %>%
  distinct() %>%
  split(.$iso3c) %>%
  map(~lm(protein_consumed_kg_percap ~ year, data = .x)) %>%
  map_df(broom::tidy, .id = 'iso3c') %>%
  filter(term == 'year') %>%
  select(iso3c, terrestrial_estimate = estimate)

# Calculate rate of change in aquatic reliance
aquatic_protein_reliance <- aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, year) %>%
  summarize(prop_animal_protein = sum(prop_animal_protein, na.rm = TRUE)) %>%
  distinct() %>%
  split(.$iso3c) %>%
  map(~lm(prop_animal_protein ~ year, data = .x)) %>%
  map_df(broom::tidy, .id = 'iso3c') %>%
  filter(term == 'year') %>%
  select(iso3c, aquatic_reliance_estimate = estimate)

# Join and compute quadrant proportions per region
quad_df <- aquatic_protein_quant %>%
  full_join(terrestrial_protein_quant, by = "iso3c") %>%
  full_join(aquatic_protein_reliance, by = "iso3c") %>%
  add_region(col = "iso3c", region.col.name = "region") %>%
  mutate(
    quad_raw = case_when(
      aquatic_estimate > 0 & terrestrial_estimate > 0 ~ "Q1",
      aquatic_estimate < 0 & terrestrial_estimate > 0 ~ "Q2",
      aquatic_estimate < 0 & terrestrial_estimate < 0 ~ "Q3",
      aquatic_estimate > 0 & terrestrial_estimate < 0 ~ "Q4"
    )
  ) %>%
  group_by(region, quad_raw) %>%
  summarise(n = n_distinct(iso3c), .groups = "drop") %>%
  group_by(region) %>%
  mutate(
    prop = n / sum(n),
    quadrant = factor(
      quad_raw,
      levels = c("Q1", "Q2", "Q3", "Q4"),
      labels = c("Aquatic + / Terrestrial +",
                 "Aquatic - / Terrestrial +",
                 "Aquatic - / Terrestrial -",
                 "Aquatic + / Terrestrial -")
    )
  ) %>%
  
  
  ungroup()

# Plot parameters
max_val   <- 0.75
half_width <- 0.2

# Build spike polygons
# r = prop * max_val * sqrt(2) ensures that at prop=1, the spike tip's
# x and y components each equal max_val (the axis limit)
spike_df <- quad_df %>%
  mutate(
    center_angle = case_when(
      quad_raw == "Q1" ~ pi / 4,
      quad_raw == "Q2" ~ 3 * pi / 4,
      quad_raw == "Q3" ~ 5 * pi / 4,
      quad_raw == "Q4" ~ 7 * pi / 4
    ),
    r       = prop * max_val * sqrt(2),
    x_left  = r * cos(center_angle - half_width),
    y_left  = r * sin(center_angle - half_width),
    x_right = r * cos(center_angle + half_width),
    y_right = r * sin(center_angle + half_width)
  ) %>%
  rowwise() %>%
  reframe(
    region   = region,
    quadrant = quadrant,
    prop     = prop,
    x        = c(0, x_left, x_right),
    y        = c(0, y_left, y_right),
    group    = paste(region, quadrant)
  )

quad_colors <- c(
  "Aquatic + / Terrestrial +" = "#2a9d8f",  # teal green
  "Aquatic - / Terrestrial +" = "#e76f51",  # burnt orange
  "Aquatic - / Terrestrial -" = "#6d3b47",  # dark plum
  "Aquatic + / Terrestrial -" = "#457b9d"   # steel blue
)

# Plot
(plot <- ggplot(spike_df) +
  geom_polygon(
    aes(x = x, y = y, fill = quadrant, group = group),
    alpha = 0.85
  ) +
  geom_hline(yintercept = 0, linewidth = 0.4, colour = "black") +
  geom_vline(xintercept = 0, linewidth = 0.4, colour = "black") +
  facet_wrap(~ region) +
  coord_equal(xlim = c(-max_val, max_val), ylim = c(-max_val, max_val)) +
  scale_x_continuous(
    breaks = c(-max_val, max_val),
    labels = c("75%", "75%"),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    breaks = c(-max_val, max_val),
    labels = c("75%", "75%"),
    expand = c(0, 0)
  ) +
  scale_fill_manual(values = quad_colors) +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE)) +
  labs(
    x    = "Percentage of countries",
    y    = "Percentage of countries",
    fill = "Quadrant"
  ) +
  theme_minimal(base_size = 8) +
  theme(
    panel.grid            = element_blank(),
    panel.background      = element_rect(fill = "transparent", color = NA),
    plot.background       = element_rect(fill = "transparent", color = NA),
    strip.background      = element_blank(),
    strip.text            = element_text(size = 7, face = "bold"),
    axis.text             = element_text(size = 6, color = "black"),
    axis.title            = element_text(size = 7, face = "bold"),
    legend.position       = "bottom",
    legend.box            = "vertical",
    legend.title          = element_text(size = 7, face = "bold"),
    legend.text           = element_text(size = 6),
    legend.key.size       = unit(0.35, "cm"),
    legend.spacing        = unit(0.1, "cm"),
    legend.background     = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    panel.spacing         = unit(1, "lines")
  ))

ggsave("../images/SI/quadrant_bar_region_SI.jpg", plot = plot, device = "jpg", units = "in", width = 5, height = 3.5)
```

## Bayesian model assessment

### Trace Plot

```{r trace plot}
######################
#### Simple Trace ####
######################
model_type = "simple"
trace_df <- data.frame()

for (i in seq_along(response_vars))  {
  
  # Load in Stan data
  fit <- rstan::extract(readRDS(paste0("../output/bayesian_output/", model_type, "/fit_", response_vars[i],".stan")), permuted = T)

  chain_1 <- data.frame(beta0 = fit$beta[1:3000,1],
           beta = fit$beta_0[1:3000],
           sigma = fit$sigma[1:3000],
           model = response_vars[i],
           iteration = 1:3000,
           chain = "1")
  
  chain_2 <- data.frame(beta0 = fit$beta[1:3000,1],
           beta = fit$beta_0[3001:6000],
           sigma = fit$sigma[3001:6000],
           model = response_vars[i],
           iteration = 1:3000,
           chain = "2")
  
  chain_3 <- data.frame(beta0 = fit$beta[1:3000,1],
           beta = fit$beta_0[6001:9000],
           sigma = fit$sigma[6001:9000],
           model = response_vars[i],
           iteration = 1:3000,
           chain = "3")
  
  chains <- chain_1 %>% bind_rows(chain_2) %>% bind_rows(chain_3)
  
  trace_df <- bind_rows(trace_df, chains)
    
}



##################################
####### Hierarchical Trace #######
##################################

# Create separate datasets for each chain
model_type = "realm"
trace_df <- data.frame()

for (i in seq_along(response_vars))  {
  
  # Load in Stan data
  fit <- rstan::extract(readRDS(paste0("../output/bayesian_output/", model_type, "/fit_", response_vars[i],".stan")), permuted = T)
  
  chain_1 <- data.frame(beta = fit$beta[1:3000,1],
           beta_0 = fit$beta_0[1:3000],
           sigma = fit$sigma[1:3000],
           beta_0i = fit$beta_0i[1:3000,1],
           z_beta_0 = fit$z_beta_0[1:3000,1],
           model = response_vars[i],
           iteration = 1:3000,
           chain = "1")
  
  chain_2 <- data.frame(beta = fit$beta[3001:6000,1],
           beta_0 = fit$beta_0[3001:6000],
           sigma = fit$sigma[3001:6000],
           beta_0i = fit$beta_0i[3001:6000,1],
           z_beta_0 = fit$z_beta_0[3001:6000,1],
           model = response_vars[i],
           iteration = 1:3000,
           chain = "2")
  
  chain_3 <- data.frame(beta = fit$beta[6001:9000,1],
           beta_0 = fit$beta_0[6001:9000],
           sigma = fit$sigma[6001:9000],
           beta_0i = fit$beta_0i[6001:9000,1],
           z_beta_0 = fit$z_beta_0[6001:9000,1],
           model = response_vars[i],
           iteration = 1:3000,
           chain = "3")
  
  chains <- chain_1 %>% bind_rows(chain_2) %>% bind_rows(chain_3)
  
  trace_df <- bind_rows(trace_df, chains)
    
}

plot <- trace_df %>%
  pivot_longer(cols = -c(model, iteration, chain), names_to = "parameter") %>%
  ggplot(aes(x = iteration, y = value, color = chain)) +
  geom_line() +
  scale_color_viridis_d(option = "plasma", end = 0.95) +
  facet_grid(model ~ parameter) +
  guides(color = "none") +
  labs(x = "Parameter", y = "Parameter value") +
  theme_pubclean()

ggsave(paste0("../images/bayesian plots/", model_type, "/trace.jpg"), plot = plot, device = "jpg", width = 18, height = 30, units = "in", dpi = 300)
```

### R\^2

```{r}
response_vars = response_vars
model_type = "simple"

#######################
#### Simple R^2    ####
#######################

R2_matrix <- matrix(nrow = 3000, ncol = length(response_vars))
for (i in seq_along(1:23)) {

# Extract posterior draws
fit <- rstan::extract(readRDS(paste0("../output/bayesian_output/", model_type, "/fit_", response_vars[i],".stan")), permuted = T)

# y_pred is iterations x N
y_pred <- fit$y_pred[1:3000,]

# y_pred: iterations × N
mu_s <- y_pred   # each row is mu_s for that iteration

var_mu_s <- apply(mu_s, 1, var)   # length = iterations
post_var_obs <- apply(y_pred, 2, var)   # length N

sigma_s <- fit$sigma[1:3000]   # length = iterations
var_resid_s <- sigma_s^2


R2_s <- var_mu_s / (var_mu_s + var_resid_s)
# print((hist(R2_s,main = response_vars[i])))
R2_matrix[,i] <- R2_s
}

colnames(R2_matrix) <- response_vars

plot_df <- R2_matrix %>%
  data.frame() %>%
  pivot_longer(everything()) %>%
  mutate(
    value  = as.numeric(value),
    suffix = sub(".*_", "", name)
  )

# compute the order of model names by suffix
ordered_names <- plot_df %>%
  distinct(name, suffix) %>%
  arrange(suffix) %>%
  pull(name)

(plot <- plot_df %>%
  mutate(name = factor(name, levels = ordered_names)) %>%
  ggplot(aes(x = value, y = name)) +
  ggridges::geom_density_ridges(fill = "lightblue") +
  theme_pubclean() +
  labs(x = bquote("Posterior" ~ R^2), y = "Model"))


ggsave("../images/bayesian plots/simple/r2.jpg", plot = plot, device = "jpg", width = 6, height = 4, units = "in", dpi = 300)

######################################
########## Hierarchical R^2 ##########
######################################

R2_df <- data.frame(
  fixed_r2    = numeric(0),
  random_r2   = numeric(0),
  combined_r2 = numeric(0),
  model       = character(0)
)

# Change to either continental / realm
model_type = "realm"
for (i in seq_along(response_vars))  {
  
fit <- rstan::extract(readRDS(paste0("../output/bayesian_output/", model_type, "/fit_", response_vars[i],".stan")), permuted = T)
  
  # Z: N x 16 realm indicator matrix (1 for realm membership)
b_s <- fit$beta_0i[1:3000,]   # iterations x 16

realm_ids <- readRDS(paste0("../output/bayesian_output/", model_type, "_ids.rds"))

X <- readRDS(paste0("../output/bayesian_output/", model_type, "/X_",response_vars[i],".rds"))
realm_ids <- readRDS(paste0("../output/bayesian_output/", model_type, "/ids_",response_vars[i],".rds"))

Z <- model.matrix(~ 0 + factor(realm_ids))

# random-only linear predictor: iterations x N
mu_random <- b_s %*% t(Z)


beta_s <- fit$beta[1:3000, ]   # 8000 × K
mu_fixed <- beta_s %*% t(X)    # 8000 × N

var_fixed_s  <- apply(mu_fixed,  1, var)   # length = iterations
var_random_s <- apply(mu_random, 1, var)   # length = iterations
var_resid_s  <- fit$sigma^2               # or fit$sigma^2 depending on your model

mu_total     <- mu_fixed + mu_random
var_total_s  <- apply(mu_total, 1, var)

R2_fixed_s   <- var_fixed_s  / (var_total_s + var_resid_s)
R2_random_s  <- var_random_s / (var_total_s + var_resid_s)
R2_cond_s    <- var_total_s  / (var_total_s + var_resid_s)

# Create new data frame with Current iteration R2 values
new_data <- data.frame(fixed_r2 = R2_fixed_s, random_r2 = R2_random_s, combined_r2 = R2_cond_s, model = response_vars[i])

# Add on R2 values to R2 data frame
R2_df <- bind_rows(R2_df, new_data)
  
}

# Plot all model R^2
(plot <- R2_df %>%
  pivot_longer(cols = c("fixed_r2", "random_r2", "combined_r2")) %>%
  mutate(model = factor(model, levels = ordered_names),
         name = case_when(name == "combined_r2" ~ "combined",
                          name == "fixed_r2" ~ "fixed",
                          name == "random_r2" ~ "random")) %>%
  ggplot(aes(x = value, fill = name)) +
  geom_histogram(position = "identity", alpha = 0.7) +
  facet_wrap(~ model) +
  theme_pubclean() +
  scale_fill_viridis_d(option = "plasma", end = 0.95) +
  labs(x = bquote("Posterior" ~ R^2), y = "# iterations", fill = "Effect type") +
  theme(legend.position = "bottom",
        strip.text.x = element_text(margin = margin(0.05,0,0.05,0, "cm"), size = 6),
        axis.text = element_text(size = 6)))

ggsave(paste0("../images/bayesian plots/", model_type, "/r2.jpg"), plot = plot, device = "jpg", width = 9, height = 5, units = "in", dpi = 300)
```

### Divergent transitions

```{r divergent transitions}
# Change to either continental / realm
model_type = "simple"

# Rename the divergent transition column names to model names
divergent_transitions <- matrix(nrow = 3, ncol = 23)
colnames(divergent_transitions) <- response_vars

for (i in seq_along(response_vars)) {
  
  # Extract number of divergent samples
  sampler_params <- get_sampler_params(readRDS(paste0("../output/bayesian_output/", model_type, "/fit_", response_vars[i],".stan")), inc_warmup = FALSE)
  num_transitions <- sapply(sampler_params, function(x) sum(x[, "divergent__"]))
  
  divergent_transitions[,i] <- num_transitions
  
}

# Create dataframe for plotting divergent transitions
plot_df <- divergent_transitions %>%
  as.data.frame() %>%
  mutate(chain = as.character(1:3)) %>%
  pivot_longer(!"chain", names_to = "model", values_to = "divergent") %>%
  mutate(
    suffix = sub(".*_", "", model)
  )

# compute the order of model names by suffix
ordered_names <- plot_df %>%
  distinct(model, suffix) %>%
  arrange(suffix) %>%
  pull(model)

# Plot # divergent transitions
(plot <- plot_df %>%
  mutate(model = factor(model, levels = ordered_names)) %>%
  ggplot(aes(y = model, x = divergent, fill = chain)) +
  geom_col() +
  scale_fill_viridis_d(end = 0.8) +
  labs(y = "Model", x = "# divergent transitions") +
  theme_light() +
    theme(legend.position = "bottom"))

ggsave(paste0("../images/bayesian plots/", model_type, "/divergent_samples.jpg"), plot = plot, device = "jpg", width = 7, height = 5, units = "in", dpi = 300)
```

### Posterior Predictive Checks

```{r posterior predictive check}
model_type = "simple"
ppc_df <- purrr::map_dfr(seq_along(response_vars), \(i)
  bayesplot::ppc_data(
    readRDS(paste0("../output/bayesian_output/", model_type,
                   "/y_", response_vars[i], ".rds")),
    extract(
      readRDS(paste0("../output/bayesian_output/", model_type,
                     "/fit_", response_vars[i], ".stan")),
      permuted = TRUE
    )$y_pred[sample(8000, 200), ]
  ) |>
    mutate(model = response_vars[i])
)

plot <- ggplot(ppc_df, ggplot2::aes(value)) +
  geom_density(
    data = filter(ppc_df, !is_y),
    aes(group = interaction(model, rep_id)),
    linewidth = .25, alpha = .6
  ) +
  geom_density(
    data = filter(ppc_df, is_y),
    linewidth = 1
  ) +
  facet_wrap(~ model, scales = "free_y", ncol = 4) +
  bayesplot_theme_get()

ggsave(paste0("../images/bayesian plots/", model_type, "/posterior_predictive.jpg"), plot = plot, device = "jpg", width = 15, height = 10, units = "in", dpi = 300)
```

### R hat

```{r r hat}
model_type = "continental"
rhat_df <- data.frame()
for (i in seq_along(response_vars)) {
  fit <- readRDS(paste0("../output/bayesian_output/", model_type,
                     "/fit_", response_vars[i], ".stan"))
  

  if (model_type == "simple") {
    
    vals <- rstan::summary(fit, pars = c("beta", "beta_0", "sigma", "lp__"))$summary[, "Rhat"]
    
  } else if (model_type != "simple") {
    
    vals <- rstan::summary(fit, pars = c("beta_0", "beta", "sigma", "beta_0i", "sigma_beta_0", "z_beta_0", "lp__"))$summary[, "Rhat"]
    
  }
  # Get R_hat values
  
  
  # Add iteration i rhat values to total dataframe for plotting
  new_data <- data.frame(rhat = vals, model = response_vars[i])
  rhat_df <- bind_rows(rhat_df, new_data)
}

(plot <- rhat_df %>%
  ggplot(aes(x = rhat)) +
  geom_histogram() +
  facet_wrap(~ model) +
  labs(x = "R-hat", y = "# parameters") +
  theme_pubclean() +
    theme(axis.text = element_text(size = 7)))

ggsave(paste0("../images/bayesian plots/", model_type, "/rhat.jpg"), plot = plot, device = "jpg", width = 7, height = 5, units = "in", dpi = 300)



```

### Multicollinearity check

```{r}
var_labels <- c(
  eez_area = "EEZ area",
  gdp_trade = "GDP from trade (%)",
  inundation_area_log1000km2 = "Inundated area (log 1000 km²)",
  log_gdp = "GDP (log scale)",
  pop_10k_pct_coast = "Population within 10 km of coast (%)",
  supermarkets = "Supermarkets per 100,000 people"
)

model_type = "realm"

X <- readRDS(paste0("../output/bayesian_output/", model_type,
                     "/X_", response_vars[1], ".rds"))
# correlation matrix
cor_mat <- cor(X, use = "pairwise.complete.obs")

# helper functions
get_lower_tri <- function(cormat){
  cormat[upper.tri(cormat)] <- NA
  cormat
}

get_upper_tri <- function(cormat){
  cormat[lower.tri(cormat)] <- NA
  cormat
}

get_diag <- function(cormat){
  cormat[!diag(nrow(cormat))] <- NA
  cormat
}

# split matrix parts
lower_tri <- get_lower_tri(cor_mat)
upper_tri <- get_upper_tri(cor_mat)
diag_mat  <- get_diag(cor_mat)

# melt
lower_melt <- melt(lower_tri, na.rm = TRUE)
upper_melt <- melt(upper_tri, na.rm = TRUE)
diag_melt  <- melt(diag_mat,  na.rm = TRUE) %>%
  mutate(clean_names= c("GDP (log scale)","GDP from trade (%)","EEZ area","Inundated area (log 1000 km²)","Supermarkets per 100,000 people","Population within 10 km of coast (%)"),
         clean_names = str_wrap(clean_names, width = 12))

# remove diagonal values from upper triangle
upper_melt <- upper_melt[upper_melt$Var1 != upper_melt$Var2, ]

(plot <- ggplot() +

  # lower triangle heatmap
  geom_tile(
    data = lower_melt,
    aes(Var2, Var1, fill = value)
  ) +

  # diagonal white tiles
  geom_tile(
    data = diag_melt,
    aes(Var2, Var1),
    fill = "white",
    color = "black"
  ) +

  # upper triangle correlation numbers
  geom_text(
    data = upper_melt,
    aes(Var2, Var1, label = sprintf("%.2f", value), color = value),
    size = 4
  ) +

  # diagonal variable names
  geom_text(
    data = diag_melt,
    aes(Var2, Var1, label = clean_names),
    fontface = "bold",
    family = "Arial Black",
    size = 2
  ) +

  scale_fill_gradient2(
  low = "darkcyan",
  high = "salmon",
  mid = "gold",
  midpoint = 0,
  limits = c(-1,1),
  guide = guide_colorbar(
    barheight = unit(.94, "npc"),
    barwidth = unit(0.8, "lines")
  )
) +

  scale_color_gradient2(
    low = "darkcyan",
    high = "salmon",
    mid = "gold",
    midpoint = 0,
    limit = c(-1,1),
    guide = "none"
  ) +

  coord_fixed() +
  scale_x_discrete(limits = rev) +
  # scale_y_discrete(limits = rev) +

  theme_minimal() +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    panel.grid = element_blank(),
    legend.title = element_blank(),
    legend.ticks = element_line(hjust = 5)
  ))


ggsave("../images/SI/collinearity_plot_SI.jpg", plot = plot, device = "jpg", units = "in", width = 5, height = 3.5)
```

# Posterior Predictive distribuitons

```{r ppd distributions}
model_type = "realm"
source("../R/clean_model_name.R")
# =========================================================
# 1. Helper: extract posterior predictive draws for a model
# =========================================================
extract_ppd <- function(model_name, model_type) {

  # Load fitted Stan model
  fit <- readRDS(
    paste0("../output/bayesian_output/", model_type, "/fit_", model_name, ".stan")
  )

  x <- rstan::extract(fit, permuted = TRUE)

  # Posterior predictive draws (S × N)
  y_pred <- x$y_pred

  # Load X and y to get N
  X <- readRDS(
    paste0("../output/bayesian_output/", model_type, "/X_", model_name, ".rds")
  )
  y <- readRDS(
    paste0("../output/bayesian_output/", model_type, "/y_", model_name, ".rds")
  )

  N <- length(y)

  # Load realm names (vector of length N)
  realm_levels <- readRDS(
    paste0("../output/bayesian_output/", model_type, "/realms_", model_name, ".rds")
  )

  # Safety check
  if (length(realm_levels) != N) {
    stop(paste0("realm_levels length mismatch for model ", model_name))
  }

  # Format long
  ppd_long <- as.data.frame(y_pred) %>%
    mutate(draw = row_number()) %>%
    pivot_longer(-draw, names_to = "obs", values_to = "ppd") %>%
    mutate(
      obs = as.integer(gsub("V", "", obs)),
      region = realm_levels[obs],
      y_obs = y[obs],
      model = model_name
    )

  return(ppd_long)
}

# =========================================================
# 2. Loop over all models and bind results
# =========================================================
ppd_all_models <- map_dfr(response_vars, ~ extract_ppd(.x, model_type)) %>%
  mutate(clean_name = clean_model_name(model)) # Add column for clean model names


# Calculate posterior predictive means
ppd_means <- ppd_all_models %>%
  group_by(model, obs, region) %>%
  summarise(
    ppd_mean = mean(ppd),
    y_obs    = first(y_obs),
    .groups  = "drop"
  )

# Save combined posterior predictive dataset
write_parquet(
  ppd_all_models,
  paste0("../output/bayesian_output/", model_type, "/posterior_predictive_all_models.parquet")
)

# Read in PPD data
ppd_all_models <- read_parquet(
  paste0("../output/bayesian_output/", model_type, "/posterior_predictive_all_models.parquet")
)

# =========================================================
# 3. Combined posterior predictive plot (all models together)
# =========================================================

library(ggplot2)
library(stringr)

draw_sample <- sample(unique(ppd_all_models$draw), size = 30)

(plot <- ppd_all_models %>%
  filter(
    !str_detect(model, "animal_cons_cap"),
    str_detect(clean_name, "capita") |
    str_detect(clean_name, "Aquatic reliance") |
    str_detect(clean_name, "reliance")
  ) %>%
  filter(draw %in% draw_sample) %>%
  ggplot(aes(x = ppd, group = draw)) +
  geom_density(alpha = 0.08, color = "grey60") +
  geom_density(aes(x = y_obs), color = "black", linewidth = 1.1) +
  facet_wrap(
    ~ clean_name,
    scales = "free",
    labeller = label_wrap_gen(width = 18)   # <-- WRAPS facet labels
  ) +
  labs(
    x = "Response value",
    y = "Density"
  ) +
  theme_minimal(base_size = 8) +
  theme(
    # Gridlines
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(color = scales::alpha("black", 0.25), linewidth = 0.3),
    panel.grid.minor.y = element_blank(),

    # Backgrounds
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),

    # Facet strips
    strip.background = element_blank(),
    strip.text = element_text(size = 7, face = "bold"),

    # Axes
    axis.text = element_text(size = 6, color = "black"),
    axis.title.y = element_text(size = 7, margin = margin(r = 5), face = "bold"),
    axis.text.x = element_text(size = 5.5, angle = 30, hjust = 1),

    # No legend here, but keep consistent styling
    legend.position = "bottom",
    legend.title = element_text(size = 7, face = "bold"),
    legend.text = element_text(size = 6),
    legend.key.size = unit(0.3, "cm"),
    legend.box = "vertical",
    legend.spacing = unit(0.1, "cm"),
    legend.margin = margin(t = 0, b = 0),
    legend.box.margin = margin(t = -2, b = -2),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA)
  ))

ggsave(paste0("../images/bayesian plots/", model_type, "/ppd.jpg"), plot = plot, device = "jpg", width = 7, height = 5, units = "in", dpi = 300)
```

```{r}
# Get letters for panel numbers
panel_tags <- letters[1:15] 

# --- 1. Function to draw a single panel ---
make_ppd_panel <- function(df, title, tag = "") {
  ggplot(df, aes(x = ppd, group = draw)) +
    geom_density(alpha = 0.08, color = "grey60") +
    geom_density(aes(x = y_obs), color = "black", linewidth = 1.1) +
    labs(y = "Density") +
    annotate(
      "text",
      x = -Inf, y = Inf,
      label = tag,
      hjust = -0.3, vjust = 1.3,
      size = 3, fontface = "bold"
    ) +
    theme_minimal(base_size = 8) +
    theme(
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.y = element_line(color = scales::alpha("black", 0.25), linewidth = 0.3),
      panel.grid.minor.y = element_blank(),
      panel.background = element_rect(fill = "transparent", color = NA),
      plot.background = element_rect(fill = "transparent", color = NA),
      strip.background = element_blank(),
      strip.text = element_blank(),
      axis.text = element_text(size = 6, color = "black"),
      axis.title.y = element_text(size = 7, margin = margin(r = 5), face = "bold"),
      axis.text.x = element_text(size = 5.5, angle = 30, hjust = 1),
      axis.title.x = element_blank()
    )
}

# --- 2. Split df_plot into 9 panels ---

xlab_plot <- ggplot() +
  theme_void() +
  annotate("text", x = 0.5, y = 0.5,
           label = "Response value",
           size = 2.5, fontface = "bold")

draw_sample <- sample(unique(ppd_all_models$draw), size = 30)

df_plot <- ppd_all_models %>%
  filter(
    model != "animal_cons_cap",
    str_detect(clean_name, "capita") |
      str_detect(clean_name, "Aquatic reliance") |
      str_detect(clean_name, "reliance")
  ) %>%
  filter(draw %in% draw_sample) %>%
  mutate(clean_name = factor(clean_name))

panel_names <- levels(df_plot$clean_name)

layout_order <- c(15, 3, 5,
                  1, 7, 9,
                  13, 11, 4,
                  2, 6, 8,
                  10, 14, 12)

panel_tags <- letters[1:15]

# assign tags in layout order
tags_by_panel_index <- rep("", length(panel_names))
tags_by_panel_index[layout_order] <- panel_tags

panels <- lapply(seq_along(panel_names), function(i) {
  nm <- panel_names[i]
  df_sub <- df_plot %>% filter(clean_name == nm)
  make_ppd_panel(df_sub, nm, tag = tags_by_panel_index[i])
})

library(patchwork)

empty <- ggplot() + theme_void()

row1 <- panels[[15]]  | panels[[3]]  | panels[[5]]
row2 <- panels[[1]]  | panels[[7]]  | panels[[9]]
row3 <- panels[[13]]  | panels[[11]]  | panels[[4]]
row4 <- panels[[2]] | panels[[6]] | panels[[8]]
row5 <- panels[[10]] | panels[[14]] | panels[[12]]
row6 <- empty | xlab_plot | empty

(final_plot <- wrap_plots(
  row1, row2, row3, row4, row5, row6,
  ncol = 1
) + plot_layout(heights = c(1,1,1,1,1,0.15)))

ggsave(paste0("../images/SI/ppd_SI.jpg"), plot = final_plot, device = "jpg", width = 4, height = 5, units = "in", dpi = 300)
```
