---
title: "02_results"
author: "Connor Quiroz, Lana Huang, Christopher D. Golden, Sarah Glaser, and Jessica A. Gephart"
date: "2025-09-05"
output: html_document
---

# Chapter 1 Results

# Load in packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(exploreARTIS)
library(ggrepel)
library(countrycode)
library(colorspace)
library(ggpubr)
library(arrow)
library(rstan)
library(sf)
library(lwgeom)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggridges)
library(bayesplot)
library(purrr)
library(patchwork)
library(dplyr)
library(ggplot2)
library(forcats)
library(stringr)
library(grid)
library(broom)
library(ggtext)
```

# Read in data

## Read in aquatic reliance data & FAO population data

```{r read in FAO derived data}
# Read in population
population_fao_wb <- read.csv("../output/population_fao_wb.csv")

# Read in aquatic reliance data
aquatic_reliance <- read.csv("../output/fao_aquatic_reliance_source.csv") %>%
  left_join(population_fao_wb,by=c("year","iso3c"))%>%
  mutate(protein_consumed_kg_percap=protein_consumed_t*1000/population_fao)

# Read in FAO sourcing data
aquatic_fbs <- read.csv("../output/fao_fbs_aquatic_source.csv")
```

## Load in biophysical and socio-economic data as covariates into the Bayesian model

```{r load data}
# Read in adaptive capacity data
ac_df <- read_parquet("../data/e_s_ac_data_total.parquet") %>%
  filter(scenario == "ssp126") %>%
  select(-aa_reliance_pct, -scenario, -change_in_stock,-live_weight_t,-pct_change) %>%
  rename(iso3c = "consumer_iso3c")
  
# Read in ecoregion data (all shapes associated with soverign)
ecoregion_full <- read.csv("../data/Spatially joined ecoregion to EEZ/ecoregion.csv") %>%
  rename("iso3c" = "ISO_SOV1",
         ecoregion = "ECOREGION",
         province = "PROVINCE",
         realm = "REALM",
         shape_area = "Shape_Area")

# Read in ecoregions for largest soverign shape
ecoregion_unique <- read.csv("../data/Spatially joined ecoregion to EEZ/ecoregion_clean.csv") %>%
  select(-FREQUENCY) %>%
  rename(iso3c = "ISO_SOV1",
         shape_area = "MAX_Shape_Area")

# Read in dietary sourcing flexibility index
dsfi_df <- read_csv("../data/covariates/food-systems-dashboard-2026-01-20 (dietary sourcing flexibility index).csv") %>%
  filter(`Start Year` == 2019 | `End Year` == 2019) %>%
  select(c("iso3c" = ISO3, "dsfi" = Value))

# Read in national indicator data
indicator_df <- read_csv("../data/covariates/all_national_indicators.csv")

# Read in coastline data
coastline_df <- read_csv("../data/covariates/countries-by-coastline-2026.csv") %>%
  mutate(iso3c = case_when(!is.na(flagCode) ~ countrycode(flagCode, origin = "iso2c", destination = "iso3c"),
                           country == "Namibia" ~ "NAM")) %>%
  select(c(iso3c, "coastline_km" = "CountryCoastlineKm"))

# Read in proportion of people within 10k of the coast
pop_10k_coast <- read.csv("../data/covariates/pop_10k_coast.csv")
```

## Clean ecoregion data

Impute countries with no denoted ecoregion as "landlocked"

```{r clean ecoregion data}
# Join largest shape for the respective soverign country to their ecoregion names
ecoregion <- left_join(ecoregion_unique, ecoregion_full, by = c("iso3c", "shape_area"))

# Assign countries not associated with a marine ecoregion to "landlocked" code
landlocked_countries <- aquatic_reliance %>%
  distinct(iso3c) %>%
  drop_na() %>%
  add_region(col = "iso3c", region.col.name = "region") %>%
  filter(!iso3c %in% ecoregion$iso3c) %>%
  # mutate(ecoregion = "Landlocked",
  #        province = "Landlocked",
  #        realm = "Landlocked") %>%
  mutate(ecoregion = paste0("Landlocked, ", region),
         province = paste0("Landlocked, ", region),
         realm = paste0("Landlocked, ", region)) %>%
  select(-region)

# Clean ecoregion data
ecoregion <- bind_rows(ecoregion, landlocked_countries) %>%
  arrange(iso3c) %>%
  select(-shape_area)

# Clean up files
rm(list = c("ecoregion_full", "ecoregion_unique"))

# Write to csv
write_csv(ecoregion, "../output/ecoregion_clean.csv")
```

## Create list of response variables & their associated data to put through Bayesian model (See RQ2 for Bayesian model code)

This code chunk is also in the main results. This code chunk extracts the name of each model name used in the Bayesian model. In code lines further down, we call the "response_vars" object to generate statistics and figures.

```{r acquire animal consumption data to be fed through analysis}
# Calculate animal consumption
terr_animal_cons_cap <- aquatic_reliance %>%
  filter(year == 2019) %>%
  group_by(iso3c) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))

animal_cons_cap <- aquatic_reliance %>%
  filter(year == 2019) %>%
  group_by(iso3c) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))

# Calculate aquatic animal consumption
aquatic_animal_cons_cap <- aquatic_reliance %>%
  filter(year == 2019, food_group == "aquatic") %>%
  group_by(iso3c) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))

# Calculate animal reliance consumption
aquatic_rel <- aquatic_animal_cons_cap %>%
  left_join(animal_cons_cap, by = "iso3c") %>%
  group_by(iso3c) %>%
  summarize(prop_animal_protein = protein_consumed_kg_percap.x / protein_consumed_kg_percap.y)

# Add items not created in the loop to the list
response_vars <- c("terr_animal_cons_cap","animal_cons_cap", "aquatic_animal_cons_cap", "aquatic_rel", "aquatic_protein_roc", "terrestrial_protein_roc")

for (i in c("capture", "aquaculture", "domestic", "foreign", "marine", "inland")) {
  
  # Calculate consumption per capita by method
  if (i %in% c("capture", "aquaculture")) {
    assign(
      paste0(i, "_cap"),
      aquatic_reliance %>%
        filter(year == 2019, food_group == "aquatic", method == i) %>%
        group_by(iso3c) %>%
        summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))
    )
  }
  
  # Calculate consumption per capita by consumption source
  if (i %in% c("domestic", "foreign")) {
    assign(
      paste0(i, "_cap"),
      aquatic_reliance %>%
        filter(year == 2019, food_group == "aquatic", consumption_source == i) %>%
        group_by(iso3c) %>%
        summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))
    )
  }
  
  # Calculate consumption per capita by habitat
  if (i %in% c("marine", "inland")) {
    assign(
      paste0(i, "_cap"),
      aquatic_reliance %>%
        filter(year == 2019, food_group == "aquatic", habitat == i) %>%
        group_by(iso3c) %>%
        summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))
    )
  }
  
  response_vars <- c(response_vars, paste0(i, "_cap"))
  
  # Calculate reliance by method, production source, or habitat
  assign(
    paste0(i, "_rel"),
    get(paste0(i, "_cap")) %>%
      left_join(animal_cons_cap, by = "iso3c") %>%
      group_by(iso3c) %>%
      summarize(prop_animal_protein = protein_consumed_kg_percap.x / protein_consumed_kg_percap.y)
  )
  
  response_vars <- c(response_vars, paste0(i, "_rel"))
  
  # Calculate production sources' proportion of total aquatic animal consumption per capita
  assign(
    paste0(i, "_prop_of_aquatic_cons"),
    get(paste0(i, "_cap")) %>%
      left_join(aquatic_animal_cons_cap, by = "iso3c") %>%
      group_by(iso3c) %>%
      summarize(aquatic_source_prop = protein_consumed_kg_percap.x / protein_consumed_kg_percap.y)
  )
  
  response_vars <- c(response_vars, paste0(i, "_prop_of_aquatic_cons"))
}

#calculate rate of change in aquatic protein quantity
aquatic_protein_roc <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, year, method) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap, na.rm = TRUE)) %>%
  distinct() %>%
  filter(method == "capture") %>%
  split(.$iso3c) %>%
  map(~lm(protein_consumed_kg_percap~year, data = .x)) %>%
  map_df(broom::tidy, .id = 'iso3c') %>%
  filter(term == 'year') %>%
  select(c("iso3c", "aquatic_estimate" = "estimate"))

#calculate rate of change in terrestrial protein quantity
terrestrial_protein_roc <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "terrestrial") %>%
  group_by(iso3c, year) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap, na.rm = TRUE)) %>%
  distinct() %>%
  split(.$iso3c) %>%
  map(~lm(protein_consumed_kg_percap~year, data = .x)) %>%
  map_df(broom::tidy, .id = 'iso3c') %>%
  filter(term == 'year') %>%
  select(c("iso3c", "terrestrial_estimate" = "estimate"))

# Save list of response variables to a .RDS file
saveRDS(response_vars, "../output/response_vars.rds")
```

# RQ 1a: How do aquatic food reliance trends disaggregated by source differ from aggregate aquatic food reliance?

## Figure: Compare aggregate to disaggregate reliance

Left plot: trends in global total aquatic protein (blue) reliance with respect to the terrestrial protein reliance (red). The jump in data at 2010 is the visualization data switching from FAO-FBS old methodology to the new methodology. Right plot: Trends in  aquatic animal protein reliance by sourcing, where each facet represents the habitat and method and the color represents the production source. Dotted lines represent forward and back casting of old / new methodology data to years as error estimates.

```{r compare aggregate to disaggregated reliance}
aa_reliance_mean_global_trends_alt <- aquatic_reliance %>%
  group_by(year, data_source, food_group) %>%
  summarise(food_group_total_protein_t = sum(protein_consumed_t)) %>%
  group_by(year, data_source) %>%
  mutate(prop_animal_protein = 100*food_group_total_protein_t/sum(food_group_total_protein_t)) %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS")) # Remove overlapping years, prioritizing new years

# Define year ranges for projections (add this before the existing projections code)
year_ranges <- tibble(
  data_source = c("old FBS", "new FBS"),
  start_year  = c(1996, 1996),
  end_year    = c(2020, 2020)
)

observed_ranges <- tibble(
  data_source = c("old FBS", "new FBS"),
  min_year = c(1996, 2010),
  max_year = c(2009, 2020)
)

# --- FIRST PLOT: aggregate global trend with projections ---

coefs_global <- aa_reliance_mean_global_trends_alt %>%
  filter(food_group == "aquatic") %>%
  group_by(data_source) %>%
  nest() %>%
  mutate(model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
         tidy  = map(model, broom::tidy)) %>%
  unnest(tidy) %>%
  select(data_source, term, estimate) %>%
  pivot_wider(values_from = estimate, names_from = term) %>%
  rename(year_coef = year)

projections_global <- coefs_global %>%
  left_join(year_ranges, by = "data_source") %>%
  rowwise() %>%
  mutate(year = list(start_year:end_year)) %>%
  unnest(year) %>%
  ungroup() %>%
  mutate(prop_animal_protein = `(Intercept)` + year_coef * year,
         type = "projected") %>%
  left_join(observed_ranges, by = "data_source") %>%
  filter(year < min_year | year > max_year)

observed_global_old <- aa_reliance_mean_global_trends_alt %>%
  filter(food_group == "aquatic", data_source == "old FBS", year <= 2009) %>%
  mutate(type = "observed")

observed_global_new <- aa_reliance_mean_global_trends_alt %>%
  filter(food_group == "aquatic", data_source == "new FBS", year >= 2010) %>%
  mutate(type = "observed")

combined_global <- bind_rows(observed_global_old, observed_global_new)

aa_reliance_mean_global_trends <- ggplot() +
  geom_area(data = combined_global,
            aes(x = year, y = prop_animal_protein),
            fill = "#1179be") +
  geom_line(data = projections_global,
            aes(x = year, y = prop_animal_protein, group = data_source),
            colour = "black", linetype = "dashed", linewidth = 0.5, alpha = 0.7) +
  labs(x = "", y = "Global Animal Protein Reliance (%)") +
  theme_minimal(base_size = 8) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(color = scales::alpha("black", 0.25), linewidth = 0.3),
    panel.grid.minor.y = element_blank(),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),
    axis.text = element_text(size = 6, color = "black"),
    axis.title.y = element_text(size = 7, margin = margin(r = 5), face = "bold"),
    legend.position = "bottom",
    legend.title = element_text(size = 7, face = "bold"),
    legend.text = element_text(size = 6),
    legend.key.size = unit(0.3, "cm"),
    legend.box = "vertical",
    legend.spacing = unit(0.1, "cm"),
    legend.margin = margin(t = 0, b = 0),
    legend.box.margin = margin(t = -2, b = -2),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    aspect.ratio = 0.4
  )


# Calculate the total terrestrial animal protein
 animal_protein_sourcing <- aquatic_reliance %>%
  group_by(year, data_source, food_group) %>%
  summarise(animal_protein = sum(protein_consumed_t)) %>%
  filter(food_group == "terrestrial")

# Calculate the total aquatic animal protein
 aquatic_protein_sourcing <- aquatic_reliance %>%
  group_by(year, data_source, food_group) %>%
  summarise(aquatic_protein = sum(protein_consumed_t)) %>%
  filter(food_group == "aquatic")
 
 # Calculate total tonnage of combined terrestrial and aquatic animal protein
 total_protein <- left_join(animal_protein_sourcing, aquatic_protein_sourcing, by = c("year", "data_source")) %>%
   group_by(year, data_source) %>%
   summarize(total_protein = sum(animal_protein,aquatic_protein)) # add total aquatic animal protein (unseparated by sourcing) to animal proteins by year and data source
 
# Join total protein to aquatic protein weights separated by protein to calculate reliance by sourcing
 aa_reliance_multiple_trends_alt <- aquatic_reliance %>%
  group_by(year, data_source, consumption_source, habitat, method, food_group) %>%
  summarise(food_group_total_protein_t = sum(protein_consumed_t)) %>%
  filter(food_group == "aquatic") %>%
  left_join(total_protein, by = c("year", "data_source")) %>%
  ungroup() %>%
  mutate(prop_animal_protein = 100 * (food_group_total_protein_t / total_protein)) %>% # Divide AA / TERRESTRIAL + AA
  filter(habitat != "unknown") %>%
   filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS")) # Remove overlapping years, prioritizing new years
 
 
 aa_reliance_multiple_trends_alt_1 <- aa_reliance_multiple_trends_alt %>%
  mutate(habitat = str_to_title(habitat)) %>%
  mutate(habitat_method = paste(habitat, method, sep = " "))
 
# Create forward / backward projections
 
 coefs <- aquatic_reliance %>%
  group_by(year, data_source, habitat, method, food_group) %>%
  summarise(food_group_total_protein_t = sum(protein_consumed_t), .groups = "drop") %>%
  filter(food_group == "aquatic") %>%
  left_join(total_protein, by = c("year", "data_source")) %>%
  mutate(prop_animal_protein = 100 * (food_group_total_protein_t / total_protein)) %>%
  group_by(data_source, habitat, method) %>%   # <- KEEP data_source
  nest() %>%
  mutate(model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
         tidy  = map(model, broom::tidy)) %>%
  unnest(tidy) %>%
  select(data_source, habitat, method, term, estimate) %>%
  pivot_wider(values_from = estimate, names_from = term) %>%
  rename(year_coef = year)

projections <- coefs %>%
  left_join(year_ranges, by = "data_source") %>%
  rowwise() %>%
  mutate(year = list(start_year:end_year)) %>%
  unnest(year) %>%
  ungroup() %>%
  mutate(
    prop_animal_protein = `(Intercept)` + year_coef * year,
    habitat = str_to_title(habitat)
  ) %>%
  filter(method != "unknown", habitat != "Unknown") %>%
  select(year, data_source, habitat, method, prop_animal_protein) %>%
  mutate(type = "projected") %>%
  left_join(observed_ranges, by = "data_source") %>%
  filter(year < min_year | year > max_year)  # Keep only years outside observed range

 
# Artis palette
artis_colors <- artis_palette(3)

artis_colors <- artis_colors[c(2,1)]

# Prepare observed subsets without overlap
observed_old <- aa_reliance_multiple_trends_alt_1 %>%
  filter(data_source == "old FBS", year <= 2009) %>%
  mutate(type = "observed")

observed_new <- aa_reliance_multiple_trends_alt_1 %>%
  filter(data_source == "new FBS", year >= 2010) %>%
  mutate(type = "observed")

# Filter projections to non-overlapping years
projections_old <- projections %>%
  filter(data_source == "old FBS", year > 2009) %>%
  mutate(type = "projected")

projections_new <- projections %>%
  filter(data_source == "new FBS", year < 2010) %>%
  mutate(type = "projected")

# Combine all cleanly
combined <- bind_rows(observed_old, observed_new, projections_old, projections_new) %>%
  mutate(habitat_method = paste(habitat, method))

# Plot
aa_reliance_multiple_trends <- ggplot() +
  geom_area(data = filter(combined, type == "observed"),
            aes(x = year, y = prop_animal_protein, fill = consumption_source),
            size = 0.75) +
  geom_line(data = filter(combined, type == "projected"),
            aes(x = year, y = prop_animal_protein, group = data_source),
            colour = "black", linetype = "dashed", linewidth = 0.5, alpha = 0.7) +
  facet_wrap(~ habitat_method) +
  scale_fill_manual(values = c("#1179be", "#b1d0ed")) +
  labs(fill = "Consumption source", x = "", y = "Aquatic Animal Reliance (%)") +
  theme_minimal(base_size = 8) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(color = scales::alpha("black", 0.25), linewidth = 0.3),
    panel.grid.minor.y = element_blank(),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),
    strip.background = element_blank(),
    strip.text = element_text(size = 7, face = "bold"),
    axis.text = element_text(size = 6, color = "black"),
    axis.title.y = element_text(size = 7, margin = margin(r = 5), face = "bold"),
    axis.text.x = element_text(size = 5.5, angle = 30, hjust = 1),
    legend.position = "bottom",
    legend.title = element_text(size = 7, face = "bold"),
    legend.text = element_text(size = 6),
    legend.key.size = unit(0.3, "cm"),
    legend.box = "vertical",
    legend.spacing = unit(0.1, "cm"),
    legend.margin = margin(t = 0, b = 0),
    legend.box.margin = margin(t = -2, b = -2),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA)
  )



(aa_reliance_mean_global_trends_plot <- cowplot::plot_grid(
  aa_reliance_mean_global_trends,
  aa_reliance_multiple_trends,
  ncol = 2,
  rel_heights = c(0.8, 2.2),  # Give more space to the faceted plot
  labels = c("A", "B"),
  label_size = 10
))

ggsave("../images/aa_reliance_mean_global_trends.jpg", plot = aa_reliance_mean_global_trends_plot, device = "jpg", units = "in", width = 5, height = 3, bg = "transparent")

?ggsave()
```

## Statistic: X Countries have dominant reliance on X source

```{r X Countries have dominant reliance on X source}
# X Countries have dominant reliance on X source
aquatic_reliance %>%
  filter(year == 2020, food_group == "aquatic") %>%
  group_by(iso3c) %>%
  mutate(largest_group = prop_animal_protein == max(prop_animal_protein)) %>%
  filter(largest_group == TRUE) %>%
  ungroup() %>%
  count(method)
```

> Foreign marine capture reliance is the most dominant across 86 countries, while domestic inland aquaculture is more concentrated at 12 countries. 

## Statistic: How many countries have received X % of their animal source protein from marine vs inland

```{r X Countries have dominant reliance on X source}
# X Countries have dominant reliance on X source
aquatic_reliance %>%
  filter(year == 2020, food_group == "aquatic") %>%
  group_by(iso3c, habitat) %>%
  summarize(prop_animal_protein = sum(prop_animal_protein)) %>%
  mutate(individual_prop = round(prop_animal_protein / sum(prop_animal_protein),2)) %>%
  group_by(habitat) %>%
  count(individual_prop >= 0.8)
```

> Marine sources remain the most widespread reliant habitat, with 102 countries obtaining least 80% of their aquatic animal proteins from marine habitats, while only 18 sources at least 80% from inland sources. 

## Figure: Look at contributing production sources to a country's rate of change in their aggregate aquatic animal reliance

This panel visualizes for each country separated by region, their average annual change in their aggregate aquatic animal reliance, with separate colors as stacked bars representing the disaggregate reliance. The regions include Africa, Asia, South America, North America, Europe, and Oceania.

```{r all 6 continents visualization}
# Fit linear models to estimate the average annual change in disaggregated
# aquatic animal reliance (2010-2020) by production source (habitat x method)
# for each country
disagg <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "aquatic") %>%
  group_by(year, iso3c, habitat, method) %>%
  summarize(prop_animal_protein = sum(prop_animal_protein)) %>%
  group_by(iso3c, habitat, method) %>%
  nest() %>%
  mutate(
    model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
    tidy  = map(model, broom::tidy)
  ) %>%
  unnest(tidy) %>%
  filter(term == "year") %>%
  mutate(
    consource_habitat_method = paste(habitat, method),
    group_label = consource_habitat_method,
    group_label = case_when(
      group_label == "unknown unknown" ~ "unknown",
      TRUE ~ group_label
    )
  ) %>%
  drop_na(estimate) %>%
  group_by(group_label) %>%
  mutate(pct_above = mean(estimate > 0) * 100) %>%
  ungroup()

# Fit linear models to estimate the average annual change in aggregate
# aquatic animal reliance (2010-2020)
agg <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, year) %>%
  summarise(prop_animal_protein = sum(prop_animal_protein), .groups = "drop") %>%
  group_by(iso3c) %>%
  nest() %>%
  mutate(
    model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
    tidy  = map(model, broom::tidy)
  ) %>%
  unnest(tidy) %>%
  filter(term == "year") %>%
  mutate(group_label = "aggregate") %>%
  drop_na(estimate) %>%
  group_by(group_label) %>%
  mutate(pct_above = mean(estimate > 0) * 100) %>%
  ungroup()

# Clean aggregate estimate data
agg_estimate <- bind_rows(agg, disagg) %>%
  group_by(iso3c) %>%
  mutate(agg_est = estimate[group_label == "aggregate"][1]) %>%
  add_region("iso3c", "region") %>%
  ungroup() %>%
  mutate(
    iso3c  = fct_reorder(iso3c, -agg_est),
    is_agg = group_label == "aggregate"
  ) %>%
  filter(group_label == "aggregate") %>%
  select(iso3c, agg_estimate = estimate)

# Top row of plots
top_row <- disagg %>%
  group_by(iso3c) %>%
  add_region("iso3c", "region") %>%
    filter(!region %in% c("Oceania", "South America")) %>%
  ungroup() %>%
    left_join(agg_estimate, by = "iso3c") %>%
  mutate(
    iso3c  = fct_reorder(iso3c, -agg_estimate),
  ) %>%
  ggplot(aes(x = estimate, y = iso3c, fill = group_label)) +

  geom_col() +
    geom_point(aes(x = agg_estimate, y = iso3c), shape = 16) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.4) +

  scale_alpha_identity() +
  scale_size_identity() +
  scale_fill_manual(
    values = c("#D8C84A","#E68A2E", "#7A5FA3", "#3A77B5", "gray")
  ) +
  labs(
    x = "Average annual change in aquatic animal reliance (2010–2020)",
    y     = "Country",
    fill = "Production source",
  ) +

  facet_wrap(~ region, scales = "free", nrow = 1) +

  theme_minimal(base_size = 8) +
  theme(
    legend.position    = "bottom",
    panel.grid.minor   = element_blank(),
    axis.text = element_text(size = 4.75),
    legend.key.size    = unit(0.3, "cm"),
    legend.box         = "vertical",
    legend.spacing     = unit(0.1, "cm"),
    legend.text        = element_text(size = 5),
    legend.margin      = margin(t = 0, b = 0),
    axis.title.y = element_blank(),
    panel.grid.major.y = element_line(color = scales::alpha("black", 0.05), linewidth = 0.3),
    panel.grid.major.x = element_line(color = scales::alpha("black", 0.25)),
      axis.title.x = element_blank()
  ) +
    guides(fill = "none")


# Bottom row of plots
bottom_row <- disagg %>%
  group_by(iso3c) %>%
  add_region("iso3c", "region") %>%
    filter(region %in% c("Oceania", "South America")) %>%
  ungroup() %>%
    left_join(agg_estimate, by = "iso3c") %>%
  mutate(
    iso3c  = fct_reorder(iso3c, -agg_estimate),
  ) %>%
  ggplot(aes(x = estimate, y = iso3c, fill = group_label)) +

  geom_col() +
    geom_point(aes(x = agg_estimate, y = iso3c), shape = 16) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.4) +

  scale_alpha_identity() +
  scale_size_identity() +
  scale_fill_manual(
    values = c("#D8C84A","#E68A2E", "#7A5FA3", "#3A77B5", "gray")
  ) +
  labs(
    x = "Average annual change in aquatic animal reliance (2010–2020)",
    y     = "Country",
    fill = "Production source",
  ) +

  facet_wrap(~ region, scales = "free") +

  theme_minimal(base_size = 8) +
  theme(
    axis.title.x = element_text(face = "bold"),
    legend.position    = "bottom",
    panel.grid.minor   = element_blank(),
    axis.text        = element_text(size = 5.35),
    legend.key.size    = unit(0.3, "cm"),
    legend.box         = "vertical",
    legend.spacing     = unit(0.1, "cm"),
    legend.text        = element_text(size = 5),
    legend.margin      = margin(t = 0, b = 0),
    axis.title.y = element_blank(),
    panel.grid.major.y = element_line(color = scales::alpha("black", 0.05), linewidth = 0.3),
    panel.grid.major.x = element_line(color = scales::alpha("black", 0.25))
  )

bottom_row_narrowed <- cowplot::plot_grid(
  NULL, bottom_row, NULL,
  ncol = 3,
  rel_widths = c(0.1, 0.3, 0.1)  # adjust 0.8 to make it narrower/wider
)

(bottom_panel <- cowplot::plot_grid(top_row, bottom_row_narrowed, nrow = 2, rel_heights = c(2,1)))

top_panel <- agg %>%
  count(num_countries = estimate > 0) %>%
  ggplot(aes(x = 1, y = n, fill = num_countries)) +
  coord_flip() +
  geom_col(position = "stack") +
  labs(y = "Number of countries", fill = "Direction of change") +
  scale_fill_manual(values = c("#b2182b", "#2166ac"),
                    labels  = c(`FALSE` = "Declining reliance",
              `TRUE` = "Increasing reliance")) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text = element_text(size = 4.75),
        axis.title.y = element_blank(),
        axis.title.x = element_text(size = axis.title.size-2, face = "bold"),
    panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_line(color = scales::alpha("black", 0.25), linewidth = 0.3),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),
    legend.position = "bottom",
    legend.title = element_text(size = 7, face = "bold"),
    legend.text = element_text(size = 6),
    legend.key.size = unit(0.25, "cm"),
    legend.box = "vertical",
    legend.spacing = unit(.05, "cm"),
    legend.margin = margin(t = 0, b = 0),
    legend.box.margin = margin(t = -2, b = -2),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    aspect.ratio = 0.05)
  
(final_plot <- cowplot::plot_grid(top_panel, bottom_panel, nrow = 2, rel_heights = c(0.35,2),
                                  labels = c("A", "B"),label_fontface = "bold", label_size = 9.5))


ggsave("../images/pct_reliance_contribution_four_sources.jpg", plot = final_plot, device = "jpg", units = "in", width = 8, height = 6, dpi = 500)
```

# RQ 1b: How does aquatic animal reliance change with the relative contribution of aquatic versus terrestrial animal consumption

## Figure: Scatterplot (terrestrial consumption ~ aquatic consumption; colored by: aquatic reliance) + bar chart overlay 

Scatterplot of average annual change in daily terrestrial animal protein consumption (kg/capita/day) 2010–2020 (y-axis) with respect to the average annual change in daily aquatic animal protein consumption (kg/capita/day) 2010–2020 colored by the annual change in daily aquatic animal protein reliance(%) 2010-2020, overlaid by a 1:1 line. The bar plots represent the number of countries in each quadrant, with the highest count of countries in any quadrant being 21. The two right hand y-axes represent these number of countries.

```{r Region counts overlayed on quadrant plot}
# Fit linear models to estimate the average annual change in aquatic protein
# consumption (2010-2020)
aquatic_protein_quant <- aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, year) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap, na.rm = TRUE)) %>%
  distinct() %>%
  split(.$iso3c) %>%
  map(~lm(protein_consumed_kg_percap~year, data = .x)) %>%
  map_df(broom::tidy, .id = 'iso3c') %>%
  filter(term == 'year') %>%
  select(c("iso3c", "aquatic_estimate" = "estimate"))


# Fit linear models to estimate the average annual change in terrestrial protein
# consumption (2010-2020)
terrestrial_protein_quant <- aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         data_source == "new FBS", food_group == "terrestrial") %>%
  group_by(iso3c, year) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap, na.rm = TRUE)) %>%
  distinct() %>%
  split(.$iso3c) %>%
  map(~lm(protein_consumed_kg_percap~year, data = .x)) %>%
  map_df(broom::tidy, .id = 'iso3c') %>%
  filter(term == 'year') %>%
  select(c("iso3c", "terrestrial_estimate" = "estimate"))

# Fit linear models to estimate the average annual change in aggregate
# global aquatic animal reliance (2010-2020) 
aquatic_protein_reliance <- aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, year) %>%
  summarize(prop_animal_protein = sum(prop_animal_protein, na.rm = TRUE)) %>%
  distinct() %>%
  split(.$iso3c) %>%
  map(~lm(prop_animal_protein~year, data = .x)) %>%
  map_df(broom::tidy, .id = 'iso3c') %>%
  filter(term == 'year') %>%
  select(c("iso3c", "aquatic_reliance_estimate" = "estimate"))
  
#join data 
plot_df<- aquatic_protein_quant %>%
  full_join(terrestrial_protein_quant, by="iso3c") %>%
  full_join(aquatic_protein_reliance, by="iso3c") %>%
  add_region(col = "iso3c", region.col.name = "region")


# Create color scale for regions
region_colors <- c(
  "Africa"        = "#b2182b",  # deep red
  "Asia"          = "#2166ac",  # deep blue
  "Europe"        = "#6baed6",  # light blue
  "North America" = "#ef8a62",  # light red/salmon
  "Oceania"       = "#67a9cf",  # mid blue
  "South America" = "#d6604d"   # mid red
)

# Get region legend from bar plot
p_bar <- plot_df %>%
  mutate(quadrant = case_when(
    aquatic_estimate > 0 & terrestrial_estimate > 0 ~ "Q1",
    aquatic_estimate < 0 & terrestrial_estimate > 0 ~ "Q2",
    aquatic_estimate < 0 & terrestrial_estimate < 0 ~ "Q3",
    aquatic_estimate > 0 & terrestrial_estimate < 0 ~ "Q4"
  )) %>%
  group_by(quadrant, region) %>%
  summarize(n = n_distinct(iso3c), .groups = "drop") %>%
  mutate(quadrant = factor(quadrant, levels = c("Q2","Q1","Q3","Q4"))) %>%
  ggplot(aes(x = region, y = n, fill = region)) +
  geom_col(width = 0.75) +
  scale_fill_manual(values = region_colors) +
  labs(x = "", y = "Number of countries", fill = "Region") +
  theme_minimal(base_size = 7) +
  theme(
    axis.text.x = element_blank(),
    legend.position = "bottom",
    legend.title = element_text(size = 6, face = "bold"),
    legend.text  = element_text(size = 5),
    legend.key.size = unit(0.6, "lines"),
    axis.title = element_text(size = 5)
  ) +
  facet_wrap(~ quadrant)

legend_regions <- get_plot_component(p_bar, "guide-box-bottom", return_all = TRUE)

# Scatterplot with legend
p_scatter_base <- plot_df %>%
  filter(iso3c != "MDV") %>%
  ggplot(aes(
    x = aquatic_estimate,
    y = terrestrial_estimate,
    fill = aquatic_reliance_estimate,
    label = iso3c
  )) +
  geom_point(shape = 21, size = 1.8, color = "black", stroke = 0.2) +
  lims(x = c(-0.25,0.25),
       y = c(-4.5,4.5)) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  geom_text_repel(size = 2.4) +
  scale_fill_gradient2(
  low = "#ca0020",     # muted red
  mid = "#f7f7f7",     # very light grey
  high = "#0571b0",    # muted blue
  midpoint = 0
) +
  labs(
    x = "Average annual change in daily aquatic animal protein consumption (kg/capita/day) 2010–2020",
    y = "Average annual change in daily\nterrestrial animal protein consumption (kg/capita/day) 2010–2020",
    fill = "Annual change in daily aquatic\nanimal protein reliance (%) 2010-2020"
    
    
  ) +
  theme_minimal(base_size = 8) +
  theme(
    panel.grid.minor = element_blank(),
    axis.text = element_text(size = 6, color = "black"),
    axis.title = element_text(size = 5, face = "bold"),
    legend.position = "bottom",
    legend.title = element_text(size = 6, face = "bold"),
    legend.text = element_text(size = 5, angle = 35),
    legend.key.size = unit(.8, "lines")
  )

# Extract aquatic reliance legend
legend_reliance <- get_plot_component(p_scatter_base, "guide-box-bottom", return_all = TRUE)

# Remove legend from scatter for main panel
p_scatter <- p_scatter_base + theme(legend.position = "none")

x_max <- max(abs(plot_df$aquatic_estimate), na.rm = TRUE)
y_max <- max(abs(plot_df$terrestrial_estimate), na.rm = TRUE)

# p_scatter <- p_scatter +
#   coord_cartesian(
#     xlim = c(-0.5,0.5),
#     ylim = c(-y_max, y_max)
#   )

# Quadrant counts + insets
quad_counts <- plot_df %>%
  mutate(quadrant = case_when(
    aquatic_estimate > 0 & terrestrial_estimate > 0 ~ "Q1",
    aquatic_estimate < 0 & terrestrial_estimate > 0 ~ "Q2",
    aquatic_estimate < 0 & terrestrial_estimate < 0 ~ "Q3",
    aquatic_estimate > 0 & terrestrial_estimate < 0 ~ "Q4"
  )) %>%
  group_by(quadrant, region) %>%
  summarise(n = n_distinct(iso3c), .groups = "drop")

max_count <- max(quad_counts$n)

make_inset <- function(q) {
  show_y <- q %in% c("Q1", "Q4")
  
  quad_counts %>%
    filter(quadrant == q) %>%
    ggplot(aes(x = region, y = n, fill = region)) +
    geom_col(width = 0.75) +
    scale_fill_manual(values = region_colors) +
    scale_y_continuous(limits = c(0, max_count),
                       breaks = c(0, max_count),
                       position = "right") +
    labs(y = if (show_y) "No. countries" else NULL) +  # label only on right insets
    theme_minimal(base_size = 6) +
    theme(
      axis.title.x  = element_blank(),
      axis.title.y  = if (show_y) element_text(size = 5, color = "black", angle = 90) else element_blank(),
      axis.text.x   = element_blank(),
      axis.text.y   = if (show_y) element_text(size = 5, color = "black") else element_blank(),
      axis.ticks.y  = if (show_y) element_line(linewidth = 0.3) else element_blank(),
      panel.grid    = element_blank(),
      legend.position = "none",
      plot.margin   = margin(0, 0, 0, 0)
    )
}

p_q1 <- make_inset("Q1")
p_q2 <- make_inset("Q2")
p_q3 <- make_inset("Q3")
p_q4 <- make_inset("Q4")

# Add insets inside scatterplot

shrink_x <- 0.18
shrink_y <- 0.18

p_with_insets <- p_scatter +
  inset_element(p_q1, left = 1 - shrink_x, right = 1, bottom = 1 - shrink_y, top = 1) +
  inset_element(p_q2, left = 0, right = shrink_x, bottom = 1 - shrink_y, top = 1) +
  inset_element(p_q3, left = 0, right = shrink_x, bottom = 0, top = shrink_y) +
  inset_element(p_q4, left = 1 - shrink_x, right = 1, bottom = 0, top = shrink_y)

scatter_with_axis <- cowplot::plot_grid(
  p_with_insets,
  ncol = 2,
  rel_widths = c(1, 0.10)
)

# Add reliance and region legends side by side

legend_row <- cowplot::plot_grid(
  legend_reliance,
  legend_regions,
  ncol = 2,
  rel_widths = c(1, 1)
)

# Compile plot

(final_plot <- cowplot::plot_grid(
  scatter_with_axis,
  legend_row,
  ncol = 1,
  rel_heights = c(1, 0.18)
))


# Test for whether the median value of terrestrial protein consumption is greater than the aquatic animal protein consumption
wilcox.test(plot_df$terrestrial_estimate, plot_df$aquatic_estimate, paired = FALSE, alternative = "greater")

# Calculate CV for the terrestrial consumption roc
cv_terr <- sd(plot_df$terrestrial_estimate) / mean(plot_df$terrestrial_estimate)
 
 # Calculate CV for the aquatic animal consumption roc
cv_aquatic <- sd(plot_df$aquatic_estimate) / mean(plot_df$aquatic_reliance_estimate)

# Calcualte CV difference
cv_diff <- abs(cv_aquatic / cv_terr) %>% round(2)


ggsave("../images/scatter_bar_overlay.jpg", plot = final_plot, device = "jpg", units = "in", width = 5, height = 3.5)
```

> The variability in aquatic animal consumption is over 26 times greater in variability than terrestrial animal consumption by a factor of `cv_diff`

# RQ 2: What biophysical and socioeconomic factors are associated with reliance on farmed and wild aquatic animals for protein and overall consumption?

## Model: Hierarchical Bayesian regression

```{r hierarchical bayesian model}
# Variable list with EEZ area instead of coastline (km)
# predictors <- c(
#   "log_gdp",
#   "gdp_trade",
#   "eez_total",
#   "inundation_area_log1000km2",
#   "supermarkets",
#   "pop_10k_pct_coast",
# )

# Variable list with coastline (km) instead of EEZ area
predictors <- c(
  "log_gdp",
  "gdp_trade",
  "coastline_km",
  "inundation_area_log1000km2",
  "supermarkets",
  "pop_10k_pct_coast",
  "eez_total"
)

model_type = "realm"


for (i in seq_along(response_vars)) {

  joined_df <- get(response_vars[i]) %>%
    left_join(ac_df, by = "iso3c") %>%
    left_join(ecoregion, by = "iso3c") %>%
    ungroup() %>%
    # add_region(col = "iso3c", region.col.name = "realm") %>%   # keep realm
    left_join(
      indicator_df %>%
        select("eez_total", "iso3c",
               "inundation_area_log1000km2" = "inland_water_max"),
      by = "iso3c"
    ) %>%
    left_join(coastline_df, by = "iso3c") %>%
    left_join(pop_10k_coast, by = "iso3c") %>%
    mutate(
      log_gdp = log(gdp),
      coastline_km = case_when(
        is.na(coastline_km) & str_detect(realm, "Landlocked") ~ 0,
        TRUE ~ coastline_km
      ),
      eez_total = case_when(
        is.na(eez_total) & str_detect(realm, "Landlocked") ~ 0,
        TRUE ~ eez_total
      )
    ) %>%
    select(iso3c, 2, all_of(predictors), realm) %>%
    drop_na() %>%
    filter(realm != "")

  # Design matrix
  X <- joined_df %>%
    select(all_of(predictors)) %>%
    mutate(across(everything(), ~ (.x - min(.x)) / (max(.x) - min(.x)))) %>%
    as.matrix()

  # Response
  y <- scale(joined_df %>% pull(2))[, 1]

  # Hierarchical grouping (province index from realm)
  province <- as.integer(factor(joined_df$realm))
  J <- length(unique(province))

  # Save X and y
  # saveRDS(X, paste0("../output/bayesian_output/", model_type, "/coastline_km_covariate/X_", response_vars[i], ".rds"))
  # saveRDS(y, paste0("../output/bayesian_output/", model_type, "/coastline_km_covariate/y_", response_vars[i], ".rds"))
  
  saveRDS(joined_df$realm,
        paste0("../output/bayesian_output/", model_type, "/coastline_km_covariate/realms_", response_vars[i], ".rds"))

  # Stan data
  stan_data <- list(
    N = nrow(X),
    K = ncol(X),
    J = J,
    X = X,
    y = y,
    province = province
  )

  # Inits matching your Stan model
  inits <- function() {
    list(
      beta_0       = 0,
      sigma_beta_0 = 0.5,
      z_beta_0     = rnorm(J, 0, 0.1),
      beta         = rep(0, ncol(X)),
      sigma        = 1
    )
  }

  # Fit Stan model
  fit <- stan(
    file = "../R/social_bayesian_model_hierarchical.stan",
    data = stan_data,
    init = inits,
    chains = 3,
    iter = 4000,
    warmup = 1000
  )

  # saveRDS(fit, paste0("../output/bayesian_output/fit_", response_vars[i], ".stan"))
  
  saveRDS(fit, paste0("../output/bayesian_output/", model_type, "/coastline_km_covariate/fit_", response_vars[i], ".rds"))
}
```

## Figure: Posterior distributions for the partial slope across models

Here, we map the posterior distributions for models of terrestrial animal consumption per capita, aquatic animal consumption per capita, capture consumption per capita, aquaculture consumption per capita, domestic consumption per capita, foreign consumption per capita, marine consumption per capita, inland consumption per capita, and aggregate aquatic animal reliance. Posterior distributions are displayed with 90%, 80%, and 50% credibility intervals, represented by the lightest, darker, and darkest colors respectively. The point represents the posterior median.

```{r hierarcical global variable posteriors}
model_type = "realm"

# Create a data frame of posteriors of partial slopes across Bayesian models
slopes_df <- data.frame()

for (i in seq_along(response_vars)) {
  
  # Extract MCMC samples
  x <- rstan::extract(readRDS(paste0("../output/bayesian_output/", model_type, "/fit_", response_vars[i],".stan")), permuted = T)
  beta_global <- x$beta       # iter × K (slopes)
  
  X <- readRDS(paste0("../output/bayesian_output/", model_type, "/X_", response_vars[i],".rds"))
  
  n_iter <- nrow(beta_global)
  K <- ncol(beta_global)
  vars <- colnames(X)

# Extract global slopes
post_long_slopes <- expand_grid(
  iter = 1:n_iter,
  k = 1:K
) %>%
  mutate(
    value = map2_dbl(iter, k, ~beta_global[.x, .y]),
    var = vars[k]
  )

# Calculate credibility intervals
ci_df <- post_long_slopes %>%
  group_by(var) %>%
  summarise(
    lo95 = quantile(value, 0.025),
    lo80 = quantile(value,0.1),
    lo50 = quantile(value, 0.25),
    med = quantile(value, 0.5),
    hi50 = quantile(value, 0.75),
    hi80 = quantile(value,0.9),
    hi95 = quantile(value, 0.975),
    .groups = "drop"
  ) %>%
  mutate(model =  response_vars[i])

slopes_df <- slopes_df %>%
  bind_rows(ci_df)

}

# Clean variable names
var_labels <- c(
  # coastline_km = "Coastline length (km)",
  eez_total = "EEZ area",
  gdp_trade = "GDP from trade (%)",
  inundation_area_log1000km2 = "Inundated area (log 1000 km²)",
  log_gdp = "GDP (log scale)",
  pop_10k_pct_coast = "Population within 10 km of coast (%)",
  supermarkets = "Supermarkets per 100,000 people"
)

# Set model colors for visualization
model_colors <- c(
  "Terrestrial animal consumption per capita" = "#8B4513",
  "Aquatic animal consumption per capita"     = "#3A77B5",
  "Aquatic reliance"                          = "#2166ac",
  "Capture consumption per capita"            = "#E68A2E",
  "Aquaculture consumption per capita"        = "#D8C84A",
  "Domestic consumption per capita"           = "#6BAED6",
  "Foreign consumption per capita"            = "#EF8A62",
  "Marine consumption per capita"             = "#7A5FA3",
  "Inland consumption per capita"             = "#D8A43A"
)

# Get colors for 90, 80, and 50% credibility intervals
col_95 <- lighten(model_colors, 0.65)
col_80 <- lighten(model_colors, 0.4)
col_50 <- lighten(model_colors, 0.10)

# Load in script to clean model names from underscores to be easily readable
source("../R/clean_model_name.R")

# 2. Add clean_name once
slopes_df <- slopes_df %>%
  mutate(clean_name = clean_model_name(model))

# 3. Build plotting data
df_plot <- slopes_df %>%
  mutate(
    clean_name = clean_model_name(model),
    nonzero = (lo95 < 0 & hi95 < 0) | (lo95 > 0 & hi95 > 0),
    suffix = sub(".*_", "", model),
    suffix_num = suppressWarnings(as.numeric(suffix))
  )  %>%
  filter(model != "animal_cons_cap") %>%
  mutate(
    nonzero    = (lo95 < 0 & hi95 < 0) | (lo95 > 0 & hi95 > 0),
    suffix     = sub(".*_", "", model),
    suffix_num = suppressWarnings(as.numeric(suffix))
  ) %>%
  filter(
    str_detect(clean_name, "capita") |
    str_detect(clean_name, "Aquatic reliance")
  ) %>%
  arrange(suffix_num, suffix) %>%
  mutate(
    clean_name = factor(clean_name, levels = unique(clean_name)),
    col95 = col_95[as.character(clean_name)],
    col80 = col_80[as.character(clean_name)],
    col50 = col_50[as.character(clean_name)],
    panel_id = as.numeric(factor(clean_name)),
    left_col = panel_id %% 3 == 1
  )

panel_tags <- letters[1:9] 

# Function to draw a single panel
make_panel <- function(df, show_yaxis = FALSE, tag = "") {
  ggplot(df, aes(y = var)) +
    geom_segment(aes(x = lo95, xend = hi95, color = col95),
                 linewidth = 2, lineend = "square") +
    geom_segment(aes(x = lo80, xend = hi80, color = col80),
                 linewidth = 2, lineend = "square") +
    geom_segment(aes(x = lo50, xend = hi50, color = col50),
                 linewidth = 2, lineend = "square") +
    geom_point(aes(x = med), size = 1.8, color = "black") +
    geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.4) +
    scale_color_identity() +
    scale_y_discrete(labels = var_labels) +

    # ← add tag in upper-left corner
    annotate("text", x = -Inf, y = Inf, label = tag,
             hjust = -0.3, vjust = 1.3, size = 3.2, fontface = "bold") +

    theme_minimal(base_size = 8) +
    theme(
      plot.title = element_blank(),
      panel.grid.major.y = element_blank(),
      panel.grid.minor   = element_blank(),
      strip.text         = element_text(size = 5, face = "bold"),
      axis.text.x        = element_text(size = 6, color = "black"),
      axis.title         = element_blank(),
      panel.spacing      = unit(1.2, "lines"),
      plot.margin        = margin(5, 5, 5, 5),
      axis.text.y        = if (show_yaxis) element_text(size = 6, color = "black") else element_blank(),
      axis.ticks.y       = if (show_yaxis) element_line() else element_blank()
    )
}

# Split df_plot into 9 panels
panel_names <- levels(df_plot$clean_name)

panels <- lapply(seq_along(panel_names), function(i) {
  name <- panel_names[i]
  df_sub <- df_plot %>% filter(clean_name == name)

  show_y <- i %in% c(1, 4, 7)
  wrapped_name <- stringr::str_wrap(name, width = 20)

  make_panel(df_sub, show_yaxis = show_y, tag = panel_tags[i]) +
    ggtitle(wrapped_name)
})

# Build the x-axis label as its own plot
xlab_plot <- ggplot() +
  theme_void() +
  annotate("text", x = 0.5, y = 0.5,
           label = "Posterior slope estimate",
           size = 2, fontface = "bold")

# Empty plot for left and right columns
empty <- ggplot() + theme_void()

# Assemble 3×3 grid + bottom row with centered x-label
final_plot_with_xlab <-
  (
    (panels[[1]] | panels[[2]] | panels[[3]]) /
    (panels[[4]] | panels[[5]] | panels[[6]]) /
    (panels[[7]] | panels[[8]] | panels[[9]]) /
    (empty | xlab_plot | empty)
  ) +
  plot_layout(heights = c(1, 1, 1, 0.12))

final_plot_with_xlab


ggsave(paste0("../images/bayesian plots/", model_type, "/posteriors_no_titles.jpg"), plot = final_plot_with_xlab, device = "jpg", width = 5, height = 3, units = "in", dpi = 300)
```

## Figure: Plot Maps of intercept median posteriors

The region specific intercept deviations from the slope line for the levels of daily animal protein consumption per capita (left) and aquatic animal reliance per capita (right). Outlined countries indicate that the 95% posterior credible interval does not include zero.
  
```{r realm_maps_robinson_final, message=FALSE, warning=FALSE}
# Get intercept posteriors
model_type = "realm"
intecrept_df <- data.frame()
for (i in seq_along(response_vars)) {
  # Extract MCMC samples
  x <- rstan::extract(readRDS(paste0("../output/bayesian_output/", model_type, "/fit_", response_vars[i],".stan")), permuted = T)
  beta_global <- x$beta       # iter × K (slopes)
  
  X <- readRDS(paste0("../output/bayesian_output/", model_type, "/X_", response_vars[i],".rds"))
  
  
  
  n_iter <- nrow(beta_global)
  K <- ncol(beta_global)
  vars <- colnames(X)
  
  # Posteriors of the realm intercepts
  beta_0i <- x$beta_0i   # iterations × J

  n_iter <- nrow(beta_0i)
  J <- ncol(beta_0i)

  # Get corresponding realm indicators (e.g., Europe, Oceania) to match with intercept values
realm_level <- realm_levels <- readRDS(
  paste0("../output/bayesian_output/", model_type, "/realms_", response_vars[i], ".rds")
)
  
  
  post_intercepts <- expand_grid(
  iter = 1:n_iter,
  realm = 1:J
) %>%
  mutate(
    value = beta_0i[cbind(iter, realm)],
    realm = realm_level[realm]
  )

  ci_intercepts <- post_intercepts %>%
  group_by(realm) %>%
  summarise(
    lo95 = quantile(value, 0.025),
    lo50 = quantile(value, 0.25),
    med  = quantile(value, 0.5),
    hi50 = quantile(value, 0.75),
    hi95 = quantile(value, 0.975),
    .groups = "drop"
  ) %>%
  mutate(model = response_vars[i])
  
  intecrept_df <- intecrept_df %>%
    bind_rows(ci_intercepts)
}


# Load MEOW ecoregions
meow <- st_read("../data/MEOW/meow_ecos.shp")

# Realm levels
realm_levels <- c(
  "Arctic","Central Indo-Pacific","Eastern Indo-Pacific",
  "Landlocked, Africa","Landlocked, Asia","Landlocked, Europe",
  "Landlocked, South America","Southern Ocean","Temperate Australasia",
  "Temperate Northern Atlantic","Temperate Northern Pacific",
  "Temperate South America","Temperate Southern Africa",
  "Tropical Atlantic","Tropical Eastern Pacific","Western Indo-Pacific"
)

marine_realms <- realm_levels[!str_detect(realm_levels, "Landlocked")]
landlocked_realms <- realm_levels[str_detect(realm_levels, "Landlocked")]

#  Dissolve MEOW ecoregions into REALM polygons
meow_realms <- meow %>%
  group_by(REALM) %>%
  summarise(geometry = st_union(geometry), .groups = "drop") %>%
  rename(realm = REALM)

# Split polygons at the antimeridian
meow_realms <- meow_realms %>%
  st_wrap_dateline(options = c("WRAPDATELINE=YES", "DATELINEOFFSET=180")) %>%
  st_collection_extract("POLYGON")


# Build landlocked realms using CONTINENTS
# but trim only the Russia sliver east of 180 degrees
countries <- ne_countries(scale = "medium", returnclass = "sf")

# Trim Russia's far-east sliver that bleeds onto the map as a horizontal line if not fixed
russia <- countries %>% filter(iso_a3 == "RUS")

russia_trimmed <- st_difference(
  russia,
  st_as_sfc(st_bbox(c(
    xmin = 180, ymin = -90, xmax = 200, ymax = 90
  ), crs = 4326))
)

countries_fixed <- countries %>%
  filter(iso_a3 != "RUS") %>%
  bind_rows(russia_trimmed)

# Clean continent basemap (for always-visible land)
continent_basemap <- countries_fixed %>%
  group_by(continent) %>%
  summarise(geometry = st_union(geometry), .groups = "drop") %>%
  st_make_valid() %>%
  st_wrap_dateline(options = c("WRAPDATELINE=YES", "DATELINEOFFSET=180")) %>%
  st_collection_extract("POLYGON") %>%
  st_transform(crs = robin)

# Build continent unions (now safe)
continents <- countries_fixed %>%
  mutate(continent = factor(continent)) %>%
  group_by(continent) %>%
  summarise(geometry = st_union(geometry), .groups = "drop")

# Map landlocked realms to continents
landlocked_lookup <- tibble(
  realm = landlocked_realms,
  continent = str_remove(landlocked_realms, "Landlocked, ")
)

landlocked_geom <- landlocked_lookup %>%
  left_join(continents, by = "continent") %>%
  select(realm, geometry)

# Combine marine + landlocked geometries
realm_geom <- bind_rows(
  meow_realms %>% filter(realm %in% marine_realms),
  landlocked_geom
)

# Project to Robinson
robin <- "+proj=robin +lon_0=0 +datum=WGS84 +units=m +no_defs"
realm_geom_robin <- st_transform(realm_geom, crs = robin)

# Mapping function with significance outlines
selected_df <- intecrept_df %>%
  filter(model %in% c("terr_animal_cons_cap",
                      "aquatic_rel",
                      "aquaculture_rel",
                      "capture_rel"
                      ))

global_min <- min(selected_df$med, na.rm = TRUE)
global_max <- max(selected_df$med, na.rm = TRUE)

global_limit <- max(abs(global_min), abs(global_max))

make_realm_map <- function(model_name, df = intecrept_df) {

  # Filter model + significance flag
  df_model <- df %>%
    filter(model == model_name) %>%
    mutate(significant = (lo95 > 0 & hi95 > 0) | (lo95 < 0 & hi95 < 0))

  # Join model values to realm geometry
  geom_joined <- realm_geom_robin %>%
    left_join(df_model, by = "realm") %>%
    filter(!is.na(st_is_empty(geometry)) & !st_is_empty(geometry))

  ggplot() +
    # Basemap (always visible)
    geom_sf(
      data = continent_basemap,
      fill = "darkgray",
      color = "grey70",
      size = 0.2
    ) +

    # Realm shading (outline only for non-NA)
    geom_sf(
      data = geom_joined,
      aes(fill = med, color = !is.na(med)),
      size = 0.2
    ) +
    scale_color_manual(
      values = c("TRUE" = "grey30", "FALSE" = NA),
      guide = "none"
    ) +

    # Significance outlines
    geom_sf(
      data = geom_joined %>% filter(significant),
      fill = NA,
      color = "#1F1F1F",
      linewidth = 0.5
    ) +

    # Global symmetric fill scale (computed outside this function)
    scale_fill_gradient2(
      low = "#b2182b",
      mid = "grey85",
      high = "#2166ac",
      midpoint = 0,
      limits = c(-global_limit, global_limit),
      oob = scales::squish
    ) +

    # Hide legend inside each subplot
    theme_realm_panel() +
    theme(legend.position = "none") +

    labs(
      title = paste(model_name),
      fill = "Posterior\nmedian\nintercept"
    )
}
  

# Produce maps for all models
all_models <- unique(intecrept_df$model)
realm_maps <- map(all_models, ~ make_realm_map(.x))
names(realm_maps) <- all_models

theme_realm_panel <- function() {
  theme_minimal(base_size = 8) +
    theme(
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.y = element_line(color = scales::alpha("black", 0.25), linewidth = 0.3),
      panel.grid.minor.y = element_blank(),
      panel.background = element_rect(fill = "transparent", color = NA),
      plot.background = element_rect(fill = "transparent", color = NA),
      axis.text = element_text(size = 6, color = "black"),
      axis.title = element_blank(),
      legend.position = "bottom",
      legend.title = element_text(size = 7, face = "bold", hjust = 0.5),
      legend.text = element_text(size = 6),
      legend.key.size = unit(0.5, "cm"),
      legend.box = "vertical",
      legend.spacing = unit(0.1, "cm"),
      legend.margin = margin(t = 0, b = 0),
      legend.box.margin = margin(t = -2, b = -2),
      legend.background = element_rect(fill = "transparent", color = NA),
      legend.box.background = element_rect(fill = "transparent", color = NA)
    )
}

# Create individual panels
p1 <- realm_maps[["animal_cons_cap"]] +
  labs(title = "Daily animal protein consumption per capita") +
  theme_realm_panel() +
  theme(plot.title = element_text(hjust = 0.5, size = 8, face = "bold"))

p2 <- realm_maps[["aquatic_rel"]] +
  labs(title = "Aquatic animal reliance per capita") +
  theme_realm_panel() +
  theme(plot.title = element_text(hjust = 0.5, size = 8, face = "bold"))

# Create combined panel plot
(plot <- (p1 | p2) +
  plot_layout(guides = "collect") &
  theme(
    legend.position = "bottom",
    aspect.ratio = 0.45,   # ← try 0.40–0.55 depending on taste
    plot.margin = margin(0, 0, 0, 0) 
  ))

# Save plot
ggsave(paste0("../images/bayesian plots/", model_type, "/intercept_map.jpg"), plot = plot, device = "jpg", width = 5, height = 4, units = "in", dpi = 300)
```