---
title: "02_results"
author: "Connor Quiroz, Lana Huang, Christopher D. Golden, Sarah Glaser, and Jessica A. Gephart"
date: "2025-09-05"
output: html_document
---

# Chapter 1 Results

# Load in packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(exploreARTIS)
library(ggrepel)
library(countrycode)
library(colorspace)
library(ggpubr)
library(arrow)
library(rstan)
library(sf)
library(lwgeom)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggridges)
library(bayesplot)
library(purrr)
library(patchwork)
library(dplyr)
library(ggplot2)
library(forcats)
library(stringr)
library(grid)
library(broom)
library(ggtext)
library(progress)
```

# Read in data

## Read in aquatic reliance data & FAO population data

```{r read in FAO derived data}
# Read in population
population_fao_wb <- read.csv("../output/population_fao_wb.csv")

# Read in aquatic reliance data
aquatic_reliance <- read.csv("../output/fao_aquatic_reliance_source.csv") %>%
  left_join(population_fao_wb,by=c("year","iso3c"))%>%
  mutate(protein_consumed_kg_percap=protein_consumed_t*1000/population_fao)

# Read in FAO sourcing data
aquatic_fbs <- read.csv("../output/fao_fbs_aquatic_source.csv")
```

## Load in biophysical and socio-economic data as covariates into the Bayesian model

```{r load data}
# Read in adaptive capacity data
ac_df <- read_parquet("../data/e_s_ac_data_total.parquet") %>%
  filter(scenario == "ssp126") %>%
  select(-aa_reliance_pct, -scenario, -change_in_stock,-live_weight_t,-pct_change) %>%
  rename(iso3c = "consumer_iso3c")
  
# Read in ecoregion data (all shapes associated with soverign)
ecoregion_full <- read.csv("../data/Spatially joined ecoregion to EEZ/ecoregion.csv") %>%
  rename("iso3c" = "ISO_SOV1",
         ecoregion = "ECOREGION",
         province = "PROVINCE",
         realm = "REALM",
         shape_area = "Shape_Area")

# Read in ecoregions for largest soverign shape
ecoregion_unique <- read.csv("../data/Spatially joined ecoregion to EEZ/ecoregion_clean.csv") %>%
  select(-FREQUENCY) %>%
  rename(iso3c = "ISO_SOV1",
         shape_area = "MAX_Shape_Area")

# Read in dietary sourcing flexibility index
dsfi_df <- read_csv("../data/covariates/food-systems-dashboard-2026-01-20 (dietary sourcing flexibility index).csv") %>%
  filter(`Start Year` == 2019 | `End Year` == 2019) %>%
  select(c("iso3c" = ISO3, "dsfi" = Value))

# Read in national indicator data
indicator_df <- read_csv("../data/covariates/all_national_indicators.csv")

# Read in coastline data
coastline_df <- read_csv("../data/covariates/countries-by-coastline-2026.csv") %>%
  mutate(iso3c = case_when(!is.na(flagCode) ~ countrycode(flagCode, origin = "iso2c", destination = "iso3c"),
                           country == "Namibia" ~ "NAM")) %>%
  select(c(iso3c, "coastline_km" = "CountryCoastlineKm"))

# Read in proportion of people within 10k of the coast
pop_10k_coast <- read.csv("../data/covariates/pop_10k_coast.csv")
```

## Clean ecoregion data

Impute countries with no denoted ecoregion as "landlocked"

```{r clean ecoregion data}
# Join largest shape for the respective soverign country to their ecoregion names
ecoregion <- left_join(ecoregion_unique, ecoregion_full, by = c("iso3c", "shape_area"))

# Assign countries not associated with a marine ecoregion to "landlocked" code
landlocked_countries <- aquatic_reliance %>%
  distinct(iso3c) %>%
  drop_na() %>%
  add_region(col = "iso3c", region.col.name = "region") %>%
  filter(!iso3c %in% ecoregion$iso3c) %>%
  # mutate(ecoregion = "Landlocked",
  #        province = "Landlocked",
  #        realm = "Landlocked") %>%
  mutate(ecoregion = paste0("Landlocked, ", region),
         province = paste0("Landlocked, ", region),
         realm = paste0("Landlocked, ", region)) %>%
  select(-region)

# Clean ecoregion data
ecoregion <- bind_rows(ecoregion, landlocked_countries) %>%
  arrange(iso3c) %>%
  select(-shape_area)

# Clean up files
rm(list = c("ecoregion_full", "ecoregion_unique"))

# Write to csv
# write_csv(ecoregion, "../output/ecoregion_clean.csv")
```

## Create list of response variables & their associated data to put through Bayesian model (See RQ2 for Bayesian model code)

This code chunk is also in the main results. This code chunk extracts the name of each model name used in the Bayesian model. In code lines further down, we call the "response_vars" object to generate statistics and figures.

```{r acquire animal consumption data to be fed through analysis}
if ( !file.exists("../output/response_vars.rds") ) {
  
  # Calculate animal consumption
terr_animal_cons_cap <- aquatic_reliance %>%
  filter(year == 2019) %>%
  group_by(iso3c) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))

# Calculate aquatic animal consumption
aquatic_animal_cons_cap <- aquatic_reliance %>%
  filter(year == 2019, food_group == "aquatic") %>%
  group_by(iso3c) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))

# Calculate animal reliance consumption
aquatic_rel <- aquatic_animal_cons_cap %>%
  left_join(animal_cons_cap, by = "iso3c") %>%
  group_by(iso3c) %>%
  summarize(prop_animal_protein = protein_consumed_kg_percap.x / protein_consumed_kg_percap.y)

# Add items not created in the loop to the list
response_vars <- c("terr_animal_cons_cap", "aquatic_animal_cons_cap", "aquatic_rel")

for (i in c("capture", "aquaculture", "domestic", "foreign", "marine", "inland")) {
  
  # Calculate consumption per capita by method
  if (i %in% c("capture", "aquaculture")) {
    assign(
      paste0(i, "_cap"),
      aquatic_reliance %>%
        filter(year == 2019, food_group == "aquatic", method == i) %>%
        group_by(iso3c) %>%
        summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))
    )
  }
  
  # Calculate consumption per capita by consumption source
  if (i %in% c("domestic", "foreign")) {
    assign(
      paste0(i, "_cap"),
      aquatic_reliance %>%
        filter(year == 2019, food_group == "aquatic", consumption_source == i) %>%
        group_by(iso3c) %>%
        summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))
    )
  }
  
  # Calculate consumption per capita by habitat
  if (i %in% c("marine", "inland")) {
    assign(
      paste0(i, "_cap"),
      aquatic_reliance %>%
        filter(year == 2019, food_group == "aquatic", habitat == i) %>%
        group_by(iso3c) %>%
        summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))
    )
  }
  
  response_vars <- c(response_vars, paste0(i, "_cap"))
  
  # Calculate reliance by method, production source, or habitat
  assign(
    paste0(i, "_rel"),
    get(paste0(i, "_cap")) %>%
      left_join(animal_cons_cap, by = "iso3c") %>%
      group_by(iso3c) %>%
      summarize(prop_animal_protein = protein_consumed_kg_percap.x / protein_consumed_kg_percap.y)
  )
  
  response_vars <- c(response_vars, paste0(i, "_rel"))
  
}

# Save response_vars to RDS format
saveRDS(response_vars, "../data/response_vars.rds")
}
```

# RQ 1a: How do aquatic food reliance trends disaggregated by source differ from aggregate aquatic food reliance?

## Figure: Compare aggregate to disaggregate reliance

Left plot: trends in global total aquatic protein (blue) reliance with respect to the terrestrial protein reliance (red). The jump in data at 2010 is the visualization data switching from FAO-FBS old methodology to the new methodology. Right plot: Trends in  aquatic animal protein reliance by sourcing, where each facet represents the habitat and method and the color represents the production source. Dotted lines represent forward and back casting of old / new methodology data to years as error estimates.

```{r compare aggregate to disaggregated reliance}
# For each year, calculate the share of global animal proteins that are sourced from aquatic animals. 
aa_reliance_mean_global_trends_alt <- aquatic_reliance %>%
  group_by(year, data_source, food_group) %>%
  summarise(food_group_total_protein_t = sum(protein_consumed_t)) %>%
  group_by(year, data_source) %>%
  mutate(prop_animal_protein = 100*food_group_total_protein_t/sum(food_group_total_protein_t)) %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS")) # Remove overlapping years, prioritizing new years

# Define year ranges for projections (add this before the existing projections code)
year_ranges <- tibble(
  data_source = c("old FBS", "new FBS"),
  start_year  = c(1996, 1996),
  end_year    = c(2020, 2020)
)

observed_ranges <- tibble(
  data_source = c("old FBS", "new FBS"),
  min_year = c(1996, 2010),
  max_year = c(2009, 2020)
)

# First plot: aggregate global trend with projections

coefs_global <- aa_reliance_mean_global_trends_alt %>%
  filter(food_group == "aquatic") %>%
  group_by(data_source) %>%
  nest() %>%
  mutate(model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
         tidy  = map(model, broom::tidy)) %>%
  unnest(tidy) %>%
  select(data_source, term, estimate) %>%
  pivot_wider(values_from = estimate, names_from = term) %>%
  rename(year_coef = year)

projections_global <- coefs_global %>%
  left_join(year_ranges, by = "data_source") %>%
  rowwise() %>%
  mutate(year = list(start_year:end_year)) %>%
  unnest(year) %>%
  ungroup() %>%
  mutate(prop_animal_protein = `(Intercept)` + year_coef * year,
         type = "projected") %>%
  left_join(observed_ranges, by = "data_source") %>%
  filter(year < min_year | year > max_year)

observed_global_old <- aa_reliance_mean_global_trends_alt %>%
  filter(food_group == "aquatic", data_source == "old FBS", year <= 2009) %>%
  mutate(type = "observed")

observed_global_new <- aa_reliance_mean_global_trends_alt %>%
  filter(food_group == "aquatic", data_source == "new FBS", year >= 2010) %>%
  mutate(type = "observed")

combined_global <- bind_rows(observed_global_old, observed_global_new)

aa_reliance_mean_global_trends <- ggplot() +
  geom_area(data = combined_global,
            aes(x = year, y = prop_animal_protein),
            fill = "#1179be") +
  geom_line(data = projections_global,
            aes(x = year, y = prop_animal_protein, group = data_source),
            colour = "black", linetype = "dashed", linewidth = 0.5, alpha = 0.7) +
  labs(x = "", y = "Global Animal Protein Reliance (%)") +
  theme_minimal(base_size = 8) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(color = scales::alpha("black", 0.25), linewidth = 0.3),
    panel.grid.minor.y = element_blank(),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),
    axis.text = element_text(size = 6, color = "black"),
    axis.title.y = element_text(size = 7, margin = margin(r = 5), face = "bold"),
    legend.position = "bottom",
    legend.title = element_text(size = 7, face = "bold"),
    legend.text = element_text(size = 6),
    legend.key.size = unit(0.3, "cm"),
    legend.box = "vertical",
    legend.spacing = unit(0.1, "cm"),
    legend.margin = margin(t = 0, b = 0),
    legend.box.margin = margin(t = -2, b = -2),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    aspect.ratio = 0.4
  )


# Calculate the total terrestrial animal protein
 animal_protein_sourcing <- aquatic_reliance %>%
  group_by(year, data_source, food_group) %>%
  summarise(animal_protein = sum(protein_consumed_t)) %>%
  filter(food_group == "terrestrial")

# Calculate the total aquatic animal protein
 aquatic_protein_sourcing <- aquatic_reliance %>%
  group_by(year, data_source, food_group) %>%
  summarise(aquatic_protein = sum(protein_consumed_t)) %>%
  filter(food_group == "aquatic")
 
 # Calculate total tonnage of combined terrestrial and aquatic animal protein
 total_protein <- left_join(animal_protein_sourcing, aquatic_protein_sourcing, by = c("year", "data_source")) %>%
   group_by(year, data_source) %>%
   summarize(total_protein = sum(animal_protein,aquatic_protein)) # add total aquatic animal protein (unseparated by sourcing) to animal proteins by year and data source
 
# Join total protein to aquatic protein weights separated by protein to calculate reliance by sourcing
 aa_reliance_multiple_trends_alt <- aquatic_reliance %>%
  group_by(year, data_source, consumption_source, habitat, method, food_group) %>%
  summarise(food_group_total_protein_t = sum(protein_consumed_t)) %>%
  filter(food_group == "aquatic") %>%
  left_join(total_protein, by = c("year", "data_source")) %>%
  ungroup() %>%
  mutate(prop_animal_protein = 100 * (food_group_total_protein_t / total_protein)) %>% # Divide AA / TERRESTRIAL + AA
  filter(habitat != "unknown") %>%
   filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS")) # Remove overlapping years, prioritizing new years
 
 
 aa_reliance_multiple_trends_alt_1 <- aa_reliance_multiple_trends_alt %>%
  mutate(habitat = str_to_title(habitat)) %>%
  mutate(habitat_method = paste(habitat, method, sep = " "))
 
# Create forward / backward projections
 coefs <- aquatic_reliance %>%
  group_by(year, data_source, habitat, method, food_group) %>%
  summarise(food_group_total_protein_t = sum(protein_consumed_t), .groups = "drop") %>%
  filter(food_group == "aquatic") %>%
  left_join(total_protein, by = c("year", "data_source")) %>%
  mutate(prop_animal_protein = 100 * (food_group_total_protein_t / total_protein)) %>%
  group_by(data_source, habitat, method) %>%   # <- KEEP data_source
  nest() %>%
  mutate(model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
         tidy  = map(model, broom::tidy)) %>%
  unnest(tidy) %>%
  select(data_source, habitat, method, term, estimate) %>%
  pivot_wider(values_from = estimate, names_from = term) %>%
  rename(year_coef = year)

projections <- coefs %>%
  left_join(year_ranges, by = "data_source") %>%
  rowwise() %>%
  mutate(year = list(start_year:end_year)) %>%
  unnest(year) %>%
  ungroup() %>%
  mutate(
    prop_animal_protein = `(Intercept)` + year_coef * year,
    habitat = str_to_title(habitat)
  ) %>%
  filter(method != "unknown", habitat != "Unknown") %>%
  select(year, data_source, habitat, method, prop_animal_protein) %>%
  mutate(type = "projected") %>%
  left_join(observed_ranges, by = "data_source") %>%
  filter(year < min_year | year > max_year)  # Keep only years outside observed range

 
# Artis palette
artis_colors <- artis_palette(3)

artis_colors <- artis_colors[c(2,1)]

# Prepare observed subsets without overlap
observed_old <- aa_reliance_multiple_trends_alt_1 %>%
  filter(data_source == "old FBS", year <= 2009) %>%
  mutate(type = "observed")

observed_new <- aa_reliance_multiple_trends_alt_1 %>%
  filter(data_source == "new FBS", year >= 2010) %>%
  mutate(type = "observed")

# Filter projections to non-overlapping years
projections_old <- projections %>%
  filter(data_source == "old FBS", year > 2009) %>%
  mutate(type = "projected")

projections_new <- projections %>%
  filter(data_source == "new FBS", year < 2010) %>%
  mutate(type = "projected")

# Combine all cleanly
combined <- bind_rows(observed_old, observed_new, projections_old, projections_new) %>%
  mutate(habitat_method = paste(habitat, method))

# Second plot: Disaggregated aquatic animal reliance
aa_reliance_multiple_trends <- ggplot() +
  geom_area(data = filter(combined, type == "observed"),
            aes(x = year, y = prop_animal_protein, fill = consumption_source),
            size = 0.75) +
  geom_line(data = filter(combined, type == "projected"),
            aes(x = year, y = prop_animal_protein, group = data_source),
            colour = "black", linetype = "dashed", linewidth = 0.5, alpha = 0.7) +
  facet_wrap(~ habitat_method) +
  scale_fill_manual(values = c("#2B6AA8", "#7AAFD4")) +
  labs(fill = "Consumption source", x = "", y = "Aquatic Animal Reliance (%)") +
  theme_minimal(base_size = 8) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(color = scales::alpha("black", 0.25), linewidth = 0.3),
    panel.grid.minor.y = element_blank(),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),
    strip.background = element_blank(),
    strip.text = element_text(size = 7, face = "bold"),
    axis.text = element_text(size = 6, color = "black"),
    axis.title.y = element_text(size = 7, margin = margin(r = 5), face = "bold"),
    axis.text.x = element_text(size = 5.5, angle = 30, hjust = 1),
    legend.position = "bottom",
    legend.title = element_text(size = 7, face = "bold"),
    legend.text = element_text(size = 6),
    legend.key.size = unit(0.3, "cm"),
    legend.box = "vertical",
    legend.spacing = unit(0.1, "cm"),
    legend.margin = margin(t = 0, b = 0),
    legend.box.margin = margin(t = -2, b = -2),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA)
  )



(aa_reliance_mean_global_trends_plot <- cowplot::plot_grid(
  aa_reliance_mean_global_trends,
  aa_reliance_multiple_trends,
  ncol = 2,
  rel_heights = c(0.8, 2.2),  # Give more space to the faceted plot
  labels = c("A", "B"),
  label_size = 10
))

ggsave(
  filename = "../images/aa_reliance_mean_global_trends.pdf",
  plot     = aa_reliance_mean_global_trends_plot,
  device   = cairo_pdf,
  width    = 7.2,   # two-column journal width
  height   = 3.2,
  units    = "in"
)
```



## Statistic: X Countries have dominant reliance on X source

```{r X Countries have dominant reliance on X source}
# X Countries have dominant reliance on X source
aquatic_reliance %>%
  filter(year == 2020, food_group == "aquatic") %>%
  group_by(iso3c) %>%
  mutate(largest_group = prop_animal_protein == max(prop_animal_protein)) %>%
  filter(largest_group == TRUE) %>%
  ungroup() %>%
  count(method)
```

> Foreign marine capture reliance is the most dominant across 86 countries, while domestic inland aquaculture is more concentrated at 12 countries. 

## Statistic: How many countries have received X % of their animal source protein from marine vs inland

```{r X Countries have dominant reliance on X source}
# X Countries have dominant reliance on X source
aquatic_reliance %>%
  filter(year == 2020, food_group == "aquatic") %>%
  group_by(iso3c, habitat) %>%
  summarize(prop_animal_protein = sum(prop_animal_protein)) %>%
  mutate(individual_prop = round(prop_animal_protein / sum(prop_animal_protein),2)) %>%
  group_by(habitat) %>%
  count(individual_prop >= 0.8)
```

> Marine sources remain the most widespread reliant habitat, with 102 countries obtaining least 80% of their aquatic animal proteins from marine habitats, while only 18 sources at least 80% from inland sources. 

## Figure: Look at contributing production sources to a country's rate of change in their aggregate aquatic animal reliance

This panel visualizes for each country separated by region, their average annual change in their aggregate aquatic animal reliance, with separate colors as stacked bars representing the disaggregate reliance. The regions include Africa, Asia, South America, North America, Europe, and Oceania.

```{r aggregate vs disaggregate change in reliance}
# Shared base theme (mirrors theme_realm_panel style)
theme_reliance <- function(base_size = 8,
                           axis_title_x = TRUE,
                           axis_title_y = FALSE,
                           legend       = TRUE) {
  base <- theme_minimal(base_size = base_size) +
    theme(
      text                    = element_text(family = "sans"),
      plot.background         = element_rect(fill = "transparent", color = NA),
      panel.background        = element_rect(fill = "transparent", color = NA),
      panel.grid.minor        = element_blank(),
      panel.grid.major.y      = element_line(
        color     = scales::alpha("black", 0.06),
        linewidth = 0.3
      ),
      panel.grid.major.x      = element_line(
        color     = scales::alpha("black", 0.25),
        linewidth = 0.3
      ),
      axis.text               = element_text(size = base_size - 3),   # 5 pt
      axis.title.x            = element_text(
        size  = base_size,
        face  = "bold",
        vjust = -0.5,
        margin = margin(t = 4)
      ),
      axis.title.y            = element_blank(),
      strip.text              = element_text(size = base_size - 1, face = "bold"),
      legend.position         = if (legend) "bottom" else "none",
      legend.title            = element_text(size = base_size - 1, face = "bold"),
      legend.text             = element_text(size = base_size - 2),
      legend.key.size         = unit(0.3, "cm"),
      legend.box              = "vertical",
      legend.spacing          = unit(0.1, "cm"),
      legend.margin           = margin(t = 1, b = 0),
      legend.box.margin       = margin(t = -2, b = -2),
      legend.background       = element_rect(fill = "transparent", color = NA),
      legend.box.background   = element_rect(fill = "transparent", color = NA),
      plot.margin             = margin(3, 3, 3, 3)
    )

  if (!axis_title_x) {
    base <- base + theme(axis.title.x = element_blank())
  }

  base
}


# Shared fill scale (consistent colour across all bar panels)
scale_fill_consource <- function(...) {
  scale_fill_manual(
    values = c(
      "inland aquaculture" = "#E8B000",
      "inland capture"     = "#D4601A",
      "marine aquaculture"    = "#9B6DC5",
      "marine capture"        = "#2B6AA8",
      "unknown"            = "#A8A8A8"
    ),
    ...,
    drop = FALSE
  )
}


# Shared geom layer list for bar panels
bar_layers <- list(
  geom_col(),
  geom_point(aes(x = agg_estimate, y = iso3c), shape = 21,
             fill = "grey20", color = "white", size = 1.2, stroke = 0.3)
)


# Fit linear models to estimate the average annual change in disaggregated
# aquatic animal reliance (2010-2020) by production source (habitat x method)
# for each country
disagg <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "aquatic") %>%
  group_by(year, iso3c, habitat, method) %>%
  summarize(prop_animal_protein = sum(prop_animal_protein), .groups = "drop") %>%
  group_by(iso3c, habitat, method) %>%
  nest() %>%
  mutate(
    model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
    tidy  = map(model, broom::tidy)
  ) %>%
  unnest(tidy) %>%
  filter(term == "year") %>%
  mutate(
    group_label = paste(habitat, method),
    group_label = if_else(group_label == "unknown unknown", "unknown", group_label)
  ) %>%
  drop_na(estimate) %>%
  group_by(group_label) %>%
  mutate(pct_above = mean(estimate > 0) * 100) %>%
  ungroup()

# Fit linear models to estimate the average annual change in aggregate
# aquatic animal reliance (2010-2020)
agg <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, year) %>%
  summarise(prop_animal_protein = sum(prop_animal_protein), .groups = "drop") %>%
  group_by(iso3c) %>%
  nest() %>%
  mutate(
    model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
    tidy  = map(model, broom::tidy)
  ) %>%
  unnest(tidy) %>%
  filter(term == "year") %>%
  mutate(group_label = "aggregate") %>%
  drop_na(estimate) %>%
  group_by(group_label) %>%
  mutate(pct_above = mean(estimate > 0) * 100) %>%
  ungroup()

# Clean aggregate estimate data
agg_estimate <- bind_rows(agg, disagg) %>%
  group_by(iso3c) %>%
  mutate(agg_est = estimate[group_label == "aggregate"][1]) %>%
  add_region("iso3c", "region") %>%
  ungroup() %>%
  filter(group_label == "aggregate") %>%
  select(iso3c, agg_estimate = estimate)


# E. Helper to build a bar panel
make_bar_panel <- function(regions,
                           axis_x_label = FALSE,
                           legend_show  = FALSE,
                           text_size    = 4.75) {
  disagg %>%
    add_region("iso3c", "region") %>%
    filter(region %in% regions) %>%
    left_join(agg_estimate, by = "iso3c") %>%
    mutate(iso3c = fct_reorder(iso3c, -agg_estimate)) %>%
    ggplot(aes(x = estimate, y = iso3c, fill = group_label)) +
    bar_layers +
    scale_fill_consource(name = "Production source") +
    facet_wrap(~ region, scales = "free", nrow = 1) +
    labs(
      x    = "Average annual change in aquatic animal reliance (2010–2020)",
      fill = "Production source"
    ) +
    theme_reliance(
      axis_title_x = axis_x_label,
      legend       = legend_show
    ) +
    theme(axis.text = element_text(size = text_size))
}


# Build panels
all_regions   <- unique(add_region(disagg, "iso3c", "region")$region)
top_regions   <- all_regions[!all_regions %in% c("Oceania", "South America")]
bot_regions   <- c("Oceania", "South America")

top_row <- make_bar_panel(top_regions, axis_x_label = FALSE, legend_show = FALSE,
                          text_size = 4.75)

bottom_row <- make_bar_panel(bot_regions, axis_x_label = TRUE, legend_show = TRUE,
                             text_size = 5.35)

bottom_row_narrowed <- cowplot::plot_grid(
  NULL, bottom_row, NULL,
  ncol       = 3,
  rel_widths = c(0.1, 0.3, 0.1)
)

bottom_panel <- cowplot::plot_grid(
  top_row,
  bottom_row_narrowed,
  nrow        = 2,
  rel_heights = c(2, 1)
)


# Top panel: # of all countries with increasing / decreasing reliance

top_panel <- agg %>%
  count(direction = estimate > 0) %>%
  ggplot(aes(x = 1, y = n, fill = direction)) +
  coord_flip() +
  geom_col(position = "stack", width = 0.65) +
  scale_fill_manual(
    values = c(`FALSE` = "#B2182B", `TRUE` = "#2166AC"),
    labels = c(`FALSE` = "Declining reliance", `TRUE` = "Increasing reliance"),
    name   = "Direction of change"
  ) +
  labs(y = "Number of countries") +
  theme_reliance(axis_title_x = TRUE, legend = TRUE) +
  theme(
    axis.text.y  = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.x = element_text(size = 6, face = "bold"),
    panel.grid.major.y = element_blank(),
    aspect.ratio = 0.05
  )


# Combine panels

final_plot <- cowplot::plot_grid(
  top_panel,
  bottom_panel,
  nrow          = 2,
  rel_heights   = c(0.35, 2),
  labels        = c("A", "B"),
  label_size    = 9.5
)


# Save plot

ggsave(
  filename = "../images/pct_reliance_contribution_four_sources.pdf",
  plot     = final_plot,
  device   = cairo_pdf,   # embeds fonts; swap to "jpg" + dpi = 500 if raster needed
  width    = 8,
  height   = 6,
  units    = "in"
)
```

## Statistic: How many countries have increased their X reliance?

```{r}
# Proportion of countries with increasing reliance
disagg %>%
  filter(estimate > 0) %>%
  group_by(habitat, method) %>%
  count()

# Proportion of countries with decreasing reliance
disagg %>%
  filter(estimate < 0) %>%
  group_by(habitat, method) %>%
  count()


disagg %>%
  add_region("iso3c", "region") %>%
  filter(estimate > 0) %>%
  group_by(habitat, method, region) %>%
  count() %>% 
  filter(habitat == "marine",
         method == "aquaculture") %>%
  arrange(-n)

disagg %>%
  add_region("iso3c", "region") %>%
  filter(estimate < 0) %>%
  group_by(habitat, method, region) %>%
  count() %>% 
  filter(habitat == "marine",
         method == "capture") %>%
  arrange(-n)
```


# RQ 1b: How does aquatic animal reliance change with the relative contribution of aquatic versus terrestrial animal consumption

## Figure: Scatterplot (terrestrial consumption ~ aquatic consumption; colored by: aquatic reliance) + bar chart overlay 

Scatterplot of average annual change in daily terrestrial animal protein consumption (kg/capita/day) 2010–2020 (y-axis) with respect to the average annual change in daily aquatic animal protein consumption (kg/capita/day) 2010–2020 colored by the annual change in daily aquatic animal protein reliance(%) 2010-2020, overlaid by a 1:1 line. The bar plots represent the number of countries in each quadrant, with the highest count of countries in any quadrant being 21. The two right hand y-axes represent these number of countries.

```{r}
# Region palette — desaturated analogues of the RdBu diverging scale
region_colors <- c(
  "Africa"        = "#6B2737",
  "North America" = "#C45C3A",
  "South America" = "#D4956A",
  "Asia"          = "#4A7FA5",
  "Europe"        = "#7BAEC4",
  "Oceania"       = "#A8C8D8"
)

# Diverging fill scale — aligned with intercept map palette
div_low  <- "#B2182B"
div_mid  <- "#FFFFFF"
div_high <- "#2166AC"

# Shared base theme
theme_scatter <- function(base_size = 8) {
  theme_minimal(base_size = base_size) +
    theme(
      text                  = element_text(family = "sans", color = "black"),
      plot.background       = element_rect(fill = "transparent", color = NA),
      panel.background      = element_rect(fill = "transparent", color = NA),
      panel.grid.minor      = element_blank(),
      panel.grid.major      = element_line(
        color = scales::alpha("black", 0.12), linewidth = 0.3
      ),
      axis.text             = element_text(size = base_size - 2, color = "black"),
      axis.title            = element_text(size = base_size - 2, face = "bold"),
      legend.position       = "bottom",
      legend.title          = element_text(size = base_size - 2, face = "bold"),
      legend.text           = element_text(size = base_size - 3),
      legend.key.size       = unit(0.3, "cm"),
      legend.box            = "horizontal",
      legend.spacing.x      = unit(0.4, "cm"),
      legend.margin         = margin(t = 0, b = 0),
      legend.box.margin     = margin(t = -2, b = -2),
      legend.background     = element_rect(fill = "transparent", color = NA),
      legend.box.background = element_rect(fill = "transparent", color = NA),
      plot.margin           = margin(3, 4, 3, 4)
    )
}

# Fit linear models to estimate the average annual change in aquatic protein
# consumption (2010-2020)
aquatic_protein_quant <- aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, year) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap, na.rm = TRUE),
            .groups = "drop") %>%
  distinct() %>%
  split(.$iso3c) %>%
  map(~ lm(protein_consumed_kg_percap ~ year, data = .x)) %>%
  map_df(broom::tidy, .id = "iso3c") %>%
  filter(term == "year") %>%
  select(iso3c, aquatic_estimate = estimate)

# Fit linear models to estimate the average annual change in terrestrial protein
# consumption (2010-2020)
terrestrial_protein_quant <- aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         data_source == "new FBS", food_group == "terrestrial") %>%
  group_by(iso3c, year) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap, na.rm = TRUE),
            .groups = "drop") %>%
  distinct() %>%
  split(.$iso3c) %>%
  map(~ lm(protein_consumed_kg_percap ~ year, data = .x)) %>%
  map_df(broom::tidy, .id = "iso3c") %>%
  filter(term == "year") %>%
  select(iso3c, terrestrial_estimate = estimate)

# Fit linear models to estimate the average annual change in aggregate
# global aquatic animal reliance (2010-2020)
aquatic_protein_reliance <- aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, year) %>%
  summarize(prop_animal_protein = sum(prop_animal_protein, na.rm = TRUE),
            .groups = "drop") %>%
  distinct() %>%
  split(.$iso3c) %>%
  map(~ lm(prop_animal_protein ~ year, data = .x)) %>%
  map_df(broom::tidy, .id = "iso3c") %>%
  filter(term == "year") %>%
  select(iso3c, aquatic_reliance_estimate = estimate)

# Join all three model outputs and add region
plot_df <- aquatic_protein_quant %>%
  full_join(terrestrial_protein_quant, by = "iso3c") %>%
  full_join(aquatic_protein_reliance,  by = "iso3c") %>%
  add_region(col = "iso3c", region.col.name = "region")

# Pre-compute quadrant membership for corner insets
quad_counts <- plot_df %>%
  mutate(quadrant = case_when(
    aquatic_estimate > 0 & terrestrial_estimate > 0 ~ "Q1",
    aquatic_estimate < 0 & terrestrial_estimate > 0 ~ "Q2",
    aquatic_estimate < 0 & terrestrial_estimate < 0 ~ "Q3",
    aquatic_estimate > 0 & terrestrial_estimate < 0 ~ "Q4"
  )) %>%
  group_by(quadrant, region) %>%
  summarise(n = n_distinct(iso3c), .groups = "drop")

max_count <- max(quad_counts$n)

# Dummy bar plot used only to extract a square-keyed region legend
# geom_col produces filled rect keys in the legend, which is cleaner
# than the point/circle keys that come from geom_point
p_region_legend_source <- plot_df %>%
  group_by(region) %>%
  summarise(n = n(), .groups = "drop") %>%
  ggplot(aes(x = region, y = n, fill = region)) +
  geom_col() +
  scale_fill_manual(
    values = region_colors,
    name   = "Region",
    guide  = guide_legend(
      title.position = "top",
      title.hjust    = 0.5,
      nrow           = 2
    )
  ) +
  theme_scatter() +
  theme(legend.key.size = unit(0.3, "cm"))

legend_regions <- cowplot::get_plot_component(
  p_region_legend_source, "guide-box", return_all = TRUE
)
legend_regions <- Filter(function(x) !inherits(x, "zeroGrob"), legend_regions)[[1]]

# Main scatterplot — fill encodes reliance change, no region border encoding
# on points (region is shown only in the inset bar charts and the square legend)
p_scatter_base <- plot_df %>%
  filter(iso3c != "MDV") %>%
  ggplot(aes(
    x     = aquatic_estimate,
    y     = terrestrial_estimate,
    fill  = aquatic_reliance_estimate,
    label = iso3c
  )) +
  geom_hline(yintercept = 0, linewidth = 0.4, color = "grey40") +
  geom_vline(xintercept = 0, linewidth = 0.4, color = "grey40") +
  geom_abline(intercept = 0, slope = 1,
              linetype = "dashed", linewidth = 0.35, color = "grey55") +
  geom_point(shape = 21, size = 1.8, color = "grey30", stroke = 0.25) +
  geom_text_repel(
    size               = 2.0,
    min.segment.length = 0.2,
    segment.size       = 0.25,
    segment.color      = "grey60",
    show.legend        = FALSE
  ) +
  lims(x = c(-0.25, 0.25), y = c(-4.5, 4.5)) +
  scale_fill_gradient2(
    low      = div_low,
    mid      = div_mid,
    high     = div_high,
    midpoint = 0,
    name     = "Annual change in aquatic animal protein reliance (%) 2010–2020",
    guide    = guide_colorbar(
      title.position = "top",
      title.hjust    = 0.5,
      barwidth       = unit(4.5, "cm"),
      barheight      = unit(0.3, "cm"),
      frame.colour   = "grey40",
      ticks.colour   = "grey40"
    )
  ) +
  labs(
    x = "Average annual change in daily aquatic animal protein consumption (kg/capita/day) 2010–2020",
    y = "Average annual change in daily\nterrestrial animal protein consumption (kg/capita/day) 2010–2020"
  ) +
  theme_scatter()

# Extract colorbar legend from scatter before stripping it
legend_reliance <- cowplot::get_plot_component(
  p_scatter_base, "guide-box", return_all = TRUE
)
legend_reliance <- Filter(function(x) !inherits(x, "zeroGrob"), legend_reliance)[[1]]

# Remove legend from scatter for the main panel
p_scatter <- p_scatter_base + theme(legend.position = "none")

# Helper to build a single quadrant inset bar chart
make_inset <- function(q) {
  show_y <- q %in% c("Q1", "Q4")

  quad_counts %>%
    filter(quadrant == q) %>%
    ggplot(aes(x = region, y = n, fill = region)) +
    geom_col(width = 0.7) +
    scale_fill_manual(values = region_colors) +
    scale_y_continuous(
      limits   = c(0, max_count),
      breaks   = c(0, max_count),
      position = "right"
    ) +
    labs(y = if (show_y) "No. countries" else NULL) +
    theme_minimal(base_size = 6) +
    theme(
      axis.title.x    = element_blank(),
      axis.title.y    = if (show_y) element_text(size = 4.5, color = "grey40") else element_blank(),
      axis.text.x     = element_blank(),
      axis.text.y     = if (show_y) element_text(size = 4.5, color = "grey40") else element_blank(),
      axis.ticks.y    = if (show_y) element_line(linewidth = 0.2, color = "grey60") else element_blank(),
      panel.grid      = element_blank(),
      plot.background = element_rect(fill = scales::alpha("white", 0.75), color = NA),
      legend.position = "none",
      plot.margin     = margin(0, 0, 0, 0)
    )
}

# Build all four corner insets
p_q1 <- make_inset("Q1")
p_q2 <- make_inset("Q2")
p_q3 <- make_inset("Q3")
p_q4 <- make_inset("Q4")

# Position insets in scatterplot corners using patchwork::inset_element
shrink_x <- 0.18
shrink_y <- 0.18

p_with_insets <- p_scatter +
  inset_element(p_q1, left = 1 - shrink_x, right = 1, bottom = 1 - shrink_y, top = 1) +
  inset_element(p_q2, left = 0,             right = shrink_x, bottom = 1 - shrink_y, top = 1) +
  inset_element(p_q3, left = 0,             right = shrink_x, bottom = 0,            top = shrink_y) +
  inset_element(p_q4, left = 1 - shrink_x, right = 1,        bottom = 0,            top = shrink_y)

# Place colorbar and region legend side by side below the main panel
legend_row <- cowplot::plot_grid(
  legend_reliance,
  legend_regions,
  ncol       = 2,
  rel_widths = c(1.2, 1)
)

# Stack main panel above legend row
final_plot <- cowplot::plot_grid(
  p_with_insets,
  legend_row,
  ncol        = 1,
  rel_heights = c(1, 0.16)
)

# Statistical test: terrestrial consumption rate of change > aquatic
wilcox.test(plot_df$terrestrial_estimate, plot_df$aquatic_estimate,
            paired = FALSE, alternative = "greater")

# Coefficient of variation for terrestrial consumption rate of change
cv_terr <- sd(plot_df$terrestrial_estimate, na.rm = TRUE) /
           mean(plot_df$terrestrial_estimate, na.rm = TRUE)

# Coefficient of variation for aquatic reliance rate of change
cv_aquatic <- sd(plot_df$aquatic_estimate, na.rm = TRUE) /
              mean(plot_df$aquatic_reliance_estimate, na.rm = TRUE)

# Ratio of CVs
cv_diff <- abs(cv_aquatic / cv_terr) %>% round(2)

ggsave(
  filename = "../images/scatter_bar_overlay.pdf",
  plot     = final_plot,
  device   = cairo_pdf,
  width    = 5,
  height   = 3.5,
  units    = "in"
)
```


> The variability in aquatic animal consumption is over 26 times greater in variability than terrestrial animal consumption by a factor of `cv_diff`

# RQ 2: What biophysical and socioeconomic factors are associated with reliance on farmed and wild aquatic animals for protein and overall consumption?

## Model: Hierarchical Bayesian regression

Only run this code chunk if all models aren't already saved

```{r hierarchical bayesian model}
# Variable list with coastline_km area instead of EEZ area
# predictors <- c(
#   "coastline_km",
#   "log_gdp",
#   "gdp_trade",
#   "inundation_area_log1000km2",
#   "supermarkets",
#   "pop_10k_pct_coast"
# )

# All response variables
predictors <- c(
  "coastline_km",
  "eez_total",
  "log_gdp",
  "gdp_trade",
  "inundation_area_log1000km2",
  "supermarkets",
  "pop_10k_pct_coast"
)

# Variable list with EEZ total area instead of coastline_km
# predictors <- c(
#   "eez_total",
#   "log_gdp",
#   "gdp_trade",
#   "inundation_area_log1000km2",
#   "supermarkets",
#   "pop_10k_pct_coast"
# )

# Get which random effect is used
model_type = "realm"

# Get which variable is selected (for unique file path)
var_used = "all"

# Create file path to save files to
base_path <- file.path(
  "..", "output", "bayesian_output",
  paste0("random_effect_", model_type),
  paste0("var_", var_used)
)

# Check whether all expected output files already exist for every response
# variable. The three outputs per variable are: X matrix, y vector, and
# the fitted Stan model. If all are present the loop is skipped entirely;
# if any single file is missing the full loop runs.
all_outputs_exist <- all(
  map_lgl(response_vars, function(rv) {
    all(file.exists(
      file.path(base_path, paste0("X_",      rv, ".rds")),
      file.path(base_path, paste0("y_",      rv, ".rds")),
      file.path(base_path, paste0("realms_", rv, ".rds")),
      file.path(base_path, paste0("fit_",    rv, ".rds"))
    ))
  })
)



# Initialize the progress bar
pb <- progress_bar$new(
  format = "  [:bar] :percent :elapsedfull",
  total = length(response_vars), # Number of models to run
  clear = FALSE, # Do not clear the bar after completion
  width = 60 # Width of the bar
)


if (all_outputs_exist) {
  message("All Stan outputs already exist for all response variables — skipping model fitting loop.")
} else {

  for (i in seq_along(response_vars)) {

    # Update progress bar
    pb$tick()

    joined_df <- get(response_vars[i]) %>%
      left_join(ac_df, by = "iso3c") %>%
      left_join(ecoregion, by = "iso3c") %>%
      ungroup() %>%
      left_join(
        indicator_df %>%
          select("eez_total", "iso3c",
                 "inundation_area_log1000km2" = "inland_water_max"),
        by = "iso3c"
      ) %>%
      left_join(coastline_df, by = "iso3c") %>%
      left_join(pop_10k_coast, by = "iso3c") %>%
      mutate(
        log_gdp = log(gdp),
        coastline_km = case_when(
          is.na(coastline_km) & str_detect(realm, "Landlocked") ~ 0,
          TRUE ~ coastline_km
        ),
        eez_total = case_when(
          is.na(eez_total) & str_detect(realm, "Landlocked") ~ 0,
          TRUE ~ eez_total
        )
      ) %>%
      select(iso3c, 2, all_of(predictors), realm) %>%
      drop_na() %>%
      filter(realm != "")
    
    # Save used countries in a vector
    saveRDS(joined_df$iso3c,"../data/model_countries")

    # Design matrix — z-scored predictors
    X <- joined_df %>%
      select(all_of(predictors)) %>%
      mutate(across(everything(), ~ scale(.x)[, 1])) %>%
      as.matrix()

    # Response — z-scored
    y <- scale(joined_df %>% pull(2))[, 1]

    # Hierarchical grouping (realm index)
    province <- as.integer(factor(joined_df$realm))
    J        <- length(unique(province))

    # Create output directory if it doesn't already exist
    if (!dir.exists(base_path)) {
      dir.create(base_path, recursive = TRUE, showWarnings = FALSE)
    }

    # Save design matrix, response, and realm vector
    saveRDS(X, file.path(base_path, paste0("X_", response_vars[i], ".rds")))
    saveRDS(y, file.path(base_path, paste0("y_", response_vars[i], ".rds")))
    saveRDS(
      joined_df$realm,
      file.path(base_path, paste0("realms_", response_vars[i], ".rds"))
    )

    # Stan data
    stan_data <- list(
      N        = nrow(X),
      K        = ncol(X),
      J        = J,
      X        = X,
      y        = y,
      province = province
    )

    # Initial values matching Stan model parameters
    inits <- function() {
      list(
        beta_0       = 0,
        sigma_beta_0 = 0.5,
        z_beta_0     = rnorm(J, 0, 0.1),
        beta         = rep(0, ncol(X)),
        sigma        = 1
      )
    }

    # Fit Stan model
    fit <- stan(
      file   = "../utils/social_bayesian_model_hierarchical.stan",
      data   = stan_data,
      init   = inits,
      chains = 3,
      iter   = 4000,
      warmup = 1000
    )

    # Save fitted model
    saveRDS(fit, file.path(base_path, paste0("fit_", response_vars[i], ".rds")))

    message("Completed: ", response_vars[i])
  }

}
```

## Figure: Posterior distributions for the partial slope across models

Here, we map the posterior distributions for models of terrestrial animal consumption per capita, aquatic animal consumption per capita, capture consumption per capita, aquaculture consumption per capita, domestic consumption per capita, foreign consumption per capita, marine consumption per capita, inland consumption per capita, and aggregate aquatic animal reliance. Posterior distributions are displayed with 90%, 80%, and 50% credibility intervals, represented by the lightest, darker, and darkest colors respectively. The point represents the posterior median.

```{r hierarcical partial slope posteriors}
model_type = "realm"

# Create a data frame of posteriors of partial slopes across Bayesian models
slopes_df <- data.frame()

for (i in seq_along(response_vars)) {

  # Extract MCMC samples
  x <- rstan::extract(
    readRDS(file.path(base_path, paste0("fit_", response_vars[i], ".rds"))),
    permuted = TRUE
  )
  beta_global <- x$beta
  X           <- readRDS(file.path(base_path, paste0("X_", response_vars[i], ".rds")))

  n_iter <- nrow(beta_global)
  K      <- ncol(beta_global)
  vars   <- colnames(X)

  # Extract global slopes and calculate credibility intervals
  ci_df <- expand_grid(iter = 1:n_iter, k = 1:K) %>%
    mutate(
      value = map2_dbl(iter, k, ~ beta_global[.x, .y]),
      var   = vars[k]
    ) %>%
    group_by(var) %>%
    summarise(
      lo95 = quantile(value, 0.025),
      lo80 = quantile(value, 0.10),
      lo50 = quantile(value, 0.25),
      med  = quantile(value, 0.50),
      hi50 = quantile(value, 0.75),
      hi80 = quantile(value, 0.90),
      hi95 = quantile(value, 0.975),
      .groups = "drop"
    ) %>%
    mutate(model = response_vars[i])

  slopes_df <- bind_rows(slopes_df, ci_df)
}

# Clean variable names for axis labels
var_labels <- c(
  coastline_km               = "Coastline length (km)",
  gdp_trade                  = "GDP from trade (%)",
  inundation_area_log1000km2 = "Inundated area (log 1000 km²)",
  log_gdp                    = "Log GDP per capita",
  pop_10k_pct_coast          = "Pop. within 10 km of coast (%)",
  supermarkets               = "Supermarkets per 100,000",
  eez_total                  = "Log EEZ area km²"
)

# Model colours — harmonized palette
model_colors <- c(
  "Terrestrial animal consumption per capita" = "#8C1A11",
  "Aquatic animal consumption per capita"     = "#2B6AA8",
  "Aquatic reliance"                          = "#1A7A6E",
  "Capture consumption per capita"            = "#D4601A",
  "Aquaculture consumption per capita"        = "#E8B000",
  "Domestic consumption per capita"           = "#7BAEC4",
  "Foreign consumption per capita"            = "#C45C3A",
  "Marine consumption per capita"             = "#9B6DC5",
  "Inland consumption per capita"             = "#D4956A"
)

# Interval colours — three lightness steps per model colour
col_95 <- colorspace::lighten(model_colors, 0.65)
col_80 <- colorspace::lighten(model_colors, 0.35)
col_50 <- colorspace::lighten(model_colors, 0.05)

# Load script to convert model name strings to readable labels
source("../utils/clean_model_name.R")

# Build plotting data
df_plot <- slopes_df %>%
  mutate(
    clean_name = clean_model_name(model),
    nonzero    = (lo95 < 0 & hi95 < 0) | (lo95 > 0 & hi95 > 0),
    suffix     = sub(".*_", "", model),
    suffix_num = suppressWarnings(as.numeric(suffix))
  ) %>%
  filter(
    str_detect(clean_name, "capita") | str_detect(clean_name, "Aquatic animal reliance"),
    !model == "animal_cons_cap"
  ) %>%
  arrange(suffix_num, suffix) %>%
  mutate(
    clean_name = factor(clean_name, levels = unique(clean_name)),
    col95 = coalesce(col_95[as.character(clean_name)], "grey80"),
    col80 = coalesce(col_80[as.character(clean_name)], "grey60"),
    col50 = coalesce(col_50[as.character(clean_name)], "grey40"),
    panel_id = as.numeric(factor(clean_name))
  )

# Shared symmetric x-axis limits so zero lines align across all panels
x_limit <- max(abs(c(df_plot$lo95, df_plot$hi95)), na.rm = TRUE) * 1.05
x_limit <- ceiling(x_limit * 10) / 10

panel_names <- levels(df_plot$clean_name)

# Function to draw a single credible interval panel
# left_col = TRUE shifts the tag inward to clear the y-axis label text
make_panel <- function(df, show_yaxis = FALSE, title = "", tag = "",
                       left_col = FALSE) {

  df_sig  <- df %>% filter(nonzero)
  df_nsig <- df %>% filter(!nonzero)

  # Left-column panels have wide y-axis text; nudge the tag right so it
  # doesn't sit directly on top of the longest label
  tag_pos <- if (left_col) c(0.52, 1) else c(0.01, 1)

  ggplot(df, aes(y = var)) +
    geom_segment(aes(x = lo95, xend = hi95, color = col95),
                 linewidth = 2.2, lineend = "butt", alpha = 1) +
    geom_segment(aes(x = lo80, xend = hi80, color = col80),
                 linewidth = 2.2, lineend = "butt", alpha = 1) +
    geom_segment(aes(x = lo50, xend = hi50, color = col50),
                 linewidth = 2.2, lineend = "butt", alpha = 1) +
    geom_point(aes(x = med), shape = 21, size = 1.6,
               fill = "white", color = "grey20", stroke = 0.4) +

    geom_vline(xintercept = 0, linetype = "dashed",
               linewidth = 0.35, color = "grey40") +
    scale_color_identity() +
    scale_x_continuous(limits = c(-x_limit, x_limit)) +
    scale_y_discrete(labels = var_labels) +
    labs(title = title, tag = tag) +
    theme_minimal(base_size = 8) +
    theme(
      plot.title         = element_text(size = 3.5, face = "bold", hjust = 0.5,
                                        margin = margin(b = 2)),
      plot.tag           = element_text(size = 7, face = "bold"),
      plot.tag.position  = tag_pos,
      panel.grid.major.y = element_blank(),
      panel.grid.minor   = element_blank(),
      panel.grid.major.x = element_line(
        color = scales::alpha("black", 0.12), linewidth = 0.3
      ),
      axis.text.x        = element_text(size = 5.5, color = "black"),
      axis.title         = element_blank(),
      plot.margin        = margin(4, 6, 4, 6),
      axis.text.y        = if (show_yaxis) element_text(size = 5.5, color = "black",
                                                         hjust = 1) else element_blank(),
      axis.ticks.y       = element_blank()
    )
}

# Build all nine panels — panels 1, 4, 7 are left column (show y-axis + shifted tag)
panels <- lapply(seq_along(panel_names), function(i) {
  name     <- panel_names[i]
  df_sub   <- df_plot %>% filter(clean_name == name)
  show_y   <- i %in% c(1, 4, 7)
  left_col <- i %in% c(1, 4, 7)
  make_panel(df_sub,
             show_yaxis = show_y,
             left_col   = left_col,
             title      = name,
             tag        = letters[i])
})

# Centered x-axis label placed below the grid
xlab_plot <- ggplot() +
  theme_void() +
  annotate("text", x = 0.5, y = 0.5,
           label    = "Posterior slope estimate",
           size     = 2.4,
           fontface = "bold",
           hjust    = 0.5,
           vjust    = 0.5)

# Assemble 3×3 grid
grid_plot <- (
    (panels[[1]] | panels[[2]] | panels[[3]]) /
    (panels[[4]] | panels[[5]] | panels[[6]]) /
    (panels[[7]] | panels[[8]] | panels[[9]])
  ) +
  plot_layout(heights = c(1, 1, 1)) &
  theme(plot.margin = margin(3, 4, 3, 4))

# Stack grid above centered x-axis label
final_plot <- cowplot::plot_grid(
  grid_plot,
  xlab_plot,
  ncol        = 1,
  rel_heights = c(1, 0.07)
)

ggsave(
  filename = "../images/posteriors_partial_slopes.pdf",
  plot     = final_plot,
  device   = cairo_pdf,
  width    = 5,
  height   = 3,
  units    = "in"
)
```

## Figure: Plot Maps of intercept median posteriors

The region specific intercept deviations from the slope line for the levels of daily animal protein consumption per capita (left) and aquatic animal reliance per capita (right). Outlined countries indicate that the 95% posterior credible interval does not include zero.
  
```{r realm_maps_robinson_final, message=FALSE, warning=FALSE}
# Get intercept posteriors
model_type = "realm"
intecrept_df <- data.frame()
for (i in seq_along(response_vars)) {
     
  # Extract MCMC samples
  x <- rstan::extract(readRDS(file.path(base_path, paste0("fit_", response_vars[i], ".rds"))), permuted = T)
  beta_global <- x$beta       # iter × K (slopes)
  X <- readRDS(file.path(base_path, paste0("X_", response_vars[i], ".rds")))  
  realm_level <- readRDS(file.path(base_path, paste0("realms_", response_vars[i], ".rds")))
  
  n_iter <- nrow(beta_global)
  K <- ncol(beta_global)
  vars <- colnames(X)
  
  # Posteriors of the realm intercepts
  beta_0i <- x$beta_0i   # iterations × J

  n_iter <- nrow(beta_0i)
  J <- ncol(beta_0i)

  # Get corresponding realm indicators (e.g., Europe, Oceania) to match with intercept values
realm_labels <- levels(factor(realm_level))  # length J,

post_intercepts <- expand_grid(
  iter = 1:n_iter,
  realm = 1:J
) %>%
  mutate(
    value = beta_0i[cbind(iter, realm)],
    realm = realm_labels[realm]
  )

  ci_intercepts <- post_intercepts %>%
  group_by(realm) %>%
  summarise(
    lo95 = quantile(value, 0.025),
    lo50 = quantile(value, 0.25),
    med  = quantile(value, 0.5),
    hi50 = quantile(value, 0.75),
    hi95 = quantile(value, 0.975),
    .groups = "drop"
  ) %>%
  mutate(model = response_vars[i])
  
  intecrept_df <- intecrept_df %>%
    bind_rows(ci_intercepts)
}


# Load MEOW ecoregions
meow <- st_read("../data/MEOW/meow_ecos.shp")

# Realm levels
realm_levels <- c(
  "Arctic","Central Indo-Pacific","Eastern Indo-Pacific",
  "Landlocked, Africa","Landlocked, Asia","Landlocked, Europe",
  "Landlocked, South America","Southern Ocean","Temperate Australasia",
  "Temperate Northern Atlantic","Temperate Northern Pacific",
  "Temperate South America","Temperate Southern Africa",
  "Tropical Atlantic","Tropical Eastern Pacific","Western Indo-Pacific"
)

marine_realms <- realm_levels[!str_detect(realm_levels, "Landlocked")]
landlocked_realms <- realm_levels[str_detect(realm_levels, "Landlocked")]

#  Dissolve MEOW ecoregions into REALM polygons
meow_realms <- meow %>%
  group_by(REALM) %>%
  summarise(geometry = st_union(geometry), .groups = "drop") %>%
  rename(realm = REALM)

# Split polygons at the antimeridian
meow_realms <- meow_realms %>%
  st_wrap_dateline(options = c("WRAPDATELINE=YES", "DATELINEOFFSET=180")) %>%
  st_collection_extract("POLYGON")


# Build landlocked realms using CONTINENTS
# but trim only the Russia sliver east of 180 degrees
countries <- ne_countries(scale = "medium", returnclass = "sf")

# Trim Russia's far-east sliver that bleeds onto the map as a horizontal line if not fixed
russia <- countries %>% filter(iso_a3 == "RUS")

russia_trimmed <- st_difference(
  russia,
  st_as_sfc(st_bbox(c(
    xmin = 180, ymin = -90, xmax = 200, ymax = 90
  ), crs = 4326))
)

countries_fixed <- countries %>%
  filter(iso_a3 != "RUS") %>%
  bind_rows(russia_trimmed)

# Define robinson projection
robin <- "+proj=robin +lon_0=0 +datum=WGS84 +units=m +no_defs"

# Clean continent basemap (for always-visible land)
continent_basemap <- countries_fixed %>%
  group_by(continent) %>%
  summarise(geometry = st_union(geometry), .groups = "drop") %>%
  st_make_valid() %>%
  st_wrap_dateline(options = c("WRAPDATELINE=YES", "DATELINEOFFSET=180")) %>%
  st_collection_extract("POLYGON") %>%
  st_transform(crs = robin)

# Build continent unions (now safe)
continents <- countries_fixed %>%
  mutate(continent = factor(continent)) %>%
  group_by(continent) %>%
  summarise(geometry = st_union(geometry), .groups = "drop")

# Map landlocked realms to continents
landlocked_lookup <- tibble(
  realm = landlocked_realms,
  continent = str_remove(landlocked_realms, "Landlocked, ")
)

landlocked_geom <- landlocked_lookup %>%
  left_join(continents, by = "continent") %>%
  select(realm, geometry)

# Combine marine + landlocked geometries
realm_geom <- bind_rows(
  meow_realms %>% filter(realm %in% marine_realms),
  landlocked_geom
)

# Project to Robinson
realm_geom_robin <- st_transform(realm_geom, crs = robin)

# Mapping function with significance outlines
selected_df <- intecrept_df %>%
  filter(model %in% c("terr_animal_cons_cap",
                      "aquatic_rel",
                      "aquaculture_rel",
                      "capture_rel"
                      ))

global_min <- min(selected_df$med, na.rm = TRUE)
global_max <- max(selected_df$med, na.rm = TRUE)

global_limit <- max(abs(global_min), abs(global_max))

# Panel theme
theme_realm_panel <- function(base_size = 8) {
  theme_void(base_size = base_size) +
    theme(
      text                  = element_text(family = "sans"),   # swap "sans" for your journal font
      plot.title            = element_text(size = 8, face = "bold", hjust = 0.5,
                                           margin = margin(b = 3)),
      plot.background       = element_rect(fill = "transparent", color = NA),
      panel.background      = element_rect(fill = "transparent", color = NA),
      legend.position       = "bottom",
      legend.title          = element_text(size = 7, face = "bold", hjust = 0.5),
      legend.text           = element_text(size = 6),
      legend.key.width      = unit(1.5, "cm"),
      legend.key.height     = unit(0.3, "cm"),
      legend.box            = "vertical",
      legend.spacing        = unit(0.1, "cm"),
      legend.margin         = margin(t = 2, b = 0),
      legend.background     = element_rect(fill = "transparent", color = NA),
      legend.box.background = element_rect(fill = "transparent", color = NA),
      plot.margin           = margin(2, 2, 2, 2)
    )
}
 
 
# Graticule layer (Robinson projection reference lines)
graticule <- st_graticule(
  lat = seq(-60, 60, 30),
  lon = seq(-180, 180, 60)
) %>%
  st_transform(crs = robin)
 
# Diverging color palette
div_low  <- "#B2182B"
div_mid  <- "#FFFFFF"
div_high <- "#2166AC"
 
# Updated map-making function
make_realm_map <- function(model_name, df = intecrept_df) {
 
  df_model <- df %>%
    filter(model == model_name) %>%
    mutate(significant = (lo95 > 0 & hi95 > 0) | (lo95 < 0 & hi95 < 0))
 
  geom_joined <- realm_geom_robin %>%
    left_join(df_model, by = "realm") %>%
    filter(!is.na(st_is_empty(geometry)) & !st_is_empty(geometry))
 
  ggplot() +
 
    # Subtle Robinson graticules for geographic anchor
    geom_sf(
      data  = graticule,
      color = "grey88",
      linewidth = 0.2,
      linetype  = "solid"
    ) +
 
    # Land basemap
    geom_sf(
      data  = continent_basemap,
      fill  = "grey82",
      color = "grey60",
      linewidth = 0.15
    ) +
 
    # Realm fill
    geom_sf(
      data = geom_joined,
      aes(fill = med),
      color = "grey40",
      linewidth = 0.2
    ) +
 
    # Significance: dual-stroke outline (white halo + dark inner)
    geom_sf(
      data      = geom_joined %>% filter(significant),
      fill      = NA,
      color     = "white",
      linewidth = 1.0
    ) +
    geom_sf(
      data      = geom_joined %>% filter(significant),
      fill      = NA,
      color     = "#1a1a1a",
      linewidth = 0.45
    ) +
 
    # Global symmetric diverging scale
    scale_fill_gradient2(
      low      = div_low,
      mid      = div_mid,
      high     = div_high,
      midpoint = 0,
      limits   = c(-global_limit, global_limit),
      oob      = scales::squish,
      name     = "Posterior median\nintercept",
      guide    = guide_colorbar(
        title.position = "top",
        title.hjust    = 0.5,
        barwidth       = unit(4, "cm"),
        barheight      = unit(0.3, "cm"),
        frame.colour   = "grey40",
        ticks.colour   = "grey40"
      )
    ) +
 
    theme_realm_panel() +
    theme(legend.position = "none")   # suppress per-panel; re-added in combined plot
}
 
 
# Build panels with clean human-readable titles
panel_titles <- c(
  terr_animal_cons_cap = "Daily terrestrial animal protein consumption (g/cap/day)",
  aquatic_rel     = "Aquatic animal protein reliance (%)",
  aquaculture_rel = "Aquaculture production reliance",
  capture_rel     = "Capture fisheries reliance"
)
 
p1 <- make_realm_map("terr_animal_cons_cap") +
  labs(title = panel_titles["terr_animal_cons_cap"])
 
p2 <- make_realm_map("aquatic_rel") +
  labs(title = panel_titles["aquatic_rel"])
 
 
# Combined figure with patchwork
# Restore shared legend on a single representative panel, then collect
add_legend <- function(p) {
  p + theme(legend.position = "bottom")
}
 
plot <- (add_legend(p1) | add_legend(p2)) +
  plot_layout(guides = "collect") +
  plot_annotation(
    tag_levels = "A",           # A, B,  panels
    tag_prefix = "",
    theme = theme(
      plot.tag = element_text(size = 9, face = "bold", hjust = 0),
      plot.background = element_rect(fill = "white", color = NA)
    )
  ) &
  theme(
    legend.position   = "bottom",
    legend.box.margin = margin(t = 4),
    aspect.ratio      = 0.45,
    plot.margin       = margin(3, 3, 3, 3)
  )
 
 
# Add significance-outline caption note
plot <- plot +
  plot_annotation(
    caption = "Bold outlines indicate realms where the 95% credible interval excludes zero.",
    theme = theme(
      plot.caption = element_text(size = 6, hjust = 0, color = "grey40",
                                  margin = margin(t = 4)),
      plot.background = element_rect(fill = "white", color = NA)
    )
  )
 
 
# Save plot
ggsave(
  filename = "../images/intercept_map.pdf",   # PDF for vector submission
  plot     = plot,
  device   = cairo_pdf,
  width    = 7.2,
  height   = 5.5,
  units    = "in"
)
```

```{r}
intecrept_df %>%
  filter((model == "terr_animal_cons_cap" & med < 0) |
         (model == "aquatic_rel" & med < 0)) %>%
  group_by(realm) %>%
  filter(all(med < 0)) %>%
  ungroup()

intecrept_df %>%
  filter(model %in% c("terr_animal_cons_cap", "aquatic_rel")) %>%
  group_by(realm) %>%
  filter(all(med[model == "terr_animal_cons_cap"] < 0) & 
         all(med[model == "aquatic_rel"] < 0)) %>%
  ungroup() %>% 
  pull(realm)
```

