---
title: "02_results"
author: "Connor Quiroz, Lana Huang, Christopher D. Golden, Sarah Glaser, and Jessica A. Gephart"
date: "2025-09-05"
output: html_document
---

# Chapter 1 Results

# Load in packages and set script inits

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(exploreARTIS)
library(ggrepel)
library(countrycode)
library(colorspace)
library(ggpubr)
library(arrow)
library(rstan)
library(sf)
library(lwgeom)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggridges)
library(bayesplot)
library(purrr)
library(patchwork)
library(dplyr)
library(ggplot2)
library(forcats)
library(stringr)
library(grid)
library(broom)
library(ggtext)
library(progress)
library(ggpattern)
library(extrafont)
library(svglite)
library(rsvg)
library(showtext)


# Set up global font family for plots
# Register a specific static weight from the variable font file
font_add(family = "Bahnschrift", regular = "/path/to/bahnschrift.ttf")
showtext_auto()
showtext_opts(dpi = 96)  # match whatever dpi your ggsave/rsvg pipeline uses

# font_import()
family = "Bahnschrift"

# Initialize directories

# Get which random effect is used (for Bayesian model)
model_type = "realm"

# Get which variable is selected (for unique file path)
var_used = "all"

# Create file path to save files to
base_path <- file.path(
  "..", "output", "bayesian_output",
  paste0("random_effect_", model_type),
  paste0("var_", var_used)
)
```
# Read in data

## Read in aquatic reliance data & FAO population data

```{r read in FAO derived data}
# Read in aquatic reliance data
aquatic_reliance <- read.csv("../output/fao_aquatic_reliance_source.csv")

# Read in FAO sourcing data
aquatic_fbs <- read.csv("../output/fao_fbs_aquatic_source.csv")
```

## Load in biophysical and socio-economic data as covariates that we use in the Bayesian model 

```{r load data}
# Read in adaptive capacity data
ac_df <- read_parquet("../data/e_s_ac_data_total.parquet") %>%
  filter(scenario == "ssp126") %>%
  select(
    -aa_reliance_pct,
    -scenario,
    -change_in_stock,
    -live_weight_t,
    -pct_change
    ) %>%
  rename(iso3c = "consumer_iso3c")
  
# Read in ecoregion data (all shapes associated with soverign)
ecoregion_full <- 
  read.csv("../data/Spatially joined ecoregion to EEZ/ecoregion.csv") %>%
  rename(
    "iso3c" = "ISO_SOV1",
    ecoregion = "ECOREGION",
    province = "PROVINCE",
    realm = "REALM",
    shape_area = "Shape_Area"
    )

# Read in ecoregions for largest sovereign shape
ecoregion_unique <- read.csv("../data/Spatially joined ecoregion to EEZ/ecoregion_clean.csv") %>%
  select(-FREQUENCY) %>%
  rename(
    iso3c = "ISO_SOV1",
    shape_area = "MAX_Shape_Area"
    )

# Read in dietary sourcing flexibility index
dsfi_df <- read_csv("../data/covariates/food-systems-dashboard-2026-01-20 (dietary sourcing flexibility index).csv") %>%
  filter(`Start Year` == 2019 | `End Year` == 2019) %>%
  select(c("iso3c" = ISO3, "dsfi" = Value))

# Read in national indicator data
indicator_df <- read_csv("../data/covariates/all_national_indicators.csv")

# Read in coastline data
coastline_df <- read_csv("../data/covariates/countries-by-coastline-2026.csv") %>%
  # Convert iso2c to iso3c and correct Namibia country name to NAM iso3c
  mutate(iso3c = case_when(
    !is.na(flagCode) ~ countrycode(
      flagCode,
      origin = "iso2c",
      destination = "iso3c"
      ),
    country == "Namibia" ~ "NAM"
    )) %>%
  select(c(iso3c, "coastline_km" = "CountryCoastlineKm"))

# Read in proportion of people within 10k of the coast
pop_10k_coast <- read.csv("../data/covariates/pop_10k_coast.csv")
```

## Clean ecoregion data

Impute countries with no denoted ecoregion as "landlocked"

```{r clean ecoregion data}
# Join largest shape for the respective sovereign country
# to their ecoregion names
ecoregion <- left_join(
  ecoregion_unique,
  ecoregion_full,
  by = c("iso3c", "shape_area")
  ) %>%
  filter(realm != "")

# Assign countries not associated with a marine ecoregion to "landlocked" code
landlocked_countries <- aquatic_reliance %>%
  distinct(iso3c) %>%
  drop_na() %>%
  add_region(col = "iso3c", region.col.name = "region") %>%
  filter(!iso3c %in% ecoregion$iso3c) %>%
  mutate(
    ecoregion = paste0("Landlocked, ", region),
    province = paste0("Landlocked, ", region),
    realm = paste0("Landlocked, ", region)) %>%
  select(-region)

# Clean ecoregion data
ecoregion <- bind_rows(
  ecoregion,
  landlocked_countries
  ) %>%
  arrange(iso3c) %>%
  select(-shape_area)

# Clean up environment
rm(list = c("ecoregion_full", "ecoregion_unique"))
```

# RQ 1a: How do aquatic food reliance trends disaggregated by source differ
# from aggregate aquatic food reliance?

## Figure 1: Compare aggregate to disaggregate reliance

Three panel plot:

Panel A: trends in global aggregate aquatic animal protein reliance.

Panel B: trends in global aquatic animal protein reliance by method of
consumption (i.e., capture, aquaculture)

Panel C: trends in global aquatic animal protein reliance disaggregated by
method (i.e. capture vs aquaculture, inland vs marine, domestic vs foreign)

Note. Because aquatic animal reliance values (1996-2023) are separated between
two data sources, with old-FBS methods covering 1996-2013 and new-FBS methods
covering the 2010-2023 of these datasets, we include projections to visualize
differences between the datasets. This involves:

1. Removing any overlapping years between datasets, prioritizing the new-FBS
2. Projecting historical FAO values forward to new FBS years & projecting
new-FBS values back to historical years. These will be represented as dashed
lines in Figure 1.


These projections include forward casting
old-FBS reliance values to 

```{r compare aggregate to disaggregated reliance}
# Define year ranges for projections
year_ranges <- tibble(
  data_source = c("old FBS", "new FBS"),
  start_year  = c(1996, 1996),
  end_year    = c(2023, 2023)
)

# Define year ranges for observed data
observed_ranges <- tibble(
  data_source = c("old FBS", "new FBS"),
  min_year = c(1996, 2010),
  max_year = c(2009, 2023)
)

##############
### PLOT A ###
##############
# Calculate aggregate global aquatic animal reliance
####################################################

# To calculate reliance across production sources, whether by disaggregated
# source or aggregated across sources, we will calculate the total protein
# consumption per year 

# Calculate total tonnage of combined terrestrial and aquatic animal protein
total_protein <- aquatic_reliance %>%
group_by(
  year,
  data_source,
  ) %>%
 # sum terrestrial and aquatic animal proteins
summarize(total_protein = sum(protein_consumed_t_peryr),
          .groups = "keep")

# For each year, calculate the share of global animal proteins
# that are sourced from aquatic animals. 
global_reliance_agg <- aquatic_reliance %>%
  filter(food_group == "aquatic") %>%
  group_by(
    year,
    data_source,
    food_group
    ) %>%
  # Calculate annual global aquatic animal protein consumption
  summarize(food_group_total_protein_t = sum(protein_consumed_t_peryr),
  .groups = "keep") %>%
  left_join(total_protein, by = c("year", "data_source")) %>%
  ungroup() %>%
  mutate(perc_animal_protein = 100 *
           (food_group_total_protein_t / total_protein)) %>%
  # Remove overlapping FBS years
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"))

# Create forward / backward projections
coefs_global <- global_reliance_agg %>%
  # this grouping allows us to calculate forward and back cast linear model
  # estimates to generate forward and back cast projections
  group_by(data_source) %>%
  nest() %>%
  # Generate out forward and back cast linear model estimates based on data
  mutate(model = map(data, ~ lm(perc_animal_protein ~ year, data = .x)),
         tidy  = map(model, broom::tidy)) %>%
  unnest(tidy) %>%
  # Convert data to clean tidy form so that we can easily pull out linear
  # model slope and intercept estimates for generating back and forward
  # cast values
  select(
    data_source,
    term,
    estimate) %>%
  pivot_wider(
    values_from = estimate,
    names_from = term
    ) %>%
  rename(year_coef = year)

# Generate out forward and back cast linear model estimates based on data
projections_global <- coefs_global %>%
  left_join(year_ranges, by = "data_source") %>%
  rowwise() %>%
  mutate(year = list(start_year:end_year)) %>%
  unnest(year) %>%
  ungroup() %>%
  mutate(perc_animal_protein = `(Intercept)` + year_coef * year,
         type = "projected") %>%
  left_join(observed_ranges, by = "data_source") %>%
  # Keep only years outside observed range
  filter(year < min_year | year > max_year)

# Create plot
aa_reliance_mean_global_trends <- ggplot() +
  geom_area(
    data = global_reliance_agg,
    aes(
      x = year,
      y = perc_animal_protein
      ),
    fill = "#8FACB4"
    ) +
  geom_line(
    data = projections_global,
    aes(
      x = year,
      y = perc_animal_protein,
      group = data_source),
    colour = "black",
    linetype = "dashed",
    linewidth = 0.3,
    alpha = 0.7
    ) +
  labs(x = "", y = "Global aquatic animal reliance (%)") +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(expand = c(0.01,0)) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(color = "black", linewidth = 0.15),
    panel.grid.minor.y = element_blank(),
    axis.text.x = element_text(face = "bold"),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),
    axis.text = element_text(size = 4.5, color = "black", family = family),
    axis.title.y = element_text(
      size = 7, margin = margin(r = 5),
      face = "bold", family = family
      ),
    aspect.ratio = 0.4,
    axis.ticks.x = element_line(linewidth = 0.1),
    axis.ticks.length.x=unit(0.5, "mm"),
    axis.ticks.y = element_blank()
  )
 
##############
### PLOT B ###
##############
# Calculate global aquatic animal reliance by production method
##############################################################

# For each year, calculate the share of global animal proteins
# that are sourced from capture and aquaculture 
aa_reliance_method_trends <- aquatic_reliance %>%
  filter(food_group == "aquatic") %>%
  group_by(
  year,
  data_source,
  method,
  food_group
  ) %>%
  # Calculate annual global aquatic animal protein consumption
  summarize(food_group_total_protein_t = sum(protein_consumed_t_peryr),
            .groups = "keep") %>%
  left_join(
    total_protein,
    by = c("year", "data_source")
    ) %>%
  ungroup() %>%
  mutate(perc_animal_protein = 100 *
           (food_group_total_protein_t / total_protein)) %>%
  filter(method != "unknown") %>%
  # Covert to title string for plot
  mutate(method = str_to_title(method)) %>%
  # Remove overlapping FBS years
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"))

# Create forward / backward projections
coefs <- aquatic_reliance %>%
# this grouping allows us to calculate forward and back cast linear model
# estimates to generate forward and back cast projections
group_by(
  year,
  data_source,
  method,
  food_group
  ) %>%
summarize(
  food_group_total_protein_t = sum(protein_consumed_t_peryr),
  .groups = "drop"
  ) %>%
filter(food_group == "aquatic") %>%
left_join(
  total_protein,
  by = c("year", "data_source")
  ) %>%
mutate(perc_animal_protein = 100 * 
         (food_group_total_protein_t / total_protein)) %>%
group_by(
  data_source,
  method
  ) %>%
nest() %>%
# Generate out forward and back cast linear model estimates based on data
mutate(model = map(data, ~ lm(perc_animal_protein ~ year, data = .x)),
       tidy  = map(model, broom::tidy)) %>%
unnest(tidy) %>%
# Convert data to clean tidy form so that we can easily pull out linear
# model slope and intercept estimates for generating back and forward
# cast values
select(
  data_source,
  method,
  term,
  estimate) %>%
pivot_wider(
  values_from = estimate,
  names_from = term
  ) %>%
rename(year_coef = year)
 
# Generate out forward and back cast linear model estimates based on data
projections <- coefs %>%
  left_join(year_ranges, by = "data_source") %>%
  rowwise() %>%
  mutate(year = list(start_year:end_year)) %>%
  unnest(year) %>%
  ungroup() %>%
  mutate(
    perc_animal_protein = `(Intercept)` + year_coef * year
  ) %>%
  filter(method != "unknown") %>%
  select(
    year,
    data_source,
    method,
    perc_animal_protein
    ) %>%
  mutate(type = "projected",
         method = str_to_title(method)) %>%
  left_join(
    observed_ranges,
    by = "data_source"
    ) %>%
  # Keep only years outside observed range
  filter(year < min_year | year > max_year)  

# Obtain observed measures across old and new FBS
obs <- aa_reliance_method_trends %>%
  filter(data_source == "old FBS" & year <= 2009 |
           data_source == "new FBS" & year >= 2010) %>%
  mutate(type = "observed")

# Obtain projected measures (i.e., back and forward cast values)

# back cast does not include 2010 nor does forward cast include 2009 because 
# there is observed data on each of those years - we want projections only for
# beyond the observed data
proj <- projections %>%
  filter(data_source == "old FBS" & year > 2009 |
           data_source == "new FBS" & year < 2010) %>%
  mutate(type = "projected")

# Combine both observed and projected data
combined_method <- bind_rows(obs, proj)

# Create plot
aa_reliance_aq_cap <- aa_reliance_method_trends %>%
  # Clean up row names for plt
  mutate(method = factor(method, levels = c("Capture", "Aquaculture"))) %>%
    ggplot() +
  geom_area(
    aes(
      x = year,
      y = perc_animal_protein,
      fill = method
      ),
    size = 0.75) +
  geom_line(
    data = filter(combined_method, type == "projected"),
            aes(
              x = year,
              y = perc_animal_protein,
              group = data_source
              ),
    colour = "black",
    linetype = "dashed",
    linewidth = 0.3,
    alpha = 0.7
    ) +
  scale_fill_manual(values = c("#337D81","#D3983E")) +
  labs(x = "", y = "Aquatic animal reliance (%)\n(by method)") +
  # theme_minimal(base_size = 8) +
    facet_grid(~ factor(method, levels=c("Capture", "Aquaculture"))) +
  guides(fill = "none") +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(expand = c(0.01,0)) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(color = "black", linewidth = 0.15),
    panel.grid.minor.y = element_blank(),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),
    strip.background = element_blank(),
    strip.text = element_text(
      size = 6, face = "bold", family = family, color = "black"
      ),
    axis.text = element_text(size = 5, color = "black", family = family),
    axis.title.y = element_text(
      size = 7, margin = margin(r = 5), face = "bold", family = family
      ),
    axis.text.x = element_text(
      size = 4.5, angle = 30, hjust = 1, face = "bold", family = family
      ),
    aspect.ratio = 0.8,
    axis.ticks.y = element_blank(),
    axis.ticks.x = element_line(linewidth = 0.1),
    axis.ticks.length.x=unit(0.5, "mm")
  )

##############
### PLOT C ###
##############
# Calculate global aquatic animal protein reliance by inland vs marine,
# aquaculture vs capture, and foreign vs domestic.
#######################################################################
  
# For each year, calculate the share of global animal proteins that are sourced
# from each disaggregate source (i.e. capture vs aquaculture, inland vs marine,
# domestic vs foreign) 
 aa_reliance_disaggregate <- aquatic_reliance %>%
  filter(food_group == "aquatic") %>%
  group_by(
  year,
  data_source,
  habitat,
  method,
  consumption_source,
  food_group
  ) %>%
  # Calculate annual global aquatic animal protein consumption
  summarize(food_group_total_protein_t = sum(protein_consumed_t_peryr),
            .groups = "keep") %>%
  left_join(
    total_protein,
    by = c("year", "data_source")
    ) %>%
  ungroup() %>%
  mutate(perc_animal_protein = 100 *
           (food_group_total_protein_t / total_protein)) %>%
  filter(method != "unknown") %>%
  # Covert to title string for plot
  # Remove overlapping FBS years
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS")) %>%
  # Clean row names for plot
  mutate(habitat = str_to_title(habitat)) %>%
  mutate(habitat_method = paste(habitat, method, sep = " "))
 
# Create forward / backward projections
coefs <- aquatic_reliance %>%
# this grouping allows us to calculate forward and back cast linear model
# estimates to generate forward and back cast projections
group_by(
  year,
  data_source,
  habitat,
  method,
  food_group
  ) %>%
summarize(
  food_group_total_protein_t = sum(protein_consumed_t_peryr),
  .groups = "drop"
  ) %>%
filter(food_group == "aquatic") %>%
left_join(
  total_protein,
  by = c("year", "data_source")
  ) %>%
mutate(perc_animal_protein = 100 * 
         (food_group_total_protein_t / total_protein)) %>%
group_by(
  data_source,
  habitat,
  method
  ) %>%
nest() %>%
# Generate out forward and back cast linear model estimates based on data
mutate(model = map(data, ~ lm(perc_animal_protein ~ year, data = .x)),
       tidy  = map(model, broom::tidy)) %>%
unnest(tidy) %>%
# Convert data to clean tidy form so that we can easily pull out linear
# model slope and intercept estimates for generating back and forward
# cast values
select(
  data_source,
  habitat,
  method,
  term,
  estimate) %>%
pivot_wider(
  values_from = estimate,
  names_from = term
  ) %>%
rename(year_coef = year)

# Generate out forward and back cast linear model estimates based on data
projections <- coefs %>%
  left_join(year_ranges, by = "data_source") %>%
  rowwise() %>%
  mutate(year = list(start_year:end_year)) %>%
  unnest(year) %>%
  ungroup() %>%
  mutate(
    perc_animal_protein = `(Intercept)` + year_coef * year
  ) %>%
  filter(method != "unknown") %>%
  select(
    year,
    data_source,
    habitat,
    method,
    perc_animal_protein
    ) %>%
  mutate(type = "projected") %>%
  mutate(habitat = str_to_title(habitat)) %>%
  left_join(
    observed_ranges,
    by = "data_source"
    ) %>%
  # Keep only years outside observed range
  filter(year < min_year | year > max_year)

# Obtain old FBS observed measures
obs <- aa_reliance_disaggregate %>%
  filter(data_source == "old FBS" & year <= 2009 |
           data_source == "new FBS" & year >= 2010) %>%
  mutate(type = "observed")

# Backcasted projections
proj <- projections %>%
  filter(data_source == "old FBS" & year > 2009 |
           data_source == "new FBS" & year < 2010) %>%
  mutate(type = "projected")


# Combine both observed and projected data
combined <- bind_rows(
  obs,
  proj
  ) %>%
  # Clean data for plotting
  mutate(habitat_method = paste(habitat, method)) %>%
  mutate(habitat_method = factor(
    habitat_method,
    levels = c("Marine capture",
               "Inland capture",
               "Marine aquaculture",
               "Inland aquaculture")
    )) %>%
  mutate(
    habitat_method_consource = paste0(
      habitat_method, " ", consumption_source
      ))

# Create dataset for foreign line
lines <- combined %>%
  mutate(foreign_val = case_when(
    consumption_source == "foreign" ~ perc_animal_protein
  ))

# To correctly visualize colors for each consumption source, we had to play
# with the mechanics of the plot. This involves creating a custom color palette
# (seen below) to then overlay a striped pattern indicating foreign consumption
# within each habitat x method source combination.
disagg_colors <- list("#114F59",
                      pattern(
                            grobTree(
                              polygonGrob(
                                x = c(0,1,1,0),
                                y = c(0,0,1,1)
                                ),
                              segmentsGrob(
                              x0 = seq(-1,1, by = 0.04),
                              x1 = seq(-1,1, by = 0.04) + 0.25,
                              y0 = 0, y1 = 1
                              )
                              ),
                              extend = "repeat",
                              gp = gpar(col = "black", fill = "#114F59",
                                        lwd = .6)
                            ),
                   "#5DA49A",
                      pattern(
                            grobTree(
                              polygonGrob(
                                x = c(0,1,1,0),
                                y = c(0,0,1,1)
                                ),
                              segmentsGrob(
                              x0 = seq(-1,1, by = 0.04),
                              x1 = seq(-1,1, by = 0.04) + 0.01,
                              y0 = 0, y1 = 1
                              )
                              ),
                              extend = "repeat",
                              gp = gpar(col = "black", fill = "#5DA49A",
                                        lwd = .6)
                            ),
                   "#C87534",
                      pattern(
                            grobTree(
                              polygonGrob(
                                x = c(0,1,1,0),
                                y = c(0,0,1,1)
                                ),
                              segmentsGrob(
                              x0 = seq(-1,1, by = 0.04),
                              x1 = seq(-1,1, by = 0.04) + 0.05,
                              y0 = 0, y1 = 1
                              )
                              ),
                              extend = "repeat",
                              gp = gpar(col = "black", fill = "#C87534",
                                        lwd = .6)
                            ),
                   "#D3AF55",
                      pattern(
                            grobTree(
                              polygonGrob(
                                x = c(0,1,1,0),
                                y = c(0,0,1,1)
                                ),
                              segmentsGrob(
                              x0 = seq(-1,1, by = 0.04),
                              x1 = seq(-1,1, by = 0.04) + 0.03,
                              y0 = 0, y1 = 1
                              )
                              ),
                              extend = "repeat",
                              gp = gpar(col = "black", fill = "#D3AF55",
                                        lwd = .6)
                            ))

# Assign colors their respective names
names(disagg_colors) <- 
  c("Marine capture domestic",
    "Marine capture foreign",
    "Inland capture domestic",
    "Inland capture foreign",
    "Marine aquaculture domestic",
    "Marine aquaculture foreign",
    "Inland aquaculture domestic",
    "Inland aquaculture foreign")

# Set the linewidth of inland capture to be lower so that the striped foreign
# lines are more visible on the plot
disagg_linewidth <- c(0.3,0.03,0.3,0.3)

# Assign line widths their respective names
names(disagg_linewidth) <- c("Marine capture",
                             "Inland capture",
                             "Marine aquaculture",
                             "Inland aquaculture")

# Create plot
aa_reliance_disaggregate <- ggplot() +
  geom_area(
    data = filter(combined, type == "observed"),
    aes(
      x = year,
      y = perc_animal_protein,
      fill = habitat_method_consource
      )) +
  geom_line(
    data = filter(combined, type == "projected"),
    aes(
      x = year,
      y = perc_animal_protein,
      group = data_source
      ),
    colour = "black",
    linetype = "dashed",
    linewidth = 0.3,
    alpha = 0.7
    ) +
  geom_line(
    data = lines %>%
      drop_na(foreign_val),
    aes(
      x = year,
      y = foreign_val,
      linewidth = habitat_method
      )) +
  facet_wrap(~ habitat_method) +
  scale_linewidth_manual(values = disagg_linewidth) +
  # Create custom legend icon and fill for geom_area (i.e., stripe pattern)
  scale_fill_manual(
    values = disagg_colors,
    breaks = c(
      "Inland aquaculture foreign",
      "Marine capture domestic"
      ),
    # Convert arbitrary two select groups on line above
    labels = c("domestic", "foreign"), 
    # to foreign / domestic
    guide = guide_legend(
      override.aes = list(
        fill = list("white",
                    pattern(
                      grobTree(
                        polygonGrob(
                          x = c(0,1,1,0),
                          y = c(0,0,1,1)
                          ),
                        segmentsGrob(
                          x0 = seq(-1,1, by = 0.25),
                          x1 = seq(-1,1, by = 0.25) + 1,
                          y0 = 0, y1 = 1
                          )),
                      extend = "repeat",
                      gp = gpar(
                        col = "black", fill = "white",lwd = 0.57
                        ))),
        color = "black",
        linewidth = 0.2
        ))) +
  labs(
    fill = "Consumption source",
    x = "",
    y = "Aquatic animal reliance (%)\n(by habitat/method/consumption source)"
    ) +
  guides(linewidth = "none") +
  # theme_minimal(base_size = 8) +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(expand = c(0.01,0)) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(color = "black", linewidth = 0.15),
    panel.grid.minor.y = element_blank(),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),
    strip.background = element_blank(),
    strip.text = element_text(
      size = 6, face = "bold", family = family, color = "black"
      ),
    axis.text = element_text(size = 5, color = "black", family = family),
    axis.title.y = element_text(
      size = 7, margin = margin(r = 5), face = "bold", family = family
      ),
    axis.text.x = element_text(
      size = 4.5, angle = 30, hjust = 1, face = "bold", family = family
      ),
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = 6, family = family),
    legend.key.size = unit(0.3, "cm"),
    legend.box = "vertical",
    legend.spacing = unit(0.1, "cm"),
    legend.margin = margin(t = 0, b = 0),
    legend.box.margin = margin(t = -2, b = -2),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    aspect.ratio = 0.9,
    axis.ticks.y = element_blank(),
    axis.ticks.x = element_line(linewidth = 0.1),
    axis.ticks.length.x=unit(0.5, "mm")
  )

# Put plots together
set_null_device(cairo_pdf)
aa_reliance_mean_global_trends_plot <- cowplot::plot_grid(
  aa_reliance_mean_global_trends,
  aa_reliance_aq_cap,
  aa_reliance_disaggregate,
  ncol = 3,
  rel_heights = c(0.8, 0.8,2.2),  # Give more space to the faceted plot
  rel_widths = c(0.9,1,1),
  labels = c("A", "B", "C"),
  label_size = 10
) %>%
  # Supress warning - this warning only has to do with fonts, no data
  # or visualization is harmed via this warning
  suppressWarnings()

# Print final plot
print(aa_reliance_mean_global_trends_plot)

# Save plot
ggsave(
  filename = "../images/aa_reliance_mean_global_trends.pdf",
  plot     = aa_reliance_mean_global_trends_plot,
  device   = cairo_pdf,
  width    = 7.2,   # two-column journal width
  height   = 3.2,
  units    = "in"
)

```

### What percent has global aquatic animal reliance stayed between during 1996-2023?

```{r how much has aggregate reliance shifted since 1996?}
# To get a conservative estimate values of the range for global aquatic animal
# reliance, we need to measure across three values: observed old-FBS values,
# observed new FBS values, and forwardcasted / backcasted values. We need to
# include forwardcasted / backcasted estimates because the dataset is not
# continuous, requiring bidirectional interpolation for years not covered in
# old-FBS and in new-FBS  

# First, we need to calculate global aquatic animal protein reliance and measure
# produce forward / back cast projections

# For each year, calculate the share of global animal proteins
# that are sourced from aquatic animals. 
global_reliance_agg <- aquatic_reliance %>%
  filter(food_group == "aquatic") %>%
  group_by(
    year,
    data_source,
    food_group
    ) %>%
  # Calculate annual global aquatic animal protein consumption
  summarize(food_group_total_protein_t = sum(protein_consumed_t_peryr),
  .groups = "keep") %>%
  left_join(total_protein, by = c("year", "data_source")) %>%
  ungroup() %>%
  mutate(perc_animal_protein = 100 *
           (food_group_total_protein_t / total_protein)) %>%
  # Remove overlapping FBS years
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"))

# Create forward / backward projections
coefs_global <- global_reliance_agg %>%
  # this grouping allows us to calculate forward and back cast linear model
  # estimates to generate forward and back cast projections
  group_by(data_source) %>%
  nest() %>%
  # Generate out forward and back cast linear model estimates based on data
  mutate(model = map(data, ~ lm(perc_animal_protein ~ year, data = .x)),
         tidy  = map(model, broom::tidy)) %>%
  unnest(tidy) %>%
  # Convert data to clean tidy form so that we can easily pull out linear
  # model slope and intercept estimates for generating back and forward
  # cast values
  select(
    data_source,
    term,
    estimate) %>%
  pivot_wider(
    values_from = estimate,
    names_from = term
    ) %>%
  rename(year_coef = year)

# Generate out forward and back cast linear model estimates based on data
projections_global <- coefs_global %>%
  left_join(year_ranges, by = "data_source") %>%
  rowwise() %>%
  mutate(year = list(start_year:end_year)) %>%
  unnest(year) %>%
  ungroup() %>%
  mutate(perc_animal_protein = `(Intercept)` + year_coef * year,
         type = "projected") %>%
  left_join(observed_ranges, by = "data_source") %>%
  # Keep only years outside observed range
  filter(year < min_year | year > max_year)

# Combine observations to projections
combined_global <- bind_rows(
  global_reliance_agg,
  projections_global
)

# Across all three measures, calculate the *minimum* global aquatic animal reliance has been from 1996-2023
max_global_reliance <- combined_global %>%
  pull(perc_animal_protein) %>%
  max() %>%
  ceiling()

# Across all three measures, calculate the *maximum* global aquatic animal reliance has been from 1996-2023
min_global_reliance <- combined_global %>%
  pull(perc_animal_protein) %>%
  min() %>%
  floor()
```

In aggregate, global aquatic animal protein reliance has remained stagnant,
ranging from `r min_global_reliance`% to `r max_global_reliance`% between 1996
and 2023 (Figure 1A).

---

### How much has aquaculture and capture reliance changed?

When combining old and new FBS data into one continuous time series, there is a
a jump in the time series from 2009 to 2010 due to methodological
differences in the datasets. In our analysis, old FBS runs from 1996 to 2009 and
new FBS runs from 2010 to 2023. To maximize reliability in our rate of change
estimates, we will project new FBS animal protein reliance data back over 1996
to 2009 (i.e., back cast) and project old FBS animal protein reliance data
forward over 2010 to 2023 (i.e., forward cast). We will then measure
the rate of change across four scenarios:

  (old FBS)   (new FBS)
1. observed  - observed
2. back cast - observed
3. observed  - forward cast
4. back cast - forward cast

For each country at the aggregate and disaggregate level, we will take the
midpoint of the max and min of these rates of changes to be their change in
aquatic animal protein reliance.

```{r change in capture fishery reliance} 
# Step 0. Calculate totals in animal protein consumption to calculate reliance

# Step 1. Calculate percent change from 1996-2023 for both domestic aquaculture and capture.

# Step 2. We need to get a conservative of the percent change because we are measuring % change across two datasets,
# old-FBS and new-FBS. So we will measure the ROC against three possible scenarios: (1). obs_oldFBS-obs_newFBS; (2) backcasted-obs_newFBS;
# (3) obs_oldFBS-forwardcasted and take the lowest magnitude of the three.

# We will reuse the combined_method section from code chunk block 7, which calculates the backcasted and forwardcasted
# projections between the FAO-FBS datasets

# Step 0.

## Calculate global annual terrestrial animal protein consumption (tonnage)
 terr_protein_sourcing <- aquatic_reliance %>%
  group_by(
    year,
    data_source,
    food_group
    ) %>%
  summarize(animal_protein = sum(protein_consumed_t_peryr)) %>%
  filter(food_group == "terrestrial")

## Calculate global annual aquatic animal protein consumption (tonnage)
 aquatic_protein_sourcing <- aquatic_reliance %>%
  group_by(
    year,
    data_source,
    food_group
    ) %>%
  summarize(aquatic_protein = sum(protein_consumed_t_peryr)) %>%
  filter(food_group == "aquatic")

## Calculate total consumption of terrestrial and aquatic animal proteins
total_protein <- left_join(
  terr_protein_sourcing,
  aquatic_protein_sourcing,
  by = c("year", "data_source")
  # this grouping allows will allow us to generate forward and back casts
  # of old and new FBS data.
  ) %>%
   group_by(
     year,
     data_source
     ) %>%
  # sum terrestrial and aquatic animal proteins
  summarize(total_protein = sum(animal_protein, aquatic_protein)) 

# Step 1.

# Define year ranges for forward and back cast projections
year_ranges <- tibble(
  data_source = c("old FBS", "new FBS"),
  start_year  = c(1996, 1996),
  end_year    = c(2023, 2023)
)

# Define year ranges for observed data
observed_ranges <- tibble(
  data_source = c("old FBS", "new FBS"),
  min_year = c(1996, 2010),
  max_year = c(2009, 2023)
)

# Calculate global aquatic animal protein reliance disaggregated by method
global_reliance_disagg_met <- aquatic_reliance %>%
  group_by(
    year,
    data_source,
    method,
    food_group
    ) %>%
  summarize(food_group_total_protein_t = sum(protein_consumed_t_peryr),
            .groups = "keep") %>%
  filter(food_group == "aquatic") %>%
  left_join(total_protein, by = c("year", "data_source")) %>%
  ungroup() %>%
  mutate(perc_animal_protein = 100 * (food_group_total_protein_t / total_protein)) %>%
  # filter out overlapping FBS years
   filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"))

 # Create forward / backward projections
 coefs <- global_reliance_disagg_met %>%
  # this grouping allows us to calculate forward and back cast linear model
  # estimates to generate forward and back cast projections
  group_by(
    data_source,
    method
    ) %>%
  # nest data and run linear model across all countries
  nest() %>%
  mutate(
    model = map(data, ~ lm(perc_animal_protein ~ year, data = .x)),
    tidy  = map(model, broom::tidy)
    ) %>%
  unnest(tidy) %>%
  # Convert data to clean tidy form so that we can easily pull out linear
  # model slope and intercept estimates for generating back and forward
  # cast values
  select(
    data_source,
    method,
    term,
    estimate
    ) %>%
  pivot_wider(values_from = estimate, names_from = term) %>%
  rename(year_coef = year)
 
# Convert into usable data to be joined onto plot 
projections <- coefs %>%
  left_join(year_ranges, by = "data_source") %>%
  rowwise() %>%
  mutate(year = list(start_year:end_year)) %>%
  unnest(year) %>%
  ungroup() %>%
  # Generate out forward and back cast linear model estimates based on data 
  mutate(
    perc_animal_protein = `(Intercept)` + year_coef * year
  ) %>%
  filter(method != "unknown") %>%
  select(year, data_source, method, perc_animal_protein) %>%
  mutate(type = "projected") %>%
  left_join(observed_ranges, by = "data_source") %>%
  # Keep only years outside observed range
  filter(year < min_year | year > max_year)

# Obtain observed measures
obs <- global_reliance_disagg_met %>%
  filter(data_source == "old FBS" &  year <= 2009 |
           data_source == "new FBS" & year >= 2010) %>%
  mutate(type = "observed")

# Get projected measures
proj <- projections %>%
  filter(data_source == "old FBS" & year > 2009 |
           data_source == "new FBS" & year < 2010) %>%
  mutate(type = "projected")



# Combine observed and projected data - this will be used to compare
# the four scenarios of projections
combined <- bind_rows(
  obs,
  proj
  ) %>%
  # remove overlapping FBS years
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"))


################################# Aquaculture
## Calculate global annual aquaculture protein reliance
aq <- global_reliance_disagg_met %>%
  filter(method == "aquaculture")

## Get observed increases in aquaculture reliance (i.e., obs - obs)
obs_increases_aq <- aq %>%
  summarize(
    ratio = perc_animal_protein[year == 2023] /
            perc_animal_protein[year == 1996]
  , .groups = "keep") %>% pull(ratio)

## Extract projections to obtain forward / back casting
aq_projections <- combined %>%
  # remove overlapping FBS years
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS")) %>%
  filter(method == "aquaculture") %>%
  select(
    year,
    data_source,
    method,
    perc_animal_protein,
    type
    )

## Get back cast measures
backcasted_aq <- bind_rows(
  
  aq_projections %>%
    filter(type == "projected" & year < 2010),
  
  aq %>%
  filter(year > 2009) %>%
  mutate(type = "observed") %>%
  select(year,
         data_source,
         method,
         perc_animal_protein,
         type
         )
  
  )

## Get forward cast measures
forwardcasted_aq <- bind_rows(
  
  aq_projections %>%
    filter(type == "projected" & year >= 2010),
  
  aq %>%
  filter(year <= 2009) %>%
  mutate(type = "observed") %>%
  select(
    year,
    data_source,
    method,
    perc_animal_protein,
    type
    )
  
  )

################################# Capture
## Calculate global annual capture protein reliance
cap <- global_reliance_disagg_met %>%
  filter(method == "capture")

## Get observed decreases in marine capture 
obs_decreases_cap <- cap %>%
  summarize(
    ratio = perc_animal_protein[year == 2023] /
            perc_animal_protein[year == 1996]
  , .groups = "keep") %>% pull(ratio)

## Extract projections to obtain forward / back casting
cap_projections <- combined %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS")) %>%
  filter(method == "capture") %>%
  select(
    year,
    data_source,
    method,
    perc_animal_protein,
    type
    )

## Get backcasted dataset
backcasted_cap <- bind_rows(
  
  cap_projections %>%
    filter(type == "projected" & year < 2010),
  
  cap %>%
  filter(year > 2009) %>%
  mutate(type = "observed") %>%
  select(
    year,
    data_source,
    method,
    perc_animal_protein,
    type)
  
  )

## Get forwardcasted dataset
forwardcasted_cap <- bind_rows(
  
  cap_projections %>%
    filter(type == "projected" & year >= 2010),
  
  cap %>%
  filter(year <= 2009) %>%
  mutate(type = "observed") %>%
  select(
    year,
    data_source,
    method,
    perc_animal_protein,
    type
    )
  
  )

# Step 2.

################################# Aquaculture
## What is the rate of increase?

## Get vector of all potential rates of increase in percentage
roi_aq_vec <- 100 * c(

# obs - obs
obs_increases_aq,

# back cast - obs
aq %>%
  filter(year == 2023) %>%
  pull(perc_animal_protein) /
  backcasted_aq %>%
  filter(year == 1996) %>%
  pull(perc_animal_protein),

# obs - forward cast
forwardcasted_aq %>%
  filter(year == 2023) %>%
  pull(perc_animal_protein) / aq %>%
  filter(year == 1996) %>%
  pull(perc_animal_protein),

# back cast - forward cast
forwardcasted_aq %>%
  filter(year == 2023) %>%
  pull(perc_animal_protein) / 
backcasted_aq %>%
  filter(year == 1996) %>%
  pull(perc_animal_protein)

)

roi_aq_lb <- min(roi_aq_vec) %>% round(0) # lower bound for rate of increase
roi_aq_ub <- max(roi_aq_vec) %>% round(0) # upper bound for rate of increase

# Rate of increase for  aquaculture is the midpoint of the lower
# and upper bound for the rate of increase
roi_aq <- (roi_aq_lb + roi_aq_ub) / 2

################################# Capture
## What is the rate of decrease? (Get the most conservative estimate)
## We subtract from 1 to get percent decrease.

## Get vector of all potential rates of decrease in percentage
rod_cap_vec <- 100 * (1 - c(

  # obs - obs
obs_decreases_cap,

# back cast - obs
cap %>%
  filter(year == 2023) %>%
  pull(perc_animal_protein) /
  backcasted_cap %>%
  filter(year == 1996) %>%
  pull(perc_animal_protein),

# obs - forward cast
forwardcasted_cap %>%
  filter(year == 2023) %>%
  pull(perc_animal_protein) / cap %>%
  filter(year == 1996) %>%
  pull(perc_animal_protein),

# back cast - forward cast
forwardcasted_cap %>%
  filter(year == 2023) %>%
  pull(perc_animal_protein) / 
backcasted_cap %>%
  filter(year == 1996) %>%
  pull(perc_animal_protein)
  
  ))

rod_cap_lb <- min(rod_cap_vec) %>% round(0) # lower bound for rate of increase
rod_cap_ub <- max(rod_cap_vec) %>% round(0) # upper bound for rate of increase

# Rate of increase for  capuaculture is the midpoint of the lower
# and upper bound for the rate of increase
rod_cap <- (rod_cap_lb + rod_cap_ub) / 2
```

However, disaggregating aquatic animal protein reliance unmasks opposing global trends: capture fishery reliance declined `r rod_cap`% (`r rod_cap_lb`-`r rod_cap_ub`%), while aquaculture reliance grew `r roi_aq` (`r roi_aq_lb`-`r roi_aq_ub`%).

---

### What are increases in aquaculture and decreases in capture fisheries mainly attributed to?

When combining old and new FBS data into one continuous time series, there is a
a jump in the time series from 2009 to 2010 due to methodological
differences in the datasets. In our analysis, old FBS runs from 1996 to 2009 and
new FBS runs from 2010 to 2023. To maximize reliability in our rate of change
estimates, we will project new FBS animal protein reliance data back over 1996
to 2009 (i.e., back cast) and project old FBS animal protein reliance data
forward over 2010 to 2023 (i.e., forward cast). We will then measure
the rate of change across four scenarios:

  (old FBS)   (new FBS)
1. observed  - observed
2. back cast - observed
3. observed  - forward cast
4. back cast - forward cast

For each country at the aggregate and disaggregate level, we will take the
midpoint of the max and min of these rates of changes to be their change in
aquatic animal protein reliance.

```{r times aq reliance has increased since 1996 and marine capture has decreased}
# Step 0. Calculate totals in animal protein consumption to calculate reliance

# Step 1. Get forward / backcasted projections and fit into dataframe.

# Step 2. Calculate observed percent change from 1996-2023 for both
# inland aquaculture and marine capture.

# Step 3. We need to get a conservative of the percent change because we are
# measuring % change across two datasets, # old-FBS and new-FBS. So we will
# measure the ROC against three possible scenarios: (1). obs_oldFBS-obs_newFBS;
# (2) backcasted-obs_newFBS; (3) obs_oldFBS-forwardcasted; and
# (4) backcasted-forwardcasted and take the midpoint of the max and min values
# of rate of change.

# Step 0.

## Calculate global annual terrestrial animal protein consumption (tonnage)
terr_protein_sourcing <- aquatic_reliance %>%
  group_by(year, data_source, food_group) %>%
  summarize(animal_protein = sum(protein_consumed_t_peryr),
            .groups = "keep") %>%
  filter(food_group == "terrestrial")

## Calculate global annual aquatic animal protein consumption (tonnage)
aquatic_protein_sourcing <- aquatic_reliance %>%
  group_by(year, data_source, food_group) %>%
  summarize(aquatic_protein = sum(protein_consumed_t_peryr),
            .groups = "keep") %>%
  filter(food_group == "aquatic")

## Calculate total consumption of terrestrial and aquatic animal proteins
total_protein <- left_join(
  terr_protein_sourcing,
  aquatic_protein_sourcing,
  by = c("year", "data_source")
  # this grouping allows will allow us to generate forward and back casts
  # of old and new FBS data.
  ) %>%
   group_by(
     year,
     data_source
     ) %>%
  # sum terrestrial and aquatic animal proteins
   summarize(total_protein = sum(animal_protein, aquatic_protein),
            .groups = "keep") 

# Step 1.

# Define year ranges for forward and back cast projections
year_ranges <- tibble(
  data_source = c("old FBS", "new FBS"),
  start_year  = c(1996, 1996),
  end_year    = c(2023, 2023)
)

# Define year ranges for observed data
observed_ranges <- tibble(
  data_source = c("old FBS", "new FBS"),
  min_year = c(1996, 2010),
  max_year = c(2009, 2023)
)

# Calculate global aquatic animal protein reliance disaggregated by
# habitat and method
 global_reliance_disagg_hab_met <- aquatic_reliance %>%
  group_by(
    year,
    data_source,
    habitat,
    method,
    food_group
    ) %>%
  summarize(food_group_total_protein_t = sum(protein_consumed_t_peryr),
            .groups = "keep") %>%
  filter(food_group == "aquatic") %>%
  left_join(total_protein, by = c("year", "data_source")) %>%
  ungroup() %>%
  mutate(perc_animal_protein = 100 *
           (food_group_total_protein_t / total_protein)) %>%
  filter(habitat != "unknown") %>%
   # remove overlapping FBS years
   filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"))


# Create forward / backward projections
 coefs <- global_reliance_disagg_hab_met %>%
  # this grouping allows us to calculate forward and back cast linear model
  # estimates to generate forward and back cast projections
  group_by(
    data_source,
    habitat,
    method
    ) %>%
  # nest data and run linear model across all countries
  nest() %>%
  mutate(model = map(data, ~ lm(perc_animal_protein ~ year, data = .x)),
         tidy  = map(model, broom::tidy)) %>%
  unnest(tidy) %>%
  # Convert data to clean tidy form so that we can easily pull out linear
  # model slope and intercept estimates for generating back and forward
  # cast values
  select(
    data_source,
    habitat,
    method,
    term,
    estimate
    ) %>%
  pivot_wider(
    values_from = estimate,
    names_from = term
    ) %>%
  rename(year_coef = year)

# Generate out forward and back cast linear model estimates based on data
projections <- coefs %>%
  left_join(year_ranges, by = "data_source") %>%
  rowwise() %>%
  mutate(year = list(start_year:end_year)) %>%
  unnest(year) %>%
  ungroup() %>%
  mutate(
    perc_animal_protein = `(Intercept)` + year_coef * year
  ) %>%
  filter(method != "unknown", habitat != "unknown") %>%
  select(year, data_source, habitat, method, perc_animal_protein) %>%
  mutate(type = "projected") %>%
  left_join(observed_ranges, by = "data_source") %>%
  # Keep only years outside observed range
  filter(year < min_year | year > max_year)

# Obtain observed measures
obs <- global_reliance_disagg_hab_met %>%
  filter(data_source == "old FBS" &  year <= 2009 |
           data_source == "new FBS" & year >= 2010) %>%
  mutate(type = "observed")

# Get projected measures
proj <- projections %>%
  filter(data_source == "old FBS" & year > 2009 |
           data_source == "new FBS" & year < 2010) %>%
  mutate(type = "projected")

# Combine observed and projected data - this will be used to compare
# the four scenarios of projections
combined <- bind_rows(
  obs,
  proj
  ) %>%
  # remove overlapping FBS years
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"))

# Step 2.

################################# Inland aquaculture
## Calculate global annual inland aquaculture protein reliance
inaq <- global_reliance_disagg_hab_met %>%
  filter(habitat == "inland", method == "aquaculture")

## Get observed increases in inland aquaculture (i.e., obs - obs)
obs_increases_inaq <- inaq %>%
  summarize(
    ratio = perc_animal_protein[year == 2023] /
            perc_animal_protein[year == 1996],
    .groups = "keep") %>%
  pull(ratio)


## Filter observations and projections to inland aquaculture
inaq_projections <- combined %>%
  filter(
    habitat == "inland",
    method == "aquaculture"
    ) %>%
  select(
    year,
    data_source,
    habitat,
    method,
    perc_animal_protein,
    type
    )

## Get back cast measures
backcasted_inaq <- bind_rows(
  inaq_projections %>%
    filter(type == "projected" & year < 2010),
  inaq %>%
  filter(year > 2009) %>%
  mutate(type = "observed") %>%
  select(
    year,
    data_source,
    habitat, method,
    perc_animal_protein,
    type)
  )

## Get forward cast measures
forwardcasted_inaq <- bind_rows(
  inaq_projections %>%
    filter(type == "projected" & year >= 2010),
  inaq %>%
  filter(year <= 2009) %>%
  mutate(type = "observed") %>%
  select(
    year,
    data_source,
    habitat, method,
    perc_animal_protein,
    type
    )
  )

################################# Marine capture
## Calculate global annual marine capture protein reliance
maca <- global_reliance_disagg_hab_met %>%
  filter(habitat == "marine", method == "capture")

## Get observed decreases in marine capture (i.e., obs - obs)
obs_decreases_maca <- maca %>%
  summarize(
    ratio = perc_animal_protein[year == 2023] /
            perc_animal_protein[year == 1996],
    .groups = "keep") %>%
  pull(ratio)

## Filter observations and projections to marine capture
maca_projections <- 
  combined %>%
    filter(
      habitat == "marine",
      method == "capture"
      ) %>%
  select(
    year,
    data_source,
    habitat,
    method,
    perc_animal_protein,
    type
    ) 

## Get back cast measures
backcasted_maca <- bind_rows(
  maca_projections %>%
    filter(type == "projected" & year < 2010),
  maca %>%
  filter(year > 2009) %>%
  mutate(type = "observed") %>%
  select(
    year,
    data_source,
    habitat,
    method,
    perc_animal_protein,
    type
    )
  )

## Get forward cast measures
forwardcasted_maca <- bind_rows(
  maca_projections %>%
    filter(type == "projected" & year >= 2010),
  maca %>%
  filter(year <= 2009) %>%
  mutate(type = "observed") %>%
  select(
    year,
    data_source,
    habitat,
    method,
    perc_animal_protein,
    type
    )
)

# Step 3.

################################# Domestic inland aquaculture
## What is the rate of increase? (Get the most conservative estimate)
## Get vector of all potential rates of increase in percentage
roi_inaq_vec <- 100 * c(

# obs - obs
obs_increases_inaq,

# back cast - obs
inaq %>%
  filter(year == 2023) %>%
  pull(perc_animal_protein) /
  backcasted_inaq %>%
  filter(year == 1996) %>%
  pull(perc_animal_protein),

# obs - forward cast
forwardcasted_inaq %>%
  filter(year == 2023) %>%
  pull(perc_animal_protein) / inaq %>%
  filter(year == 1996) %>%
  pull(perc_animal_protein),

# back cast - forward cast
forwardcasted_inaq %>%
  filter(year == 2023) %>%
  pull(perc_animal_protein) / 
backcasted_inaq %>%
  filter(year == 1996) %>%
  pull(perc_animal_protein)

)

roi_inaq_lb <- min(roi_inaq_vec) %>% round(0) # lower bound for rate of increase
roi_inaq_ub <- max(roi_inaq_vec) %>% round(0) # upper bound for rate of increase

# Rate of increase for inland aquaculture is the midpoint of the lower
# and upper bound for the rate of increase
roi_inaq <- (roi_inaq_lb + roi_inaq_ub) / 2


## Figure out the start point and end point for the rate of increase - since
## obs-forwardcasted and obs-obs
## were produced equally identical rate of changes, giving the most conservative
## estimates on the percent change,
## we use the observed start and end points.

################################# Domestic marine capture
## What is the rate of decrease? (Get the most conservative estimate)
## Since it's rate of decrease, we subtract from 1 to get percent decrease.
# Get vector of all potential rates of increase
rod_maca_vec <- (1 - c(
  
# obs - obs
obs_decreases_maca,

# back cast - obs
maca %>%
  filter(year == 2023) %>%
  pull(perc_animal_protein) /
  backcasted_maca %>%
  filter(year == 1996) %>%
  pull(perc_animal_protein),

# obs - forward cast
forwardcasted_maca %>%
  filter(year == 2023) %>%
  pull(perc_animal_protein) /
  maca %>%
  filter(year == 1996) %>%
  pull(perc_animal_protein),

# back cast - forward cast
forwardcasted_maca %>%
  filter(year == 2023) %>%
  pull(perc_animal_protein) /
  backcasted_maca %>%
  filter(year == 1996) %>%
  pull(perc_animal_protein)
  



)) * 100

rod_maca_lb <- min(rod_maca_vec) %>% round(0) # lower bound for rate of decrease
rod_maca_ub <- max(rod_maca_vec) %>% round(0) # upper bound for rate of decrease

# Rate of increase for inland aquaculture is the midpoint of the lower
# and upper bound for the rate of increase
rod_maca <- (rod_maca_lb + rod_maca_ub) / 2
```

These diverging trends are driven mainly by declines in animal protein reliance
on marine capture of `r rod_maca`% (`r rod_maca_lb`-`r rod_maca_ub`%) and growth
in animal protein reliance on inland aquaculture of `r roi_inaq`%
(`r roi_inaq_lb`-`r roi_inaq_ub`%; Figure 1C).

---

## Figure 2: Look at contributing production sources to a country's rate of
change in their aggregate and disaggregate aquatic animal reliance between
1996 and 2023.

The following plot will have two panels, A at the top and B at the bottom.

A: a bar plot measuring the number of countries with decreasing and increasing
aggregate reliance between 1996-2023.

B. a stacked bar plot, faceted by region (i.e., North America, South America, 
Africa, Asia, Europe, and Oceania) quantifying the rate of change in aquatic
animal protein reliance disaggregated by aquatic food source (e.g., inland
aquaculture; marine capture) overlaid by the rate of change in aggregate
animal reliance.

On a data note (addressed in the comments in below code chunk), new and old FBS
create a jump in the time series from 2009 to 2010 due to methodological
differences in the datasets. In our analysis, old FBS runs from 1996 to 2009 and
new FBS runs from 2010 to 2023. To maximize reliability in our rate of change
estimates, we will project new FBS animal protein reliance data back over 1996
to 2009 (i.e., back cast) and project old FBS animal protein reliance data
forward over 2010 to 2023 (i.e., forward cast). We will then measure
the rate of change across four scenarios:

  (old FBS)   (new FBS)
1. observed  - observed
2. back cast - observed
3. observed  - forward cast
4. back cast - forward cast

For each country at the aggregate and disaggregate level, we will take the
midpoint of the max and min of these rates of changes to be their change in
aquatic animal protein reliance.

```{r aggregate vs disaggregate change in reliance}
# Shared base theme (mirrors theme_realm_panel style)
theme_reliance <- function(base_size = 8,
                           axis_title_x = TRUE,
                           axis_title_y = FALSE,
                           legend       = TRUE) {
  base <- theme_minimal(base_size = base_size) +
    theme(
      text                    = element_text(family = family),
      plot.background         = element_rect(fill = "transparent", color = NA),
      panel.background        = element_rect(fill = "transparent", color = NA),
      panel.grid.minor        = element_blank(),
      panel.grid.major.x = element_line(
        color = scales::alpha("black", 0.3), linewidth = 0.15
      ),
      panel.grid.major.y = element_line(
        color = scales::alpha("black", 0.06), linewidth = 0.15
      ),
      axis.text               = element_text(
        size = base_size - 3, family = family
        ),
      axis.title.x            = element_text(
        size  = base_size,
        face  = "bold",
        vjust = -0.5,
        margin = margin(t = 4),
        family = family
      ),
      axis.title.y            = element_blank(),
      strip.text              = element_text(
        size = base_size - 1, face = "bold", family = family, color = "black"
        ),
      legend.position         = if (legend) "bottom" else "none",
      legend.title            = element_text(
        size = base_size - 1, face = "bold", family = family
        ),
      legend.text             = element_text(
        size = base_size - 2, family = family
        ),
      legend.key.size         = unit(0.3, "cm"),
      legend.box              = "vertical",
      legend.spacing          = unit(0.1, "cm"),
      legend.margin           = margin(t = 1, b = 0),
      legend.box.margin       = margin(t = -2, b = -2),
      legend.background       = element_rect(fill = "transparent", color = NA),
      legend.box.background   = element_rect(fill = "transparent", color = NA),
      plot.margin             = margin(3, 3, 3, 3)
    )

  if (!axis_title_x) {
    base <- base + theme(axis.title.x = element_blank())
  }

  base
}


bar_layers <- list(
  geom_col(),
  geom_vline(xintercept = 0, linewidth = 0.4, color = "black"),
  geom_point(aes(x = agg_estimate, y = iso3c, color = term), shape = 21,
             fill = "grey20", size = 1.2, stroke = 0.3)
)

##### -------------------------------- PROBLEM:

# FAOSTAT Food balance Sheets are divided into old methodology (-2013)
# and new methodology (2010-). This creates a jump in the data sources in
# aquatic animal protein reliance. To account for this, we will forward cast
# old FAO reliance values to new years and back cast new FAO reliance
# values to old years. Then, we will calculate the rate of change across:

# 1. obs - obs
# 2. obs - forwardcast
# 3. back cast - obs
# 4. back cast - forward cast

# We will take the maximum and minimum values of these as the ranges for
# rate of change in aquatic animal reliance, and will plot the midpoint of these
# ranges.

# Define year ranges for projections (add this before the existing
# projections code)
year_ranges <- tibble(
  data_source = c("old FBS", "new FBS"),
  start_year  = c(1996, 1996),
  end_year    = c(2023, 2023)
)
artis_palette(10)
# Define year ranges for observed data
observed_ranges <- tibble(
  data_source = c("old FBS", "new FBS"),
  min_year = c(1996, 2010),
  max_year = c(2009, 2023)
)

############ RUN THIS ANALYSIS ON DISAGGREGATE RELIANCE

# First we will exclude any countries that have short time series - these short
# time series may have exaggerated rates of change in the aquatic animal protein
# reliance
iso3c_with_at_least_5_yrs <- aquatic_reliance %>%
  # Remove overlapping FBS years
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         food_group == "aquatic") %>%
  group_by(iso3c, year) %>%
  summarize(prop_animal_protein = sum(prop_animal_protein)) %>%
  # Calculate the number of years for each country
  group_by(iso3c) %>%
  summarize(n_years = n()) %>%
  filter(n_years > 5) %>%
  # Only get countries with over five years of data
  pull(iso3c)

# Calculate country aquatic animal protein reliance by habitat and method
disagg_hab_met <- aquatic_reliance %>%
   # Remove overlapping years prioritizing new years
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         food_group == "aquatic") %>%
  # Filter to only countries with greater than 5 year time series
  filter(iso3c %in% iso3c_with_at_least_5_yrs) %>%
  group_by(year, iso3c, habitat, method, data_source) %>%
  summarize(prop_animal_protein = sum(prop_animal_protein), .groups = "drop")


# Obtain linear model intercept and slope estimates to
# create forward / backcasted projections
coefs_hab_met <- disagg_hab_met %>%
  group_by(iso3c, habitat, method, data_source) %>%
  nest() %>%
  mutate(model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
         tidy  = map(model, broom::tidy)) %>%
  unnest(tidy) %>%
  select(data_source, iso3c, habitat, method, term, estimate) %>%
  pivot_wider(values_from = estimate, names_from = term) %>%
  rename(year_coef = year)

# Calculate estimates via forward and back cast from linear model's
# intercept and slope estimates
proj_hab_met <- coefs_hab_met %>%
  left_join(year_ranges, by = "data_source") %>%
  rowwise() %>%
  mutate(year = list(start_year:end_year)) %>%
  unnest(year) %>%
  ungroup() %>%
  mutate(prop_animal_protein = `(Intercept)` + year_coef * year,
         type = case_when(
           
           year < 2010 ~ "backcast",
           year >= 2010 ~ "forwardcast"
           
         )) %>%
  left_join(observed_ranges, by = "data_source") %>%
  filter(year < min_year | year > max_year) %>%
  select(
    data_source,
    iso3c,
    habitat,
    method,
    year_coef,
    year,
    prop_animal_protein,
    type
    )

# Obtain observed disaggregate aquatic animal protein reliance data
obs_hab_met <- disagg_hab_met %>%
  filter(data_source == "old FBS" & year <= 2009 |
           data_source == "new FBS" & year >= 2010) %>%
  mutate(type = "observed")

# Time series notations:
# left side of the _ indicates old FBS years, whereas the right side of the 
# _ indicates new FBS years

# obs_obs - observed to observed
# back_obs - back casted to observed
# obs_forward - observed to forward casted 
# back_forward - back casted to forward casted
calc_roc_disagg <- function(data, type = type) {
  
  data %>%
    group_by(year, iso3c, habitat, method) %>%
    summarize(prop_animal_protein = sum(prop_animal_protein),
              .groups = "drop") %>%
    group_by(iso3c, habitat, method) %>%
    nest() %>%
    mutate(model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
           tidy  = map(model, broom::tidy)) %>%
    unnest(tidy) %>%
    filter(term == "year") %>%
    mutate(
      group_label = paste(habitat, method),
      group_label = case_when(
        group_label == "unknown unknown" ~ "unknown",
        TRUE ~ group_label
      )) %>%
    drop_na(estimate) %>%
    group_by(group_label) %>%
    ungroup() %>%
    mutate(group_label = factor(
      group_label,
      levels = c(
        "marine capture",
        "inland capture",
        "marine aquaculture",
        "inland aquaculture",
        "unknown"
      )
    ),
    type = type)
  
}

# Calculate four different types of rate of change for disaggregate:

# 1. observed - observed
obs_obs <- bind_rows(
  obs_hab_met,
  proj_hab_met
  ) %>%
  filter(type == "observed") %>%
  calc_roc_disagg(type = "obs_obs")

# 2. back casted to observed
back_obs <- bind_rows(
  obs_hab_met,
  proj_hab_met
  ) %>%
  filter(type == "backcast" |
           type == "observed" & data_source == "new FBS") %>%
  calc_roc_disagg(type = "back_obs")
  
# 3. observed to forward casted
obs_forward <- bind_rows(
  obs_hab_met,
  proj_hab_met
  ) %>%
  filter(type == "forwardcast" |
           type == "observed" & data_source == "old FBS") %>%
  calc_roc_disagg(type = "obs_forward")

# 4. observed to forward casted
back_forward <- bind_rows(
  obs_hab_met,
  proj_hab_met
  ) %>%
  filter(type == "forwardcast" |
           type == "backcast") %>%
  calc_roc_disagg(type = "back_forward")

# For each countries' trend in aquatic animal protein reliance
# across habitat and method, we will take the midpoint of the
# min and max rate of change and use that in the plot.
disagg <- bind_rows(
  obs_obs,
  back_obs,
  obs_forward,
  back_forward
  ) %>%
  group_by(iso3c,
           habitat,
           method
           ) %>%
  filter(estimate == min(estimate) |
         estimate == max(estimate)) %>%
  summarize(
    estimate = 100 *(min(estimate) + max(estimate) / 2),
    .groups = "keep"
    ) %>%
  mutate(group_label = paste0(habitat, " ", method),
         term = "year")


############ AGGREGATE RELIANCE

# Because slopes are additive to the aggregate reliance slope, we add the best
# fitting disaggregate slopes for each country to calculate the aggregate
# change in aquatic animal protein reliance

# Calculate country aggregate aquatic animal protein reliance
agg_estimate <- disagg %>%
  group_by(iso3c) %>%
  summarize(agg_estimate = sum(estimate))


# Helper to build a bar panel
make_bar_panel <- function(regions,
                           axis_x_label = FALSE,
                           legend_show  = FALSE,
                           text_size    = 4.75) {
  
  disagg %>%
    add_region("iso3c", "region") %>%
    filter(region %in% regions) %>%
    left_join(agg_estimate, by = "iso3c") %>%
    ggplot(
      aes(
        x = estimate,
        y = fct_reorder(iso3c, -agg_estimate),
        fill = group_label
        )) +
    bar_layers +
    scale_fill_manual(
      values = c(
        "inland aquaculture"    = "#D3AF55",
        "marine aquaculture"    = "#C87534",
        "inland capture"        = "#5DA49A",
        "marine capture"        = "#114F59",
        "unknown unknown"       = "#A8A8A8"
        ),
      labels = c("Marine capture", "Inland capture",
                 "Marine aquaculture", "Inland aquaculture",
                 "Unknown"),
      name = "Source",
      guide = guide_legend(
        order = 1,
        override.aes = list(
          fill = list(
            "#114F59",
            "#5DA49A",
            "#C87534",
            "#D3AF55",
            "#A8A8A8"
          )))) +
    facet_wrap(~ region, scales = "free", nrow = 1) +
    labs(
      x    = "Average annual change in aquatic animal protein reliance (%; 1996–2023)",
      fill = "Source",
      color = ""
    ) +
    theme_reliance(
      axis_title_x = axis_x_label,
      legend       = legend_show
    ) +
    theme(axis.text = element_text(size = text_size, family = family),
          strip.text = element_text(family = family)) +
    scale_color_manual(
      values = "white",
      label = "aggregate reliance",
      guide = guide_legend(
        order = 2,
        override.aes = list(
        fill = list(
          pattern(
            grobTree(
              circleGrob(
                x = 0.5, 
                y = 0.5, 
                r = 0.3,
                gp = gpar(fill = "grey20")
                )),
            extend = "repeat",
            gp = gpar(col = "grey20", fill = "white", lwd = 0.25)
            )),
        color = "grey20",
        linewidth = 0.2,
        size = 3.5
        )))
}


# Build panels - they are separated into two rows because including all in one
# would make the countries too small to see
all_regions   <- unique(add_region(disagg, "iso3c", "region")$region)
top_regions   <- all_regions[!all_regions %in% c("Oceania", "South America")]
bot_regions   <- c("Oceania", "South America")

# Top row of bars (i.e.. North America, Asia, Africa, Europe)
top_row <- make_bar_panel(
  top_regions,
  axis_x_label = FALSE,
  legend_show = FALSE,
  text_size = 4.75) +
  theme(plot.margin = margin(t = 0,  # Top margin
                             r = 5,  # Right margin
                             b = 0,  # Bottom margin
                             l = 5))

# Top row of bars (i.e.. Africa and Oceania)
bottom_row <- make_bar_panel(
  bot_regions,
  axis_x_label = TRUE,
  legend_show = TRUE,
  text_size = 5.15)

# Since there are less plots in the bottom row, add spacers in to make
# the aspect ratio of the bottom plots look similar in size to the top plots
bottom_row_narrowed <- cowplot::plot_grid(
  NULL, bottom_row, NULL,
  ncol       = 3,
  rel_widths = c(0.1, 0.3, 0.1)
) %>%
  # Supress warning - this warning only has to do with fonts, no data
  # or visualization is harmed via this warning
  suppressWarnings()

# combine the bottom and top bar plots (this will make up the bottom panel of
# the plot)
bottom_panel <- cowplot::plot_grid(
  top_row,
  bottom_row_narrowed,
  nrow        = 2,
  rel_heights = c(2, 1.25)
) %>%
  # Supress warning - this warning only has to do with fonts, no data
  # or visualization is harmed via this warning
  suppressWarnings()


# Calculate the number of countries used in analysis
num_countries <- agg_estimate %>%
  ungroup() %>%
  distinct(iso3c) %>%
  nrow()

# Top panel: # of all countries with increasing / decreasing reliance
top_panel <- agg_estimate %>%
  mutate(direction = case_when(
    agg_estimate < 0 ~ "Decreasing reliance",
    agg_estimate > 0 ~ "Increasing reliance"
  )) %>%
  count(direction) %>%
  mutate(direction = factor(
    direction,
    levels = c("Increasing reliance", "Decreasing reliance")
    )) %>%
  ggplot(aes(x = 1, y = n, fill = direction)) +
  coord_flip() +
  geom_col(position = "stack", width = 0.65) +
  scale_fill_manual(
    values = c("#8FACB4", "#b4c461"),
    name   = "Change in aggregate aquatic animal protein reliance (1996-2023)",
    labels = c("Decrease in reliance", "Increase in reliance"),
    guide = guide_legend(override.aes = list(
                        fill = list(
                          "#b4c461",
                          "#8FACB4"
                          ),
                        linewidth = 0.15,
                        size = 3.5
                        
                      ))) +
  labs(y = paste0("Number of countries (", num_countries, " total)")) +
  theme_reliance(axis_title_x = TRUE, legend = TRUE) +
  theme(
    axis.text.y  = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.x = element_text(size = 8, face = "bold", family = family),
    axis.text.x = element_text(family = family),
    legend.title = element_text(family = family, size = 7, hjust = 1.25),
    legend.text = element_text(family = family, size = 6),

    panel.grid.major.y = element_blank(),
    aspect.ratio = 0.05
  )


# Combine panels
final_plot <- cowplot::plot_grid(
  top_panel,
  bottom_panel,
  nrow          = 2,
  rel_heights   = c(0.35, 2),
  labels        = c("A", "B"),
  label_size    = 9.5
) %>%
  # Supress warning - this warning only has to do with fonts, no data
  # or visualization is harmed via this warning
  suppressWarnings()

# Print final plot
print(final_plot)


# Save plot
ggsave(
  filename = "../images/pct_reliance_contribution_four_sources.pdf",
  plot     = final_plot,
  device   = cairo_pdf,
  width    = 8,
  height   = 6
)
```

### How many countries have increased / decreased aggregate aquatic animal reliance?

```{r}
# Calculate the total number of countries in our analysis
num_countries <- agg_estimate %>% distinct(iso3c) %>% nrow()

# Calculate the number of countries 
num_decreasing_global <- agg_estimate %>%
  count(is_neg = agg_estimate > 0) %>%
  filter(is_neg == FALSE) %>%
  pull(n)
```

Our disaggregation of aquatic animal protein reliance also exposes a mosaic of trends across countries and regions (Figure 2B, SI Figure 9).
In aggregate, aquatic animal protein reliance declined in `r num_decreasing_global` out of `r num_countries` countries between 1996 and 2023.

### What is the percent contribution of inland and marine capture to countries' declining aggregate reliance?

```{r}
# How many countries with decreases in aggregate aquatic animal protein reliance
# have over 80% of their declines attributed to capture?
num_countries_cap <- disagg %>%
  select(
    -group_label,
    -term
    ) %>%
  rename(disagg_estimate = estimate) %>%
  group_by(iso3c, method) %>%
  summarize(disagg_estimate = sum(disagg_estimate)) %>%
  left_join(agg_estimate) %>%
  mutate(rel_is_neg = disagg_estimate < 0) %>%
  filter(agg_estimate < 0, rel_is_neg == TRUE) %>%
  group_by(iso3c) %>%
  mutate(pct_contribution = 100 *( disagg_estimate / sum(disagg_estimate))) %>%
  filter(method == "capture") %>%
  filter(pct_contribution > 80) %>%
  filter(agg_estimate < -0.15) %>%
  nrow()
  
# How many countries with moderate to steep increases in aggregate aquatic
# animal protein reliance have over 60% of their growth attributed to
# aquaculture?
num_countries_aq <- disagg %>%
  select(
    -group_label,
    -term
    ) %>%
  rename(disagg_estimate = estimate) %>%
  group_by(iso3c, method) %>%
  summarize(disagg_estimate = sum(disagg_estimate)) %>%
  left_join(agg_estimate) %>%
  mutate(rel_is_pos = disagg_estimate > 0) %>%
  filter(agg_estimate > 0, rel_is_pos == TRUE) %>%
  group_by(iso3c) %>%
  mutate(pct_contribution = 100 *( disagg_estimate / sum(disagg_estimate))) %>%
  filter(method == "aquaculture") %>%
  filter(pct_contribution > 60) %>%
  arrange(-agg_estimate) %>%
  filter(agg_estimate > 0.15) %>%
  nrow()
```

With the disaggregate reliance, we expose a major attributor for the shift away from aquatic animal proteins: widespread declines in animal protein reliance on capture fisheries have contributed over 80% towards negative shifts in the aggregate aquatic animal protein reliance for `r num_countries_cap` countries.
In contrast, substantial growth in animal protein reliance on aquaculture has been concentrated, where only `r num_countries_aq` countries that moderately to sharply increased their aggregate aquatic animal protein reliance reliance attributed over 60\% of their growth to aquaculture.
<!---data obtained from the following sentences through 03_Supplementary_Information.Rmd --->
We find nations throughout much of Asia, Africa, South America, and Oceania decreased their reliance on capture fisheries, whereas North American nations widely increased their reliance on aquaculture (SI Figures 10-15).
Uniquely, nearly every nation in Europe, 38 out of 40, increased their reliance on marine aquaculture, underscoring the widely varying aquatic animal protein reliance (SI Figures 10-15).

# RQ 1b: How does aquatic animal reliance change with the relative contribution of aquatic versus terrestrial animal consumption

## Figure 3: Scatterplot (terrestrial consumption ~ aquatic consumption; colored by: aquatic reliance) + bar chart overlay 

Scatterplot of average annual change in daily terrestrial animal protein consumption (kg/capita/day) 2010–2023 (y-axis) with respect to the average annual change in daily aquatic animal protein consumption (kg/capita/day) 2010–2023 colored by the annual change in daily aquatic animal protein reliance(%) 2010-2023, overlaid by a 1:1 line. The bar plots represent the number of countries in each quadrant, with the highest count of countries in any quadrant being 21. The two right hand y-axes represent these number of countries.

```{r}
# Region palette — desaturated analogues of the RdBu diverging scale
region_colors <- c(
  "Africa"        = "#6B2737",
  "North America" = "#C45C3A",
  "South America" = "#D4956A",
  "Asia"          = "#4A7FA5",
  "Europe"        = "#7BAEC4",
  "Oceania"       = "#A8C8D8"
)

# Shared base theme
theme_scatter <- function(base_size = 8) {
  theme_minimal(base_size = base_size) +
    theme(
      text                  = element_text(color = "black", family = family),
      plot.background       = element_rect(fill = "transparent", color = NA),
      panel.background      = element_rect(fill = "transparent", color = NA),
      panel.grid.minor      = element_blank(),
      panel.grid.major      = element_line(
        color = scales::alpha("black", 0.12), linewidth = 0.3
      ),
      axis.text             = element_text(size = base_size - 2, color = "black", family = family),
      axis.title            = element_text(size = base_size - 2, face = "bold", family = family),
      legend.position       = "bottom",
      legend.title          = element_text(size = base_size - 2, face = "bold", family = family),
      legend.text           = element_text(size = base_size - 3, family = family),
      legend.key.size       = unit(0.3, "cm"),
      legend.box            = "horizontal",
      legend.spacing.x      = unit(0.4, "cm"),
      legend.margin         = margin(t = 0, b = 0),
      legend.box.margin     = margin(t = -2, b = -2),
      legend.background     = element_rect(fill = "transparent", color = NA),
      legend.box.background = element_rect(fill = "transparent", color = NA),
      plot.margin           = margin(3, 4, 3, 4)
    )
}

##### -------------------------------- PROBLEM:

# FAOSTAT Food balance Sheets are divided into old methodology (-2013)
# and new methodology (2010-). This creates a jump in the data sources in
# aquatic animal protein reliance. To account for this, we will forward cast
# old FAO reliance values to new years and back cast new FAO reliance
# values to old years. Then, we will calculate the rate of change across:

# 1. obs - obs
# 2. obs - forwardcast
# 3. back cast - obs
# 4. back cast - forward cast

# We will take the maximum and minimum values of these as the ranges for
# rate of change in aquatic animal reliance, and will plot the midpoint of these
# ranges.

# Define year ranges for projections
year_ranges <- tibble(
  data_source = c("old FBS", "new FBS"),
  start_year  = c(1996, 1996),
  end_year    = c(2023, 2023)
)

# Define year ranges for observed data
observed_ranges <- tibble(
  data_source = c("old FBS", "new FBS"),
  min_year = c(1996, 2010),
  max_year = c(2009, 2023)
)

# First we will exclude any countries that have short time series - these short
# time series may have exaggerated rates of change in the aquatic animal protein
# reliance
iso3c_with_at_least_5_yrs <- aquatic_reliance %>%
  # Remove overlapping FBS years
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         food_group == "aquatic") %>%
  group_by(iso3c, year) %>%
  summarize(prop_animal_protein = sum(prop_animal_protein)) %>%
  # Calculate the number of years for each country
  group_by(iso3c) %>%
  summarize(n_years = n()) %>%
  filter(n_years > 5) %>%
  # Only get countries with over five years of data
  pull(iso3c)

######### AQUATIC ANIMAL PROTEIN CONSUMPTION RATE OF CHANGE (1996-2023)

# Calculate country aggregate aquatic animal protein reliance
aquatic_protein_quant <-  aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         food_group == "aquatic") %>%
  group_by(data_source, iso3c, year) %>%
  summarize(protein_consumed_kg_percap_peryr = sum(protein_consumed_kg_percap_peryr, na.rm = TRUE),
            .groups = "drop") %>%
  # Filter to only countries with greater than 5 year time series
  filter(iso3c %in% iso3c_with_at_least_5_yrs)
  
# Obtain linear model intercept and slope estimates to create forward / backcasted projections
coefs_aq_cons <- aquatic_protein_quant %>%
  group_by(data_source, iso3c) %>%
  nest() %>%
  mutate(model = map(data, ~ lm(protein_consumed_kg_percap_peryr ~ year, data = .x)),
         tidy  = map(model, broom::tidy)) %>%
  unnest(tidy) %>%
  select(data_source, iso3c, term, estimate) %>%
  pivot_wider(values_from = estimate, names_from = term) %>%
  rename(year_coef = year)

# Calculate forwardcasts and backcasts from linear model intercept and slope estimates
proj_aq_cons <- coefs_aq_cons %>%
  left_join(year_ranges, by = "data_source") %>%
  rowwise() %>%
  mutate(year = list(start_year:end_year)) %>%
  unnest(year) %>%
  ungroup() %>%
  mutate(protein_consumed_kg_percap_peryr = `(Intercept)` + year_coef * year,
         type = case_when(
           
           year < 2010 ~ "backcast",
           year >= 2010 ~ "forwardcast"
           
         )) %>%
  left_join(observed_ranges, by = "data_source") %>%
  filter(year < min_year | year > max_year)

# Obtain observed aggregate aquatic animal protein reliance data
obs_aq_cons <- aquatic_protein_quant %>%
  filter(data_source == "old FBS" & year <= 2009 |
           data_source == "new FBS" & year >= 2010) %>%
  mutate(type = "observed")

# Time series notations:
# left side of the _ indicates old FBS years, whereas the right side of the 
# _ indicates new FBS years

# obs_obs - observed to observed
# back_obs - back casted to observed
# obs_forward - observed to forward casted 
# back_forward - back casted to forward casted

# Helper function to calculate the rate of change in a variable over time
calc_roc <- function(data, type = type, var_name = "prop_animal_protein") {
  
  data %>%
  group_by(iso3c) %>%
  nest() %>%
  mutate(
    model = map(data, ~ lm(!!sym(var_name) ~ year, data = .x)),
    tidy  = map(model, broom::tidy)
  ) %>%
  unnest(tidy) %>%
  filter(term == "year") %>%
  drop_na(estimate) %>%
  ungroup() %>%
    mutate(type = type)
  
}

# 1. observed - observed
obs_obs <- bind_rows(
  obs_aq_cons,
  proj_aq_cons
  ) %>%
  filter(type == "observed") %>%
  calc_roc(type = "obs_obs", var_name = "protein_consumed_kg_percap_peryr")

# 2. back casted to observed
back_obs <- bind_rows(
  obs_aq_cons,
  proj_aq_cons
  ) %>%
  filter(type == "backcast" |
           type == "observed" & data_source == "new FBS") %>%
  calc_roc(type = "back_obs", var_name = "protein_consumed_kg_percap_peryr")
  
# 3. observed to forward casted
obs_forward <- bind_rows(
  obs_aq_cons,
  proj_aq_cons
  ) %>%
  filter(type == "forwardcast" |
           type == "observed" & data_source == "old FBS") %>%
  calc_roc(type = "obs_forward", var_name = "protein_consumed_kg_percap_peryr")

# 4. observed to forward casted
back_forward <- bind_rows(
  obs_aq_cons,
  proj_aq_cons
  ) %>%
  filter(type == "forwardcast" |
           type == "backcast") %>%
  calc_roc(type = "back_forward", var_name = "protein_consumed_kg_percap_peryr")


# For each countries' trend in aggregate aquatic animal protein reliance
# we will take the midpoint of the min and max rate of
# change and use that in the plot.
aquatic_protein_quant <- bind_rows(
  obs_obs,
  back_obs,
  obs_forward,
  back_forward
  ) %>%
  group_by(iso3c) %>%
  summarize(aquatic_estimate = min(estimate) + max(estimate) / 2)

######### TERRESTRIAL ANIMAL PROTEIN CONSUMPTION RATE OF CHANGE (1996-2023)

# Calculate country aggregate aquatic animal protein reliance
terrestrial_protein_quant <-  aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"), food_group == "terrestrial") %>%
  group_by(data_source, iso3c, year) %>%
  summarize(protein_consumed_kg_percap_peryr = sum(protein_consumed_kg_percap_peryr, na.rm = TRUE),
            .groups = "drop") %>%
  # Filter to only countries with greater than 5 year time series
  filter(iso3c %in% iso3c_with_at_least_5_yrs)

# Obtain linear model intercept and slope estimates to create forward / backcasted projections
coefs_terr_cons <- terrestrial_protein_quant %>%
  group_by(data_source, iso3c) %>%
  nest() %>%
  mutate(model = map(data, ~ lm(protein_consumed_kg_percap_peryr ~ year, data = .x)),
         tidy  = map(model, broom::tidy)) %>%
  unnest(tidy) %>%
  select(data_source, iso3c, term, estimate) %>%
  pivot_wider(values_from = estimate, names_from = term) %>%
  rename(year_coef = year)

# Calculate forwardcasts and backcasts from linear model intercept and slope estimates
proj_terr_cons <- coefs_terr_cons %>%
  left_join(year_ranges, by = "data_source") %>%
  rowwise() %>%
  mutate(year = list(start_year:end_year)) %>%
  unnest(year) %>%
  ungroup() %>%
  mutate(protein_consumed_kg_percap_peryr = `(Intercept)` + year_coef * year,
         type = case_when(
           
           year < 2010 ~ "backcast",
           year >= 2010 ~ "forwardcast"
           
         )) %>%
  left_join(observed_ranges, by = "data_source") %>%
  filter(year < min_year | year > max_year)

# Obtain observed aggregate aquatic animal protein reliance data
obs_terr_cons <- terrestrial_protein_quant %>%
  filter(data_source == "old FBS" & year <= 2009 |
           data_source == "new FBS" & year >= 2010) %>%
  mutate(type = "observed")

# Time series notations:
# left side of the _ indicates old FBS years, whereas the right side of the 
# _ indicates new FBS years

# obs_obs - observed to observed
# back_obs - back casted to observed
# obs_forward - observed to forward casted 
# back_forward - back casted to forward casted

# Calculate four different types of rate of change for disaggregate:

# 1. observed - observed
obs_obs <- bind_rows(
  obs_terr_cons,
  proj_terr_cons
  ) %>%
  filter(type == "observed") %>%
  calc_roc(type = "obs_obs", var_name = "protein_consumed_kg_percap_peryr")

# 2. back casted to observed
back_obs <- bind_rows(
  obs_terr_cons,
  proj_terr_cons
  ) %>%
  filter(type == "backcast" |
           type == "observed" & data_source == "new FBS") %>%
  calc_roc(type = "back_obs", var_name = "protein_consumed_kg_percap_peryr")
  
# 3. observed to forward casted
obs_forward <- bind_rows(
  obs_terr_cons,
  proj_terr_cons
  ) %>%
  filter(type == "forwardcast" |
           type == "observed" & data_source == "old FBS") %>%
  calc_roc(type = "obs_forward", var_name = "protein_consumed_kg_percap_peryr")

# 4. observed to forward casted
back_forward <- bind_rows(
  obs_terr_cons,
  proj_terr_cons
  ) %>%
  filter(type == "forwardcast" |
           type == "backcast") %>%
  calc_roc(type = "back_forward", var_name = "protein_consumed_kg_percap_peryr")


# For each countries' trend in aggregate aquatic animal protein reliance
# we will take the midpoint of the min and max rate of
# change and use that in the plot.
terrestrial_protein_quant <- bind_rows(
  obs_obs,
  back_obs,
  obs_forward,
  back_forward) %>%
  group_by(iso3c) %>%
  summarize(terrestrial_estimate = min(estimate) + max(estimate) / 2)


######### PLOTTING

# Join all three model outputs and add region
plot_df <- full_join(
  aquatic_protein_quant,
  terrestrial_protein_quant,
  by = "iso3c"
  ) %>%
  add_region("iso3c", "region")

# Pre-compute quadrant membership for corner insets
quad_counts <- plot_df %>%
  mutate(quadrant = case_when(
    aquatic_estimate > 0 & terrestrial_estimate > 0 ~ "Q1",
    aquatic_estimate < 0 & terrestrial_estimate > 0 ~ "Q2",
    aquatic_estimate < 0 & terrestrial_estimate < 0 ~ "Q3",
    aquatic_estimate > 0 & terrestrial_estimate < 0 ~ "Q4"
  )) %>%
  group_by(quadrant, region) %>%
  summarize(n = n_distinct(iso3c), .groups = "drop")

max_count <- max(quad_counts$n)

# Dummy bar plot used only to extract a square-keyed region legend
# geom_col produces filled rect keys in the legend, which is cleaner
# than the point/circle keys that come from geom_point
p_region_legend_source <- plot_df %>%
  group_by(region) %>%
  summarize(n = n(), .groups = "drop") %>%
  ggplot(aes(x = region, y = n, fill = region)) +
  geom_col() +
  scale_fill_manual(
    values = artis_palette(6),
    name   = "Region",
    guide  = guide_legend(
      title.position = "top",
      title.hjust    = 0.5,
      nrow           = 2
    )
  ) +
  theme_scatter() +
  theme(legend.key.size = unit(0.3, "cm"))

legend_regions <- cowplot::get_plot_component(
  p_region_legend_source, "guide-box", return_all = TRUE
)
legend_regions <- Filter(function(x) !inherits(x, "zeroGrob"), legend_regions)[[1]]

# Main scatterplot — fill encodes reliance change, no region border encoding
# on points (region is shown only in the inset bar charts and the square legend)
p_scatter_base <- plot_df %>%
  mutate(aquatic_outpaces_terrestrial = aquatic_estimate > terrestrial_estimate) %>%
  # filter(iso3c != "MDV") %>%
  ggplot(aes(
    x     = aquatic_estimate,
    y     = terrestrial_estimate,
    fill  = aquatic_outpaces_terrestrial,
    label = iso3c
  )) +
  scale_fill_manual(labels = c("Terr. outpaces aq. animal protein cons.", "Aq. outpaces terr animal protein cons."), values = c("#b4c461", "#8FACB4")) +
  # scale_fill_manual(values = artis_palette(6)) +
  geom_hline(yintercept = 0, linewidth = 0.4, color = "black") +
  geom_vline(xintercept = 0, linewidth = 0.4, color = "black") +
  geom_abline(intercept = 0, slope = 1,
              linetype = "dashed", linewidth = 0.35, color = "black") +
  geom_point(shape = 21, size = 1.8, stroke = 0.25, alpha = 0.7
  ) +
  geom_text_repel(
    size               = 2.0,
    min.segment.length = 0.2,
    segment.size       = 0.25,
    segment.color      = "grey60",
    show.legend        = FALSE,
    family = family
  ) +
  lims(x = c(-0.25, 0.25), y = c(-4.5, 4.5)) +
  labs(
    x = "Average annual change in aquatic animal protein consumption (kg/capita/year) 1996–2023",
    y = "Average annual change in terrestrial animal protein consumption\n(kg/capita/year) 1996–2023") +
  theme_scatter() +
  guides(fill=guide_legend(nrow=2)) +
  theme(legend.title = element_blank())


# Extract colorbar legend from scatter before stripping it
legend_reliance <- cowplot::get_plot_component(
  p_scatter_base, "guide-box", return_all = TRUE
)
legend_reliance <- Filter(function(x) !inherits(x, "zeroGrob"), legend_reliance)[[1]]

# Remove legend from scatter for the main panel
p_scatter <- p_scatter_base + theme(legend.position = "none")

# Helper to build a single quadrant inset bar chart
make_inset <- function(q) {
  show_y <- q %in% c("Q1", "Q4")

  quad_counts %>%
    filter(quadrant == q) %>%
    ggplot(aes(x = region, y = n, fill = region)) +
    geom_col(width = 0.7) +
    scale_fill_manual(values = artis_palette(6)) +
    scale_y_continuous(
      limits   = c(0, max_count),
      breaks   = c(0, max_count),
      position = "right"
    ) +
    labs(y = if (show_y) "No. countries" else NULL) +
    theme_minimal(base_size = 6) +
    theme(
      axis.title.x    = element_blank(),
      axis.title.y    = if (show_y) element_text(size = 4.5, color = "black", family = family) else element_blank(),
      axis.text.x     = element_blank(),
      axis.text.y     = if (show_y) element_text(size = 4.5, color = "black", family = family) else element_blank(),
      axis.ticks.y    = if (show_y) element_line(linewidth = 0.2, color = "black") else element_blank(),
      panel.grid      = element_blank(),
      plot.background = element_rect(fill = scales::alpha("white", 0.75), color = NA),
      legend.position = "none",
      plot.margin     = margin(0, 0, 0, 0)
    )
}

# Build all four corner insets
p_q1 <- make_inset("Q1")
p_q2 <- make_inset("Q2")
p_q3 <- make_inset("Q3")
p_q4 <- make_inset("Q4")

# Position insets in scatterplot corners using patchwork::inset_element
shrink_x <- 0.18
shrink_y <- 0.18

p_with_insets <- p_scatter +
  inset_element(p_q1, left = 1 - shrink_x, right = 1, bottom = 1 - shrink_y, top = 1) +
  inset_element(p_q2, left = 0,             right = shrink_x, bottom = 1 - shrink_y, top = 1) +
  inset_element(p_q3, left = 0,             right = shrink_x, bottom = 0,            top = shrink_y) +
  inset_element(p_q4, left = 1 - shrink_x, right = 1,        bottom = 0,            top = shrink_y)

# Place colorbar and region legend side by side below the main panel
legend_row <- cowplot::plot_grid(
  legend_reliance, legend_regions
) %>%
  # Supress warning - this warning only has to do with fonts, no data
  # or visualization is harmed via this warning
  suppressWarnings()

# Stack main panel above legend row
final_plot <- cowplot::plot_grid(
  p_with_insets,
  legend_row,
  ncol        = 1,
  rel_heights = c(1, 0.16)
) %>%
  # Supress warning - this warning only has to do with fonts, no data
  # or visualization is harmed via this warning
  suppressWarnings()

# Print final plot
print(final_plot)

# Save plot
ggsave(
  filename = "../images/scatter_bar_overlay.pdf",
  plot     = final_plot,
  device   = cairo_pdf,
  width    = 5,
  height   = 3.5,
  units    = "in"
)
```

### How many countries increased their animal protein consumption between 1996 and 2023?

```{r}
plot_df %>%
  mutate(quadrant = case_when(
    aquatic_estimate > 0 & terrestrial_estimate > 0 ~ "Q1",
    aquatic_estimate < 0 & terrestrial_estimate > 0 ~ "Q2",
    aquatic_estimate < 0 & terrestrial_estimate < 0 ~ "Q3",
    aquatic_estimate > 0 & terrestrial_estimate < 0 ~ "Q4"
  )) %>%
  filter(quadrant == "Q3",
         region == "Africa") %>%
  mutate(aq_outpaces_terr = case_when(
    aquatic_estimate > terrestrial_estimate ~ TRUE,
    aquatic_estimate < terrestrial_estimate ~ FALSE,
  )) %>%
  count(aq_outpaces_terr)
```


```{r}
######### ANIMAL PROTEIN CONSUMPTION RATE OF CHANGE (1996-2023)

# Calculate country aggregate aquatic animal protein reliance
animal_protein_quant <-  aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS")) %>%
  group_by(data_source, iso3c, year) %>%
  summarize(
    protein_consumed_kg_percap_peryr = sum(protein_consumed_kg_percap_peryr,
                                           na.rm = TRUE),
            .groups = "drop"
    ) %>%
  # Filter time series so that each country has at least 5 data points
  group_by(iso3c) %>%
  mutate(n_years = n()) %>%
  ungroup() %>%
  filter(n_years > 5) %>%
  distinct() 

# Obtain linear model intercept and slope estimates to create forward
# backcasted projections
coefs_aq_cons <- animal_protein_quant %>%
  group_by(data_source, iso3c) %>%
  nest() %>%
  mutate(
    model = map(data, ~ lm(protein_consumed_kg_percap_peryr ~ year, data = .x)),
    tidy  = map(model, broom::tidy)
    ) %>%
  unnest(tidy) %>%
  select(data_source, iso3c, term, estimate) %>%
  pivot_wider(values_from = estimate, names_from = term) %>%
  rename(year_coef = year)

# Calculate forwardcasts and backcasts from linear model intercept
# and slope estimates
proj_animal_cons <- coefs_aq_cons %>%
  left_join(year_ranges, by = "data_source") %>%
  rowwise() %>%
  mutate(year = list(start_year:end_year)) %>%
  unnest(year) %>%
  ungroup() %>%
  mutate(protein_consumed_kg_percap_peryr = `(Intercept)` + year_coef * year,
         type = case_when(
           
           year < 2010 ~ "backcast",
           year >= 2010 ~ "forwardcast"
           
         )) %>%
  left_join(observed_ranges, by = "data_source") %>%
  filter(year < min_year | year > max_year)

# Obtain observed aggregate aquatic animal protein reliance data
obs_animal_cons <- animal_protein_quant %>%
  filter(data_source == "old FBS" & year <= 2009 |
           data_source == "new FBS" & year >= 2010) %>%
  mutate(type = "observed")

# Time series notations:
# left side of the _ indicates old FBS years, whereas the right side of the 
# _ indicates new FBS years

# obs_obs - observed to observed
# back_obs - back casted to observed
# obs_forward - observed to forward casted 
# back_forward - back casted to forward casted

# Calculate four different types of rate of change for disaggregate:

# 1. observed - observed
obs_obs <- bind_rows(
  obs_animal_cons,
  proj_animal_cons
  ) %>%
  filter(type == "observed") %>%
  calc_roc(type = "obs_obs", var_name = "protein_consumed_kg_percap_peryr")

# 2. back casted to observed
back_obs <- bind_rows(
  obs_animal_cons,
  proj_animal_cons
  ) %>%
  filter(type == "backcast" |
           type == "observed" & data_source == "new FBS") %>%
  calc_roc(type = "back_obs", var_name = "protein_consumed_kg_percap_peryr")
  
# 3. observed to forward casted
obs_forward <- bind_rows(
  obs_animal_cons,
  proj_animal_cons
  ) %>%
  filter(type == "forwardcast" |
           type == "observed" & data_source == "old FBS") %>%
  calc_roc(type = "obs_forward", var_name = "protein_consumed_kg_percap_peryr")

# 4. observed to forward casted
back_forward <- bind_rows(
  obs_animal_cons,
  proj_animal_cons
  ) %>%
  filter(type == "forwardcast" |
           type == "backcast") %>%
  calc_roc(type = "back_forward", var_name = "protein_consumed_kg_percap_peryr")


# For each countries' trend in aggregate aquatic animal protein reliance
# we will take the midpoint of the min and max rate of
# change and use that in the plot.
animal_protein_quant <- bind_rows(
  obs_obs,
  back_obs,
  obs_forward,
  back_forward
  ) %>%
  group_by(iso3c) %>%
  summarize(animal_protein_estimate = min(estimate) + max(estimate) / 2)

# Get the number of countries that increased their animal protein consumption
num_increasing_animal <- animal_protein_quant %>%
  count(roc_is_pos = animal_protein_estimate > 0) %>%
  filter(roc_is_pos == TRUE) %>%
  pull(n)


mean_increase <- animal_protein_quant %>%
  mutate(roc_is_pos = animal_protein_estimate > 0) %>%
  filter(roc_is_pos == TRUE) %>%
  summarize(mean_increase = mean(animal_protein_estimate))
```

### What is the global annual change in animal protein consumption?

```{r}
# Load in population data
population <- read_csv("../data/world_population.csv")

# Calculate annual global population
pop <- population %>%
  mutate(
    # remove spaces from population rows and convert to string
    population_thousands = gsub("[[:space:]]", "", population_thousands),
    population_thousands = as.numeric(population_thousands),
    # Convert population to  units of one
    population = population_thousands * 1000
    ) %>%
  select(-population_thousands)
  
# Calculate the global animal protein consumption in kilograms / year / cap
animal_cons <- aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS")) %>%
  group_by(year) %>%
  summarize(protein_consumed_t_peryr = sum(protein_consumed_t_peryr)) %>%
  left_join(pop) %>%
  # Convert to kg / year / cap
  mutate(kg_year_cap = (protein_consumed_t_peryr / 0.001) / population) %>%
  select(year, kg_year_cap)

# Calculate global rate of change per year
lm <- lm(kg_year_cap ~ year, data = animal_cons)
animal_cons_global_roc <- lm$coefficients[2] %>% round(3) %>% unname()
```

In total, animal protein intake has broadly increased since 1996, with `r num_increasing_animal` nations increasing their animal protein consumption and a global increase of `r animal_cons_global_roc` kg cap$^{-1}$ yr$^{-1}$

```{r}
plot_df %>%
  mutate(aquatic_outpaces_terrestrial = aquatic_estimate > terrestrial_estimate) %>%
  View()
```


```{r}
# Are a majority of North American countries' terrestrial food consumption
# outpacing their aquatic food consumption?
  left_join(terrestrial_protein_quant,aquatic_protein_quant) %>%
  left_join(aquatic_protein_reliance) %>%
  exploreARTIS::add_region("iso3c", "region") %>%
    filter(region == "North America") %>%
    count(terr_greater_than_aquatic = terrestrial_estimate > aquatic_estimate)
# Answer: yes
  
# Are a majority of European countries' terrestrial food consumption outpacing
# their aquatic food consumption?
  left_join(terrestrial_protein_quant,aquatic_protein_quant) %>%
  exploreARTIS::add_region("iso3c", "region") %>%
    filter(region == "South America") %>%
    count(terr_greater_than_aquatic = terrestrial_estimate > aquatic_estimate)
# Answer: yes

# Are a majority of South American countries' terrestrial food consumption
# outpacing their aquatic food consumption?
  left_join(terrestrial_protein_quant,aquatic_protein_quant) %>%
  exploreARTIS::add_region("iso3c", "region") %>%
    filter(region == "South America") %>%
    count(terr_greater_than_aquatic = terrestrial_estimate > aquatic_estimate)
# Answer: yes
  
# Case study 1 example: Marshall Islands
mhl_vals <- plot_df %>%
  filter(iso3c == "MHL") %>%
  pivot_longer(cols = c("aquatic_estimate", "terrestrial_estimate", "aquatic_reliance_estimate")) %>%
  pull(value) %>%
  round(2)
# Case study 2 example: Senegal
sen_vals <- plot_df %>%
  filter(iso3c == "SEN") %>%
  pivot_longer(cols = c("aquatic_estimate", "terrestrial_estimate", "aquatic_reliance_estimate")) %>%
  pull(value) %>%
  round(2)
```

For example, while annual consumption of aquatic and terrestrial animal proteins have increased within a majority of European, North American, and South American nations, a majority of these nations’ terrestrial consumption has outpaced their aquatic food consumption, highlighting a broader shift toward terrestrial reliance (SI Figure).

> Of these `r num_countries_increasing_animal_cons` 61% decreased in aquatic animal reliance, indicating a shift toward terrestrial animal protein consumption.

> These changes were in animal consumption were highly attributed to increasing and declining terrestrial animal consumption, as 61% of countries that increased their animal protein consumption, decreased reliance on aquatic foods

> Only 44 X countries have only increased their terrestrial animal protein consumption, Y countries decreased both their terrestrial and aquatic animal protein consumptionand Z countries only increased their aquatic animal protein consumption

Of those countries, 86 have increased both in their terrestrial and aquatic animal consumption, 38 have only increased their terrestrial animal consumption, and 4 have only increased their aquatic animal consumption.


# RQ 2: What biophysical and socioeconomic factors are associated with reliance on farmed and wild aquatic animals for protein and overall consumption?

## Create list of response variables & their associated data to put
## through Bayesian model

This reads in and grabs the response variable data used across Bayesian models.
The list of response variables will be held in **response_vars**, and the length
of response_vars will represent the number of Bayesian models run.

```{r acquire animal consumption data to be fed through analysis}
years = 2019:2023
# Calculate animal consumption
terr_animal_cons_cap <- aquatic_reliance %>%
  filter(year %in% years) %>%
  group_by(iso3c, year) %>%
  summarize(
    protein_consumed_kg_percap_peryr = sum(protein_consumed_kg_percap_peryr),
    .groups = "drop"
    ) %>%
  group_by(iso3c) %>%
  summarize(protein_consumed_kg_percap_peryr = mean(protein_consumed_kg_percap_peryr))

animal_cons_cap <- terr_animal_cons_cap

# Calculate aquatic animal consumption
aquatic_animal_cons_cap <- aquatic_reliance %>%
  filter(year %in% years, food_group == "aquatic") %>%
  group_by(iso3c, year) %>%
  summarize(
    protein_consumed_kg_percap_peryr = sum(protein_consumed_kg_percap_peryr),
    .groups = "drop"
    ) %>%
  group_by(iso3c) %>%
  summarize(protein_consumed_kg_percap_peryr = mean(protein_consumed_kg_percap_peryr))

# Calculate countries' aggregate aquatic animal protein reliance in 2023
aquatic_rel <- aquatic_animal_cons_cap %>%
  left_join(animal_cons_cap, by = "iso3c") %>%
  group_by(iso3c) %>%
  summarize(
    prop_animal_protein = protein_consumed_kg_percap_peryr.x / protein_consumed_kg_percap_peryr.y,
    .groups = "drop"
    )

# We want to run a

# Start list of response variables that are used in the Bayesian model
response_vars <- c("terr_animal_cons_cap", "aquatic_animal_cons_cap", "aquatic_rel")

for (i in c("capture", "aquaculture", "domestic", "foreign", "marine", "inland")) {
  
  # Calculate consumption per capita by method
  if (i %in% c("capture", "aquaculture")) {
    assign(
      paste0(i, "_cap"),
      aquatic_reliance %>%
        filter(year %in% years, food_group == "aquatic", method == i) %>%
        group_by(iso3c) %>%
        summarize(protein_consumed_kg_percap_peryr = sum(protein_consumed_kg_percap_peryr)) %>%
        group_by(iso3c) %>%
        summarize(protein_consumed_kg_percap_peryr = mean(protein_consumed_kg_percap_peryr))
    )
  }
  
  # Calculate consumption per capita by consumption source
  if (i %in% c("domestic", "foreign")) {
    assign(
      paste0(i, "_cap"),
      aquatic_reliance %>%
        filter(year %in% years, food_group == "aquatic", consumption_source == i) %>%
        group_by(iso3c) %>%
        summarize(protein_consumed_kg_percap_peryr = sum(protein_consumed_kg_percap_peryr)) %>%
        group_by(iso3c) %>%
        summarize(protein_consumed_kg_percap_peryr = mean(protein_consumed_kg_percap_peryr))
    )
  }
  
  # Calculate consumption per capita by habitat
  if (i %in% c("marine", "inland")) {
    assign(
      paste0(i, "_cap"),
      aquatic_reliance %>%
        filter(year %in% years, food_group == "aquatic", habitat == i) %>%
        group_by(iso3c) %>%
        summarize(protein_consumed_kg_percap_peryr = sum(protein_consumed_kg_percap_peryr)) %>%
        group_by(iso3c) %>%
        summarize(protein_consumed_kg_percap_peryr = mean(protein_consumed_kg_percap_peryr))
    )
  }
  
  response_vars <- c(response_vars, paste0(i, "_cap"))
  
  # Calculate reliance by method, production source, or habitat
  assign(
    paste0(i, "_rel"),
    get(paste0(i, "_cap")) %>%
      left_join(animal_cons_cap, by = "iso3c") %>%
      group_by(iso3c) %>%
      summarize(
        prop_animal_protein = protein_consumed_kg_percap_peryr.x /protein_consumed_kg_percap_peryr.y,
        .groups = "keep")
  )
  
  response_vars <- c(response_vars, paste0(i, "_rel"))
  
}

# Save response_vars to RDS format
saveRDS(response_vars, "../data/response_vars.rds")
```

## Bayesian Model

Only run this code chunk if all models aren't already saved

```{r hierarchical bayesian model}
# All response variables
predictors <- c(
  "coastline_km",
  "eez_total",
  "log_gdp",
  "gdp_trade",
  "inundation_area_log1000km2",
  "supermarkets",
  "pop_10k_pct_coast"
)

# Check whether all expected output files already exist for every response
# variable. The three outputs per variable are: X matrix, y vector, and
# the fitted Stan model. If all are present the loop is skipped entirely;
# if any single file is missing the full loop runs.
all_outputs_exist <- all(
  map_lgl(response_vars, function(rv) {
    all(file.exists(
      file.path(base_path, paste0("X_",      rv, ".rds")),
      file.path(base_path, paste0("y_",      rv, ".rds")),
      file.path(base_path, paste0("realms_", rv, ".rds")),
      file.path(base_path, paste0("fit_",    rv, ".rds"))
    ))
  })
)



# Initialize the progress bar
pb <- progress_bar$new(
  format = "  [:bar] :percent :elapsedfull",
  total = length(response_vars), # Number of models to run
  clear = FALSE, # Do not clear the bar after completion
  width = 60 # Width of the bar
)


if (all_outputs_exist) {
  message("All Stan outputs already exist for all response variables — skipping model fitting loop.")
} else {

  for (i in seq_along(response_vars)) {

    # Update progress bar
    pb$tick()

    joined_df <- get(response_vars[i]) %>%
      left_join(ac_df, by = "iso3c") %>%
      left_join(ecoregion, by = "iso3c") %>%
      ungroup() %>%
      left_join(
        indicator_df %>%
          select("eez_total", "iso3c",
                 "inundation_area_log1000km2" = "inland_water_max"),
        by = "iso3c"
      ) %>%
      left_join(coastline_df, by = "iso3c") %>%
      left_join(pop_10k_coast, by = "iso3c") %>%
      mutate(
        log_gdp = log(gdp),
        coastline_km = case_when(
          is.na(coastline_km) & str_detect(realm, "Landlocked") ~ 0,
          TRUE ~ coastline_km
        ),
        eez_total = case_when(
          is.na(eez_total) & str_detect(realm, "Landlocked") ~ 0,
          TRUE ~ eez_total
        )
      ) %>%
      select(iso3c, 2, all_of(predictors), realm) %>%
      drop_na() %>%
      filter(realm != "")
    
    # Save used countries in a vector
    saveRDS(joined_df$iso3c,"../data/model_countries")

    # Design matrix — z-scored predictors
    X <- joined_df %>%
      select(all_of(predictors)) %>%
      mutate(across(everything(), ~ scale(.x)[, 1])) %>%
      as.matrix()

    # Response — z-scored
    y <- scale(joined_df %>% pull(2))[, 1]

    # Hierarchical grouping (realm index)
    province <- as.integer(factor(joined_df$realm))
    J        <- length(unique(province))

    # Create output directory if it doesn't already exist
    if (!dir.exists(base_path)) {
      dir.create(base_path, recursive = TRUE, showWarnings = FALSE)
    }

    # Save design matrix, response, and realm vector
    saveRDS(X, file.path(base_path, paste0("X_", response_vars[i], ".rds")))
    saveRDS(y, file.path(base_path, paste0("y_", response_vars[i], ".rds")))
    saveRDS(
      joined_df$realm,
      file.path(base_path, paste0("realms_", response_vars[i], ".rds"))
    )

    # Stan data
    stan_data <- list(
      N        = nrow(X),
      K        = ncol(X),
      J        = J,
      X        = X,
      y        = y,
      province = province
    )

    # Initial values matching Stan model parameters
    inits <- function() {
      list(
        beta_0       = 0,
        sigma_beta_0 = 0.5,
        z_beta_0     = rnorm(J, 0, 0.1),
        beta         = rep(0, ncol(X)),
        sigma        = 1
      )
    }

    # Fit Stan model
    fit <- stan(
      file   = "../utils/social_bayesian_model_hierarchical.stan",
      data   = stan_data,
      init   = inits,
      chains = 3,
      iter   = 4000,
      warmup = 1000
    )

    # Save fitted model
    saveRDS(fit, file.path(base_path, paste0("fit_", response_vars[i], ".rds")))

    message("Completed: ", response_vars[i])
  }

}
```

## Figure 4: Posterior distributions for the partial slope across models

Here, we map the posterior distributions for models of terrestrial animal consumption per capita, aquatic animal consumption per capita, capture consumption per capita, aquaculture consumption per capita, domestic consumption per capita, foreign consumption per capita, marine consumption per capita, inland consumption per capita, and aggregate aquatic animal reliance. Posterior distributions are displayed with 90%, 80%, and 50% credibility intervals, represented by the lightest, darker, and darkest colors respectively. The point represents the posterior median.

```{r hierarcical partial slope posteriors}
model_type = "realm"

response_vars <- readRDS("../data/response_vars.rds")

# Create a data frame of posteriors of partial slopes across Bayesian models
slopes_df <- data.frame()

for (i in seq_along(response_vars)) {

  # Extract MCMC samples
  x <- rstan::extract(
    readRDS(file.path(base_path, paste0("fit_", response_vars[i], ".rds"))),
    permuted = TRUE
  )
  beta_global <- x$beta
  X           <- readRDS(file.path(base_path, paste0("X_", response_vars[i], ".rds")))

  n_iter <- nrow(beta_global)
  K      <- ncol(beta_global)
  vars   <- colnames(X)

  # Extract global slopes and calculate credibility intervals
  ci_df <- expand_grid(iter = 1:n_iter, k = 1:K) %>%
    mutate(
      value = map2_dbl(iter, k, ~ beta_global[.x, .y]),
      var   = vars[k]
    ) %>%
    group_by(var) %>%
    summarize(
      lo95 = quantile(value, 0.025),
      lo80 = quantile(value, 0.10),
      lo50 = quantile(value, 0.25),
      med  = quantile(value, 0.50),
      hi50 = quantile(value, 0.75),
      hi80 = quantile(value, 0.90),
      hi95 = quantile(value, 0.975),
      .groups = "drop"
    ) %>%
    mutate(model = response_vars[i])

  slopes_df <- bind_rows(slopes_df, ci_df)
}

# Clean variable names for axis labels
var_labels <- c(
  coastline_km               = "Coastline length (km)",
  gdp_trade                  = "GDP from trade (%)",
  inundation_area_log1000km2 = "Inundated area (log 1000 km²)",
  log_gdp                    = "Log GDP per capita",
  pop_10k_pct_coast          = "Pop. within 10 km of coast (%)",
  supermarkets               = "Supermarkets per 100,000",
  eez_total                  = "Log EEZ area km²"
)

# Model colours — harmonized palette
model_colors <- c(
  "Terrestrial animal consumption per capita" = "#8C1A11",
  "Aquatic animal consumption per capita"     = "#2B6AA8",
  "Aquatic reliance"                          = "#1A7A6E",
  "Capture consumption per capita"            = "#D4601A",
  "Aquaculture consumption per capita"        = "#E8B000",
  "Domestic consumption per capita"           = "#7BAEC4",
  "Foreign consumption per capita"            = "#C45C3A",
  "Marine consumption per capita"             = "#9B6DC5",
  "Inland consumption per capita"             = "#D4956A"
)

# Interval colours — three lightness steps per model colour
col_95 <- colorspace::lighten(model_colors, 0.65)
col_80 <- colorspace::lighten(model_colors, 0.35)
col_50 <- colorspace::lighten(model_colors, 0.05)

# Load script to convert model name strings to readable labels
source("../utils/clean_model_name.R")

# Build plotting data
df_plot <- slopes_df %>%
  mutate(
    clean_name = clean_model_name(model),
    nonzero    = (lo95 < 0 & hi95 < 0) | (lo95 > 0 & hi95 > 0),
    suffix     = sub(".*_", "", model),
    suffix_num = suppressWarnings(as.numeric(suffix))
  ) %>%
  filter(
    str_detect(clean_name, "capita") | str_detect(clean_name, "Aquatic animal protein reliance"),
    !model == "animal_cons_cap"
  ) %>%
  arrange(suffix_num, suffix) %>%
  mutate(
    clean_name = factor(clean_name, levels = unique(clean_name)),
    col95 = coalesce(col_95[as.character(clean_name)], "grey80"),
    col80 = coalesce(col_80[as.character(clean_name)], "grey60"),
    col50 = coalesce(col_50[as.character(clean_name)], "grey40"),
    panel_id = as.numeric(factor(clean_name))
  )

# Shared symmetric x-axis limits so zero lines align across all panels
x_limit <- max(abs(c(df_plot$lo95, df_plot$hi95)), na.rm = TRUE) * 1.05
x_limit <- ceiling(x_limit * 10) / 10

panel_names <- levels(df_plot$clean_name)

# Function to draw a single credible interval panel
# left_col = TRUE shifts the tag inward to clear the y-axis label text
make_panel <- function(df, show_yaxis = FALSE, title = "", tag = "",
                       left_col = FALSE) {

  df_sig  <- df %>% filter(nonzero)
  df_nsig <- df %>% filter(!nonzero)

  # Left-column panels have wide y-axis text; nudge the tag right so it
  # doesn't sit directly on top of the longest label
  tag_pos <- if (left_col) c(0.52, 1) else c(0.01, 1)

  ggplot(df, aes(y = var)) +
        geom_vline(xintercept = 0, linetype = "dashed",
               linewidth = 0.3, color = "black") +
    geom_segment(aes(x = lo95, xend = hi95, color = col95),
                 linewidth = 2.2, lineend = "butt", alpha = 1) +
    geom_segment(aes(x = lo80, xend = hi80, color = col80),
                 linewidth = 2.2, lineend = "butt", alpha = 1) +
    geom_segment(aes(x = lo50, xend = hi50, color = col50),
                 linewidth = 2.2, lineend = "butt", alpha = 1) +
    geom_point(aes(x = med), shape = 21, size = 1.6,
               fill = "white", color = "grey20", stroke = 0.4) +
    scale_color_identity() +
    scale_x_continuous(limits = c(-x_limit, x_limit)) +
    scale_y_discrete(labels = var_labels) +
    labs(title = title, tag = tag) +
    theme_minimal(base_size = 8) +
    theme(
      plot.title         = element_text(size = 3.5, face = "bold", hjust = 0.5,
                                        margin = margin(b = 2), family = family),
      plot.tag           = element_text(size = 7, face = "bold", family = family),
      plot.tag.position  = tag_pos,
      panel.grid.major.y = element_blank(),
      panel.grid.minor   = element_blank(),
      panel.grid.major.x = element_line(
        color = scales::alpha("black", 0.15), linewidth = 0.15
      ),
      axis.text.x        = element_text(size = 5.5, color = "black", family = family),
      axis.title         = element_blank(),
      plot.margin        = margin(4, 6, 4, 6),
      axis.text.y        = if (show_yaxis) element_text(size = 5.5, color = "black", family = family,
                                                         hjust = 1) else element_blank(),
      axis.ticks.y       = element_blank(),
      axis.ticks.x = element_line(linewidth = 0.15, color = scales::alpha("black", 0.15))
    )
}

# Build all nine panels — panels 1, 4, 7 are left column (show y-axis + shifted tag)
panels <- lapply(seq_along(panel_names), function(i) {
  name     <- panel_names[i]
  df_sub   <- df_plot %>% filter(clean_name == name)
  show_y   <- i %in% c(1, 4, 7)
  left_col <- i %in% c(1, 4, 7)
  make_panel(df_sub,
             show_yaxis = show_y,
             left_col   = left_col,
             title      = name)
})

# Centered x-axis label placed below the grid
xlab_plot <- ggplot() +
  theme_void() +
  annotate("text", x = 0.5, y = 0.5,
           label    = "Posterior slope estimate",
           size     = 2.4,
           fontface = "bold",
           hjust    = 0.5,
           vjust    = 0.5, 
           family = family)


# Create custom legend for plot
legend_bayesian_plot <- data.frame(group = c("ci50", "ci80", "ci95"), type = c("med", "med", "med"), x = c(1,1,1), y = c(2,2,2)) %>%
  ggplot(aes(x = x, y = y, fill = group)) +
  geom_col(aes()) +
  geom_point(aes(color = type), shape = 21,
             fill = "grey20", size = 1.2, stroke = 0.3) +
  theme_minimal() +
    scale_color_manual(values = "grey20",
                     name = "",
                     label = "posterior median",
                     guide = guide_legend(
                       order = 2,
                       override.aes = list(
                        fill = list(pattern(
                                      grobTree(
                                        circleGrob(x = 0.5, 
                                                   y = 0.5, 
                                                   r = 0.67,
                                                   gp = gpar(fill = "white"))
                                        ),
                                        extend = "repeat",
                                        gp = gpar(col = "grey20", fill = "grey20",
                                                  lwd = 5)

                                      )
                                    ),
                        color = "grey20",
                        linewidth = 0.2,
                        size = 3.5
                        
                      ))) +
  theme(legend.position       = "bottom",
      legend.title            = element_text(size = 7, face = "bold", family = family),
      legend.text             = element_text(size = 6, family = family),
      legend.key.size         = unit(0.3, "cm"),
      legend.box              = "vertical",
      legend.spacing          = unit(0.1, "cm"),
      legend.margin           = margin(t = 1, b = 0),
      legend.box.margin       = margin(t = -2, b = -2),
      legend.background       = element_rect(fill = "transparent", color = NA),
      legend.box.background   = element_rect(fill = "transparent", color = NA)) +
  # labs(fill = "Credibility intervals") +
  scale_fill_manual(labels = c("50%", "80%", "95%"),
                    name = "Bayesian credibility intervals",
                    values = c(scales::alpha("black", 0.05),
                                    scales::alpha("black", 0.35),
                                    scales::alpha("black", 0.65)),
                    guide = guide_legend(
                      order = 1,
                      override.aes = list(
                        fill = list(scales::alpha("black", 0.15),
                                    scales::alpha("black", 0.35),
                                    scales::alpha("black", 0.65))
                    )))

# Extract colorbar legend from scatter before stripping it
legend_bayesian <- cowplot::get_plot_component(
  legend_bayesian_plot, "guide-box", return_all = TRUE
)
legend_bayesian <- Filter(function(x) !inherits(x, "zeroGrob"), legend_bayesian)[[1]]

# Assemble 3×3 grid
grid_plot <- (
    (panels[[1]] | panels[[2]] | panels[[3]]) /
    (panels[[4]] | panels[[5]] | panels[[6]]) /
    (panels[[7]] | panels[[8]] | panels[[9]])
  ) +
  plot_layout(heights = c(1, 1, 1)) &
  theme(plot.margin = margin(3, 4, 3, 4))

# Stack grid above centered x-axis label
plot <- cowplot::plot_grid(
  grid_plot,
  xlab_plot,
  ncol        = 1,
  rel_heights = c(1, 0.07)
) %>%
  # Supress warning - this warning only has to do with fonts, no data
  # or visualization is harmed via this warning
  suppressWarnings()

# add on legend
final_plot_w_legend <- cowplot::plot_grid(
  plot,
  legend_bayesian,
  ncol        = 1,
  rel_heights = c(1, 0.1)
) %>%
  # Supress warning - this warning only has to do with fonts, no data
  # or visualization is harmed via this warning
  suppressWarnings()

# Print final plot
print(final_plot_w_legend)

ggsave(
  filename = "../images/posteriors_partial_slopes_average.pdf",
  plot     = final_plot_w_legend,
  device   = cairo_pdf,
  width    = 5,
  height   = 4,
  units    = "in"
)
```
