---
title: "02_results"
author: "Connor Quiroz and Jessica Gephart"
date: "2025-09-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(exploreARTIS)
library(cowplot)
library(ggrepel)
library(countrycode)
library(colorspace)
library(ggpubr)
library(arrow)
library(rstan)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggridges)
library(bayesplot)
library(purrr)
library(patchwork)
```

# Chapter 1 Results

## Read in data

### Read in aquatic reliance data & FAO population data

```{r read in FAO derived data}
# Read in population
population_fao_wb <- read.csv("../output/population_fao_wb.csv")

# Read in aquatic reliance data
aquatic_reliance <- read.csv("../output/fao_aquatic_reliance_source.csv") %>%
  left_join(population_fao_wb,by=c("year","iso3c"))%>%
  mutate(protein_consumed_kg_percap=protein_consumed_t*1000/population_fao)

# Read in FAO sourcing data
aquatic_fbs <- read.csv("../output/fao_fbs_aquatic_source.csv")
```

### Load in biophysical and socio-economic data as covariates into the Bayesian model

```{r load data}
# Read in adaptive capacity data
ac_df <- read_parquet("../data/e_s_ac_data_total.parquet") %>%
  filter(scenario == "ssp126") %>%
  select(-aa_reliance_pct, -scenario, -change_in_stock,-live_weight_t,-pct_change) %>%
  rename(iso3c = "consumer_iso3c")
  
# Read in ecoregion data (all shapes associated with soverign)
ecoregion_full <- read.csv("../data/Spatially joined ecoregion to EEZ/ecoregion.csv") %>%
  rename("iso3c" = "ISO_SOV1",
         ecoregion = "ECOREGION",
         province = "PROVINCE",
         realm = "REALM",
         shape_area = "Shape_Area")

# Read in ecoregions for largest soverign shape
ecoregion_unique <- read.csv("../data/Spatially joined ecoregion to EEZ/ecoregion_clean.csv") %>%
  select(-FREQUENCY) %>%
  rename(iso3c = "ISO_SOV1",
         shape_area = "MAX_Shape_Area")

# Read in consumption data
# consumption <- read_parquet("C:/Users/cjqui/Desktop/SGL/Chapter 2 (trade climate risk)/trade-climate-risk/data/example_consumption_eez_2024_12_06.parquet") %>%
#   filter(year == 2019)

# Read in dietary sourcing flexibility index
dsfi_df <- read_csv("../data/covariates/food-systems-dashboard-2026-01-20 (dietary sourcing flexibility index).csv") %>%
  filter(`Start Year` == 2019 | `End Year` == 2019) %>%
  select(c("iso3c" = ISO3, "dsfi" = Value))

# Read in national indicator data
indicator_df <- read_csv("../data/covariates/all_national_indicators.csv")

# Read in coastline data
coastline_df <- read_csv("../data/covariates/countries-by-coastline-2026.csv") %>%
  mutate(iso3c = case_when(!is.na(flagCode) ~ countrycode(flagCode, origin = "iso2c", destination = "iso3c"),
                           country == "Namibia" ~ "NAM")) %>%
  select(c(iso3c, "coastline_km" = "CountryCoastlineKm"))

# Read in proportion of people within 10k of the coast
pop_10k_coast <- read.csv("../data/covariates/pop_10k_coast.csv")

```

### Clean data

```{r clean  data}
# Join largest shape for the respective soverign country to their ecoregion names
ecoregion <- left_join(ecoregion_unique, ecoregion_full, by = c("iso3c", "shape_area"))

# Assign countries not associated with a marine ecoregion to "landlocked" code
landlocked_countries <- aquatic_reliance %>%
  distinct(iso3c) %>%
  drop_na() %>%
  add_region(col = "iso3c", region.col.name = "region") %>%
  filter(!iso3c %in% ecoregion$iso3c) %>%
  # mutate(ecoregion = "Landlocked",
  #        province = "Landlocked",
  #        realm = "Landlocked") %>%
  mutate(ecoregion = paste0("Landlocked, ", region),
         province = paste0("Landlocked, ", region),
         realm = paste0("Landlocked, ", region)) %>%
  select(-region)

# Clean ecoregion data
ecoregion <- bind_rows(ecoregion, landlocked_countries) %>%
  arrange(iso3c) %>%
  select(-shape_area)

# Clean up files
rm(list = c("ecoregion_full", "ecoregion_unique"))


write_csv(ecoregion, "../output/ecoregion_clean.csv")

# Calculate total import weight for 2019
# imports_df <- consumption %>%
#   filter(year == 2019,
#          producer_iso3c != consumer_iso3c) %>%
#   group_by(consumer_iso3c) %>%
#   summarize(consumption_t = sum(consumption_t))
# 
# rm(consumption)
# gc()
```

### Get list of response variables to put through Bayesian model

```{r acquire animal consumption data to be fed through analysis}
# Calculate animal consumption
terr_animal_cons_cap <- aquatic_reliance %>%
  filter(year == 2019) %>%
  group_by(iso3c) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))

animal_cons_cap <- aquatic_reliance %>%
  filter(year == 2019) %>%
  group_by(iso3c) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))

# Calculate aquatic animal consumption
aquatic_animal_cons_cap <- aquatic_reliance %>%
  filter(year == 2019, food_group == "aquatic") %>%
  group_by(iso3c) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))

# Calculate animal reliance consumption
aquatic_rel <- aquatic_animal_cons_cap %>%
  left_join(animal_cons_cap, by = "iso3c") %>%
  group_by(iso3c) %>%
  summarize(prop_animal_protein = protein_consumed_kg_percap.x / protein_consumed_kg_percap.y)

# Add items not created in the loop to the list
response_vars <- c("terr_animal_cons_cap","animal_cons_cap", "aquatic_animal_cons_cap", "aquatic_rel", "aquatic_protein_roc", "terrestrial_protein_roc")

for (i in c("capture", "aquaculture", "domestic", "foreign", "marine", "inland")) {
  
  # Calculate consumption per capita by method
  if (i %in% c("capture", "aquaculture")) {
    assign(
      paste0(i, "_cap"),
      aquatic_reliance %>%
        filter(year == 2019, food_group == "aquatic", method == i) %>%
        group_by(iso3c) %>%
        summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))
    )
  }
  
  # Calculate consumption per capita by consumption source
  if (i %in% c("domestic", "foreign")) {
    assign(
      paste0(i, "_cap"),
      aquatic_reliance %>%
        filter(year == 2019, food_group == "aquatic", consumption_source == i) %>%
        group_by(iso3c) %>%
        summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))
    )
  }
  
  # Calculate consumption per capita by habitat
  if (i %in% c("marine", "inland")) {
    assign(
      paste0(i, "_cap"),
      aquatic_reliance %>%
        filter(year == 2019, food_group == "aquatic", habitat == i) %>%
        group_by(iso3c) %>%
        summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))
    )
  }
  
  response_vars <- c(response_vars, paste0(i, "_cap"))
  
  # Calculate reliance by method, production source, or habitat
  assign(
    paste0(i, "_rel"),
    get(paste0(i, "_cap")) %>%
      left_join(animal_cons_cap, by = "iso3c") %>%
      group_by(iso3c) %>%
      summarize(prop_animal_protein = protein_consumed_kg_percap.x / protein_consumed_kg_percap.y)
  )
  
  response_vars <- c(response_vars, paste0(i, "_rel"))
  
  # Calculate production sources' proportion of total aquatic animal consumption per capita
  assign(
    paste0(i, "_prop_of_aquatic_cons"),
    get(paste0(i, "_cap")) %>%
      left_join(aquatic_animal_cons_cap, by = "iso3c") %>%
      group_by(iso3c) %>%
      summarize(aquatic_source_prop = protein_consumed_kg_percap.x / protein_consumed_kg_percap.y)
  )
  
  response_vars <- c(response_vars, paste0(i, "_prop_of_aquatic_cons"))
}

#calculate rate of change in aquatic protein quantity
aquatic_protein_roc <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, year, method) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap, na.rm = TRUE)) %>%
  distinct() %>%
  filter(method == "capture") %>%
  split(.$iso3c) %>%
  map(~lm(protein_consumed_kg_percap~year, data = .x)) %>%
  map_df(broom::tidy, .id = 'iso3c') %>%
  filter(term == 'year') %>%
  select(c("iso3c", "aquatic_estimate" = "estimate"))

#calculate rate of change in terrestrial protein quantity
terrestrial_protein_roc <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "terrestrial") %>%
  group_by(iso3c, year) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap, na.rm = TRUE)) %>%
  distinct() %>%
  split(.$iso3c) %>%
  map(~lm(protein_consumed_kg_percap~year, data = .x)) %>%
  map_df(broom::tidy, .id = 'iso3c') %>%
  filter(term == 'year') %>%
  select(c("iso3c", "terrestrial_estimate" = "estimate"))

# Save list of response variables to a .RDS file
saveRDS(response_vars, "../output/response_vars.rds")
```


### Simple Bayesian Model

```{r simple bayesian model}
for (i in seq_along(response_vars)) {
joined_df <- get(response_vars[i]) %>%
  left_join(ac_df, by = "iso3c") %>%
  left_join(ecoregion, by = "iso3c") %>%
  ungroup() %>%
  add_region(col = "iso3c", region.col.name = "region") %>%
  left_join(dsfi_df, by = "iso3c") %>%
  # left_join(imports_df %>% rename(iso3c = consumer_iso3c), by = "iso3c") %>%
  # mutate(log_gdp = log(gdp),
  #        log_imports = log(consumption_t)) %>%
  mutate(log_gdp = log(gdp)) %>%
  left_join(indicator_df %>%
  select("eez_total", "iso3c", "inundation_area_log1000km2" = "inland_water_max"), by = "iso3c") %>%
  left_join(coastline_df, by = "iso3c")

df <- joined_df %>%
  select(
    names(.)[2], log_gdp, gdp_trade, eez_total,
    inundation_area_log1000km2,
    coastline_km, realm
  ) %>%
    mutate( # Impute NA values with 0
      coastline_km = case_when(
      is.na(coastline_km) & str_detect(realm, "Landlocked") ~ 0,
      TRUE ~ coastline_km
    ),
    
    eez_total = case_when(
      is.na(eez_total) & str_detect(realm, "Landlocked") ~ 0,
      TRUE ~ eez_total
    )
    ) %>% 
  drop_na() %>%
  filter(realm != "") %>%
    mutate(
    across(
      c(log_gdp, gdp_trade, eez_total,
        inundation_area_log1000km2, coastline_km),
      ~ (.x - min(.x, na.rm = TRUE)) /
        (max(.x, na.rm = TRUE) - min(.x, na.rm = TRUE)),
      .names = "{.col}"
    ))

# Separate explanatory variables (no realm, no random effect)
X <- df %>%
  select(-names(.)[1], -realm) %>%
  scale()

# Response
y <- scale(df %>% pull(1))[, 1]

N <- nrow(X)
K <- ncol(X)

# Save X and y (for diagnostic statistics to be used later on)
  saveRDS(X, paste0("../output/bayesian_output/", model_type, "/X_",response_vars[i],".rds"))
  saveRDS(y, paste0("../output/bayesian_output/", model_type, "/y_",response_vars[i],".rds"))

inits <- function() {
  list(
    beta_0 = 0,
    sigma = 1,
    beta = rep(0, K)
  )
}

stan_data <- list(
  N = N,
  K = K,
  X = as.matrix(X),
  y = y
)

assign(
  paste0("fit_", response_vars[i]),
  stan(
    file = "../R/social_bayesian_model_simple.stan",
    data = stan_data,
    init = inits,
    chains = 3,
    iter = 10000,
    warmup = 2000
  )
)

saveRDS(
  get(paste0("fit_", response_vars[i])),
  paste0("../output/bayesian_output/simple/fit_", response_vars[i], ".stan")
)

}
```

### Hierarchical Bayesian model

```{r test before using}
# Variable list with EEZ area instead of coastline (km)
# predictors <- c(
#   "log_gdp",
#   "gdp_trade",
#   "eez_total",
#   "inundation_area_log1000km2",
#   "supermarkets",
#   "pop_10k_pct_coast",
# )

# Variable list with coastline (km) instead of EEZ area
predictors <- c(
  "log_gdp",
  "gdp_trade",
  "coastline_km",
  "inundation_area_log1000km2",
  "supermarkets",
  "pop_10k_pct_coast"
)

model_type = "realm"


for (i in seq_along(response_vars)) {

  joined_df <- get(response_vars[i]) %>%
    left_join(ac_df, by = "iso3c") %>%
    left_join(ecoregion, by = "iso3c") %>%
    ungroup() %>%
    # add_region(col = "iso3c", region.col.name = "realm") %>%   # keep realm
    left_join(
      indicator_df %>%
        select("eez_total", "iso3c",
               "inundation_area_log1000km2" = "inland_water_max"),
      by = "iso3c"
    ) %>%
    left_join(coastline_df, by = "iso3c") %>%
    left_join(pop_10k_coast, by = "iso3c") %>%
    mutate(
      log_gdp = log(gdp),
      coastline_km = case_when(
        is.na(coastline_km) & str_detect(realm, "Landlocked") ~ 0,
        TRUE ~ coastline_km
      ),
      eez_total = case_when(
        is.na(eez_total) & str_detect(realm, "Landlocked") ~ 0,
        TRUE ~ eez_total
      )
    ) %>%
    drop_na() %>%
    filter(realm != "")

  # Design matrix
  X <- joined_df %>%
    select(all_of(predictors)) %>%
    mutate(across(everything(), ~ (.x - min(.x)) / (max(.x) - min(.x)))) %>%
    as.matrix()

  # Response
  y <- scale(joined_df %>% pull(2))[, 1]

  # Hierarchical grouping (province index from realm)
  province <- as.integer(factor(joined_df$realm))
  J <- length(unique(province))

  # Save X and y
  saveRDS(X, paste0("../output/bayesian_output/", model_type, "/coastline_km_covariate/X_", response_vars[i], ".rds"))
  saveRDS(y, paste0("../output/bayesian_output/", model_type, "/coastline_km_covariate/y_", response_vars[i], ".rds"))
  
  saveRDS(joined_df$realm,
        paste0("../output/bayesian_output/", model_type, "/coastline_km_covariate/realms_", response_vars[i], ".rds"))

  # Stan data
  stan_data <- list(
    N = nrow(X),
    K = ncol(X),
    J = J,
    X = X,
    y = y,
    province = province
  )

  # Inits matching your Stan model
  inits <- function() {
    list(
      beta_0       = 0,
      sigma_beta_0 = 0.5,
      z_beta_0     = rnorm(J, 0, 0.1),
      beta         = rep(0, ncol(X)),
      sigma        = 1
    )
  }

  fit <- stan(
    file = "../R/social_bayesian_model_hierarchical.stan",
    data = stan_data,
    init = inits,
    chains = 3,
    iter = 4000,
    warmup = 1000
  )

  # saveRDS(fit, paste0("../output/bayesian_output/fit_", response_vars[i], ".stan"))
  
  saveRDS(fit, paste0("../output/bayesian_output/", model_type, "/coastline_km_covariate/fit_", response_vars[i], ".rds"))
}
```


```{r run model for non landlocked countries}
model_type <- "realm"

for (i in seq_along(response_vars)) {
  # Join target response to its explanatory variables
  joined_df <- get(response_vars[i]) %>%
  left_join(ac_df, by = "iso3c") %>%
  left_join(ecoregion, by = "iso3c") %>%
  ungroup() %>%
  add_region(col = "iso3c", region.col.name = "region") %>%
  left_join(dsfi_df, by = "iso3c") %>%
  # left_join(imports_df %>% rename(iso3c = consumer_iso3c), by = "iso3c") %>%
  mutate(log_gdp = log(gdp),
         log_imports = log(consumption_t)) %>%
  mutate(log_gdp = log(gdp)) %>%
  left_join(indicator_df %>%
  select("eez_total", "iso3c", "inundation_area_log1000km2" = "inland_water_max"), by = "iso3c") %>%
  left_join(coastline_df, by = "iso3c") 
  
  # Clean joined data to be used in Bayesian model
  
  if (model_type == "continental") {
    joined_df <- joined_df %>%
    select(-realm) %>%
    rename(realm = region)
  }
  
  df <- joined_df %>%
  select(
    names(.)[2], log_gdp, gdp_trade, eez_total,
    inundation_area_log1000km2,
    coastline_km, realm
  ) %>%
    mutate( # Impute NA values with 0
      coastline_km = case_when(
      is.na(coastline_km) & str_detect(realm, "Landlocked") ~ 0,
      TRUE ~ coastline_km
    ),
    
    eez_total = case_when(
      is.na(eez_total) & str_detect(realm, "Landlocked") ~ 0,
      TRUE ~ eez_total
    )
    ) %>% 
  drop_na() %>%
  filter(realm != "") %>%
    mutate(
    across(
      c(log_gdp, gdp_trade, eez_total,
        inundation_area_log1000km2, coastline_km),
      ~ (.x - min(.x, na.rm = TRUE)) /
        (max(.x, na.rm = TRUE) - min(.x, na.rm = TRUE)),
      .names = "{.col}"
    ))


 # Separate explanatory variables (not including the factor level)
  X <- df %>%
  select(-names(.)[1], -realm) %>%
             scale()

 # Separate response variable
  y_raw <- df %>% pull(1)
  y <- scale(y_raw)[,1]

  # Create province index
  realm_ids <- as.numeric(factor(df$realm))
  n_realms <- length(unique(realm_ids))
  
  # Save Realm IDs (to be used in exploratory plots)
  realm_levels <- levels(factor(df$realm))
  saveRDS(
  realm_levels,
  paste0("../output/bayesian_output/", model_type, "/realms_", response_vars[i], ".rds")
)
  
  # Save Realm IDs
  saveRDS(
  realm_ids,
  paste0("../output/bayesian_output/", model_type, "/ids_", response_vars[i], ".rds")
)
  
  
  # Save X and y (for diagnostic statistics to be used later on)
  saveRDS(X, paste0("../output/bayesian_output/", model_type, "/X_",response_vars[i],".rds"))
  saveRDS(y, paste0("../output/bayesian_output/", model_type, "/y_",response_vars[i],".rds"))
 
  
  N <- nrow(X)
  K <- ncol(X)
  
  inits <- function() {
    list(
      beta_0       = 0,
      sigma_beta_0 = 1,
      z_beta_0     = rnorm(n_realms, 0, 0.1),
      beta         = rep(0, K),
      sigma        = 1
    )
  }
  
  stan_data <- list(
    N = N,
    K = K,
    J = n_realms,  # number of provinces
    X = as.matrix(X),
    y = y,
    province = realm_ids  # province indicator for each observation
  )
  
  # Fit Bayesian model
  assign(
    paste0("fit_",response_vars[i]),
    stan(
    file = "../R/social_bayesian_model_hierarchical.stan",
    data = stan_data,
    init = inits,
    chains = 3,
    iter = 4000,
    warmup = 1000,
    control = list(adapt_delta = 0.95, max_treedepth = 12)
  )
  )
  
  
  # Save Bayesian model
  saveRDS(
    get(paste0("fit_", response_vars[i])),
    paste0("../output/bayesian_output/", model_type, "/fit_", response_vars[i],".stan")
  )
}


continental_levels <- c("Africa", "Asia", "Europe", "North America", "Oceania", "South America")
realm_levels <- c("Arctic", "Central Indo-Pacific", "Eastern Indo-Pacific", "Landlocked, Africa", "Landlocked, Asia", "Landlocked, Europe", "Landlocked, South America", "Southern Ocean", "Temperate Australasia", "Temperate Northern Atlantic", "Temperate Northern Pacific", "Temperate South America", "Temperate Southern Africa", "Tropical Atlantic", "Tropical Eastern Pacific", "Western Indo-Pacific")
```


## RQ 2: What biophysical and socioeconomic factors are associated with reliance on farmed and wild aquatic animals for protein and overall consumption?

### Global slope posteriors

```{r hierarcical global variable posteriors}
model_type = "realm"

# Create data frame to hold all posteriors across variables
slopes_df <- data.frame()

for (i in seq_along(response_vars)) {
  
  # Extract MCMC samples
  x <- rstan::extract(readRDS(paste0("../output/bayesian_output/", model_type, "/fit_", response_vars[i],".stan")), permuted = T)
  beta_global <- x$beta       # iter × K (slopes)
  
  X <- readRDS(paste0("../output/bayesian_output/", model_type, "/X_", response_vars[i],".rds"))
  
  n_iter <- nrow(beta_global)
  K <- ncol(beta_global)
  vars <- colnames(X)

# Extract global slopes
post_long_slopes <- expand_grid(
  iter = 1:n_iter,
  k = 1:K
) %>%
  mutate(
    value = map2_dbl(iter, k, ~beta_global[.x, .y]),
    var = vars[k]
  )

# Calculate credibility intervals
ci_df <- post_long_slopes %>%
  group_by(var) %>%
  summarise(
    lo95 = quantile(value, 0.025),
    lo80 = quantile(value,0.1),
    lo50 = quantile(value, 0.25),
    med = quantile(value, 0.5),
    hi50 = quantile(value, 0.75),
    hi80 = quantile(value,0.9),
    hi95 = quantile(value, 0.975),
    .groups = "drop"
  ) %>%
  mutate(model =  response_vars[i])

slopes_df <- slopes_df %>%
  bind_rows(ci_df)

}

var_labels <- c(
  coastline_km = "Coastline length (km)",
  gdp_trade = "GDP from trade (%)",
  inundation_area_log1000km2 = "Inundated area (log 1000 km²)",
  log_gdp = "GDP (log scale)",
  prop_pop_coast = "Population within 10 km of coast (%)",
  supermarkets = "Supermarkets per 100,000 people"
)

theme_tile_pub <- function() {
  theme_minimal(base_size = 8) +
    theme(
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.y = element_line(color = scales::alpha("black", 0.25), linewidth = 0.3),
      panel.grid.minor.y = element_blank(),
      panel.background = element_rect(fill = "transparent", color = NA),
      plot.background = element_rect(fill = "transparent", color = NA),
      axis.text.x = element_text(size = 6,angle = 45, vjust = 0.5, color = "black"),
      axis.text.y = element_text(size = 6, color = "black"),
      axis.title = element_text(size = 7, face = "bold"),
      legend.title = element_text(size = 7, face = "bold"),
      legend.text = element_text(size = 6),
      legend.key.size = unit(0.3, "cm"),
      legend.box = "vertical",
      legend.spacing = unit(0.1, "cm"),
      legend.margin = margin(t = 0, b = 0),
      legend.box.margin = margin(t = -2, b = -2),
      legend.background = element_rect(fill = "transparent", color = NA),
      legend.box.background = element_rect(fill = "transparent", color = NA),
      plot.title = element_text(size = 8, face = "bold", hjust = 0.5)
    )
}

library(colorspace)
library(dplyr)
library(ggplot2)
library(stringr)
library(colorspace)


# 1. Base colors per model
# 1. Base colors per model
model_colors <- c(
  "Terrestrial animal consumption consumption per capita" = "#b15928",
  "Aquatic animal consumption consumption per capita"     = "#1f78b4",
  "Aquatic reliance"                                      = "#e31a1c",
  "Capture consumption per capita"                        = "#ff7f00",
  "Aquaculture consumption per capita"                    = "#6a3d9a",
  "Domestic consumption per capita"                       = "#33a02c",
  "Foreign consumption per capita"                        = "#a6cee3",
  "Marine consumption per capita"                         = "#b2df8a",
  "Inland consumption per capita"                         = "#fb9a99"
)

col_95 <- lighten(model_colors, 0.65)
col_80 <- lighten(model_colors, 0.4)
col_50 <- lighten(model_colors, 0.10)

# Load in script to clean model names from underscores to be easily readable
source("../R/clean_model_name.R")

# 2. Add clean_name once
slopes_df <- slopes_df %>%
  mutate(clean_name = clean_model_name(model))

# 3. Build plotting data
df_plot <- slopes_df %>%
  mutate(
    clean_name = clean_model_name(model),
    nonzero = (lo95 < 0 & hi95 < 0) | (lo95 > 0 & hi95 > 0),
    suffix = sub(".*_", "", model),
    suffix_num = suppressWarnings(as.numeric(suffix))
  )  %>%
  filter(model != "animal_cons_cap") %>%
  mutate(
    nonzero    = (lo95 < 0 & hi95 < 0) | (lo95 > 0 & hi95 > 0),
    suffix     = sub(".*_", "", model),
    suffix_num = suppressWarnings(as.numeric(suffix))
  ) %>%
  filter(
    str_detect(clean_name, "capita") |
    str_detect(clean_name, "Aquatic reliance")
  ) %>%
  arrange(suffix_num, suffix) %>%
  mutate(
    clean_name = factor(clean_name, levels = unique(clean_name)),
    col95 = col_95[as.character(clean_name)],
    col80 = col_80[as.character(clean_name)],
    col50 = col_50[as.character(clean_name)],
    panel_id = as.numeric(factor(clean_name)),
    left_col = panel_id %% 3 == 1
  )

library(dplyr)
library(ggplot2)
library(patchwork)

panel_tags <- letters[1:9] 


# --- 1. Function to draw a single panel ---
make_panel <- function(df, show_yaxis = FALSE, tag = "") {
  ggplot(df, aes(y = var)) +
    geom_segment(aes(x = lo95, xend = hi95, color = col95),
                 linewidth = 2, lineend = "square") +
    geom_segment(aes(x = lo80, xend = hi80, color = col80),
                 linewidth = 2, lineend = "square") +
    geom_segment(aes(x = lo50, xend = hi50, color = col50),
                 linewidth = 2, lineend = "square") +
    geom_point(aes(x = med), size = 1.8, color = "black") +
    geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.4) +
    scale_color_identity() +
    scale_y_discrete(labels = var_labels) +

    # ← add tag in upper-left corner
    annotate("text", x = -Inf, y = Inf, label = tag,
             hjust = -0.3, vjust = 1.3, size = 3.2, fontface = "bold") +

    theme_minimal(base_size = 8) +
    theme(
      plot.title = element_blank(),
      panel.grid.major.y = element_blank(),
      panel.grid.minor   = element_blank(),
      strip.text         = element_text(size = 5, face = "bold"),
      axis.text.x        = element_text(size = 6, color = "black"),
      axis.title         = element_blank(),
      panel.spacing      = unit(1.2, "lines"),
      plot.margin        = margin(5, 5, 5, 5),
      axis.text.y        = if (show_yaxis) element_text(size = 6, color = "black") else element_blank(),
      axis.ticks.y       = if (show_yaxis) element_line() else element_blank()
    )
}

# --- 2. Split df_plot into 9 panels ---
panel_names <- levels(df_plot$clean_name)

panels <- lapply(seq_along(panel_names), function(i) {
  name <- panel_names[i]
  df_sub <- df_plot %>% filter(clean_name == name)

  show_y <- i %in% c(1, 4, 7)
  wrapped_name <- stringr::str_wrap(name, width = 20)

  make_panel(df_sub, show_yaxis = show_y, tag = panel_tags[i]) +
    ggtitle(wrapped_name)
})

library(ggplot2)
library(patchwork)

# Build the x-axis label as its own plot
xlab_plot <- ggplot() +
  theme_void() +
  annotate("text", x = 0.5, y = 0.5,
           label = "Posterior slope estimate",
           size = 2, fontface = "bold")

# Empty plot for left and right columns
empty <- ggplot() + theme_void()

# Assemble 3×3 grid + bottom row with centered x-label
final_plot_with_xlab <-
  (
    (panels[[1]] | panels[[2]] | panels[[3]]) /
    (panels[[4]] | panels[[5]] | panels[[6]]) /
    (panels[[7]] | panels[[8]] | panels[[9]]) /
    (empty | xlab_plot | empty)
  ) +
  plot_layout(heights = c(1, 1, 1, 0.12))

final_plot_with_xlab


ggsave(paste0("../images/bayesian plots/", model_type, "/posteriors_no_titles.jpg"), plot = final_plot_with_xlab, device = "jpg", width = 5, height = 3, units = "in", dpi = 300)
```

### Plot Maps of intercept median posteriors 
  
```{r realm_maps_robinson_final, message=FALSE, warning=FALSE}
model_type = "realm"
intecrept_df <- data.frame()
for (i in seq_along(response_vars)) {
  # Extract MCMC samples
  x <- rstan::extract(readRDS(paste0("../output/bayesian_output/", model_type, "/fit_", response_vars[i],".stan")), permuted = T)
  beta_global <- x$beta       # iter × K (slopes)
  
  X <- readRDS(paste0("../output/bayesian_output/", model_type, "/X_", response_vars[i],".rds"))
  
  
  
  n_iter <- nrow(beta_global)
  K <- ncol(beta_global)
  vars <- colnames(X)
  
  # Posteriors of the realm intercepts
  beta_0i <- x$beta_0i   # iterations × J

  n_iter <- nrow(beta_0i)
  J <- ncol(beta_0i)

  # Get corresponding realm indicators (e.g., Europe, Oceania) to match with intercept values
realm_level <- realm_levels <- readRDS(
  paste0("../output/bayesian_output/", model_type, "/realms_", response_vars[i], ".rds")
)
  
  
  post_intercepts <- expand_grid(
  iter = 1:n_iter,
  realm = 1:J
) %>%
  mutate(
    value = beta_0i[cbind(iter, realm)],
    realm = realm_level[realm]
  )

  ci_intercepts <- post_intercepts %>%
  group_by(realm) %>%
  summarise(
    lo95 = quantile(value, 0.025),
    lo50 = quantile(value, 0.25),
    med  = quantile(value, 0.5),
    hi50 = quantile(value, 0.75),
    hi95 = quantile(value, 0.975),
    .groups = "drop"
  ) %>%
  mutate(model = response_vars[i])
  
  intecrept_df <- intecrept_df %>%
    bind_rows(ci_intercepts)
}

library(sf)
library(lwgeom)
library(dplyr)
library(ggplot2)
library(stringr)
library(purrr)
library(rnaturalearth)
library(rnaturalearthdata)

#------------------------------------------------------------
# 1. Load MEOW ecoregions
#------------------------------------------------------------

meow <- st_read("../data/MEOW/meow_ecos.shp")

#------------------------------------------------------------
# 2. Realm levels
#------------------------------------------------------------

realm_levels <- c(
  "Arctic","Central Indo-Pacific","Eastern Indo-Pacific",
  "Landlocked, Africa","Landlocked, Asia","Landlocked, Europe",
  "Landlocked, South America","Southern Ocean","Temperate Australasia",
  "Temperate Northern Atlantic","Temperate Northern Pacific",
  "Temperate South America","Temperate Southern Africa",
  "Tropical Atlantic","Tropical Eastern Pacific","Western Indo-Pacific"
)

marine_realms <- realm_levels[!str_detect(realm_levels, "Landlocked")]
landlocked_realms <- realm_levels[str_detect(realm_levels, "Landlocked")]

#------------------------------------------------------------
# 3. Dissolve MEOW ecoregions into REALM polygons
#------------------------------------------------------------

meow_realms <- meow %>%
  group_by(REALM) %>%
  summarise(geometry = st_union(geometry), .groups = "drop") %>%
  rename(realm = REALM)

#------------------------------------------------------------
# 4. Split polygons at the antimeridian
#------------------------------------------------------------

meow_realms <- meow_realms %>%
  st_wrap_dateline(options = c("WRAPDATELINE=YES", "DATELINEOFFSET=180")) %>%
  st_collection_extract("POLYGON")

#------------------------------------------------------------
# 5. Build landlocked realms using CONTINENTS
#    but trim only the Russia sliver east of 180°
#------------------------------------------------------------

countries <- ne_countries(scale = "medium", returnclass = "sf")

# Trim Russia's far-east sliver
russia <- countries %>% filter(iso_a3 == "RUS")

russia_trimmed <- st_difference(
  russia,
  st_as_sfc(st_bbox(c(
    xmin = 180, ymin = -90, xmax = 200, ymax = 90
  ), crs = 4326))
)

countries_fixed <- countries %>%
  filter(iso_a3 != "RUS") %>%
  bind_rows(russia_trimmed)

# Clean continent basemap (for always-visible land)
continent_basemap <- countries_fixed %>%
  group_by(continent) %>%
  summarise(geometry = st_union(geometry), .groups = "drop") %>%
  st_make_valid() %>%
  st_wrap_dateline(options = c("WRAPDATELINE=YES", "DATELINEOFFSET=180")) %>%
  st_collection_extract("POLYGON") %>%
  st_transform(crs = robin)

# Build continent unions (now safe)
continents <- countries_fixed %>%
  mutate(continent = factor(continent)) %>%
  group_by(continent) %>%
  summarise(geometry = st_union(geometry), .groups = "drop")

# Map landlocked realms to continents
landlocked_lookup <- tibble(
  realm = landlocked_realms,
  continent = str_remove(landlocked_realms, "Landlocked, ")
)

landlocked_geom <- landlocked_lookup %>%
  left_join(continents, by = "continent") %>%
  select(realm, geometry)


#------------------------------------------------------------
# 6. Combine marine + landlocked geometries
#------------------------------------------------------------

realm_geom <- bind_rows(
  meow_realms %>% filter(realm %in% marine_realms),
  landlocked_geom
)

#------------------------------------------------------------
# 7. Project to Robinson
#------------------------------------------------------------

robin <- "+proj=robin +lon_0=0 +datum=WGS84 +units=m +no_defs"
realm_geom_robin <- st_transform(realm_geom, crs = robin)

#------------------------------------------------------------
# 8. Mapping function with significance outlines
#------------------------------------------------------------

selected_df <- intecrept_df %>%
  filter(model %in% c("terr_animal_cons_cap",
                      "aquatic_rel",
                      "aquaculture_rel",
                      "capture_rel"
                      ))

global_min <- min(selected_df$med, na.rm = TRUE)
global_max <- max(selected_df$med, na.rm = TRUE)

global_limit <- max(abs(global_min), abs(global_max))

make_realm_map <- function(model_name, df = intecrept_df) {

  # Filter model + significance flag
  df_model <- df %>%
    filter(model == model_name) %>%
    mutate(significant = (lo95 > 0 & hi95 > 0) | (lo95 < 0 & hi95 < 0))

  # Join model values to realm geometry
  geom_joined <- realm_geom_robin %>%
    left_join(df_model, by = "realm") %>%
    filter(!is.na(st_is_empty(geometry)) & !st_is_empty(geometry))

  ggplot() +
    # Basemap (always visible)
    geom_sf(
      data = continent_basemap,
      fill = "darkgray",
      color = "grey70",
      size = 0.2
    ) +

    # Realm shading (outline only for non-NA)
    geom_sf(
      data = geom_joined,
      aes(fill = med, color = !is.na(med)),
      size = 0.2
    ) +
    scale_color_manual(
      values = c("TRUE" = "grey30", "FALSE" = NA),
      guide = "none"
    ) +

    # Significance outlines
    geom_sf(
      data = geom_joined %>% filter(significant),
      fill = NA,
      color = "#1F1F1F",
      linewidth = 0.5
    ) +

    # Global symmetric fill scale (computed outside this function)
    scale_fill_gradient2(
      low = "#b2182b",
      mid = "grey85",
      high = "#2166ac",
      midpoint = 0,
      limits = c(-global_limit, global_limit),
      oob = scales::squish
    ) +

    # Hide legend inside each subplot
    theme_realm_panel() +
    theme(legend.position = "none") +

    labs(
      title = paste(model_name),
      fill = "Posterior\nmedian\nintercept"
    )
}
  

#------------------------------------------------------------
# 9. Produce maps for all models
#------------------------------------------------------------

all_models <- unique(intecrept_df$model)
realm_maps <- map(all_models, ~ make_realm_map(.x))
names(realm_maps) <- all_models

# Example:
# realm_maps[["animal_cons_cap"]]

theme_realm_panel <- function() {
  theme_minimal(base_size = 8) +
    theme(
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.y = element_line(color = scales::alpha("black", 0.25), linewidth = 0.3),
      panel.grid.minor.y = element_blank(),
      panel.background = element_rect(fill = "transparent", color = NA),
      plot.background = element_rect(fill = "transparent", color = NA),
      axis.text = element_text(size = 6, color = "black"),
      axis.title = element_blank(),
      legend.position = "bottom",
      legend.title = element_text(size = 7, face = "bold", hjust = 0.5),
      legend.text = element_text(size = 6),
      legend.key.size = unit(0.5, "cm"),
      legend.box = "vertical",
      legend.spacing = unit(0.1, "cm"),
      legend.margin = margin(t = 0, b = 0),
      legend.box.margin = margin(t = -2, b = -2),
      legend.background = element_rect(fill = "transparent", color = NA),
      legend.box.background = element_rect(fill = "transparent", color = NA)
    )
}

# p1 <- realm_maps[["terr_animal_cons_cap"]] +
#   labs(title = "Daily terrestrial animal consumption per capita") +
#   theme_realm_panel() +
#   theme(plot.title = element_text(hjust = 0.5, size = 8, face = "bold"))

p1 <- realm_maps[["animal_cons_cap"]] +
  labs(title = "Daily animal consumption per capita") +
  theme_realm_panel() +
  theme(plot.title = element_text(hjust = 0.5, size = 8, face = "bold"))

p2 <- realm_maps[["aquatic_rel"]] +
  labs(title = "Aquatic animal reliance per capita") +
  theme_realm_panel() +
  theme(plot.title = element_text(hjust = 0.5, size = 8, face = "bold"))
# Create combined panel plot
p1 <- realm_maps[["animal_cons_cap"]] +
  labs(title = "Daily animal protein consumption per capita") +
  theme_realm_panel() +
  theme(plot.title = element_text(hjust = 0.5, size = 8, face = "bold"))

p2 <- realm_maps[["aquatic_rel"]] +
  labs(title = "Aquatic animal reliance per capita") +
  theme_realm_panel() +
  theme(plot.title = element_text(hjust = 0.5, size = 8, face = "bold"))
# Create combined panel plot
(plot <- (p1 | p2) +
  plot_layout(guides = "collect") &
  theme(
    legend.position = "bottom",
    aspect.ratio = 0.45,   # ← try 0.40–0.55 depending on taste
    plot.margin = margin(0, 0, 0, 0) 
  ))

ggsave(paste0("../images/bayesian plots/", model_type, "/intercept_map.jpg"), plot = plot, device = "jpg", width = 5, height = 4, units = "in", dpi = 300)
```

## RQ 1a: How do aquatic food reliance trends disaggregated by source differ from aggregate aquatic food reliance?

### Country level trends in delta 2010-2020 aquatic animal reliance

```{r Country level trends in delta 2010-2020 aquatic animal reliance}
library(dplyr)
library(purrr)
library(broom)
library(ggplot2)
library(ggridges)
library(forcats)
library(stringr)

# --- 1. Disaggregated slopes (your existing pipeline) ---
disagg <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c) %>%
  mutate(n = n_distinct(year)) %>%
  filter(n >= 3) %>%
  group_by(iso3c, habitat, method) %>%
  filter(n() >= 3) %>%
  nest() %>%
  mutate(
    model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
    tidy  = map(model, broom::tidy)
  ) %>%
  unnest(tidy) %>%
  filter(term == "year") %>%
  mutate(
    consource_habitat_method = paste(habitat, method),
    sqrt_signed = sign(estimate) * sqrt(abs(estimate)),
    group_label = consource_habitat_method
  ) %>%
  drop_na(estimate) %>%
  filter(!str_detect(consource_habitat_method, "unknown")) %>%
  group_by(group_label) %>%
  mutate(pct_above = mean(estimate > 0) * 100) %>%
  ungroup()

# --- 2. Aggregate slopes (overall aquatic reliance per country) ---
agg <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c) %>%
  mutate(n = n_distinct(year)) %>%
  filter(n >= 3) %>% # Get countries with only 3 years or more of data (so lm can properly fit points)
  group_by(iso3c, year, data_source, food_group) %>%
  summarise(prop_animal_protein = sum(prop_animal_protein)) %>%
  group_by(iso3c) %>%
  nest() %>%
  mutate(
    model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
    tidy  = map(model, broom::tidy)
  ) %>%
  unnest(tidy) %>%
  filter(term == "year") %>%
  mutate(
    group_label = "aggregate"
  ) %>%
  drop_na(estimate) %>%
  group_by(group_label) %>%
  mutate(pct_above = mean(estimate > 0) * 100) %>%
  ungroup()

# --- 3. Combine disaggregated + aggregate ---
all_slopes <- bind_rows(agg, disagg)


all_slopes <- all_slopes %>%
  mutate(
    pct_below = 100 - pct_above,
    label_majority = if_else(
      pct_above >= 50,
      sprintf("%.0f%%", pct_above),
      sprintf("%.0f%%", pct_below)
    ),
    label_x = if_else(
      pct_above >= 50,
      0.002,   # right side, close to center
      -0.002   # left side, close to center
    ),
    label_hjust = if_else(pct_above >= 50, 0, 1),
    color = case_when(pct_above >= 50 ~ "#2166ac",
                      TRUE ~ "#b2182b")
  )


(plot <- ggplot(all_slopes, aes(
  x = estimate,
  y = fct_reorder(group_label, pct_above)
)) +
  geom_density_ridges_gradient(
    aes(fill = after_stat(x > 0)),
    scale = 1.4,
    rel_min_height = 0.01,
    color = "black"
  ) +
  scale_fill_manual(
    values = c(`TRUE` = "#2166ac", `FALSE` = "#b2182b")
  ) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_text(
    aes(x = label_x, label = label_majority, hjust = label_hjust),
    color = all_slopes$color,
    vjust = -2,
    size = 2.5
  ) +
  guides(fill = "none") +
  labs(x = expression(bold(Delta[2010-2020] ~ "Aquatic animal protein reliance %")), y = "Production source") +
  # xlim(c(-0.0105, 0.0105)) +

  # --- Matching theme from aa_reliance_multiple_trends ---
  theme_minimal(base_size = 8) +
  theme(
    # Grid and background
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(color = scales::alpha("black", 0.25), linewidth = 0.3),
    panel.grid.minor.y = element_blank(),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),

    # Axis styling
    axis.text = element_text(size = 6, color = "black"),
    axis.title.x = element_text(size = 7, face = "bold", margin = margin(t = 5)),
    axis.title.y = element_text(size = 7, face = "bold"),

    # Ridge spacing
    axis.text.y = element_text(size = 6),

    # No legend
    legend.position = "none"
  ))



# ggsave("../images/pct_ichange_reliance.jpg", plot = plot, device = "jpg", units = "in", width = 5, height = 4)

cor(joined_df$eez_total,joined_df$coastline_km)

library(plotly)

x <- joined_df %>%
  ggplot(aes(x = eez_total, y = coastline_km)) +
  geom_point()

ggplotly(x)
```

```{r}
library(dplyr)
library(purrr)
library(broom)
library(ggplot2)
library(ggtext)
library(forcats)
library(stringr)
disagg_years
agg_years

disagg_years <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, habitat, method) %>%
  summarise(years = list(sort(unique(year[!is.na(prop_animal_protein)]))))

agg_years <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c) %>%
  summarise(years = list(sort(unique(year))))

left_join(disagg_years, agg_years, by = "iso3c", suffix = c("_disagg", "_agg"))

# --- 1. Disaggregated slopes ---
disagg <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "aquatic") %>%
  # group_by(iso3c) %>%
  # mutate(n = n_distinct(year)) %>%
  # filter(n >= 3) %>%
  group_by(year, iso3c, habitat, method) %>%
  mutate(prop_animal_protein = sum(prop_animal_protein)) %>%
  group_by(iso3c, habitat, method) %>%
  # filter(n() >= 3) %>%
  nest() %>%
  mutate(
    model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
    tidy  = map(model, broom::tidy)
  ) %>%
  unnest(tidy) %>%
  filter(term == "year") %>%
  mutate(
    consource_habitat_method = paste(habitat, method),
    sqrt_signed = sign(estimate) * sqrt(abs(estimate)),
    group_label = consource_habitat_method
  ) %>%
  drop_na(estimate) %>%
  filter(!str_detect(consource_habitat_method, "unknown")) %>%
  group_by(group_label) %>%
  mutate(pct_above = mean(estimate > 0) * 100) %>%
  # mutate(estimate = round(estimate, 2)) %>%
  ungroup()

# --- 2. Aggregate slopes ---
agg <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "aquatic") %>%
  # group_by(iso3c) %>%
  # mutate(n = n_distinct(year)) %>%
  # filter(n >= 3) %>%
  group_by(iso3c, year) %>%
  summarise(prop_animal_protein = sum(prop_animal_protein), .groups = "drop") %>%
  group_by(iso3c) %>%
  nest() %>%
  mutate(
    model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
    tidy  = map(model, broom::tidy)
  ) %>%
  unnest(tidy) %>%
  filter(term == "year") %>%
  mutate(group_label = "aggregate") %>%
  drop_na(estimate) %>%
  group_by(group_label) %>%
  mutate(pct_above = mean(estimate > 0) * 100) %>%
  ungroup()

# --- 3. Build label lookup ONCE before plotting ---
label_lookup <- bind_rows(agg, disagg) %>%
  add_region("iso3c", "region") %>%
  mutate(
    size_group = case_when(
      region %in% c("South America", "Oceania") ~ 7,
      region == "North America" ~ 6,
      TRUE ~ 3
    ),
    iso3c_lab = paste0("<span style='font-size:", size_group, "pt;'>", iso3c, "</span>")
  ) %>%
  distinct(iso3c, iso3c_lab)

# Clean aggregate estimate data
agg_estimate <- bind_rows(agg, disagg) %>%
  group_by(iso3c) %>%
  mutate(agg_est = estimate[group_label == "aggregate"][1]) %>%
  add_region("iso3c", "region") %>%
  ungroup() %>%
  mutate(
    iso3c  = fct_reorder(iso3c, -agg_est),
    is_agg = group_label == "aggregate"
  ) %>%
  filter(group_label == "aggregate") %>%
  select(iso3c, agg_estimate = estimate)


# --- 4. Plot ---
(plot <- disagg %>%
  group_by(iso3c) %>%
  # mutate(agg_est = estimate[group_label == "aggregate"][1]) %>%
  add_region("iso3c", "region") %>%
  ungroup() %>%
    left_join(agg_estimate, by = "iso3c") %>%
  mutate(
    iso3c  = fct_reorder(iso3c, -agg_estimate),
    # is_agg = group_label == "aggregate"
  ) %>%
  ggplot(aes(x = estimate, y = iso3c, fill = group_label)) +

  geom_col(
    # aes(
    #   size  = if_else(is_agg, 2.3, 1.95),
    #   alpha = if_else(is_agg, 1, 0.55),
    #   stroke = if_else(is_agg, 0.8, 0.2)
    # ),
    # fill = "white"
  ) +
    geom_point(aes(x = agg_estimate, y = iso3c), shape = 16) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.4) +

  scale_alpha_identity() +
  scale_size_identity() +
  scale_fill_manual(
    values = c("#C73E3A", "#E68A2E", "#D8C84A", "#7A5FA3", "#3A77B5")
  ) +

  scale_y_discrete(labels = function(x) {
    label_lookup$iso3c_lab[match(x, label_lookup$iso3c)]
  }) +

  labs(
    x     = expression(bold(Delta[2010 - 2020] ~ "Aquatic animal protein reliance %")),
    y     = "Country",
    color = "Production source"
  ) +

  facet_wrap(~ region, scales = "free_y") +

  theme_minimal(base_size = 8) +
  theme(
    legend.position    = "bottom",
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    axis.text.x        = element_text(size = 6),
    # axis.text.y        = element_markdown(),
    axis.text.y = element_blank(),
    legend.key.size    = unit(0.3, "cm"),
    legend.box         = "vertical",
    legend.spacing     = unit(0.1, "cm"),
    legend.text        = element_text(size = 5),
    legend.margin      = margin(t = 0, b = 0)
  ))





# ggsave("../images/pct_reliance_contribution_four_sources.jpg", plot = plot, device = "jpg", units = "in", width = 5, height = 4, dpi = 2000)
```

```{r}
extract_legend <- function(p) {
  g <- cowplot::get_plot_component(p, "guide-box", return_all = TRUE)
  g
}

(plot <- disagg %>%
  group_by(iso3c) %>%
  # mutate(agg_est = estimate[group_label == "aggregate"][1]) %>%
  add_region("iso3c", "region") %>%
  ungroup() %>%
    left_join(agg_estimate, by = "iso3c") %>%
  mutate(
    iso3c  = fct_reorder(iso3c, -agg_estimate),
    # is_agg = group_label == "aggregate"
  ) %>%
  filter(region == "Africa") %>%
  ggplot(aes(x = estimate, y = iso3c, fill = group_label)) +

  geom_col(
    # aes(
    #   size  = if_else(is_agg, 2.3, 1.95),
    #   alpha = if_else(is_agg, 1, 0.55),
    #   stroke = if_else(is_agg, 0.8, 0.2)
    # ),
    # fill = "white"
  ) +
    geom_point(aes(x = agg_estimate, y = iso3c), shape = 16) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.4) +

  scale_alpha_identity() +
  scale_size_identity() +
  scale_fill_manual(
    values = c("#C73E3A", "#E68A2E", "#D8C84A", "#7A5FA3", "#3A77B5")
  ))

region_counts <- Counts_by_continent %>%
  filter(method_habitat== "total", continent == "Asia") %>%
ggplot(aes(x = continent, y = count, fill = direction)) +
  geom_bar(stat = "identity") +
  labs(title = "Total", x = "Continent", y = "number of countries", fill = "Direction of Change") +
  coord_flip()

pA <- disagg %>%
  group_by(iso3c) %>%
  add_region("iso3c", "region") %>%
  ungroup() %>%
  left_join(agg_estimate, by = "iso3c") %>%
  mutate(iso3c = fct_reorder(iso3c, -agg_estimate)) %>%
  ggplot(aes(x = estimate, y = iso3c, fill = group_label)) +
  geom_col() +
  geom_point(aes(x = agg_estimate, y = iso3c), shape = 16) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.4) +
  scale_fill_manual(values = c("#C73E3A", "#E68A2E", "#D8C84A", "#7A5FA3", "#3A77B5")) +
  labs(
    x = expression(bold(Delta[2010 - 2020] ~ "Aquatic animal protein reliance %")),
    y = "Country",
    fill = "Production source"
  ) +
  facet_wrap(~ region, scales = "free_y") +
  theme_minimal(base_size = 8) +
  theme(
    legend.position = "bottom",
    legend.key.size = unit(0.3, "cm"),
    legend.text = element_text(size = 5)
  )




regional_reliance <- aquatic_reliance %>%
  filter(year == 2020, food_group == "aquatic") %>%
  group_by(iso3c) %>%
  summarize(prop_animal_protein = sum(prop_animal_protein)) %>%
  add_region("iso3c", "region") %>%
  filter(region == "Asia") %>%
  ggplot(aes(x = fct_reorder(iso3c, -prop_animal_protein), prop_animal_protein)) +
  geom_col()

final_plot <- plot_grid(region_counts, regional_reliance, plot, ncol = 1, rel_heights = c(1,1,2.5))

# ggsave("../images/test.jpg", plot = final_plot, device = "jpg", units = "in", width = 5, height = 6, dpi = 2000)
```

```{r}
library(dplyr)
library(ggplot2)
library(forcats)
library(cowplot)

# -----------------------------
# 1. YOUR THREE ORIGINAL PLOTS
# -----------------------------

plot <- disagg %>%
  group_by(iso3c) %>%
  add_region("iso3c", "region") %>%
  ungroup() %>%
  left_join(agg_estimate, by = "iso3c") %>%
  mutate(iso3c = fct_reorder(iso3c, -agg_estimate)) %>%
  filter(region == "Africa") %>%
  ggplot(aes(x = estimate, y = iso3c, fill = group_label)) +
  geom_col() +
  geom_point(aes(x = agg_estimate, y = iso3c), shape = 16) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.4) +
  scale_alpha_identity() +
  scale_size_identity() +
  scale_fill_manual(values = c("#C73E3A", "#E68A2E", "#D8C84A", "#7A5FA3", "#3A77B5"))

region_counts <- Counts_by_continent %>%
  filter(method_habitat == "total", continent == "Asia") %>%
  ggplot(aes(x = continent, y = count, fill = direction)) +
  geom_bar(stat = "identity") +
  labs(title = "Total", x = "Continent", y = "number of countries", fill = "Direction of Change") +
  coord_flip()

regional_reliance <- aquatic_reliance %>%
  filter(year == 2020, food_group == "aquatic") %>%
  group_by(iso3c) %>%
  summarize(prop_animal_protein = sum(prop_animal_protein)) %>%
  add_region("iso3c", "region") %>%
  filter(region == "Asia") %>%
  ggplot(aes(x = fct_reorder(iso3c, -prop_animal_protein), prop_animal_protein)) +
  geom_col()

# -----------------------------
# 2. EXTRACT LEGENDS
# -----------------------------

extract_legend <- function(p) {
  cowplot::get_plot_component(p, "guide-box", return_all = TRUE)
}

legend_plot <- extract_legend(plot)
legend_region_counts <- extract_legend(region_counts)

# -----------------------------
# 3. REMOVE LEGENDS FROM PANELS
# -----------------------------

plot_nolegend <- plot + theme(legend.position = "none")
region_counts_nolegend <- region_counts + theme(legend.position = "none")

# -----------------------------
# 4. COMBINE LEGENDS
# -----------------------------

legend_row <- cowplot::plot_grid(
  legend_region_counts,
  legend_plot,
  ncol = 2,
  rel_widths = c(1, 1)
)

# -----------------------------
# 5. FINAL PLOT WITH LEGENDS AT BOTTOM
# -----------------------------

final_plot <- cowplot::plot_grid(
  region_counts_nolegend,
  regional_reliance,
  plot_nolegend,
  legend_row,
  ncol = 1,
  rel_heights = c(1, 1, 2.5, 0.35)
)

final_plot
```


### Compare aggregate to disaggregate reliance

```{r compare aggregate to disaggregated reliance}
aa_reliance_mean_global_trends_alt <- aquatic_reliance %>%
  group_by(year, data_source, food_group) %>%
  summarise(food_group_total_protein_t = sum(protein_consumed_t)) %>%
  group_by(year, data_source) %>%
  mutate(prop_animal_protein = 100*food_group_total_protein_t/sum(food_group_total_protein_t)) %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS")) # Remove overlapping years, prioritizing new years

# Define year ranges for projections (add this before the existing projections code)
year_ranges <- tibble(
  data_source = c("old FBS", "new FBS"),
  start_year  = c(1996, 1996),
  end_year    = c(2020, 2020)
)

observed_ranges <- tibble(
  data_source = c("old FBS", "new FBS"),
  min_year = c(1996, 2010),
  max_year = c(2009, 2020)
)

# --- FIRST PLOT: aggregate global trend with projections ---

coefs_global <- aa_reliance_mean_global_trends_alt %>%
  filter(food_group == "aquatic") %>%
  group_by(data_source) %>%
  nest() %>%
  mutate(model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
         tidy  = map(model, broom::tidy)) %>%
  unnest(tidy) %>%
  select(data_source, term, estimate) %>%
  pivot_wider(values_from = estimate, names_from = term) %>%
  rename(year_coef = year)

projections_global <- coefs_global %>%
  left_join(year_ranges, by = "data_source") %>%
  rowwise() %>%
  mutate(year = list(start_year:end_year)) %>%
  unnest(year) %>%
  ungroup() %>%
  mutate(prop_animal_protein = `(Intercept)` + year_coef * year,
         type = "projected") %>%
  left_join(observed_ranges, by = "data_source") %>%
  filter(year < min_year | year > max_year)

observed_global_old <- aa_reliance_mean_global_trends_alt %>%
  filter(food_group == "aquatic", data_source == "old FBS", year <= 2009) %>%
  mutate(type = "observed")

observed_global_new <- aa_reliance_mean_global_trends_alt %>%
  filter(food_group == "aquatic", data_source == "new FBS", year >= 2010) %>%
  mutate(type = "observed")

combined_global <- bind_rows(observed_global_old, observed_global_new)

aa_reliance_mean_global_trends <- ggplot() +
  geom_area(data = combined_global,
            aes(x = year, y = prop_animal_protein),
            fill = "#1179be") +
  geom_line(data = projections_global,
            aes(x = year, y = prop_animal_protein, group = data_source),
            colour = "black", linetype = "dashed", linewidth = 0.5, alpha = 0.7) +
  labs(x = "", y = "Global Animal Protein Reliance (%)") +
  theme_minimal(base_size = 8) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(color = scales::alpha("black", 0.25), linewidth = 0.3),
    panel.grid.minor.y = element_blank(),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),
    axis.text = element_text(size = 6, color = "black"),
    axis.title.y = element_text(size = 7, margin = margin(r = 5), face = "bold"),
    legend.position = "bottom",
    legend.title = element_text(size = 7, face = "bold"),
    legend.text = element_text(size = 6),
    legend.key.size = unit(0.3, "cm"),
    legend.box = "vertical",
    legend.spacing = unit(0.1, "cm"),
    legend.margin = margin(t = 0, b = 0),
    legend.box.margin = margin(t = -2, b = -2),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    aspect.ratio = 0.4
  )


# Calculate the total terrestrial animal protein
 animal_protein_sourcing <- aquatic_reliance %>%
  group_by(year, data_source, food_group) %>%
  summarise(animal_protein = sum(protein_consumed_t)) %>%
  filter(food_group == "terrestrial")

# Calculate the total aquatic animal protein
 aquatic_protein_sourcing <- aquatic_reliance %>%
  group_by(year, data_source, food_group) %>%
  summarise(aquatic_protein = sum(protein_consumed_t)) %>%
  filter(food_group == "aquatic")
 
 # Calculate total tonnage of combined terrestrial and aquatic animal protein
 total_protein <- left_join(animal_protein_sourcing, aquatic_protein_sourcing, by = c("year", "data_source")) %>%
   group_by(year, data_source) %>%
   summarize(total_protein = sum(animal_protein,aquatic_protein)) # add total aquatic animal protein (unseparated by sourcing) to animal proteins by year and data source
 
# Join total protein to aquatic protein weights separated by protein to calculate reliance by sourcing
 aa_reliance_multiple_trends_alt <- aquatic_reliance %>%
  group_by(year, data_source, consumption_source, habitat, method, food_group) %>%
  summarise(food_group_total_protein_t = sum(protein_consumed_t)) %>%
  filter(food_group == "aquatic") %>%
  left_join(total_protein, by = c("year", "data_source")) %>%
  ungroup() %>%
  mutate(prop_animal_protein = 100 * (food_group_total_protein_t / total_protein)) %>% # Divide AA / TERRESTRIAL + AA
  filter(habitat != "unknown") %>%
   filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS")) # Remove overlapping years, prioritizing new years
 
 
 aa_reliance_multiple_trends_alt_1 <- aa_reliance_multiple_trends_alt %>%
  mutate(habitat = str_to_title(habitat)) %>%
  mutate(habitat_method = paste(habitat, method, sep = " "))
 
# Create forward / backward projections
 
 coefs <- aquatic_reliance %>%
  group_by(year, data_source, habitat, method, food_group) %>%
  summarise(food_group_total_protein_t = sum(protein_consumed_t), .groups = "drop") %>%
  filter(food_group == "aquatic") %>%
  left_join(total_protein, by = c("year", "data_source")) %>%
  mutate(prop_animal_protein = 100 * (food_group_total_protein_t / total_protein)) %>%
  group_by(data_source, habitat, method) %>%   # <- KEEP data_source
  nest() %>%
  mutate(model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
         tidy  = map(model, broom::tidy)) %>%
  unnest(tidy) %>%
  select(data_source, habitat, method, term, estimate) %>%
  pivot_wider(values_from = estimate, names_from = term) %>%
  rename(year_coef = year)

projections <- coefs %>%
  left_join(year_ranges, by = "data_source") %>%
  rowwise() %>%
  mutate(year = list(start_year:end_year)) %>%
  unnest(year) %>%
  ungroup() %>%
  mutate(
    prop_animal_protein = `(Intercept)` + year_coef * year,
    habitat = str_to_title(habitat)
  ) %>%
  filter(method != "unknown", habitat != "Unknown") %>%
  select(year, data_source, habitat, method, prop_animal_protein) %>%
  mutate(type = "projected") %>%
  left_join(observed_ranges, by = "data_source") %>%
  filter(year < min_year | year > max_year)  # Keep only years outside observed range

 
# Artis palette
artis_colors <- artis_palette(3)

artis_colors <- artis_colors[c(2,1)]

# Prepare observed subsets without overlap
observed_old <- aa_reliance_multiple_trends_alt_1 %>%
  filter(data_source == "old FBS", year <= 2009) %>%
  mutate(type = "observed")

observed_new <- aa_reliance_multiple_trends_alt_1 %>%
  filter(data_source == "new FBS", year >= 2010) %>%
  mutate(type = "observed")

# Filter projections to non-overlapping years
projections_old <- projections %>%
  filter(data_source == "old FBS", year > 2009) %>%
  mutate(type = "projected")

projections_new <- projections %>%
  filter(data_source == "new FBS", year < 2010) %>%
  mutate(type = "projected")

# Combine all cleanly
combined <- bind_rows(observed_old, observed_new, projections_old, projections_new) %>%
  mutate(habitat_method = paste(habitat, method))

# Plot
aa_reliance_multiple_trends <- ggplot() +
  geom_area(data = filter(combined, type == "observed"),
            aes(x = year, y = prop_animal_protein, fill = consumption_source),
            size = 0.75) +
  geom_line(data = filter(combined, type == "projected"),
            aes(x = year, y = prop_animal_protein, group = data_source),
            colour = "black", linetype = "dashed", linewidth = 0.5, alpha = 0.7) +
  facet_wrap(~ habitat_method) +
  scale_fill_manual(values = c("#1179be", "#b1d0ed")) +
  labs(fill = "Consumption source", x = "", y = "Aquatic Animal Reliance (%)") +
  theme_minimal(base_size = 8) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(color = scales::alpha("black", 0.25), linewidth = 0.3),
    panel.grid.minor.y = element_blank(),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),
    strip.background = element_blank(),
    strip.text = element_text(size = 7, face = "bold"),
    axis.text = element_text(size = 6, color = "black"),
    axis.title.y = element_text(size = 7, margin = margin(r = 5), face = "bold"),
    axis.text.x = element_text(size = 5.5, angle = 30, hjust = 1),
    legend.position = "bottom",
    legend.title = element_text(size = 7, face = "bold"),
    legend.text = element_text(size = 6),
    legend.key.size = unit(0.3, "cm"),
    legend.box = "vertical",
    legend.spacing = unit(0.1, "cm"),
    legend.margin = margin(t = 0, b = 0),
    legend.box.margin = margin(t = -2, b = -2),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA)
  )



(aa_reliance_mean_global_trends_plot <- plot_grid(
  aa_reliance_mean_global_trends,
  aa_reliance_multiple_trends,
  ncol = 2,
  rel_heights = c(0.8, 2.2),  # Give more space to the faceted plot
  labels = c("A", "B"),
  label_size = 10
))

# ggsave("../images/aa_reliance_mean_global_trends.png", plot = aa_reliance_mean_global_trends_plot, device = "png", units = "in", width = 5, height = 3, bg = "transparent")
```

### X Countries have dominant reliance on X source

```{r X Countries have dominant reliance on X source}
# X Countries have dominant reliance on X source
aquatic_reliance %>%
  filter(year == 2020, food_group == "aquatic") %>%
  group_by(iso3c) %>%
  mutate(largest_group = prop_animal_protein == max(prop_animal_protein)) %>%
  filter(largest_group == TRUE) %>%
  ungroup() %>%
  count(method)
```

> Foreign marine capture reliance is the most dominant across 86 countries, while domestic inland aquaculture is more concentrated at 12 countries. 

### How many countries have received X % of their animal source protein from marine vs inland

```{r X Countries have dominant reliance on X source}
# X Countries have dominant reliance on X source
aquatic_reliance %>%
  filter(year == 2020, food_group == "aquatic") %>%
  group_by(iso3c, habitat) %>%
  summarize(prop_animal_protein = sum(prop_animal_protein)) %>%
  mutate(individual_prop = round(prop_animal_protein / sum(prop_animal_protein),2)) %>%
  group_by(habitat) %>%
  count(individual_prop >= 0.8)
```

> Marine sources remain the most widespread reliant habitat, with 102 countries obtaining least 80% of their aquatic animal proteins from marine habitats, while only 18 sources at least 80% from inland sources. 


## RQ 1b: How does aquatic animal reliance change with the relative contribution of aquatic versus terrestrial animal consumption

### calculate rates of change between terr / aquatic animal consumption

### Region counts overlayed on quadrant plot

```{r Region counts overlayed on quadrant plot}
library(dplyr)
library(ggplot2)
library(ggrepel)
library(patchwork)
library(cowplot)
library(grid)

#calculate rate of change in aquatic protein quantity
aquatic_protein_quant <- aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, year) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap, na.rm = TRUE)) %>%
  distinct() %>%
  split(.$iso3c) %>%
  map(~lm(protein_consumed_kg_percap~year, data = .x)) %>%
  map_df(broom::tidy, .id = 'iso3c') %>%
  filter(term == 'year') %>%
  select(c("iso3c", "aquatic_estimate" = "estimate"))

#calculate rate of change in aquatic protein quantity
aquatic_protein_quant <- aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, year) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap, na.rm = TRUE)) %>%
  distinct() %>%
  split(.$iso3c) %>%
  map(~lm(protein_consumed_kg_percap~year, data = .x)) %>%
  map_df(broom::tidy, .id = 'iso3c') %>%
  filter(term == 'year') %>%
  select(c("iso3c", "aquatic_estimate" = "estimate"))
#calculate rate of change in terrestrial protein quantity
terrestrial_protein_quant <- aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         data_source == "new FBS", food_group == "terrestrial") %>%
  group_by(iso3c, year) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap, na.rm = TRUE)) %>%
  distinct() %>%
  split(.$iso3c) %>%
  map(~lm(protein_consumed_kg_percap~year, data = .x)) %>%
  map_df(broom::tidy, .id = 'iso3c') %>%
  filter(term == 'year') %>%
  select(c("iso3c", "terrestrial_estimate" = "estimate"))
#calculate rate of change in aquatic reliance
aquatic_protein_reliance <- aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, year) %>%
  summarize(prop_animal_protein = sum(prop_animal_protein, na.rm = TRUE)) %>%
  distinct() %>%
  split(.$iso3c) %>%
  map(~lm(prop_animal_protein~year, data = .x)) %>%
  map_df(broom::tidy, .id = 'iso3c') %>%
  filter(term == 'year') %>%
  select(c("iso3c", "aquatic_reliance_estimate" = "estimate"))
  
#join data 
plot_df<- aquatic_protein_quant %>%
  full_join(terrestrial_protein_quant, by="iso3c") %>%
  full_join(aquatic_protein_reliance, by="iso3c") %>%
  add_region(col = "iso3c", region.col.name = "region")

# ============================================================
# 0. REGION LEGEND FROM BAR PLOT
# ============================================================

p_bar <- plot_df %>%
  mutate(quadrant = case_when(
    aquatic_estimate > 0 & terrestrial_estimate > 0 ~ "Q1",
    aquatic_estimate < 0 & terrestrial_estimate > 0 ~ "Q2",
    aquatic_estimate < 0 & terrestrial_estimate < 0 ~ "Q3",
    aquatic_estimate > 0 & terrestrial_estimate < 0 ~ "Q4"
  )) %>%
  group_by(quadrant, region) %>%
  summarize(n = n_distinct(iso3c), .groups = "drop") %>%
  mutate(quadrant = factor(quadrant, levels = c("Q2","Q1","Q3","Q4"))) %>%
  ggplot(aes(x = region, y = n, fill = region)) +
  geom_col(width = 0.75) +
  scale_fill_manual(values = artis_palette(6)) +
  labs(x = "", y = "Number of countries", fill = "Region") +
  theme_minimal(base_size = 7) +
  theme(
    axis.text.x = element_blank(),
    legend.position = "bottom",
    legend.title = element_text(size = 6, face = "bold"),
    legend.text  = element_text(size = 5),
    legend.key.size = unit(0.6, "lines")
  ) +
  facet_wrap(~ quadrant)

legend_regions <- get_plot_component(p_bar, "guide-box-bottom", return_all = TRUE)

# ============================================================
# 1. SCATTERPLOT (KEEP AQUATIC RELIANCE LEGEND)
# ============================================================

p_scatter_base <- plot_df %>%
  filter(iso3c != "MDV") %>%
  ggplot(aes(
    x = aquatic_estimate,
    y = terrestrial_estimate,
    fill = aquatic_reliance_estimate,
    label = iso3c
  )) +
  geom_point(shape = 21, size = 1.8, color = "black", stroke = 0.2) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  geom_text_repel(size = 2.4) +
  scale_fill_gradient2(
  low = "#ca0020",     # muted red
  mid = "#f7f7f7",     # very light grey
  high = "#0571b0",    # muted blue
  midpoint = 0
) +
  labs(
    x = expression(bold(Delta[2010-2020] ~ "Aquatic animal protein (kg/cap/day)")),
    y = expression(bold(Delta[2010-2020] ~ "Terrestrial animal protein (kg/cap/day)")),
    fill = expression(atop(bold(Delta[2010-2020]), bold("Aquatic animal reliance")))
  ) +
  theme_minimal(base_size = 8) +
  theme(
    panel.grid.minor = element_blank(),
    axis.text = element_text(size = 6, color = "black"),
    axis.title = element_text(size = 7, face = "bold"),
    legend.position = "bottom",
    legend.title = element_text(size = 6),
    legend.text = element_text(size = 5, angle = 35),
    legend.key.size = unit(.8, "lines")
  )

# Extract aquatic reliance legend
legend_reliance <- get_plot_component(p_scatter_base, "guide-box-bottom", return_all = TRUE)

# Remove legend from scatter for main panel
p_scatter <- p_scatter_base + theme(legend.position = "none")

x_max <- max(abs(plot_df$aquatic_estimate), na.rm = TRUE)
y_max <- max(abs(plot_df$terrestrial_estimate), na.rm = TRUE)

p_scatter <- p_scatter +
  coord_cartesian(
    xlim = c(-x_max, x_max),
    ylim = c(-y_max, y_max)
  )

# Quadrant counts + insets

quad_counts <- plot_df %>%
  mutate(quadrant = case_when(
    aquatic_estimate > 0 & terrestrial_estimate > 0 ~ "Q1",
    aquatic_estimate < 0 & terrestrial_estimate > 0 ~ "Q2",
    aquatic_estimate < 0 & terrestrial_estimate < 0 ~ "Q3",
    aquatic_estimate > 0 & terrestrial_estimate < 0 ~ "Q4"
  )) %>%
  group_by(quadrant, region) %>%
  summarise(n = n_distinct(iso3c), .groups = "drop")

max_count <- max(quad_counts$n)

make_inset <- function(q) {
  quad_counts %>%
    filter(quadrant == q) %>%
    ggplot(aes(x = region, y = n, fill = region)) +
    geom_col(width = 0.75) +
    scale_fill_manual(values = artis_palette(6)) +
    scale_y_continuous(limits = c(0, max_count)) +
    theme_minimal(base_size = 6) +
    theme(
      axis.title = element_blank(),
      axis.text = element_blank(),
      panel.grid = element_blank(),
      legend.position = "none",
      plot.margin = margin(0, 0, 0, 0)
    )
}

p_q1 <- make_inset("Q1")
p_q2 <- make_inset("Q2")
p_q3 <- make_inset("Q3")
p_q4 <- make_inset("Q4")

# add insets inside scatterplot

shrink_x <- 0.18
shrink_y <- 0.18

p_with_insets <- p_scatter +
  inset_element(p_q1, left = 1 - shrink_x, right = 1, bottom = 1 - shrink_y, top = 1) +
  inset_element(p_q2, left = 0, right = shrink_x, bottom = 1 - shrink_y, top = 1) +
  inset_element(p_q3, left = 0, right = shrink_x, bottom = 0, top = shrink_y) +
  inset_element(p_q4, left = 1 - shrink_x, right = 1, bottom = 0, top = shrink_y)

# Add right y-axis

fake_axis <- ggplot() +
  scale_y_continuous(
    limits = c(0, max_count),
    breaks = seq(0, max_count, by = 5)
  ) +
  theme_void() +
  theme(
    axis.text.y  = element_text(size = 6),
    axis.title.y = element_text(size = 7, face = "bold", angle = 90),
    axis.line.x  = element_blank(),
    axis.ticks.x = element_blank(),
    plot.margin  = margin(0, 5, 0, 0)
  ) +
  labs(y = "Number of countries")

scatter_with_axis <- plot_grid(
  p_with_insets,
  fake_axis,
  ncol = 2,
  rel_widths = c(1, 0.10)
)

# Add reliance and region legends side by side

legend_row <- plot_grid(
  legend_reliance,
  legend_regions,
  ncol = 2,
  rel_widths = c(1, 1)
)

# Compile plot

(final_plot <- plot_grid(
  scatter_with_axis,
  legend_row,
  ncol = 1,
  rel_heights = c(1, 0.18)
))

wilcox.test(plot_df$terrestrial_estimate, plot_df$aquatic_estimate, paired = FALSE, alternative = "greater")

# Calculate CV for the terrestrial consumption roc
sd(plot_df$terrestrial_estimate) / mean(plot_df$terrestrial_estimate)
 
 # Calculate CV for the aquatic animal consumption roc
sd(plot_df$aquatic_estimate) / mean(plot_df$aquatic_reliance_estimate)


# ggsave("../images/scatter_bar_overlay.jpg", plot = final_plot, device = "png", units = "in", width = 5, height = 3.5)
```
