---
title: "02.1_Supplementary_Information"
authors: "Connor Quiroz, Lana Huang, Christopher D. Golden, Sarah Glaser, and Jessica A. Gephart"
date: "2026-03-04"
output: html_document
---

# Chapter 1 Supplementary information

# Load in packages & functions

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(exploreARTIS)
library(cowplot)
library(ggrepel)
library(countrycode)
library(colorspace)
library(ggpubr)
library(arrow)
library(rstan)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggridges)
library(bayesplot)
library(purrr)
library(patchwork)
library(reshape2)
library(loo)

# Load script to convert model name strings to readable labels
source("../utils/clean_model_name.R")
```

# Read in data

### Read in aquatic reliance data & FAO population data

```{r read in FAO derived data}
# Read in population
population_fao_wb <- read.csv("../output/population_fao_wb.csv")

# Read in aquatic reliance data
aquatic_reliance <- read.csv("../output/fao_aquatic_reliance_source.csv") %>%
  left_join(population_fao_wb,by=c("year","iso3c"))%>%
  mutate(protein_consumed_kg_percap=protein_consumed_t*1000/population_fao)

# Read in FAO sourcing data
aquatic_fbs <- read.csv("../output/fao_fbs_aquatic_source.csv")
```

## Get list of response variables (models)

This code chunk is also in the main results. This code chunk extracts the name of each model name used in the Bayesian model. In code lines further down, we call the "response_vars" object to generate statistics and figures. This code chunk will only run if the response_vars object has not been created.

```{r acquire animal consumption data to be fed through analysis}
if ( !file.exists("../output/response_vars.rds") ) {
  
  # Calculate animal consumption
terr_animal_cons_cap <- aquatic_reliance %>%
  filter(year == 2019) %>%
  group_by(iso3c) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))

# Calculate aquatic animal consumption
aquatic_animal_cons_cap <- aquatic_reliance %>%
  filter(year == 2019, food_group == "aquatic") %>%
  group_by(iso3c) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))

# Calculate animal reliance consumption
aquatic_rel <- aquatic_animal_cons_cap %>%
  left_join(animal_cons_cap, by = "iso3c") %>%
  group_by(iso3c) %>%
  summarize(prop_animal_protein = protein_consumed_kg_percap.x / protein_consumed_kg_percap.y)

# Add items not created in the loop to the list
response_vars <- c("terr_animal_cons_cap", "aquatic_animal_cons_cap", "aquatic_rel")

for (i in c("capture", "aquaculture", "domestic", "foreign", "marine", "inland")) {
  
  # Calculate consumption per capita by method
  if (i %in% c("capture", "aquaculture")) {
    assign(
      paste0(i, "_cap"),
      aquatic_reliance %>%
        filter(year == 2019, food_group == "aquatic", method == i) %>%
        group_by(iso3c) %>%
        summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))
    )
  }
  
  # Calculate consumption per capita by consumption source
  if (i %in% c("domestic", "foreign")) {
    assign(
      paste0(i, "_cap"),
      aquatic_reliance %>%
        filter(year == 2019, food_group == "aquatic", consumption_source == i) %>%
        group_by(iso3c) %>%
        summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))
    )
  }
  
  # Calculate consumption per capita by habitat
  if (i %in% c("marine", "inland")) {
    assign(
      paste0(i, "_cap"),
      aquatic_reliance %>%
        filter(year == 2019, food_group == "aquatic", habitat == i) %>%
        group_by(iso3c) %>%
        summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))
    )
  }
  
  response_vars <- c(response_vars, paste0(i, "_cap"))
  
  # Calculate reliance by method, production source, or habitat
  assign(
    paste0(i, "_rel"),
    get(paste0(i, "_cap")) %>%
      left_join(animal_cons_cap, by = "iso3c") %>%
      group_by(iso3c) %>%
      summarize(prop_animal_protein = protein_consumed_kg_percap.x / protein_consumed_kg_percap.y)
  )
  
  response_vars <- c(response_vars, paste0(i, "_rel"))
  
}

# Save response_vars to RDS format
saveRDS(response_vars, "../data/response_vars.rds")
}
```

## Read in list of response variables (models)

```{r read in response variables}
# Read in list of response variables used in Bayesian models
response_vars <- readRDS("../data/response_vars.rds")
```

## Instantiate Bayesian model used

This code chunk establishes the directory to pull Bayesian output from to generate statistics & fiugres

```{r}
# Declare model type
model_type = "realm"

# Variable list with coastline_km area instead of EEZ area
predictors <- c(
  "coastline_km",
  "log_gdp",
  "gdp_trade",
  "inundation_area_log1000km2",
  "supermarkets",
  "pop_10k_pct_coast"
)

# Get which variable is selected (for unique file path)
var_used = "all"

# Create file path to save files to
base_path <- file.path(
  "..", "output", "bayesian_output",
  paste0("random_effect_", model_type),
  paste0("var_", var_used)
)
```


# RQ 1a: How do aquatic food reliance trends disaggregated by source differ from aggregate aquatic food reliance?

## SI figure: Multi metric plot with 4 panels created for each individual region

The first row of the figure has 2 panels, and the following two rows have 1 panel, making a 4 panel plot. The regions include Africa, Asia, South America, North America, Europe, and Oceania.

The top left panel highlights the number of countries with increasing / decreasing *aggregate* aquatic animal reliance

The top right panel highlights the number of countries with increasing / decreasing *disaggregate* aquatic animal reliance. Where the y-axis represents the disaggregated production sources for aquatic foods.

The middle panel measures the 2020 aggregate aquatic animal reliance for each country.

The bottom panel visualizes for each country in each region the average annual change in their aggregate aquatic animal reliance, with separate colors as stacked bars representing the disaggregate reliance.

```{r individual continent production contribution}
# shared color palettes

# Production-source palette (matches main figure)
pal_consource <- c(
  "inland aquaculture" = "#E8B000",
  "inland capture"     = "#D4601A",
  "marine aquaculture" = "#9B6DC5",
  "marine capture"     = "#2B6AA8",
  "unknown"            = "#A8A8A8"
)

# Diverging palette for direction-of-change panels
pal_direction <- c(
  "decreasing" = "#B2182B",
  "increasing" = "#2166AC"
)

# Neutral fill for reliance bar chart
col_reliance_bar <- "#606060"


# Shared base theme

theme_si <- function(base_size = 8,
                     show_x_title = TRUE,
                     show_y_title = TRUE,
                     show_legend  = TRUE) {

  base <- theme_minimal(base_size = base_size) +
    theme(
      text                  = element_text(family = "sans", color = "black"),
      plot.background       = element_rect(fill = "transparent", color = NA),
      panel.background      = element_rect(fill = "transparent", color = NA),
      panel.grid.minor      = element_blank(),
      panel.grid.major.y    = element_line(
        color = scales::alpha("black", 0.06), linewidth = 0.3
      ),
      panel.grid.major.x    = element_line(
        color = scales::alpha("black", 0.25), linewidth = 0.3
      ),
      axis.text             = element_text(size = base_size - 2.5, color = "black"),
      axis.title.x          = element_text(
        size = base_size - 1, face = "bold",
        margin = margin(t = 3)
      ),
      axis.title.y          = element_text(
        size = base_size - 1, face = "bold",
        margin = margin(r = 3)
      ),
      strip.text            = element_text(size = base_size - 1, face = "bold"),
      legend.position       = if (show_legend) "bottom" else "none",
      legend.title          = element_text(size = base_size - 1.5, face = "bold"),
      legend.text           = element_text(size = base_size - 2.5),
      legend.key.size       = unit(0.25, "cm"),
      legend.box            = "vertical",
      legend.spacing        = unit(0.05, "cm"),
      legend.margin         = margin(t = 0, b = 0),
      legend.box.margin     = margin(t = -2, b = -2),
      legend.background     = element_rect(fill = "transparent", color = NA),
      legend.box.background = element_rect(fill = "transparent", color = NA),
      plot.margin           = margin(3, 4, 3, 4)
    )

  if (!show_x_title) base <- base + theme(axis.title.x = element_blank())
  if (!show_y_title) base <- base + theme(axis.title.y = element_blank())
  base
}

# suppress y-axis grid (used for horizontal count bars
theme_si_hbar <- function(base_size = 8, show_legend = TRUE) {
  theme_si(base_size = base_size, show_legend = show_legend) +
    theme(
      panel.grid.major.y = element_blank(),
      axis.text.y        = element_blank(),
      axis.ticks.y       = element_blank(),
      axis.title.y       = element_blank()
    )
}


# Shared geom helpers

agg_point <- function() {
  geom_point(
    aes(x = agg_estimate, y = iso3c),
    shape  = 21,
    fill   = "grey20",
    color  = "white",
    size   = 1.2,
    stroke = 0.3
  )
}


# Fit linear models to estimate the average annual change in disaggregated
# aquatic animal reliance (2010-2020) by production source (habitat x method)
# for each country
disagg <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "aquatic") %>%
  group_by(year, iso3c, habitat, method) %>%
  summarize(prop_animal_protein = sum(prop_animal_protein), .groups = "drop") %>%
  group_by(iso3c, habitat, method) %>%
  nest() %>%
  mutate(
    model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
    tidy  = map(model, broom::tidy)
  ) %>%
  unnest(tidy) %>%
  filter(term == "year") %>%
  mutate(
    group_label = paste(habitat, method),
    group_label = if_else(group_label == "unknown unknown", "unknown", group_label)
  ) %>%
  drop_na(estimate) %>%
  group_by(group_label) %>%
  mutate(pct_above = mean(estimate > 0) * 100) %>%
  ungroup()

# Fit linear models to estimate the average annual change in aggregate
# aquatic animal reliance (2010-2020)
agg <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, year) %>%
  summarise(prop_animal_protein = sum(prop_animal_protein), .groups = "drop") %>%
  group_by(iso3c) %>%
  nest() %>%
  mutate(
    model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
    tidy  = map(model, broom::tidy)
  ) %>%
  unnest(tidy) %>%
  filter(term == "year") %>%
  mutate(group_label = "aggregate") %>%
  drop_na(estimate) %>%
  group_by(group_label) %>%
  mutate(pct_above = mean(estimate > 0) * 100) %>%
  ungroup()

# Clean aggregate estimate data
agg_estimate <- bind_rows(agg, disagg) %>%
  group_by(iso3c) %>%
  mutate(agg_est = estimate[group_label == "aggregate"][1]) %>%
  add_region("iso3c", "region") %>%
  ungroup() %>%
  filter(group_label == "aggregate") %>%
  select(iso3c, agg_estimate = estimate)

# Extract rate of change in aggregate aquatic animal reliance from 2010-2020 
roc_aquatic_total <- aquatic_reliance %>%
  filter(
    !(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
    data_source == "new FBS",
    food_group  == "aquatic"
  ) %>%
  group_by(iso3c, year) %>%
  summarize(prop_animal_protein = sum(prop_animal_protein, na.rm = TRUE),
            .groups = "drop") %>%
  distinct() %>%
  split(list(.$iso3c), drop = TRUE) %>%
  map(~ lm(prop_animal_protein ~ year, data = .x)) %>%
  map_df(broom::tidy, .id = "combined_id") %>%
  filter(term == "year") %>%
  select(iso3c = combined_id, roc = estimate) %>%
  mutate(
    direction      = if_else(roc > 0, "increasing", "decreasing"),
    method_habitat = "total"
  ) %>%
  add_region(col = "iso3c", region.col.name = "continent")

counts_by_continent <- roc_aquatic_total %>%
  group_by(continent, method_habitat) %>%
  count(direction) %>%
  group_by(continent) %>%
  mutate(prop = n / sum(n))


# Extract legend

extract_legend <- function(p) {
  legs <- cowplot::get_plot_component(p, "guide-box", return_all = TRUE)
  good <- Filter(function(x) !inherits(x, "zeroGrob"), legs)
  if (length(good) == 0) return(NULL)
  good[[1]]
}


# Loop across regions

regions <- c("Asia", "Africa", "Europe", "North America", "South America", "Oceania")

for (i in regions) {

  # Panel 1: aggregate + disaggregated rate-of-change bars
  p_bars <- disagg %>%
    add_region("iso3c", "region") %>%
    filter(region == i) %>%
    left_join(agg_estimate, by = "iso3c") %>%
    mutate(iso3c = fct_reorder(iso3c, -agg_estimate)) %>%
    ggplot(aes(x = estimate, y = iso3c, fill = group_label)) +
    geom_col() +
    agg_point() +
    scale_fill_manual(
      values = pal_consource,
      name   = "Production source",
      drop   = FALSE
    ) +
    guides(fill = guide_legend(ncol = 2)) +
    labs(x = "Average annual change in aquatic animal reliance (2010–2020)") +
    theme_si(show_y_title = FALSE, show_legend = TRUE)

  # Panel 2: aggregate direction counts (horizontal stacked bar)
  p_agg_counts <- counts_by_continent %>%
    filter(method_habitat == "total", continent == i) %>%
    ggplot(aes(y = continent, x = n, fill = direction)) +
    geom_col(width = 0.55) +
    scale_fill_manual(
      values = pal_direction,
      labels = c(
        "decreasing" = "Declining reliance",
        "increasing" = "Increasing reliance"
      ),
      name = "Direction\nof change"
    ) +
    guides(fill = guide_legend(nrow = 2)) +
    labs(x = "No. countries (aggregate)") +
    theme_si_hbar(show_legend = TRUE)

  # Panel 3: disaggregated direction counts
  num_regional_countries <- disagg %>%
    add_region("iso3c", "region") %>%
    filter(region == i) %>%
    distinct(iso3c) %>%
    nrow()

  p_disagg_counts <- disagg %>%
    add_region("iso3c", "region") %>%
    filter(region == i) %>%
    group_by(region, group_label) %>%
    count(increasing = estimate > 0) %>%
    ungroup() %>%
    mutate(direction = if_else(increasing, "increasing", "decreasing")) %>%
    ggplot(aes(y = group_label, x = n, fill = direction)) +
    geom_col(width = 0.55) +
    scale_fill_manual(values = pal_direction, guide = "none") +
    labs(
      x = paste0(
        "No. countries (disaggregate, n = ", num_regional_countries, ")"
      )
    ) +
    theme_si(show_y_title = FALSE, show_legend = FALSE) +
    theme(panel.grid.major.y = element_blank())

  # Panel 4: 2020 reliance, ordered by rate of change
  p_reliance <- aquatic_reliance %>%
    filter(year == 2020, food_group == "aquatic") %>%
    group_by(iso3c) %>%
    summarize(prop_animal_protein = sum(prop_animal_protein), .groups = "drop") %>%
    left_join(agg_estimate, by = "iso3c") %>%
    add_region("iso3c", "region") %>%
    filter(region == i) %>%
    ggplot(aes(
      x = fct_reorder(iso3c, agg_estimate),
      y = prop_animal_protein * 100
    )) +
    geom_col(fill = col_reliance_bar, width = 0.7) +
    labs(y = "Aquatic animal\nreliance, 2020 (%)") +
    theme_si(show_x_title = FALSE, show_legend = FALSE) +
    theme(
      axis.text.x        = element_text(size = 4.5, angle = 45, hjust = 1),
      panel.grid.major.x = element_blank(),
      panel.grid.major.y = element_line(
        color = scales::alpha("black", 0.25), linewidth = 0.3
      )
    )

  # Assemble legends
  legend_bars       <- extract_legend(p_bars)
  legend_agg_counts <- extract_legend(p_agg_counts)

  p_bars_nl       <- p_bars       + theme(legend.position = "none")
  p_agg_counts_nl <- p_agg_counts + theme(legend.position = "none")

  legend_row <- cowplot::plot_grid(
    legend_agg_counts,
    legend_bars,
    ncol       = 2,
    rel_widths = c(1, 1.45)
  )

  # Title and divider
  title_panel <- ggplot() +
    labs(title = i) +
    theme_void() +
    theme(
      plot.title = element_text(
        hjust = 0.5, face = "bold", size = 9
      )
    )

  divider_panel <- ggplot() +
    theme_void() +
    geom_hline(yintercept = 0.5, linewidth = 0.4, color = "grey60")

  # Combine top count panels 
  top_counts <- cowplot::plot_grid(
    p_agg_counts_nl,
    p_disagg_counts,
    ncol       = 2,
    rel_widths = c(1, 1)
  )

  # Final assembly
  final_plot <- cowplot::plot_grid(
    title_panel,
    divider_panel,
    top_counts,
    p_reliance,
    p_bars_nl,
    legend_row,
    ncol        = 1,
    rel_heights = c(0.12, 0.04, 0.6, 0.8, 2.5, 0.32)
  )

  # Save plot
  ggsave(
    filename = paste0("../images/SI/multi_metrics_", i, "_SI.pdf"),
    plot     = final_plot,
    device   = cairo_pdf,
    width    = 5,
    height   = 7,
    units    = "in"
  )

}
```


# RQ 1b: How does aquatic animal reliance change with the relative contribution of aquatic versus terrestrial animal consumption

## SI Figure: Proportion of countries in each quadrant by region

This plot identifies the proportion of countries that reside in each quadrant by region. The top right quadrant indicates an increasing terrestrial animal & aquatic animal consumption. The top left quadrant indicates an increasing terrestrial animal & decreasing aquatic animal consumption. The bottom left indicates both decreasing terrestrial & aquatic animal consumption. The bottom right quadrant indicates a decreasing terrestrial & increasing aquatic animal consumption.


```{r proportion of countries in each quadrant by region}
# Calculate rate of change in aquatic protein quantity
aquatic_protein_quant <- aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, year) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap, na.rm = TRUE),
            .groups = "drop") %>%
  distinct() %>%
  split(.$iso3c) %>%
  map(~ lm(protein_consumed_kg_percap ~ year, data = .x)) %>%
  map_df(broom::tidy, .id = "iso3c") %>%
  filter(term == "year") %>%
  select(iso3c, aquatic_estimate = estimate)

# Calculate rate of change in terrestrial protein quantity
terrestrial_protein_quant <- aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         data_source == "new FBS", food_group == "terrestrial") %>%
  group_by(iso3c, year) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap, na.rm = TRUE),
            .groups = "drop") %>%
  distinct() %>%
  split(.$iso3c) %>%
  map(~ lm(protein_consumed_kg_percap ~ year, data = .x)) %>%
  map_df(broom::tidy, .id = "iso3c") %>%
  filter(term == "year") %>%
  select(iso3c, terrestrial_estimate = estimate)

# Calculate rate of change in aquatic reliance
aquatic_protein_reliance <- aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, year) %>%
  summarize(prop_animal_protein = sum(prop_animal_protein, na.rm = TRUE),
            .groups = "drop") %>%
  distinct() %>%
  split(.$iso3c) %>%
  map(~ lm(prop_animal_protein ~ year, data = .x)) %>%
  map_df(broom::tidy, .id = "iso3c") %>%
  filter(term == "year") %>%
  select(iso3c, aquatic_reliance_estimate = estimate)

# Join and compute quadrant proportions per region
quad_df <- aquatic_protein_quant %>%
  full_join(terrestrial_protein_quant, by = "iso3c") %>%
  full_join(aquatic_protein_reliance,  by = "iso3c") %>%
  add_region(col = "iso3c", region.col.name = "region") %>%
  mutate(
    quad_raw = case_when(
      aquatic_estimate > 0 & terrestrial_estimate > 0 ~ "Q1",
      aquatic_estimate < 0 & terrestrial_estimate > 0 ~ "Q2",
      aquatic_estimate < 0 & terrestrial_estimate < 0 ~ "Q3",
      aquatic_estimate > 0 & terrestrial_estimate < 0 ~ "Q4"
    )
  ) %>%
  group_by(region, quad_raw) %>%
  summarise(n = n_distinct(iso3c), .groups = "drop") %>%
  group_by(region) %>%
  mutate(
    prop     = n / sum(n),
    quadrant = factor(
      quad_raw,
      levels = c("Q1", "Q2", "Q3", "Q4"),
      labels = c(
        "Aquatic \u2191 / Terrestrial \u2191",
        "Aquatic \u2193 / Terrestrial \u2191",
        "Aquatic \u2193 / Terrestrial \u2193",
        "Aquatic \u2191 / Terrestrial \u2193"
      )
    )
  ) %>%
  ungroup()

# Plot parameters
max_val    <- 0.75   # axis limit - corresponds to 75% proportion
half_width <- 0.20   # angular half-width of each spike in radians

# Quadrant colours - harmonized with project palette
# Q1 (both up) = teal: growing together
# Q2 (aquatic down, terr up) = dark red: aquatic declining
# Q3 (both down) = dark wine: both declining
# Q4 (aquatic up, terr down) = blue: aquatic gaining share
quad_colors <- c(
  "Aquatic \u2191 / Terrestrial \u2191" = "#7BAEC4",   # mid blue - ambiguous
  "Aquatic \u2193 / Terrestrial \u2191" = "#B2182B",   # strong red - reliance clearly ↓
  "Aquatic \u2193 / Terrestrial \u2193" = "#D4956A",   # warm sand - ambiguous
  "Aquatic \u2191 / Terrestrial \u2193" = "#2166AC"    # strong blue - reliance clearly ↑
)
# Build spike polygons - each spike is a triangle from the origin
# r = prop * max_val * sqrt(2) scales so that prop = 1 fills to the axis limit
# Number of arc points - more = smoother curve, 20 is plenty at this scale
n_arc <- 20

spike_df <- quad_df %>%
  mutate(
    center_angle = case_when(
      quad_raw == "Q1" ~  pi / 4,
      quad_raw == "Q2" ~ 3*pi / 4,
      quad_raw == "Q3" ~ 5*pi / 4,
      quad_raw == "Q4" ~ 7*pi / 4
    ),
    r     = prop * max_val * sqrt(2),
    group = paste(region, quad_raw)
  ) %>%
  rowwise() %>%
  reframe(
    region   = region,
    quadrant = quadrant,
    prop     = prop,
    group    = group,
    # Fan of arc angles from left edge to right edge
    angle = c(NA, seq(center_angle - half_width,
                      center_angle + half_width,
                      length.out = n_arc)),
    # Origin point first (NA angle → 0,0), then arc points at radius r
    x = c(0, r * cos(seq(center_angle - half_width,
                         center_angle + half_width,
                         length.out = n_arc))),
    y = c(0, r * sin(seq(center_angle - half_width,
                         center_angle + half_width,
                         length.out = n_arc)))
  )

# Reference circle data - drawn at 50% and 100% of max_val
# Gives readers two anchors for judging spike length without cluttering axis labels
circle_df <- expand.grid(
  region = unique(spike_df$region),
  theta  = seq(0, 2 * pi, length.out = 200)
) %>%
  as_tibble() %>%
  crossing(r_frac = c(0.33, 1.0)) %>%
  mutate(
    x = r_frac * max_val * cos(theta),
    y = r_frac * max_val * sin(theta)
  )

# Axis quadrant label positions - placed just outside max_val in each quadrant
quad_axis_labels <- tibble(
  x     = c( 0.58, -0.58, -0.58,  0.58),
  y     = c( 0.58,  0.58, -0.58, -0.58),
  label = c("A\u2191 T\u2191", "A\u2193 T\u2191", "A\u2193 T\u2193", "A\u2191 T\u2193")
)

# Single reference label - annotates the outer circle in top-right of each facet
ref_label_df <- tibble(
  region = unique(spike_df$region)
) %>%
  bind_rows(
    tibble(region = unique(spike_df$region)) %>%
      mutate(x = -max_val * 0.5 * cos(pi/4) - 0.2,
             y =  max_val * 0.5 * sin(pi/4),
             label = "25%"),
    tibble(region = unique(spike_df$region)) %>%
      mutate(x =  max_val * cos(pi/4) + 0.04,
             y =  max_val * sin(pi/4) + 0.04,
             label = "75%")
  )

# Assemble plot
plot <- ggplot() +

  # Reference circles - dashed, at 50% and 100% of axis range
  geom_path(
    data    = circle_df %>% filter(r_frac == 1.0),
    aes(x = x, y = y, group = interaction(region, r_frac)),
    colour    = "grey70",
    linewidth = 0.25,
    linetype  = "dashed"
  ) +
  geom_path(
    data    = circle_df %>% filter(r_frac == 0.33),
    aes(x = x, y = y, group = interaction(region, r_frac)),
    colour    = "grey80",
    linewidth = 0.18,
    linetype  = "dashed"
  ) +

  # Axis lines
  geom_hline(yintercept = 0, linewidth = 0.35, colour = "grey30") +
  geom_vline(xintercept = 0, linewidth = 0.35, colour = "grey30") +

  # Spikes
  geom_polygon(
    data = spike_df,
    aes(x = x, y = y, fill = quadrant, group = group),
    alpha = 0.88
  ) +

  # 75% reference label - top-right corner only, one per facet
  geom_text(
    data  = ref_label_df,
    aes(x = x, y = y, label = label),
    size  = 1.8,
    color = "grey55",
    hjust = 0,
    vjust = 0
  ) +

  facet_wrap(~ region, nrow = 2) +
  coord_equal(
    xlim   = c(-max_val - 0.12, max_val + 0.12),
    ylim   = c(-max_val - 0.12, max_val + 0.12)
  ) +
  scale_x_continuous(breaks = NULL) +
  scale_y_continuous(breaks = NULL) +
  scale_fill_manual(
    values = quad_colors,
    name   = NULL
  ) +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE,
                              override.aes = list(alpha = 1))) +
  labs(
    x = "Aquatic protein consumption \u2192",
    y = "Terrestrial protein consumption \u2192"
  ) +
  theme_minimal(base_size = 8) +
  theme(
    panel.grid            = element_blank(),
    panel.background      = element_rect(fill = "transparent", color = NA),
    plot.background       = element_rect(fill = "transparent", color = NA),
    strip.background      = element_blank(),
    strip.text            = element_text(size = 6.5, face = "bold"),
    axis.text             = element_blank(),
    axis.title.x          = element_text(size = 6, face = "bold",
                                          margin = margin(t = 2)),
    axis.title.y          = element_text(size = 6, face = "bold",
                                          margin = margin(r = 2)),
    legend.position       = "bottom",
    legend.box            = "vertical",
    legend.title          = element_blank(),
    legend.text           = element_text(size = 5.5),
    legend.key.size       = unit(0.3, "cm"),
    legend.spacing        = unit(0.1, "cm"),
    legend.margin         = margin(t = 2, b = 0),
    legend.background     = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    panel.spacing         = unit(0.8, "lines")
  )

ggsave(
  filename = "../images/SI/quadrant_spike_region_SI.pdf",
  plot     = plot,
  device   = cairo_pdf,
  width    = 5,
  height   = 3.5,
  units    = "in"
)
```


# RQ 2: What biophysical and socioeconomic factors are associated with reliance on farmed and wild aquatic animals for protein and overall consumption?

## SI Figure: Directed Acylcic Diagram (DAG). The DAG was not created in R, but in LucidChart.

## SI Figure: Extra posterior distributions that show the partial slopes for models that use disaggregated reliance as the response.

Posterior distributions are displayed with 90%, 80%, and 50% credibility intervals, represented by the lightest, darker, and darkest colors respectively. The point represents the posterior median.

```{r Extra posteriors}
# Create data frame to hold all posteriors across variables
slopes_df <- data.frame()

for (i in seq_along(response_vars)) {

  # Extract MCMC samples
  x <- rstan::extract(
    readRDS(file.path(base_path, paste0("fit_", response_vars[i], ".rds"))),
    permuted = TRUE
  )
  beta_global <- x$beta
  X           <- readRDS(file.path(base_path, paste0("X_", response_vars[i], ".rds")))

  n_iter <- nrow(beta_global)
  K      <- ncol(beta_global)
  vars   <- colnames(X)

  # Extract global slopes and calculate credibility intervals
  ci_df <- expand_grid(iter = 1:n_iter, k = 1:K) %>%
    mutate(
      value = map2_dbl(iter, k, ~ beta_global[.x, .y]),
      var   = vars[k]
    ) %>%
    group_by(var) %>%
    summarise(
      lo95 = quantile(value, 0.025),
      lo80 = quantile(value, 0.10),
      lo50 = quantile(value, 0.25),
      med  = quantile(value, 0.50),
      hi50 = quantile(value, 0.75),
      hi80 = quantile(value, 0.90),
      hi95 = quantile(value, 0.975),
      .groups = "drop"
    ) %>%
    mutate(model = response_vars[i])

  slopes_df <- bind_rows(slopes_df, ci_df)
}

# Clean variable names for axis labels
var_labels <- c(
  coastline_km               = "Coastline length (km)",
  gdp_trade                  = "GDP from trade (%)",
  inundation_area_log1000km2 = "Inundated area (log 1000 km²)",
  log_gdp                    = "Log GDP per capita",
  pop_10k_pct_coast          = "Pop. within 10 km of coast (%)",
  supermarkets               = "Supermarkets per 100,000",
  eez_total                  = "Log EEZ area km²"
)

# Model colours - harmonized palette, reliance models only
model_colors <- c(
  "Aquatic reliance"      = "#1A7A6E",
  "Aquaculture reliance"  = "#E8B000",
  "Capture reliance"      = "#D4601A",
  "Domestic reliance"     = "#7BAEC4",
  "Foreign reliance"      = "#C45C3A",
  "Marine reliance"       = "#9B6DC5",
  "Inland reliance"       = "#D4956A"
)

# Interval colours - three lightness steps per model colour
col_95 <- colorspace::lighten(model_colors, 0.65)
col_80 <- colorspace::lighten(model_colors, 0.35)
col_50 <- colorspace::lighten(model_colors, 0.05)

# Build plotting data
df_plot <- slopes_df %>%
  mutate(
    clean_name = clean_model_name(model),
    nonzero    = (lo95 < 0 & hi95 < 0) | (lo95 > 0 & hi95 > 0),
    suffix     = sub(".*_", "", model),
    suffix_num = suppressWarnings(as.numeric(suffix))
  ) %>%
  filter(
    str_detect(clean_name, "reliance"),
    model != "animal_cons_cap"
  ) %>%
  arrange(suffix_num, suffix) %>%
  mutate(
    clean_name = factor(clean_name, levels = unique(clean_name)),
    col95      = coalesce(col_95[as.character(clean_name)], "grey80"),
    col80      = coalesce(col_80[as.character(clean_name)], "grey60"),
    col50      = coalesce(col_50[as.character(clean_name)], "grey40"),
    panel_id   = as.numeric(factor(clean_name))
  )

# Shared symmetric x-axis limits so zero lines align across all panels
x_limit <- max(abs(c(df_plot$lo95, df_plot$hi95)), na.rm = TRUE) * 1.05
x_limit <- ceiling(x_limit * 10) / 10

panel_names <- levels(df_plot$clean_name)

# Function to draw a single credible interval panel
# left_col = TRUE shifts the tag inward to clear the y-axis label text
make_panel <- function(df, show_yaxis = FALSE, title = "", tag = "",
                       left_col = FALSE) {

  df_sig  <- df %>% filter(nonzero)
  df_nsig <- df %>% filter(!nonzero)

  # Left-column panels have wide y-axis text; nudge the tag right so it
  # doesn't sit directly on top of the longest label
  tag_pos <- if (left_col) c(0.52, 1) else c(0.01, 1)

  ggplot(df, aes(y = var)) +
    geom_segment(aes(x = lo95, xend = hi95, color = col95),
                 linewidth = 2.2, lineend = "butt", alpha = 1) +
    geom_segment(aes(x = lo80, xend = hi80, color = col80),
                 linewidth = 2.2, lineend = "butt", alpha = 1) +
    geom_segment(aes(x = lo50, xend = hi50, color = col50),
                 linewidth = 2.2, lineend = "butt", alpha = 1) +
    geom_point(aes(x = med), shape = 21, size = 1.6,
               fill = "white", color = "grey20", stroke = 0.4) +

    geom_vline(xintercept = 0, linetype = "dashed",
               linewidth = 0.35, color = "grey40") +
    scale_color_identity() +
    scale_x_continuous(limits = c(-x_limit, x_limit)) +
    scale_y_discrete(labels = var_labels) +
    labs(title = title, tag = tag) +
    theme_minimal(base_size = 8) +
    theme(
      plot.title         = element_text(size = 5.5, face = "bold", hjust = 0.5,
                                        margin = margin(b = 2)),
      plot.tag           = element_text(size = 7, face = "bold"),
      plot.tag.position  = tag_pos,
      panel.grid.major.y = element_blank(),
      panel.grid.minor   = element_blank(),
      panel.grid.major.x = element_line(
        color = scales::alpha("black", 0.12), linewidth = 0.3
      ),
      axis.text.x        = element_text(size = 5.5, color = "black"),
      axis.title         = element_blank(),
      plot.margin        = margin(4, 6, 4, 6),
      axis.text.y        = if (show_yaxis) element_text(size = 5.5, color = "black",
                                                         hjust = 1) else element_blank(),
      axis.ticks.y       = element_blank()
    )
}

# Build all seven panels
# Panels 1 and 4 are left-column (show y-axis + shifted tag)
# Panel 7 is the lone bottom panel - also left-column for y-axis and tag
panels <- lapply(seq_along(panel_names), function(i) {
  name     <- panel_names[i]
  df_sub   <- df_plot %>% filter(clean_name == name)
  show_y   <- i %in% c(1, 4, 7)
  left_col <- i %in% c(1, 4, 7)
  make_panel(df_sub,
             show_yaxis = show_y,
             left_col   = left_col,
             title      = name,
             tag        = letters[i])
})

# Centered x-axis label placed below the grid
xlab_plot <- ggplot() +
  theme_void() +
  annotate("text", x = 0.5, y = 0.5,
           label    = "Posterior slope estimate",
           size     = 2.4,
           fontface = "bold",
           hjust    = 0.5,
           vjust    = 0.5)

empty <- ggplot() + theme_void()

# Rows 1 and 2 are full 3-panel; row 3 has one centered panel flanked by
# empty spacers - widths kept equal so the single panel aligns to column 2
row1 <- panels[[1]] | panels[[2]] | panels[[3]]
row2 <- panels[[4]] | panels[[5]] | panels[[6]]
row3 <- (empty | panels[[7]] | empty) + plot_layout(widths = c(.0002, 1, 1.118))

grid_plot <- wrap_plots(row1, row2, row3, ncol = 1) +
  plot_layout(heights = c(1, 1, 1)) &
  theme(plot.margin = margin(3, 4, 3, 4))

# Stack grid above centered x-axis label
final_plot <- cowplot::plot_grid(
  grid_plot,
  xlab_plot,
  ncol        = 1,
  rel_heights = c(1, 0.07)
)

ggsave(
  filename = paste0("../images/SI/posteriors_partial_slopes_SI.pdf"),
  plot     = final_plot,
  device   = cairo_pdf,
  width    = 5,
  height   = 3,
  units    = "in"
)
```


## SI Figure: Collinearity check

Create a correlation matrix for the explanatory variables to test for collinearity. On the plot, the upper triangle represents the shade of correlation on the color scale, while the bottom triangle displays the correlation between variables.
 
```{r collinearity check}
# Clean variable names
var_labels <- c(
  coastline_km               = "Coastline length (km)",
  gdp_trade                  = "GDP from trade (%)",
  inundation_area_log1000km2 = "Inundated area (log 1000 km²)",
  log_gdp                    = "GDP (log scale)",
  pop_10k_pct_coast          = "Pop. within 10 km of coast (%)",
  supermarkets               = "Supermarkets per 100,000",
  eez_total                  = "Log EEZ area km²"
)

# Use the realm model setup
model_type = "realm"

# Read in explanatory variables to assess collinearity
X <- readRDS(file.path(base_path, paste0("X_", response_vars[i], ".rds")))

# Rename columns to clean labels before computing correlation
# so axis-adjacent labels are publication-ready throughout
colnames(X) <- var_labels[colnames(X)]

# Create correlation matrix
cor_mat <- cor(X, use = "pairwise.complete.obs")

# Helper functions to split matrix into triangles and diagonal
get_lower_tri <- function(m) { m[upper.tri(m)] <- NA; m }
get_upper_tri <- function(m) { m[lower.tri(m)] <- NA; m }
get_diag      <- function(m) { m[!diag(nrow(m))] <- NA; m }

# Melt each part
lower_melt <- melt(get_lower_tri(cor_mat), na.rm = TRUE)
upper_melt <- melt(get_upper_tri(cor_mat), na.rm = TRUE) %>%
  filter(Var1 != Var2) %>%
  mutate(label     = sprintf("%.2f", value))
diag_melt  <- melt(get_diag(cor_mat), na.rm = TRUE) %>%
  mutate(label = str_wrap(as.character(Var1), width = 14))

# Collinearity plot
plot <- ggplot() +

  # Lower triangle - filled heatmap tiles
  geom_tile(
    data  = lower_melt,
    aes(x = Var2, y = Var1, fill = value),
    color = "white",
    linewidth = 0.4
  ) +

  # Diagonal - white tiles with border, holds variable name labels
  geom_tile(
    data      = diag_melt,
    aes(x = Var2, y = Var1),
    fill      = "white",
    color     = "grey40",
    linewidth = 0.4
  ) +

  # Upper triangle - correlation values, coloured to match heatmap
  geom_text(
    data  = upper_melt,
    aes(x = Var2, y = Var1, label = label, color = value),
    size  = 2.2,
    fontface = "plain"
  ) +

  # Diagonal variable name labels
  geom_text(
    data     = diag_melt,
    aes(x = Var2, y = Var1, label = label),
    size     = 1.9,
    fontface = "bold",
    color    = "grey15",
    lineheight = 0.9
  ) +

  # Fill scale - aligned with project diverging palette
  scale_fill_gradient2(
    low      = "#2B6AA8",   # blue - strong negative correlation
    mid      = "#F7F7F7",   # near-white midpoint
    high     = "#B2182B",   # red - strong positive correlation
    midpoint = 0,
    limits   = c(-1, 1),
    name     = "Pearson\ncorrelation",
    guide    = guide_colorbar(
      title.position = "top",
      title.hjust    = 0.5,
      barheight      = unit(0.88, "npc"),
      barwidth       = unit(0.35, "cm"),
      frame.colour   = "grey40",
      ticks.colour   = "grey40"
    )
  ) +

  # Text colour scale - matches fill so values read clearly against tiles
  scale_color_gradient2(
    low      = "#2B6AA8",
    mid      = "#888888",   # mid-grey so near-zero values are still legible
    high     = "#B2182B",
    midpoint = 0,
    limits   = c(-1, 1),
    guide    = "none"
  ) +

  coord_fixed() +
  scale_x_discrete(limits = rev) +
  scale_y_discrete(limits = levels(factor(lower_melt$Var1))) +
  theme_minimal(base_size = 8) +
  theme(
    axis.title        = element_blank(),
    axis.text         = element_blank(),
    axis.ticks        = element_blank(),
    panel.grid        = element_blank(),
    legend.title      = element_text(size = 6.5, face = "bold"),
    legend.text       = element_text(size = 6),
    plot.background   = element_rect(fill = "transparent", color = NA),
    panel.background  = element_rect(fill = "transparent", color = NA),
    plot.margin       = margin(4, 4, 4, 4)
  )

# Save
ggsave(
  filename = "../images/SI/collinearity_plot_SI.pdf",
  plot     = plot,
  device   = cairo_pdf,
  width    = 5,
  height   = 4,
  units    = "in"
)
```

# Bayesian model assessment

## Model convergence checks

### Trace Plot

```{r trace plot}
# Create separate datasets that combines Markov chains
trace_df <- data.frame()

for (i in seq_along(response_vars))  {
  
  # Load in Stan data
  fit <- rstan::extract(readRDS(file.path(base_path, paste0("fit_", response_vars[i], ".rds"))), permuted = T)
  
  chain_1 <- data.frame(beta = fit$beta[1:3000,1],
           beta_0 = fit$beta_0[1:3000],
           sigma = fit$sigma[1:3000],
           beta_0i = fit$beta_0i[1:3000,1],
           model = response_vars[i],
           iteration = 1:3000,
           chain = "1")
  
  chain_2 <- data.frame(beta = fit$beta[3001:6000,1],
           beta_0 = fit$beta_0[3001:6000],
           sigma = fit$sigma[3001:6000],
           beta_0i = fit$beta_0i[3001:6000,1],
           model = response_vars[i],
           iteration = 1:3000,
           chain = "2")
  
  chain_3 <- data.frame(beta = fit$beta[6001:9000,1],
           beta_0 = fit$beta_0[6001:9000],
           sigma = fit$sigma[6001:9000],
           beta_0i = fit$beta_0i[6001:9000,1],
           model = response_vars[i],
           iteration = 1:3000,
           chain = "3")
  
  chains <- chain_1 %>% bind_rows(chain_2) %>% bind_rows(chain_3)
  
  trace_df <- bind_rows(trace_df, chains)
    
}

plot <- trace_df %>%
  pivot_longer(cols = -c(model, iteration, chain), names_to = "parameter") %>%
  ggplot(aes(x = iteration, y = value, color = chain)) +
  geom_line(linewidth = 0.05) +
  scale_color_viridis_d(option = "plasma", end = 0.95) +
  facet_grid(model ~ parameter) +
  guides(color = "none") +
  labs(x = "Parameter", y = "Parameter value") +
  theme_pubclean()

ggsave(
  filename = paste0("../images/SI/trace.pdf"),
  plot     = plot,
  device   = cairo_pdf,
  width    = 9,
  height   = 5,
  units    = "in"
)

```

> Result: All parameters have overlap on the Markov Chain, indicating model convergence


### R hat

```{r r hat}
library(dplyr)
library(rstan)

# Initialize empty data frame
rhat_df <- data.frame()

# Parameters we care about for convergence
params_to_check <- c("beta_0", "beta", "sigma", "beta_0i", "sigma_beta_0", "lp__")

for (i in seq_along(response_vars)) {
  # Load the fitted model
  fit <- readRDS(file.path(base_path, paste0("fit_", response_vars[i], ".rds")))
  
  # Get summary for only the parameters of interest
  fit_summary <- rstan::summary(fit, pars = params_to_check)$summary
  
  # Extract Rhat values
  vals <- fit_summary[, "Rhat"]
  
  # Extract parameter names
  param_names <- rownames(fit_summary)
  
  # Combine into a data frame
  new_data <- data.frame(
    rhat = vals,
    param = param_names,
    model = response_vars[i],
    stringsAsFactors = FALSE
  )
  
  # Append to main data frame
  rhat_df <- bind_rows(rhat_df, new_data)
}

# Optional: view summary
summary(rhat_df$rhat)

# Create rhat plot
(plot <- rhat_df %>%
    mutate(clean_name = clean_model_name(model)) %>%
    filter(
    model != "animal_cons_cap",
    str_detect(clean_name, "capita")         |
      str_detect(clean_name, "Aquatic reliance") |
      str_detect(clean_name, "reliance")
  ) %>%
  ggplot(aes(x = rhat)) +
  geom_histogram() +
  facet_wrap(~ model) +
  labs(x = "R-hat", y = "# parameters") +
  theme_pubclean() +
    theme(axis.text = element_text(size = 7)))
```

> Result: No R hat values exceed 1.01 - The Markov Chains have converged

### Divergent transitions

```{r divergent transitions}
# Rename the divergent transition column names to model names
divergent_transitions <- matrix(nrow = 3, ncol = length(response_vars))
colnames(divergent_transitions) <- response_vars

for (i in seq_along(response_vars)) {
  
  # Extract number of divergent samples
  sampler_params <- get_sampler_params(readRDS(file.path(base_path, paste0("fit_", response_vars[i], ".rds"))), inc_warmup = FALSE)
  num_transitions <- sapply(sampler_params, function(x) sum(x[, "divergent__"]))
  
  divergent_transitions[,i] <- num_transitions
  
}

# Create dataframe for plotting divergent transitions
plot_df <- divergent_transitions %>%
  as.data.frame() %>%
  mutate(chain = as.character(1:3)) %>%
  pivot_longer(!"chain", names_to = "model", values_to = "divergent") %>%
  mutate(
    suffix = sub(".*_", "", model)
  ) %>%
  mutate(clean_name = clean_model_name(model)) %>%
    filter(
    model != "animal_cons_cap",
    str_detect(clean_name, "capita")         |
      str_detect(clean_name, "Aquatic reliance") |
      str_detect(clean_name, "reliance"))

# compute the order of model names by suffix
ordered_names <- plot_df %>%
  distinct(model, suffix) %>%
  arrange(suffix) %>%
  pull(model)

# Plot # divergent transitions
(plot <- plot_df %>%
    
  mutate(model = factor(model, levels = ordered_names)) %>%
  ggplot(aes(y = model, x = divergent, fill = chain)) +
  geom_col() +
  scale_fill_viridis_d(end = 0.8) +
  labs(y = "Model", x = "# divergent transitions") +
  theme_light() +
    theme(legend.position = "bottom"))

plot_df %>% group_by(model) %>% summarise(sum = sum(divergent))
``` 

> Result: Out of 3000 samples, the highest number of divergent transitions was 5 - divergence is not an issue. 

## Model fit checks

### Generate PPD

This code chunk will only run if not run before as it takes awhile to run. This plots the posterior predictive distributions (PPDs) across each model used in the paper. 

```{r extract ppd}
if (!file.exists(file.path(base_path,"ppd.parquet"))) {

  # Load function to clean model output names to publication-friendly labels

  # Extract posterior predictive draws for a single model
  # model_name is used consistently for all file reads - no global i dependency
  extract_ppd <- function(model_name, model_type) {

    # Load fitted Stan model and extract posterior samples
    fit   <- readRDS(file.path(base_path, paste0("fit_",    model_name, ".rds")))
    x     <- rstan::extract(fit, permuted = TRUE)
    y_pred <- x$y_pred   # S × N matrix of posterior predictive draws

    # Load design matrix, response, and realm labels for this model
    X           <- readRDS(file.path(base_path, paste0("X_",      model_name, ".rds")))
    y           <- readRDS(file.path(base_path, paste0("y_",      model_name, ".rds")))
    realm_levels <- readRDS(file.path(base_path, paste0("realms_", model_name, ".rds")))

    N <- length(y)

    # Validate that realm labels match the number of observations
    if (length(realm_levels) != N) {
      stop(paste0("realm_levels length mismatch for model: ", model_name,
                  " (expected ", N, ", got ", length(realm_levels), ")"))
    }

    # Reshape to long format: one row per draw × observation
    ppd_long <- as.data.frame(y_pred) %>%
      mutate(draw = row_number()) %>%
      pivot_longer(-draw, names_to = "obs", values_to = "ppd") %>%
      mutate(
        obs    = as.integer(gsub("V", "", obs)),
        region = realm_levels[obs],
        y_obs  = y[obs],
        model  = model_name
      )

    return(ppd_long)
  }

  # Extract PPD for all models and combine
  ppd_all_models <- map_dfr(response_vars, ~ extract_ppd(.x, model_type)) %>%
    mutate(clean_name = clean_model_name(model))

  # Calculate posterior predictive means per observation
  ppd_means <- ppd_all_models %>%
    group_by(model, obs, region) %>%
    summarise(
      ppd_mean = mean(ppd),
      y_obs    = first(y_obs),
      .groups  = "drop"
    )

  # Save combined posterior predictive dataset
  write_parquet(
    ppd_all_models,
    file.path(base_path,"ppd.parquet")
  )

} else {
  message("Posterior predictive values have already been generated - skipping model posterior predictive acquisition")
}
```

### Create PPD Plot

```{r plot ppd}
# Read in posterior predictive distribution data
ppd_all_models <- read_parquet(file.path(base_path,"ppd.parquet"))

# Sample 30 posterior draws to plot - enough to show the envelope without
# overplotting, consistent across all panels
draw_sample <- sample(unique(ppd_all_models$draw), size = 30)

# Filter to models of interest and sampled draws
df_plot <- ppd_all_models %>%
  filter(
    model != "animal_cons_cap",
    str_detect(clean_name, "capita")         |
      str_detect(clean_name, "Aquatic reliance") |
      str_detect(clean_name, "reliance")
  ) %>%
  filter(draw %in% draw_sample) %>%
  mutate(clean_name = factor(clean_name))

panel_names <- levels(df_plot$clean_name)

# 4×4 layout - 15 panels means bottom row has 3 panels centered with spacers
# Panels are ordered sequentially by clean_name factor level (A–O)
# Left column: panels 1, 5, 9, 13 - show y-axis title
# Bottom row:  panels 13, 14, 15  - show x-axis title
# Left column also gets inward-nudged tag to clear y-axis label text

# Function to draw a single PPD panel
# title    = model name shown above panel
# tag      = uppercase letter (A–O)
# left_col = TRUE nudges tag right and shows y-axis title
# bot_row  = TRUE shows x-axis title
make_ppd_panel <- function(df, title = "", tag = "",
                           left_col = FALSE, bot_row = FALSE) {

  tag_pos <- if (left_col) c(0.18, 1) else c(0.01, 1)

  p <- ggplot(df, aes(x = ppd, group = draw)) +

    # Posterior predictive envelope - faint coloured lines show spread
    geom_density(
      colour    = "#7BAEC4",
      linewidth = 0.25
    ) +

    # Observed data density - drawn on top, clearly distinct
    geom_density(
      aes(x = y_obs, group = NULL),
      colour    = "grey20",
      linewidth = 0.7
    ) +

    labs(title = title, tag = tag) +
    theme_minimal(base_size = 8) +
    theme(
      text              = element_text(family = "sans", color = "black"),
      plot.background   = element_rect(fill = "transparent", color = NA),
      panel.background  = element_rect(fill = "transparent", color = NA),
      panel.grid.minor  = element_blank(),
      panel.grid.major.x = element_blank(),
      panel.grid.major.y = element_line(
        color = scales::alpha("black", 0.12), linewidth = 0.3
      ),
      plot.title        = element_text(size = 4, face = "bold", hjust = 0.5,
                                        margin = margin(b = 2)),
      plot.tag          = element_text(size = 7, face = "bold"),
      plot.tag.position = tag_pos,
      axis.text         = element_text(size = 5.5, color = "black"),
      axis.title.x      = element_blank(),
      axis.title.y      = element_blank(),
      plot.margin       = margin(3, 4, 3, 4)
    )

  if (left_col) p <- p +
    labs(y = "Density") +
    theme(axis.title.y = element_text(size = 6.5, face = "bold",
                                       margin = margin(r = 3)))

  if (bot_row) p <- p +
    labs(x = "Response value (standardised)") +
    theme(axis.title.x = element_text(size = 5, face = "bold",
                                       margin = margin(t = 3)))
  p
}

# Build all 15 panels in sequential order (A–O)
panels <- lapply(seq_along(panel_names), function(i) {
  nm     <- panel_names[i]
  df_sub <- df_plot %>% filter(clean_name == nm)
  make_ppd_panel(
    df_sub,
    title    = nm,
    tag      = letters[i],
    left_col = i %in% c(1, 5, 9, 13),
    bot_row  = i %in% c(13, 14, 15)
  )
})

# Assemble 4×4 grid - bottom row centered with half-column empty spacers
empty <- ggplot() + theme_void()

row1 <- panels[[1]]  | panels[[2]]  | panels[[3]]  | panels[[4]]
row2 <- panels[[5]]  | panels[[6]]  | panels[[7]]  | panels[[8]]
row3 <- panels[[9]]  | panels[[10]] | panels[[11]] | panels[[12]]
row4 <- (empty | panels[[13]] | panels[[14]] | panels[[15]] | empty) +
  plot_layout(widths = c(0.5, 1, 1, 1, 0.5))

final_plot <- wrap_plots(row1, row2, row3, row4, ncol = 1) +
  plot_layout(heights = c(1, 1, 1, 1)) &
  theme(plot.margin = margin(2, 3, 2, 3))
```

### PSIS-LOO model performance

To see how well each model performs on unseen data, we can run a cross validation on each model. Additionally, we can calculate the Pareto k values to see if there are influential points that may have an impact on the LOO values.

```{r PSIS-LOO plot}
# Paths and model info
combined_model_path <- base_path
n_models <- length(response_vars)

# Extract log-likelihood matrices for all models
log_lik_list <- vector("list", n_models)

for (i in seq_along(response_vars)) {
  fit <- readRDS(file.path(combined_model_path, paste0("fit_", response_vars[i], ".rds")))
  # fit <- readRDS(paste0("../bayesian output backup/var_all/fit_",response_vars[i],".rds"))
  
  # Extract log-likelihood matrix; make sure your Stan model has 'log_lik' in generated quantities
  log_lik_list[[i]] <- rstan::extract(fit, permuted = TRUE)$log_lik
}

names(log_lik_list) <- response_vars

# Compute LOO for each model
loo_list <- vector("list", n_models)

for (i in seq_along(log_lik_list)) {
  loo_list[[i]] <- loo(log_lik_list[[i]], cores = 2)
}

names(loo_list) <- response_vars


# Compare models using ELPD
loo_compare_matrix <- loo_compare(loo_list)

model_data <- as.data.frame(loo_compare_matrix) %>%
  mutate(model = rownames(loo_compare_matrix)) %>%
  mutate(clean_name = clean_model_name(model)) %>%
  select(clean_name, elpd_diff, se_diff) %>%
  mutate(
    lo95 = elpd_diff - 2 * se_diff,
    hi95 = elpd_diff + 2 * se_diff,
    # Flag models whose CI excludes zero (significantly worse than best)
    nonzero = (lo95 < 0 & hi95 < 0) | (lo95 > 0 & hi95 > 0),
    # Best model (elpd_diff == 0) gets a distinct fill
    is_best = elpd_diff == 0
  ) %>%
  arrange(elpd_diff) %>%
  mutate(clean_name = factor(clean_name, levels = clean_name))


# Assign colours; fall back to grey for any model not in the palette
model_data <- model_data %>%
  mutate(
    pt_color = coalesce(model_colors[as.character(clean_name)], "grey50"),
    pt_fill  = if_else(is_best, "white", pt_color),
    # Non-significant intervals (CI crosses zero) rendered lighter
    bar_color = if_else(nonzero,
                        colorspace::lighten(pt_color, 0.25),
                        colorspace::lighten(pt_color, 0.60))
  )

# create plot
psis_plot <- ggplot(model_data,
                    aes(y = clean_name, x = elpd_diff)) +

  # +- 2 SE error bars
  geom_errorbarh(
    aes(xmin = lo95, xmax = hi95),
    height    = 0.35,
    linewidth = 0.55
  ) +

  # Point estimates
  geom_point(fill = "black",
             shape  = 21,
    size   = 2.2,
    stroke = 0.55
  ) +

  # Reference line at zero (best model)
  geom_vline(xintercept = 0, linetype = "dashed",
             linewidth = 0.35, color = "grey40") +

  scale_color_identity() +
  scale_fill_identity() +

  labs(
    x   = "ELPD difference (higher = better)",
    y   = "Model"
  ) +

  theme_minimal(base_size = 8) +
  theme(
    plot.tag           = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_line(
      color     = scales::alpha("black", 0.12),
      linewidth = 0.3
    ),
    axis.text.x  = element_text(size = 5.5, color = "black"),
    axis.text.y  = element_text(size = 5.5, color = "black", hjust = 1),
    axis.title.x = element_text(size = 6, face = "bold",
                                margin = margin(t = 4)),
    axis.ticks.y = element_blank(),
    plot.margin  = margin(6, 8, 4, 6)
  )
```

### Pareto k

#### SI Figure: PPD / LOO / Pareto K

Pareto k is used to assess the validity of PSIS-LOO values. We ideally want these values to all be below 0.7.

```{r Pareto k}
# Pareto k diagnostic plot
pareto_k_df <- bind_rows(
  lapply(seq_along(loo_list), function(i) {
    data.frame(
      model    = names(loo_list)[i],
      point    = seq_along(loo_list[[i]]$diagnostics$pareto_k),
      pareto_k = loo_list[[i]]$diagnostics$pareto_k
    )
  }),
  .id = "model_index"
) %>%
  mutate(
    clean_name = clean_model_name(model),
    flag       = pareto_k > 0.7,
    # Threshold bands for background shading
    band = case_when(
      pareto_k <= 0.5 ~ "good",
      pareto_k <= 0.7 ~ "ok",
      TRUE            ~ "bad"
    )
  )

# Load country labels
model_countries <- readRDS("../data/model_countries")


# pareto_k_df with country labels
pareto_k_df <- bind_rows(
  lapply(seq_along(loo_list), function(i) {
    data.frame(
      model    = names(loo_list)[i],
      point    = seq_along(loo_list[[i]]$diagnostics$pareto_k),
      pareto_k = loo_list[[i]]$diagnostics$pareto_k,
      country  = model_countries   # ISO3 label per observation
    )
  }),
  .id = "model_index"
) %>%
  mutate(
    clean_name = clean_model_name(model),
    flag       = pareto_k > 0.7
  )

# Facet order matches ELPD ranking
facet_order <- levels(model_data$clean_name)
pareto_k_df <- pareto_k_df %>%
  mutate(clean_name = factor(clean_name,
                             levels = if (length(facet_order) > 0) rev(facet_order)
                                      else unique(clean_name)))

pt_ok  <- "grey60"
pt_bad <- "#C45C3A"

# Build facet tag data — one row per facet, letters assigned in horizontal order
# facet_wrap with ncol = 3 fills left-to-right, top-to-bottom
facet_levels <- levels(pareto_k_df$clean_name)  # already in your desired order
n_facets     <- length(facet_levels)

facet_tags <- pareto_k_df %>%
  group_by(clean_name) %>%
  summarise(
    x_max    = max(point),
    x_min    = min(point),
    y_max    = max(pareto_k),
    .groups  = "drop"
  ) %>%
  # Match factor order so letters align with panel positions
  mutate(
    clean_name = factor(clean_name, levels = facet_levels),
    tag        = letters[as.numeric(clean_name)]
  )

pareto_plot <- ggplot(pareto_k_df,
                      aes(x = point, y = pareto_k)) +

  geom_hline(yintercept = 0.5, linetype = "dotted",
             linewidth = 0.3, color = "grey55") +
  geom_hline(yintercept = 0.7, linetype = "dashed",
             linewidth = 0.35, color = "grey30") +

  geom_point(data = filter(pareto_k_df, !flag),
             color = pt_ok, size = 0.7, alpha = 0.55, shape = 16) +
  geom_point(data = filter(pareto_k_df,  flag),
             color = pt_bad, size = 1.1, alpha = 0.85, shape = 16) +

  ggrepel::geom_text_repel(
    data             = filter(pareto_k_df, flag),
    aes(label        = country),
    color            = pt_bad,
    size             = 1.8,
    fontface         = "bold",
    segment.color    = pt_bad,
    segment.size     = 0.25,
    segment.alpha    = 0.6,
    box.padding      = 0.25,
    point.padding    = 0.15,
    min.segment.length = 0.2,
    max.overlaps     = Inf,
    force            = 2
  ) +

  facet_wrap(~ clean_name, scales = "free_x", ncol = 3) +

  # Per-facet letter tags — top-left, using -Inf/Inf to anchor to panel edges
  geom_text(
    data        = facet_tags,
    aes(x = -Inf, y = Inf, label = tag),
    inherit.aes = FALSE,
    hjust       = -0.3,   # small inset from left edge
    vjust       = 1.4,    # small inset from top edge
    size        = 2.5,
    fontface    = "bold",
    color       = "black"
  ) +

  # k = 0.7 annotation — top-right, anchored to each panel's x_max
  geom_text(
    data        = facet_tags,
    aes(x = Inf, y = 1, label = "k = 0.7"),
    inherit.aes = FALSE,
    hjust       = 1.1,
    vjust       = 0,
    size        = 1.8,
    color       = "grey30",
    fontface    = "italic"
  ) +

  scale_y_continuous(
    breaks = c(0, 0.5, 0.7, 1),
    labels = c("0", "0.5", "0.7", "1")
  ) +

  labs(
    x = "Observation index",
    y = expression(paste("Pareto ", hat(k)))
  ) +

  theme_minimal(base_size = 8) +
  theme(
    plot.tag            = element_blank(),
    strip.text          = element_text(size = 5.5, face = "bold", hjust = 0.5,
                                       margin = margin(b = 2)),
    panel.grid.major.y  = element_line(color = scales::alpha("black", 0.10),
                                       linewidth = 0.25),
    panel.grid.minor    = element_blank(),
    panel.grid.major.x  = element_blank(),
    axis.text.x         = element_text(size = 4.5, color = "black"),
    axis.text.y         = element_text(size = 5, color = "black"),
    axis.title.x        = element_text(size = 6, face = "bold", margin = margin(t = 4)),
    axis.title.y        = element_text(size = 6, face = "bold", margin = margin(r = 4)),
    axis.ticks          = element_blank(),
    panel.spacing       = unit(4, "pt"),
    plot.margin         = margin(6, 8, 4, 6)
  )

# Combine PSIS and Pareto k as panels A and B

# Add panel tags directly in each plot's labs()
psis_plot   <- psis_plot   + labs(tag = "A")
pareto_plot <- pareto_plot + labs(tag = "B") +
  theme(plot.tag.position = c(0.01, 1))  # consistent with psis_plot

# Stack side by side with patchwork
combined_plot <- plot_grid(psis_plot, pareto_plot, rel_widths = c(1,1.5), labels = c("B", "C"))



final <- plot_grid(final_plot, combined_plot, nrow = 1, labels = "A", rel_widths = c(1,1.6))

ggsave(
  filename = "../images/SI/LOO_diagnostics_SI.pdf",
  plot     = final,
  device   = cairo_pdf,
  width    = 15,   # wider to accommodate both panels comfortably
  height   = 4,
  units    = "in"
)
```

> Result: Terrestrial animal conusmption performs substantially better in model performance than all the aquatic animal consumption & reliance models.

## Sensitivity Testing

### Main paper posteriors but using only the coastline variable. This is run as a sensitivity testing measure as for whether we can include both coastline and EEZ area. 

```{r hierarcical partial slope posteriors}
# Get file path for Bayesian model with the coastline length only variable model
separated_model_path <- "../output/bayesian_output/random_effect_realm/var_coastline_km"

# Create a data frame of posteriors of partial slopes across Bayesian models
slopes_df <- data.frame()

for (i in seq_along(response_vars)) {

  # Extract MCMC samples
  x <- rstan::extract(
    readRDS(file.path(separated_model_path, paste0("fit_", response_vars[i], ".rds"))),
    permuted = TRUE
  )
  beta_global <- x$beta
  X           <- readRDS(file.path(separated_model_path, paste0("X_", response_vars[i], ".rds")))

  n_iter <- nrow(beta_global)
  K      <- ncol(beta_global)
  vars   <- colnames(X)

  # Extract global slopes and calculate credibility intervals
  ci_df <- expand_grid(iter = 1:n_iter, k = 1:K) %>%
    mutate(
      value = map2_dbl(iter, k, ~ beta_global[.x, .y]),
      var   = vars[k]
    ) %>%
    group_by(var) %>%
    summarise(
      lo95 = quantile(value, 0.025),
      lo80 = quantile(value, 0.10),
      lo50 = quantile(value, 0.25),
      med  = quantile(value, 0.50),
      hi50 = quantile(value, 0.75),
      hi80 = quantile(value, 0.90),
      hi95 = quantile(value, 0.975),
      .groups = "drop"
    ) %>%
    mutate(model = response_vars[i])

  slopes_df <- bind_rows(slopes_df, ci_df)
}

# Clean variable names for axis labels
var_labels <- c(
  coastline_km               = "Coastline length (km)",
  gdp_trade                  = "GDP from trade (%)",
  inundation_area_log1000km2 = "Inundated area (log 1000 km²)",
  log_gdp                    = "Log GDP per capita",
  pop_10k_pct_coast          = "Pop. within 10 km of coast (%)",
  supermarkets               = "Supermarkets per 100,000",
  eez_total                  = "Log EEZ area km²"
)

# Model colours - harmonized palette
model_colors <- c(
  "Terrestrial animal consumption per capita" = "#8C1A11",
  "Aquatic animal consumption per capita"     = "#2B6AA8",
  "Aquatic reliance"                          = "#1A7A6E",
  "Capture consumption per capita"            = "#D4601A",
  "Aquaculture consumption per capita"        = "#E8B000",
  "Domestic consumption per capita"           = "#7BAEC4",
  "Foreign consumption per capita"            = "#C45C3A",
  "Marine consumption per capita"             = "#9B6DC5",
  "Inland consumption per capita"             = "#D4956A"
)

# Interval colours - three lightness steps per model colour
col_95 <- colorspace::lighten(model_colors, 0.65)
col_80 <- colorspace::lighten(model_colors, 0.35)
col_50 <- colorspace::lighten(model_colors, 0.05)

# Build plotting data
df_plot <- slopes_df %>%
  mutate(
    clean_name = clean_model_name(model),
    nonzero    = (lo95 < 0 & hi95 < 0) | (lo95 > 0 & hi95 > 0),
    suffix     = sub(".*_", "", model),
    suffix_num = suppressWarnings(as.numeric(suffix))
  ) %>%
  filter(
    str_detect(clean_name, "capita") | str_detect(clean_name, "Aquatic animal reliance"),
    !model == "animal_cons_cap"
  ) %>%
  arrange(suffix_num, suffix) %>%
  mutate(
    clean_name = factor(clean_name, levels = unique(clean_name)),
    col95 = coalesce(col_95[as.character(clean_name)], "grey80"),
    col80 = coalesce(col_80[as.character(clean_name)], "grey60"),
    col50 = coalesce(col_50[as.character(clean_name)], "grey40"),
    panel_id = as.numeric(factor(clean_name))
  )

# Shared symmetric x-axis limits so zero lines align across all panels
x_limit <- max(abs(c(df_plot$lo95, df_plot$hi95)), na.rm = TRUE) * 1.05
x_limit <- ceiling(x_limit * 10) / 10

panel_names <- levels(df_plot$clean_name)

# Function to draw a single credible interval panel
# left_col = TRUE shifts the tag inward to clear the y-axis label text
make_panel <- function(df, show_yaxis = FALSE, title = "", tag = "",
                       left_col = FALSE) {

  df_sig  <- df %>% filter(nonzero)
  df_nsig <- df %>% filter(!nonzero)

  # Left-column panels have wide y-axis text; nudge the tag right so it
  # doesn't sit directly on top of the longest label
  tag_pos <- if (left_col) c(0.52, 1) else c(0.01, 1)

  ggplot(df, aes(y = var)) +
    geom_segment(aes(x = lo95, xend = hi95, color = col95),
                 linewidth = 2.2, lineend = "butt", alpha = 1) +
    geom_segment(aes(x = lo80, xend = hi80, color = col80),
                 linewidth = 2.2, lineend = "butt", alpha = 1) +
    geom_segment(aes(x = lo50, xend = hi50, color = col50),
                 linewidth = 2.2, lineend = "butt", alpha = 1) +
    geom_point(aes(x = med), shape = 21, size = 1.6,
               fill = "white", color = "grey20", stroke = 0.4) +

    geom_vline(xintercept = 0, linetype = "dashed",
               linewidth = 0.35, color = "grey40") +
    scale_color_identity() +
    scale_x_continuous(limits = c(-x_limit, x_limit)) +
    scale_y_discrete(labels = var_labels) +
    labs(title = title, tag = tag) +
    theme_minimal(base_size = 8) +
    theme(
      plot.title         = element_text(size = 3.5, face = "bold", hjust = 0.5,
                                        margin = margin(b = 2)),
      plot.tag           = element_text(size = 7, face = "bold"),
      plot.tag.position  = tag_pos,
      panel.grid.major.y = element_blank(),
      panel.grid.minor   = element_blank(),
      panel.grid.major.x = element_line(
        color = scales::alpha("black", 0.12), linewidth = 0.3
      ),
      axis.text.x        = element_text(size = 5.5, color = "black"),
      axis.title         = element_blank(),
      plot.margin        = margin(4, 6, 4, 6),
      axis.text.y        = if (show_yaxis) element_text(size = 5.5, color = "black",
                                                         hjust = 1) else element_blank(),
      axis.ticks.y       = element_blank()
    )
}

# Build all nine panels - panels 1, 4, 7 are left column (show y-axis + shifted tag)
panels <- lapply(seq_along(panel_names), function(i) {
  name     <- panel_names[i]
  df_sub   <- df_plot %>% filter(clean_name == name)
  show_y   <- i %in% c(1, 4, 7)
  left_col <- i %in% c(1, 4, 7)
  make_panel(df_sub,
             show_yaxis = show_y,
             left_col   = left_col,
             title      = name,
             tag        = letters[i])
})

# Centered x-axis label placed below the grid
xlab_plot <- ggplot() +
  theme_void() +
  annotate("text", x = 0.5, y = 0.5,
           label    = "Posterior slope estimate",
           size     = 2.4,
           fontface = "bold",
           hjust    = 0.5,
           vjust    = 0.5)

# Assemble 3×3 grid
grid_plot <- (
    (panels[[1]] | panels[[2]] | panels[[3]]) /
    (panels[[4]] | panels[[5]] | panels[[6]]) /
    (panels[[7]] | panels[[8]] | panels[[9]])
  ) +
  plot_layout(heights = c(1, 1, 1)) &
  theme(plot.margin = margin(3, 4, 3, 4))

# Stack grid above centered x-axis label
final_plot <- cowplot::plot_grid(
  grid_plot,
  xlab_plot,
  ncol        = 1,
  rel_heights = c(1, 0.07)
)

ggsave(
  filename = "../images/SI/posteriors_partial_slopes_coastline_SI.pdf",
  plot     = final_plot,
  device   = cairo_pdf,
  width    = 5,
  height   = 3,
  units    = "in"
)
```

> Result: Posteriors between coastline & EEZ area model and the coastline only model are very similar in significance and direction.

### PSIS-LOOCV

With the collinearity matrix seen above, there are concerns that this collinearity might influence model performance. We checked posterior distribution shpaes in the previous code chunk. To additionally gague model performance and whether the collinearity is having a confounding effect on the results, we will compare the Pareto-smoothed importance sampling leave one out cross-validation expected log pointwise predictive density (PSIS-LOOCV) of the model with both coastline length and EEZ area and with the model including only coastline length.

```{r PSIS-LOOCV}
# Extract log-likelihood matrices for the combined coastline length-eez area model
combined_model_path <- base_path

# Extract log-likelihood matrices
log_lik_combined  <- rstan::extract(
  readRDS(file.path(combined_model_path,  paste0("fit_", response_vars[i], ".rds"))),
  permuted = TRUE
)$log_likr

log_lik_separated <- rstan::extract(
  readRDS(file.path(separated_model_path, paste0("fit_", response_vars[i], ".rds"))),
  permuted = TRUE
)$log_lik


# Compute LOO for each model
loo_combined  <- loo(log_lik_combined,  cores = 2)
loo_separated <- loo(log_lik_separated, cores = 2)

# Compare - positive elpd_diff favours the first model listed
loo_compare(loo_combined, loo_separated)
```

> Result: Models are indistinguishable in performance looking at the LOO ELPD CV scores. And given the partial slopes do not differ substantially across models when using removing one variable, the collinearity between the coastline length and EEZ area is not 
