---
title: "98_Appendix"
author: "Connor Quiroz"
date: "2025-07-19"
output: html_document
---


```{r look at how aa reliance changes over time}
# Look at slopes between aa_reliancec and year
aa_reliance_slopes <- supply_importance_long %>%
  group_by(consumer_iso3c, year) %>%
  summarize(aa_reliance_pct = mean(aa_reliance_pct)) %>%
  split(.$consumer_iso3c) %>% 
  map(~lm(aa_reliance_pct~year, data = .x)) %>% 
  map_df(broom::tidy, .id = 'consumer_iso3c') %>%
  filter(term == 'year') %>%
  select(consumer_iso3c, "slope" = "estimate") %>%
  mutate(fill_color = case_when(
    slope > 0 ~ "1",
    slope < 0 ~ "0",
    TRUE ~ NA
  )) %>%
  add_region(col = "consumer_iso3c", region.col.name = "region")

aa_reliance_slopes %>%
  filter(!is.na(region)) %>%
  ggplot(aes(x = reorder(consumer_iso3c, -slope), 
             y = slope, fill = region)) +
  geom_col() +
  theme_minimal_hgrid() +
  theme(axis.text.x = element_blank()) +
  labs(x = "Consuming country", y = "Slope between aa reliance and time", fill = "Region") +
  scale_fill_manual(values = artis_palette(7)) +
  ylim(-0.005, 0.005)

# Violin plot by region of foreign dependency slopes
aa_reliance_slopes %>%
  filter(!is.na(region)) %>%
  ggplot(aes(x = region, y = slope, fill = region)) +
  geom_violin(alpha = 0.5) +
  geom_point() +
  scale_fill_manual(values = artis_palette(7)) +
  theme_minimal_hgrid(12) +
  labs(x = "Consuming country", y = "Slope between aa reliance and time") +
  guides(fill = "none")


# Look at individual countries' trends in aa contribution to protein
# Get list of countries
countries <- aa_reliance_slopes %>%
  arrange(-slope) %>%
  distinct(consumer_iso3c) %>%
  pull(consumer_iso3c)

# Create new directory for foreign dependencies
if (!dir.exists("../images/aa_reliance_trends/")) {
  dir.create("../images/aa_reliance_trends/")
}

for (i in countries) {
  
  # Generate data
  data <- supply_importance_long %>%
  # filter(consumer_iso3c == i, region == "Oceania") %>%
    filter(consumer_iso3c == i) %>%
    filter(str_detect(element, "Protein")) %>%
    group_by(year, consumer_iso3c)
  
  # Create image
  
  if (nrow(data) > 0) { # Check to see if filtered data has available observations
    img <- data  %>%
      group_by(year, consumer_iso3c) %>%
  ggplot(aes(x = year, y = aa_reliance_pct, color = consumption_source)) +
  geom_line() +
  geom_point() +
  facet_grid(habitat ~ method, scales = "free_x", space = "free_x") +
    theme_cowplot() +
    labs(x = "", y = "AA contribution to protein consumption", title = i, color = "Consumption source") +
    theme(plot.title = element_text(hjust = 0.5),
          axis.text.x = element_text(size = 7, angle = 45)) +
    geom_vline(xintercept = 2009, linetype = "dashed", color = "red")
  
  # Save image
  ggsave(paste0("../images/aa_reliance_trends/", which(i == countries), ".jpg"), plot = img, device = "jpeg", height = 4, width = 6, units = "in")
  }
  
}

aa_reliance_slopes %>%
  arrange(-slope)
```


```{r look at fao dataset discrepencies}
# Get list of FAO countries
countries <- fao_imputed %>%
  ungroup() %>%
  distinct(iso3c) %>%
  pull(iso3c)

# Create new directory for foreign dependencies
if (!dir.exists("../images/fao_dataset_comparison/")) {
  dir.create("../images/fao_dataset_comparison/")
}

for (i in countries) {
  
  # Generate data
  data <- fao_imputed %>%
    filter(iso3c == i) %>%
    filter(str_detect(Element, "Protein")) %>%
    group_by(year, iso3c)
  
  # Create image
  
  if (nrow(data) > 0) { # Check to see if filtered data has available observations
  img <- data %>%
  ggplot(aes(x = year, y = prop_aquatic_animal, color = new_historical)) +
  geom_line() +
  geom_point() +
  theme_cowplot() +
  scale_color_manual(values = c("#4254b8", "#EE5A54", "#52BEDB")) +
  labs(x = "", y = "Prop aquatic animal protein", color = "FAO FBS dataset", title = i) +
    theme(plot.title = element_text(hjust = 0.5))
  

  
  # Save image
  ggsave(paste0("../images/fao_dataset_comparison/", i, ".jpg"), plot = img, device = "jpeg", height = 4, width = 6, units = "in")
  }
  
}
```


```{r look at dependency on europe production & consumption}
# Total Europe consumption that comes from Europe
prod_eur_cons_eur <- consumption_full %>%
  exploreARTIS::add_region(col = "consumer_iso3c", region.col.name = "consumer_region") %>%
  exploreARTIS::add_region(col = "source_country_iso3c", region.col.name = "producer_region") %>%
  relocate(consumer_region, .before = "source_country_iso3c") %>%
  relocate(producer_region, .after = consumer_region) %>%
  group_by(consumer_iso3c, year) %>%
  filter(consumer_region == "Europe",
         producer_region == "Europe") %>%
  summarize(sum_weight = sum(consumption_live_t, na.rm = TRUE))
  
# Europe total consumption
eur_total_cons <- consumption_full %>%
  exploreARTIS::add_region(col = "consumer_iso3c", region.col.name = "consumer_region") %>%
  exploreARTIS::add_region(col = "source_country_iso3c", region.col.name = "producer_region") %>%
  relocate(consumer_region, .before = "source_country_iso3c") %>%
  relocate(producer_region, .after = consumer_region) %>%
  group_by(consumer_iso3c, year) %>%
  filter(consumer_region == "Europe") %>%
  summarize(total_cons = sum(consumption_live_t, na.rm = TRUE))

# Proportion of Europe consumption that is produced in Europe
prop_europe_prod_cons <- left_join(prod_eur_cons_eur, eur_total_cons) %>%
  group_by(consumer_iso3c, year) %>%
  summarize(prop_cons = sum_weight / total_cons)

# Look at average change in foreign dependency from 2015-2019
foreign_dependency_slopes <- prop_europe_prod_cons %>%
  split(.$consumer_iso3c) %>% 
  map(~lm(prop_cons~year, data = .x)) %>% 
  map_df(broom::tidy, .id = 'consumer_iso3c') %>%
  filter(term == 'year') %>%
  select(consumer_iso3c, "slope" = "estimate") %>%
  mutate(fill_color = case_when(
    slope > 0 ~ "1",
    slope < 0 ~ "0",
    TRUE ~ NA
  )) %>%
  add_region(col = "consumer_iso3c", region.col.name = "region")

# Plot out slopes in a barplot
foreign_dependency_slopes %>%
  filter(!is.na(region)) %>%
  ggplot(aes(x = reorder(consumer_iso3c, -slope), 
             y = slope, fill = fill_color)) +
  geom_col() +
  theme_minimal_hgrid() +
  theme(axis.text.x = element_blank()) +
  labs(x = "Consuming country", y = "Slope between foreign\ndependency and time", fill = "+/- slope") +
  scale_fill_manual(values = c("red", "blue")) +
  ylim(-0.03, 0.03)

# Get greatest increases --> greatest decreases in fd by eez
fd_ordered <- foreign_dependency_slopes %>%
  arrange(-slope) %>%
  pull(consumer_iso3c)

# Create new directory for foreign dependencies
if (!dir.exists("../images/europe pc ordered/")) {
  dir.create("../images/europe pc ordered/")
}

for (i in fd_ordered) {
  # Create image
  img <- prop_europe_prod_cons %>%
    filter(consumer_iso3c == i) %>%
    ggplot(aes(x = year, y = prop_cons)) +
    geom_line() +
    geom_point() +
    theme_minimal_hgrid() +
    labs(x = "", y = "Proportion consumption\nproduced + consumed in Europe", title = i) +
    theme(plot.title = element_text(hjust = 0.5))
  
  # Save image
  ggsave(paste0("../images/europe pc ordered/", which(i == fd_ordered), ".jpg"), plot = img, device = "jpeg", height = 4, width = 6, units = "in")
}

# Prop consumption that was produced + consumed in Europe
prop_europe_prod_cons %>%
  filter(year == 2019) %>%
  ggplot(aes(x = fct_reorder(consumer_iso3c, -prop_cons), prop_cons)) +
  theme_cowplot() +
  geom_col() +
  theme(axis.text.x = element_blank()) +
  labs(x = "Europe country", y = "Proportion consumption\nproduced + consumed in Europe")
```

```{r}
# Calcualte total foreign weight consumption
domestic_consumption_weights <- consumption_full %>%
  group_by(consumer_iso3c, year) %>%
  filter(consumer_iso3c == source_country_iso3c) %>%
  summarize(domestic_weight = sum(consumption_live_t)) %>%
  arrange(consumer_iso3c)

# Calculate total foreign weight consumption
foreign_consumption_weights <- consumption_full %>%
  group_by(consumer_iso3c, year) %>%
  filter(consumer_iso3c != source_country_iso3c) %>%
  summarize(foreign_weight = sum(consumption_live_t)) %>%
  arrange(consumer_iso3c)

# Calculate trade dependency
consumer_foreign_dependencies <- foreign_consumption_weights %>%
  filter(consumer_iso3c %in% domestic_consumption_weights$consumer_iso3c) %>%
  left_join(domestic_consumption_weights) %>%
  group_by(consumer_iso3c, year) %>%
  summarize(foreign_dependency = round(foreign_weight /
                                       (foreign_weight +
                                          domestic_weight), 3)) %>%
  exploreARTIS::add_region("consumer_iso3c", region.col.name = "region")

# Look at average change in foreign dependency from 2015-2019
foreign_dependency_slopes <- consumer_foreign_dependencies %>%
  split(.$consumer_iso3c) %>% 
  map(~lm(foreign_dependency~year, data = .x)) %>% 
  map_df(broom::tidy, .id = 'consumer_iso3c') %>%
  filter(term == 'year') %>%
  select(consumer_iso3c, "slope" = "estimate") %>%
  mutate(fill_color = case_when(
    slope > 0 ~ "1",
    slope < 0 ~ "0",
    TRUE ~ NA
  )) %>%
  add_region(col = "consumer_iso3c", region.col.name = "region")

# Create barplot of foreign dependency slopes in europe over time
foreign_dependency_slopes %>%
  filter(!is.na(region)) %>%
  ggplot(aes(x = reorder(consumer_iso3c, -slope), 
             y = slope, fill = region)) +
  geom_col() +
  theme_minimal_hgrid() +
  theme(axis.text.x = element_blank()) +
  labs(x = "Consuming country", y = "Slope between foreign\ndependency and time", fill = "Region") +
  scale_fill_manual(values = artis_palette(7)) +
  ylim(-0.05, 0.05)

# Violin plot by region of foreign dependency slopes
foreign_dependency_slopes %>%
  filter(!is.na(region)) %>%
  ggplot(aes(x = region, y = slope, fill = region)) +
  geom_violin(alpha = 0.5) +
  geom_point() +
  scale_fill_manual(values = artis_palette(7)) +
  theme_minimal_hgrid(12) +
  labs(x = "Consuming country", y = "Slope between foreign\ndependency and time") +
  guides(fill = "none")

# Get greatest increases --> greatest decreases in fd by eez
fd_ordered <- foreign_dependency_slopes %>%
  arrange(-slope) %>%
  pull(consumer_iso3c)

# Create new directory for foreign dependencies
if (!dir.exists("../images/foreign dependency ordered/")) {
  dir.create("../images/foreign dependency ordered/")
}

for (i in fd_ordered) {
  
  # Create image
  img <- consumer_foreign_dependencies %>%
    filter(consumer_iso3c == i) %>%
    ggplot(aes(x = year, y = foreign_dependency)) +
    geom_line() +
    geom_point() +
    theme_minimal_hgrid() +
    labs(x = "", y = "Prop foreign dependency", title = i) +
    theme(plot.title = element_text(hjust = 0.5))
  
  # Save image
  ggsave(paste0("../images/foreign dependency ordered/", which(i == fd_ordered), ".jpg"), plot = img, device = "jpeg", height = 4, width = 6, units = "in")
}
```

```{r}
```{r create graphs}
# Look at change of importance on protein consumption by source, across time

# Try to do weighted mean instead
(f_d_protein <- supply_importance_long %>%
   filter(element == "Protein supply quantity (g/capita/day)") %>%
   group_by(consumer_iso3c) %>%
   filter(n_distinct(year) > 1) %>% # Filter out countries with only 1 year of data
   group_by(year, region, consumption_source) %>%
  summarize(mean_importance = mean(aa_reliance_pct, na.rm = TRUE)) %>%
  mutate(mean_importance = 100 * mean_importance) %>%
  ggplot(aes(x = year, y = mean_importance, color = consumption_source)) +
  geom_point(size = 0.4) +
  geom_line() +
  facet_wrap(~ region) +
  theme_cowplot(10) +
  theme(legend.position="bottom",
        legend.justification = "center") +
  labs(color = "Consumption source", x = "", y = "% importance on protein consumption") +
  scale_x_continuous(breaks = seq(1995, 2020, by = 10)))

# Global sourcing importance, by habitat, by method, by source
(sourcing_importance_global_protein <- supply_importance_long %>%
   filter(element == "Protein supply quantity (g/capita/day)") %>%
  group_by(year, habitat, method, consumption_source) %>%
  summarize(mean_prop_consumption = mean(prop_consumption)) %>%
    mutate(mean_prop_consumption = 100 * mean_prop_consumption) %>%
  ggplot(aes(x = year, y = mean_prop_consumption, color = consumption_source)) +
  geom_point(size = 0.4) +
  geom_line() +
  facet_grid(habitat ~ method, scales = "free_x", space = "free_x") +
  theme_cowplot(10) +
  theme(legend.position = "bottom", legend.justification = "center") +
  labs(x = "", y = "Mean % of consumption\n(across all countries)", color = "Consumption source") +
    scale_x_continuous(breaks = seq(1995, 2020, by = 10)))

# Continental importance by method
(sourcing_importance_continental <- supply_importance_long %>%
  filter(element == "Protein supply quantity (g/capita/day)") %>%
  group_by(year, habitat, method, region) %>%
  summarize(mean_prop_consumption = mean(prop_consumption)) %>%
  mutate(mean_prop_consumption = 100 * mean_prop_consumption) %>%
  ggplot(aes(x = year, y = mean_prop_consumption, color = region)) +
  geom_point(size = 0.4) +
  geom_line() +
  facet_grid(habitat ~ method, scales = "free_x", space = "free_x") +
  theme_cowplot(10) +
  scale_color_manual(values = artis_palette(6)) +
  scale_x_continuous(breaks = seq(1995, 2020, by = 10)) +
  labs(x = "", y = "Mean % of consumption\n(across countries in a region)", color = "Region"))

# Annual mean aquatic animal % contribution to protein supply by continent
supply_importance_long %>%
  filter(element == "Protein supply quantity (g/capita/day)") %>%
  group_by(consumer_iso3c, year, region) %>%
  summarize(sum = sum(aa_reliance_pct)) %>%
  group_by(year, region) %>%
  summarize(mean_importance = 100 * mean(sum)) %>%
  ggplot(aes(x = year, y = mean_importance, color = region)) +
  geom_point() +
  geom_line() +
  theme_cowplot() +
  scale_color_manual(values = artis_palette(6)) +
  labs(x = "", y = "Mean % contribution aquatic animal\nhas to protein consumption")

# Annual mean aquatic animal % contribution to protein supply globally
supply_importance_long %>%
  filter(element == "Protein supply quantity (g/capita/day)") %>%
  group_by(iso3c, year) %>%
  summarize(sum = sum(aa_reliance_pct)) %>%
  group_by(year) %>%
  summarize(mean_importance = 100 * mean(sum, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = mean_importance)) +
  geom_point() +
  geom_line() +
  theme_cowplot() +
  labs(x = "", y = "Mean % contribution aquatic animal\nhas to protein consumption")

# Looking at pct consumption x element variable (food / protein supply) by region
supply_importance_long %>%
  group_by(iso3c) %>%
  filter(n_distinct(year) > 1) %>% # Filter out countries with only 1 year of data
  group_by(region, year, element) %>%
  summarize(aa_reliance_pct = mean(aa_reliance_pct, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = aa_reliance_pct, color = region)) +
  geom_point() +
  geom_line() +
  theme_cowplot() +
  facet_grid(rows = vars(element),
             labeller = labeller(element = function(x) str_wrap(x, width = 15))) +
  scale_color_manual(values = artis_palette(6)) +
  labs(x = "", color = "Region", y = "Aquati animal % contirbution") +
  theme(strip.text = element_text(size = 8))
```
```

