---
title: "98_Appendix"
author: "Connor Quiroz"
date: "2025-07-19"
output: html_document
---

```{r create figure four in 3d}
library(plotly)
library(dplyr)

# Join and base scatter3d
fig_four_3d <- left_join(terrestrial_roc, capture_aquaculture_data, by = c("iso3c" = "consumer_iso3c")) %>%
  add_region(col = "iso3c", region.col.name = "region") %>%
  plot_ly() %>%
  add_markers(
    x = ~estimate.x,
    y = ~estimate.y,
    z = ~estimate,
    color = ~estimate,
    colors = "plasma",
    marker = list(size = 3)
  )

# Define grid
z_range <- seq(-0.2, 0.4, length.out = 2)
xy_range <- seq(-0.4, 0.4, length.out = 2)

# Plane at x = 0 (vary y and z)
x0_surface <- plot_ly(
  showscale = FALSE
) %>% add_surface(
  x = matrix(0, nrow = 2, ncol = 2),
  y = matrix(rep(xy_range, each = 2), nrow = 2),
  z = matrix(rep(z_range, 2), nrow = 2),
  opacity = 0.1,
  colorscale = list(c(0, 1), c("gray", "gray")),
  showscale = FALSE
)

# Plane at y = 0 (vary x and z)
y0_surface <- plot_ly(
  showscale = FALSE
) %>% add_surface(
  x = matrix(rep(xy_range, each = 2), nrow = 2),
  y = matrix(0, nrow = 2, ncol = 2),
  z = matrix(rep(z_range, 2), nrow = 2),
  opacity = 0.1,
  colorscale = list(c(0, 1), c("gray", "gray")),
  showscale = FALSE
)

# Combine everything
fig_four_3d <- subplot(fig_four_3d, x0_surface, y0_surface, shareX = TRUE, shareY = TRUE) %>%
  layout(
    scene = list(
      xaxis = list(title = "Change in capture fishery prevalence", range = c(-0.4, 0.4)),
      yaxis = list(title = "Aquaculture prevalence", range = c(-0.4, 0.4)),
      zaxis = list(title = "Change in meat prevalence")
    ),
    legend = list(orientation = "h", x = 0.3, y = -0.1),
    margin = list(l = 0, r = 0, b = 0, t = 0)
  )

fig_four_3d

```

```{r look at how aa reliance changes over time}
# Look at slopes between aa_reliancec and year
aa_reliance_slopes <- supply_importance_long %>%
  group_by(consumer_iso3c, year) %>%
  summarize(aa_reliance_pct = mean(aa_reliance_pct)) %>%
  split(.$consumer_iso3c) %>% 
  map(~lm(aa_reliance_pct~year, data = .x)) %>% 
  map_df(broom::tidy, .id = 'consumer_iso3c') %>%
  filter(term == 'year') %>%
  select(consumer_iso3c, "slope" = "estimate") %>%
  mutate(fill_color = case_when(
    slope > 0 ~ "1",
    slope < 0 ~ "0",
    TRUE ~ NA
  )) %>%
  add_region(col = "consumer_iso3c", region.col.name = "region")

aa_reliance_slopes %>%
  filter(!is.na(region)) %>%
  ggplot(aes(x = reorder(consumer_iso3c, -slope), 
             y = slope, fill = region)) +
  geom_col() +
  theme_minimal_hgrid() +
  theme(axis.text.x = element_blank()) +
  labs(x = "Consuming country", y = "Slope between aa reliance and time", fill = "Region") +
  scale_fill_manual(values = artis_palette(7)) +
  ylim(-0.005, 0.005)

# Violin plot by region of foreign dependency slopes
aa_reliance_slopes %>%
  filter(!is.na(region)) %>%
  ggplot(aes(x = region, y = slope, fill = region)) +
  geom_violin(alpha = 0.5) +
  geom_point() +
  scale_fill_manual(values = artis_palette(7)) +
  theme_minimal_hgrid(12) +
  labs(x = "Consuming country", y = "Slope between aa reliance and time") +
  guides(fill = "none")


# Look at individual countries' trends in aa contribution to protein
# Get list of countries
countries <- aa_reliance_slopes %>%
  arrange(-slope) %>%
  distinct(consumer_iso3c) %>%
  pull(consumer_iso3c)

# Create new directory for foreign dependencies
if (!dir.exists("../images/aa_reliance_trends/")) {
  dir.create("../images/aa_reliance_trends/")
}

for (i in countries) {
  
  # Generate data
  data <- supply_importance_long %>%
  # filter(consumer_iso3c == i, region == "Oceania") %>%
    filter(consumer_iso3c == i) %>%
    filter(str_detect(element, "Protein")) %>%
    group_by(year, consumer_iso3c)
  
  # Create image
  
  if (nrow(data) > 0) { # Check to see if filtered data has available observations
    img <- data  %>%
      group_by(year, consumer_iso3c) %>%
  ggplot(aes(x = year, y = aa_reliance_pct, color = consumption_source)) +
  geom_line() +
  geom_point() +
  facet_grid(habitat ~ method, scales = "free_x", space = "free_x") +
    theme_cowplot() +
    labs(x = "", y = "AA contribution to protein consumption", title = i, color = "Consumption source") +
    theme(plot.title = element_text(hjust = 0.5),
          axis.text.x = element_text(size = 7, angle = 45)) +
    geom_vline(xintercept = 2009, linetype = "dashed", color = "red")
  
  # Save image
  ggsave(paste0("../images/aa_reliance_trends/", which(i == countries), ".jpg"), plot = img, device = "jpeg", height = 4, width = 6, units = "in")
  }
  
}

aa_reliance_slopes %>%
  arrange(-slope)
```


```{r look at fao dataset discrepencies}
# Get list of FAO countries
countries <- fao_imputed %>%
  ungroup() %>%
  distinct(iso3c) %>%
  pull(iso3c)

# Create new directory for foreign dependencies
if (!dir.exists("../images/fao_dataset_comparison/")) {
  dir.create("../images/fao_dataset_comparison/")
}

for (i in countries) {
  
  # Generate data
  data <- fao_imputed %>%
    filter(iso3c == i) %>%
    filter(str_detect(Element, "Protein")) %>%
    group_by(year, iso3c)
  
  # Create image
  
  if (nrow(data) > 0) { # Check to see if filtered data has available observations
  img <- data %>%
  ggplot(aes(x = year, y = prop_aquatic_animal, color = new_historical)) +
  geom_line() +
  geom_point() +
  theme_cowplot() +
  scale_color_manual(values = c("#4254b8", "#EE5A54", "#52BEDB")) +
  labs(x = "", y = "Prop aquatic animal protein", color = "FAO FBS dataset", title = i) +
    theme(plot.title = element_text(hjust = 0.5))
  

  
  # Save image
  ggsave(paste0("../images/fao_dataset_comparison/", i, ".jpg"), plot = img, device = "jpeg", height = 4, width = 6, units = "in")
  }
  
}
```


```{r look at dependency on europe production & consumption}
# Total Europe consumption that comes from Europe
prod_eur_cons_eur <- consumption_full %>%
  exploreARTIS::add_region(col = "consumer_iso3c", region.col.name = "consumer_region") %>%
  exploreARTIS::add_region(col = "source_country_iso3c", region.col.name = "producer_region") %>%
  relocate(consumer_region, .before = "source_country_iso3c") %>%
  relocate(producer_region, .after = consumer_region) %>%
  group_by(consumer_iso3c, year) %>%
  filter(consumer_region == "Europe",
         producer_region == "Europe") %>%
  summarize(sum_weight = sum(consumption_live_t, na.rm = TRUE))
  
# Europe total consumption
eur_total_cons <- consumption_full %>%
  exploreARTIS::add_region(col = "consumer_iso3c", region.col.name = "consumer_region") %>%
  exploreARTIS::add_region(col = "source_country_iso3c", region.col.name = "producer_region") %>%
  relocate(consumer_region, .before = "source_country_iso3c") %>%
  relocate(producer_region, .after = consumer_region) %>%
  group_by(consumer_iso3c, year) %>%
  filter(consumer_region == "Europe") %>%
  summarize(total_cons = sum(consumption_live_t, na.rm = TRUE))

# Proportion of Europe consumption that is produced in Europe
prop_europe_prod_cons <- left_join(prod_eur_cons_eur, eur_total_cons) %>%
  group_by(consumer_iso3c, year) %>%
  summarize(prop_cons = sum_weight / total_cons)

# Look at average change in foreign dependency from 2015-2019
foreign_dependency_slopes <- prop_europe_prod_cons %>%
  split(.$consumer_iso3c) %>% 
  map(~lm(prop_cons~year, data = .x)) %>% 
  map_df(broom::tidy, .id = 'consumer_iso3c') %>%
  filter(term == 'year') %>%
  select(consumer_iso3c, "slope" = "estimate") %>%
  mutate(fill_color = case_when(
    slope > 0 ~ "1",
    slope < 0 ~ "0",
    TRUE ~ NA
  )) %>%
  add_region(col = "consumer_iso3c", region.col.name = "region")

# Plot out slopes in a barplot
foreign_dependency_slopes %>%
  filter(!is.na(region)) %>%
  ggplot(aes(x = reorder(consumer_iso3c, -slope), 
             y = slope, fill = fill_color)) +
  geom_col() +
  theme_minimal_hgrid() +
  theme(axis.text.x = element_blank()) +
  labs(x = "Consuming country", y = "Slope between foreign\ndependency and time", fill = "+/- slope") +
  scale_fill_manual(values = c("red", "blue")) +
  ylim(-0.03, 0.03)

# Get greatest increases --> greatest decreases in fd by eez
fd_ordered <- foreign_dependency_slopes %>%
  arrange(-slope) %>%
  pull(consumer_iso3c)

# Create new directory for foreign dependencies
if (!dir.exists("../images/europe pc ordered/")) {
  dir.create("../images/europe pc ordered/")
}

for (i in fd_ordered) {
  # Create image
  img <- prop_europe_prod_cons %>%
    filter(consumer_iso3c == i) %>%
    ggplot(aes(x = year, y = prop_cons)) +
    geom_line() +
    geom_point() +
    theme_minimal_hgrid() +
    labs(x = "", y = "Proportion consumption\nproduced + consumed in Europe", title = i) +
    theme(plot.title = element_text(hjust = 0.5))
  
  # Save image
  ggsave(paste0("../images/europe pc ordered/", which(i == fd_ordered), ".jpg"), plot = img, device = "jpeg", height = 4, width = 6, units = "in")
}

# Prop consumption that was produced + consumed in Europe
prop_europe_prod_cons %>%
  filter(year == 2019) %>%
  ggplot(aes(x = fct_reorder(consumer_iso3c, -prop_cons), prop_cons)) +
  theme_cowplot() +
  geom_col() +
  theme(axis.text.x = element_blank()) +
  labs(x = "Europe country", y = "Proportion consumption\nproduced + consumed in Europe")
```

```{r}
# Calcualte total foreign weight consumption
domestic_consumption_weights <- consumption_full %>%
  group_by(consumer_iso3c, year) %>%
  filter(consumer_iso3c == source_country_iso3c) %>%
  summarize(domestic_weight = sum(consumption_live_t)) %>%
  arrange(consumer_iso3c)

# Calculate total foreign weight consumption
foreign_consumption_weights <- consumption_full %>%
  group_by(consumer_iso3c, year) %>%
  filter(consumer_iso3c != source_country_iso3c) %>%
  summarize(foreign_weight = sum(consumption_live_t)) %>%
  arrange(consumer_iso3c)

# Calculate trade dependency
consumer_foreign_dependencies <- foreign_consumption_weights %>%
  filter(consumer_iso3c %in% domestic_consumption_weights$consumer_iso3c) %>%
  left_join(domestic_consumption_weights) %>%
  group_by(consumer_iso3c, year) %>%
  summarize(foreign_dependency = round(foreign_weight /
                                       (foreign_weight +
                                          domestic_weight), 3)) %>%
  exploreARTIS::add_region("consumer_iso3c", region.col.name = "region")

# Look at average change in foreign dependency from 2015-2019
foreign_dependency_slopes <- consumer_foreign_dependencies %>%
  split(.$consumer_iso3c) %>% 
  map(~lm(foreign_dependency~year, data = .x)) %>% 
  map_df(broom::tidy, .id = 'consumer_iso3c') %>%
  filter(term == 'year') %>%
  select(consumer_iso3c, "slope" = "estimate") %>%
  mutate(fill_color = case_when(
    slope > 0 ~ "1",
    slope < 0 ~ "0",
    TRUE ~ NA
  )) %>%
  add_region(col = "consumer_iso3c", region.col.name = "region")

# Create barplot of foreign dependency slopes in europe over time
foreign_dependency_slopes %>%
  filter(!is.na(region)) %>%
  ggplot(aes(x = reorder(consumer_iso3c, -slope), 
             y = slope, fill = region)) +
  geom_col() +
  theme_minimal_hgrid() +
  theme(axis.text.x = element_blank()) +
  labs(x = "Consuming country", y = "Slope between foreign\ndependency and time", fill = "Region") +
  scale_fill_manual(values = artis_palette(7)) +
  ylim(-0.05, 0.05)

# Violin plot by region of foreign dependency slopes
foreign_dependency_slopes %>%
  filter(!is.na(region)) %>%
  ggplot(aes(x = region, y = slope, fill = region)) +
  geom_violin(alpha = 0.5) +
  geom_point() +
  scale_fill_manual(values = artis_palette(7)) +
  theme_minimal_hgrid(12) +
  labs(x = "Consuming country", y = "Slope between foreign\ndependency and time") +
  guides(fill = "none")

# Get greatest increases --> greatest decreases in fd by eez
fd_ordered <- foreign_dependency_slopes %>%
  arrange(-slope) %>%
  pull(consumer_iso3c)

# Create new directory for foreign dependencies
if (!dir.exists("../images/foreign dependency ordered/")) {
  dir.create("../images/foreign dependency ordered/")
}

for (i in fd_ordered) {
  
  # Create image
  img <- consumer_foreign_dependencies %>%
    filter(consumer_iso3c == i) %>%
    ggplot(aes(x = year, y = foreign_dependency)) +
    geom_line() +
    geom_point() +
    theme_minimal_hgrid() +
    labs(x = "", y = "Prop foreign dependency", title = i) +
    theme(plot.title = element_text(hjust = 0.5))
  
  # Save image
  ggsave(paste0("../images/foreign dependency ordered/", which(i == fd_ordered), ".jpg"), plot = img, device = "jpeg", height = 4, width = 6, units = "in")
}
```


```{r create graphs}
# Look at change of importance on protein consumption by source, across time

# Try to do weighted mean instead
(f_d_protein <- supply_importance_long %>%
   filter(element == "Protein supply quantity (g/capita/day)") %>%
   group_by(consumer_iso3c) %>%
   filter(n_distinct(year) > 1) %>% # Filter out countries with only 1 year of data
   group_by(year, region, consumption_source) %>%
  summarize(mean_importance = mean(aa_reliance_pct, na.rm = TRUE)) %>%
  mutate(mean_importance = 100 * mean_importance) %>%
  ggplot(aes(x = year, y = mean_importance, color = consumption_source)) +
  geom_point(size = 0.4) +
  geom_line() +
  facet_wrap(~ region) +
  theme_cowplot(10) +
  theme(legend.position="bottom",
        legend.justification = "center") +
  labs(color = "Consumption source", x = "", y = "% importance on protein consumption") +
  scale_x_continuous(breaks = seq(1995, 2020, by = 10)))

# Global sourcing importance, by habitat, by method, by source
(sourcing_importance_global_protein <- supply_importance_long %>%
   filter(element == "Protein supply quantity (g/capita/day)") %>%
  group_by(year, habitat, method, consumption_source) %>%
  summarize(mean_prop_consumption = mean(prop_consumption)) %>%
    mutate(mean_prop_consumption = 100 * mean_prop_consumption) %>%
  ggplot(aes(x = year, y = mean_prop_consumption, color = consumption_source)) +
  geom_point(size = 0.4) +
  geom_line() +
  facet_grid(habitat ~ method, scales = "free_x", space = "free_x") +
  theme_cowplot(10) +
  theme(legend.position = "bottom", legend.justification = "center") +
  labs(x = "", y = "Mean % of consumption\n(across all countries)", color = "Consumption source") +
    scale_x_continuous(breaks = seq(1995, 2020, by = 10)))

# Continental importance by method
(sourcing_importance_continental <- supply_importance_long %>%
  filter(element == "Protein supply quantity (g/capita/day)") %>%
  group_by(year, habitat, method, region) %>%
  summarize(mean_prop_consumption = mean(prop_consumption)) %>%
  mutate(mean_prop_consumption = 100 * mean_prop_consumption) %>%
  ggplot(aes(x = year, y = mean_prop_consumption, color = region)) +
  geom_point(size = 0.4) +
  geom_line() +
  facet_grid(habitat ~ method, scales = "free_x", space = "free_x") +
  theme_cowplot(10) +
  scale_color_manual(values = artis_palette(6)) +
  scale_x_continuous(breaks = seq(1995, 2020, by = 10)) +
  labs(x = "", y = "Mean % of consumption\n(across countries in a region)", color = "Region"))

# Annual mean aquatic animal % contribution to protein supply by continent
supply_importance_long %>%
  filter(element == "Protein supply quantity (g/capita/day)") %>%
  group_by(consumer_iso3c, year, region) %>%
  summarize(sum = sum(aa_reliance_pct)) %>%
  group_by(year, region) %>%
  summarize(mean_importance = 100 * mean(sum)) %>%
  ggplot(aes(x = year, y = mean_importance, color = region)) +
  geom_point() +
  geom_line() +
  theme_cowplot() +
  scale_color_manual(values = artis_palette(6)) +
  labs(x = "", y = "Mean % contribution aquatic animal\nhas to protein consumption")

# Annual mean aquatic animal % contribution to protein supply globally
supply_importance_long %>%
  filter(element == "Protein supply quantity (g/capita/day)") %>%
  group_by(iso3c, year) %>%
  summarize(sum = sum(aa_reliance_pct)) %>%
  group_by(year) %>%
  summarize(mean_importance = 100 * mean(sum, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = mean_importance)) +
  geom_point() +
  geom_line() +
  theme_cowplot() +
  labs(x = "", y = "Mean % contribution aquatic animal\nhas to protein consumption")

# Looking at pct consumption x element variable (food / protein supply) by region
supply_importance_long %>%
  group_by(iso3c) %>%
  filter(n_distinct(year) > 1) %>% # Filter out countries with only 1 year of data
  group_by(region, year, element) %>%
  summarize(aa_reliance_pct = mean(aa_reliance_pct, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = aa_reliance_pct, color = region)) +
  geom_point() +
  geom_line() +
  theme_cowplot() +
  facet_grid(rows = vars(element),
             labeller = labeller(element = function(x) str_wrap(x, width = 15))) +
  scale_color_manual(values = artis_palette(6)) +
  labs(x = "", color = "Region", y = "Aquati animal % contirbution") +
  theme(strip.text = element_text(size = 8))
```

```{r}
# Calculate total protein consumed by country
total_protein <- aquatic_reliance %>%
  filter(data_source == "new FBS") %>%
  group_by(year, iso3c) %>%
  summarize(total_animal_protein_t = sum(protein_consumed_t))

# Count the number of countries that have increased aquaculture / marine capture reliance
aquatic_reliance %>%
  filter(data_source == "new FBS") %>%
  group_by(year, iso3c, method) %>%
  summarize(protein_consumed_t = sum(protein_consumed_t), .groups = "drop") %>%
  filter(method %in% c("aquaculture", "capture")) %>%
  left_join(total_protein, by = c("iso3c", "year")) %>%
  mutate(prop_animal_protein = protein_consumed_t / total_animal_protein_t) %>%
  group_by(iso3c, method) %>%      # <-- keep method here
  nest() %>%
  mutate(
    model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
    tidy  = map(model, tidy)
  ) %>%
  unnest(tidy) %>%
  filter(term == "year") %>%
  select(iso3c, method, estimate) %>%
  ungroup() %>%
  mutate(positive = estimate > 0) %>%
  group_by(method) %>%
  count(positive)
```

### Simple Bayesian Model

```{r simple bayesian model}
for (i in seq_along(response_vars)) {
joined_df <- get(response_vars[i]) %>%
  left_join(ac_df, by = "iso3c") %>%
  left_join(ecoregion, by = "iso3c") %>%
  ungroup() %>%
  add_region(col = "iso3c", region.col.name = "region") %>%
  left_join(dsfi_df, by = "iso3c") %>%
  # left_join(imports_df %>% rename(iso3c = consumer_iso3c), by = "iso3c") %>%
  # mutate(log_gdp = log(gdp),
  #        log_imports = log(consumption_t)) %>%
  mutate(log_gdp = log(gdp)) %>%
  left_join(indicator_df %>%
  select("eez_total", "iso3c", "inundation_area_log1000km2" = "inland_water_max"), by = "iso3c") %>%
  left_join(coastline_df, by = "iso3c")

df <- joined_df %>%
  select(
    names(.)[2], log_gdp, gdp_trade, eez_total,
    inundation_area_log1000km2,
    coastline_km, realm
  ) %>%
    mutate( # Impute NA values with 0
      coastline_km = case_when(
      is.na(coastline_km) & str_detect(realm, "Landlocked") ~ 0,
      TRUE ~ coastline_km
    ),
    
    eez_total = case_when(
      is.na(eez_total) & str_detect(realm, "Landlocked") ~ 0,
      TRUE ~ eez_total
    )
    ) %>% 
  drop_na() %>%
  filter(realm != "") %>%
    mutate(
    across(
      c(log_gdp, gdp_trade, eez_total,
        inundation_area_log1000km2, coastline_km),
      ~ (.x - min(.x, na.rm = TRUE)) /
        (max(.x, na.rm = TRUE) - min(.x, na.rm = TRUE)),
      .names = "{.col}"
    ))

# Separate explanatory variables (no realm, no random effect)
X <- df %>%
  select(-names(.)[1], -realm) %>%
  scale()

# Response
y <- scale(df %>% pull(1))[, 1]

N <- nrow(X)
K <- ncol(X)

# Save X and y (for diagnostic statistics to be used later on)
  saveRDS(X, paste0("../output/bayesian_output/", model_type, "/X_",response_vars[i],".rds"))
  saveRDS(y, paste0("../output/bayesian_output/", model_type, "/y_",response_vars[i],".rds"))

inits <- function() {
  list(
    beta_0 = 0,
    sigma = 1,
    beta = rep(0, K)
  )
}

stan_data <- list(
  N = N,
  K = K,
  X = as.matrix(X),
  y = y
)

assign(
  paste0("fit_", response_vars[i]),
  stan(
    file = "../R/social_bayesian_model_simple.stan",
    data = stan_data,
    init = inits,
    chains = 3,
    iter = 10000,
    warmup = 2000
  )
)

saveRDS(
  get(paste0("fit_", response_vars[i])),
  paste0("../output/bayesian_output/simple/fit_", response_vars[i], ".stan")
)

}
```

```{r hierarchical model for non landlocked countries (old data cleaning process)}
model_type <- "realm"

for (i in seq_along(response_vars)) {
  # Join target response to its explanatory variables
  joined_df <- get(response_vars[i]) %>%
  left_join(ac_df, by = "iso3c") %>%
  left_join(ecoregion, by = "iso3c") %>%
  ungroup() %>%
  add_region(col = "iso3c", region.col.name = "region") %>%
  left_join(dsfi_df, by = "iso3c") %>%
  # left_join(imports_df %>% rename(iso3c = consumer_iso3c), by = "iso3c") %>%
  mutate(log_gdp = log(gdp),
         log_imports = log(consumption_t)) %>%
  mutate(log_gdp = log(gdp)) %>%
  left_join(indicator_df %>%
  select("eez_total", "iso3c", "inundation_area_log1000km2" = "inland_water_max"), by = "iso3c") %>%
  left_join(coastline_df, by = "iso3c") 
  
  # Clean joined data to be used in Bayesian model
  
  if (model_type == "continental") {
    joined_df <- joined_df %>%
    select(-realm) %>%
    rename(realm = region)
  }
  
  df <- joined_df %>%
  select(
    names(.)[2], log_gdp, gdp_trade, eez_total,
    inundation_area_log1000km2,
    coastline_km, realm
  ) %>%
    mutate( # Impute NA values with 0
      coastline_km = case_when(
      is.na(coastline_km) & str_detect(realm, "Landlocked") ~ 0,
      TRUE ~ coastline_km
    ),
    
    eez_total = case_when(
      is.na(eez_total) & str_detect(realm, "Landlocked") ~ 0,
      TRUE ~ eez_total
    )
    ) %>% 
  drop_na() %>%
  filter(realm != "") %>%
    mutate(
    across(
      c(log_gdp, gdp_trade, eez_total,
        inundation_area_log1000km2, coastline_km),
      ~ (.x - min(.x, na.rm = TRUE)) /
        (max(.x, na.rm = TRUE) - min(.x, na.rm = TRUE)),
      .names = "{.col}"
    ))


 # Separate explanatory variables (not including the factor level)
  X <- df %>%
  select(-names(.)[1], -realm) %>%
             scale()

 # Separate response variable
  y_raw <- df %>% pull(1)
  y <- scale(y_raw)[,1]

  # Create province index
  realm_ids <- as.numeric(factor(df$realm))
  n_realms <- length(unique(realm_ids))
  
  # Save Realm IDs (to be used in exploratory plots)
  realm_levels <- levels(factor(df$realm))
  saveRDS(
  realm_levels,
  paste0("../output/bayesian_output/", model_type, "/realms_", response_vars[i], ".rds")
)
  
  # Save Realm IDs
  saveRDS(
  realm_ids,
  paste0("../output/bayesian_output/", model_type, "/ids_", response_vars[i], ".rds")
)
  
  
  # Save X and y (for diagnostic statistics to be used later on)
  saveRDS(X, paste0("../output/bayesian_output/", model_type, "/X_",response_vars[i],".rds"))
  saveRDS(y, paste0("../output/bayesian_output/", model_type, "/y_",response_vars[i],".rds"))
 
  
  N <- nrow(X)
  K <- ncol(X)
  
  inits <- function() {
    list(
      beta_0       = 0,
      sigma_beta_0 = 1,
      z_beta_0     = rnorm(n_realms, 0, 0.1),
      beta         = rep(0, K),
      sigma        = 1
    )
  }
  
  stan_data <- list(
    N = N,
    K = K,
    J = n_realms,  # number of provinces
    X = as.matrix(X),
    y = y,
    province = realm_ids  # province indicator for each observation
  )
  
  # Fit Bayesian model
  assign(
    paste0("fit_",response_vars[i]),
    stan(
    file = "../R/social_bayesian_model_hierarchical.stan",
    data = stan_data,
    init = inits,
    chains = 3,
    iter = 4000,
    warmup = 1000,
    control = list(adapt_delta = 0.95, max_treedepth = 12)
  )
  )
  
  
  # Save Bayesian model
  saveRDS(
    get(paste0("fit_", response_vars[i])),
    paste0("../output/bayesian_output/", model_type, "/fit_", response_vars[i],".stan")
  )
}


continental_levels <- c("Africa", "Asia", "Europe", "North America", "Oceania", "South America")
realm_levels <- c("Arctic", "Central Indo-Pacific", "Eastern Indo-Pacific", "Landlocked, Africa", "Landlocked, Asia", "Landlocked, Europe", "Landlocked, South America", "Southern Ocean", "Temperate Australasia", "Temperate Northern Atlantic", "Temperate Northern Pacific", "Temperate South America", "Temperate Southern Africa", "Tropical Atlantic", "Tropical Eastern Pacific", "Western Indo-Pacific")
```

```{r old function for setting plot themes}
theme_tile_pub <- function() {
  theme_minimal(base_size = 8) +
    theme(
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.y = element_line(color = scales::alpha("black", 0.25), linewidth = 0.3),
      panel.grid.minor.y = element_blank(),
      panel.background = element_rect(fill = "transparent", color = NA),
      plot.background = element_rect(fill = "transparent", color = NA),
      axis.text.x = element_text(size = 6,angle = 45, vjust = 0.5, color = "black"),
      axis.text.y = element_text(size = 6, color = "black"),
      axis.title = element_text(size = 7, face = "bold"),
      legend.title = element_text(size = 7, face = "bold"),
      legend.text = element_text(size = 6),
      legend.key.size = unit(0.3, "cm"),
      legend.box = "vertical",
      legend.spacing = unit(0.1, "cm"),
      legend.margin = margin(t = 0, b = 0),
      legend.box.margin = margin(t = -2, b = -2),
      legend.background = element_rect(fill = "transparent", color = NA),
      legend.box.background = element_rect(fill = "transparent", color = NA),
      plot.title = element_text(size = 8, face = "bold", hjust = 0.5)
    )
}
```

> Posterior predictive distributions

```{r ppd}
# Combined posterior predictive plot (all models together)
draw_sample <- sample(unique(ppd_all_models$draw), size = 30)

(plot <- ppd_all_models %>%
  filter(
    !str_detect(model, "animal_cons_cap"),
    str_detect(clean_name, "capita") |
    str_detect(clean_name, "Aquatic reliance") |
    str_detect(clean_name, "reliance")
  ) %>%
  filter(draw %in% draw_sample) %>%
  ggplot(aes(x = ppd, group = draw)) +
  geom_density(alpha = 0.08, color = "grey60") +
  geom_density(aes(x = y_obs), color = "black", linewidth = 1.1) +
  facet_wrap(
    ~ clean_name,
    scales = "free",
    labeller = label_wrap_gen(width = 18)   # wrap facet labels so they don't stay on one line
  ) +
  labs(
    x = "Response value",
    y = "Density"
  ) +
  theme_minimal(base_size = 8) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(color = scales::alpha("black", 0.25), linewidth = 0.3),
    panel.grid.minor.y = element_blank(),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),
    strip.background = element_blank(),
    strip.text = element_text(size = 7, face = "bold"),
    axis.text = element_text(size = 6, color = "black"),
    axis.title.y = element_text(size = 7, margin = margin(r = 5), face = "bold"),
    axis.text.x = element_text(size = 5.5, angle = 30, hjust = 1),
    legend.position = "bottom",
    legend.title = element_text(size = 7, face = "bold"),
    legend.text = element_text(size = 6),
    legend.key.size = unit(0.3, "cm"),
    legend.box = "vertical",
    legend.spacing = unit(0.1, "cm"),
    legend.margin = margin(t = 0, b = 0),
    legend.box.margin = margin(t = -2, b = -2),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA)
  ))

# Save ppd plot
ggsave(paste0("../images/bayesian plots/", model_type, "/ppd.jpg"), plot = plot, device = "jpg", width = 7, height = 5, units = "in", dpi = 300)
```

### Posterior Predictive Checks (didn't use this version for ppd checks)

```{r posterior predictive check}
model_type = "simple"
ppc_df <- purrr::map_dfr(seq_along(response_vars), \(i)
  bayesplot::ppc_data(
    readRDS(paste0("../output/bayesian_output/", model_type,
                   "/y_", response_vars[i], ".rds")),
    extract(
      readRDS(paste0("../output/bayesian_output/", model_type,
                     "/fit_", response_vars[i], ".stan")),
      permuted = TRUE
    )$y_pred[sample(8000, 200), ]
  ) |>
    mutate(model = response_vars[i])
)

plot <- ggplot(ppc_df, ggplot2::aes(value)) +
  geom_density(
    data = filter(ppc_df, !is_y),
    aes(group = interaction(model, rep_id)),
    linewidth = .25, alpha = .6
  ) +
  geom_density(
    data = filter(ppc_df, is_y),
    linewidth = 1
  ) +
  facet_wrap(~ model, scales = "free_y", ncol = 4) +
  bayesplot_theme_get()

ggsave(paste0("../images/bayesian plots/", model_type, "/posterior_predictive.jpg"), plot = plot, device = "jpg", width = 15, height = 10, units = "in", dpi = 300)
```

### Simple bayesian model R^

```{r simple bayesian model R^2}
response_vars = response_vars
model_type = "simple"

#######################
#### Simple R^2    ####
#######################

R2_matrix <- matrix(nrow = 3000, ncol = length(response_vars))
for (i in seq_along(1:23)) {

# Extract posterior draws
fit <- rstan::extract(readRDS(paste0("../output/bayesian_output/", model_type, "/fit_", response_vars[i],".stan")), permuted = T)

# y_pred is iterations x N
y_pred <- fit$y_pred[1:3000,]

# y_pred: iterations × N
mu_s <- y_pred   # each row is mu_s for that iteration

var_mu_s <- apply(mu_s, 1, var)   # length = iterations
post_var_obs <- apply(y_pred, 2, var)   # length N

sigma_s <- fit$sigma[1:3000]   # length = iterations
var_resid_s <- sigma_s^2


R2_s <- var_mu_s / (var_mu_s + var_resid_s)
# print((hist(R2_s,main = response_vars[i])))
R2_matrix[,i] <- R2_s
}

colnames(R2_matrix) <- response_vars

plot_df <- R2_matrix %>%
  data.frame() %>%
  pivot_longer(everything()) %>%
  mutate(
    value  = as.numeric(value),
    suffix = sub(".*_", "", name)
  )

# compute the order of model names by suffix
ordered_names <- plot_df %>%
  distinct(name, suffix) %>%
  arrange(suffix) %>%
  pull(name)

(plot <- plot_df %>%
  mutate(name = factor(name, levels = ordered_names)) %>%
  ggplot(aes(x = value, y = name)) +
  ggridges::geom_density_ridges(fill = "lightblue") +
  theme_pubclean() +
  labs(x = bquote("Posterior" ~ R^2), y = "Model"))


ggsave("../images/bayesian plots/simple/r2.jpg", plot = plot, device = "jpg", width = 6, height = 4, units = "in", dpi = 300)

```

### Create trace plots for the simple Bayesian model

```{r}
# Simple Trace
model_type = "simple"
trace_df <- data.frame()

for (i in seq_along(response_vars))  {
  
  # Load in Stan data
  fit <- rstan::extract(readRDS(paste0("../output/bayesian_output/", model_type, "/fit_", response_vars[i],".stan")), permuted = T)

  chain_1 <- data.frame(beta0 = fit$beta[1:3000,1],
           beta = fit$beta_0[1:3000],
           sigma = fit$sigma[1:3000],
           model = response_vars[i],
           iteration = 1:3000,
           chain = "1")
  
  chain_2 <- data.frame(beta0 = fit$beta[1:3000,1],
           beta = fit$beta_0[3001:6000],
           sigma = fit$sigma[3001:6000],
           model = response_vars[i],
           iteration = 1:3000,
           chain = "2")
  
  chain_3 <- data.frame(beta0 = fit$beta[1:3000,1],
           beta = fit$beta_0[6001:9000],
           sigma = fit$sigma[6001:9000],
           model = response_vars[i],
           iteration = 1:3000,
           chain = "3")
  
  chains <- chain_1 %>% bind_rows(chain_2) %>% bind_rows(chain_3)
  
  trace_df <- bind_rows(trace_df, chains)
    
}
```

### Country level trends in delta 2010-2020 aquatic animal reliance

```{r Country level trends in delta 2010-2020 aquatic animal reliance}
# Fit linear models to estimate the average annual change in disaggregated
# aquatic animal reliance (2010-2020) by production source (habitat x method)
# for each country
disagg <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c) %>%
  mutate(n = n_distinct(year)) %>%
  filter(n >= 3) %>%
  group_by(iso3c, habitat, method) %>%
  filter(n() >= 3) %>%
  nest() %>%
  mutate(
    model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
    tidy  = map(model, broom::tidy)
  ) %>%
  unnest(tidy) %>%
  filter(term == "year") %>%
  mutate(
    consource_habitat_method = paste(habitat, method),
    group_label = consource_habitat_method
  ) %>%
  drop_na(estimate) %>%
  filter(!str_detect(consource_habitat_method, "unknown")) %>%
  group_by(group_label) %>%
  mutate(pct_above = mean(estimate > 0) * 100) %>%
  ungroup()

# Fit linear models to estimate the average annual change in aggregate
# aquatic animal reliance (2010-2020)
agg <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c) %>%
  mutate(n = n_distinct(year)) %>%
  filter(n >= 3) %>% # Get countries with only 3 years or more of data (so lm can properly fit points)
  group_by(iso3c, year, data_source, food_group) %>%
  summarise(prop_animal_protein = sum(prop_animal_protein)) %>%
  group_by(iso3c) %>%
  nest() %>%
  mutate(
    model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
    tidy  = map(model, broom::tidy)
  ) %>%
  unnest(tidy) %>%
  filter(term == "year") %>%
  mutate(
    group_label = "aggregate"
  ) %>%
  drop_na(estimate) %>%
  group_by(group_label) %>%
  mutate(pct_above = mean(estimate > 0) * 100) %>%
  ungroup()

# Combine aggregate and disaggregate slopes
all_slopes <- bind_rows(agg, disagg)

# Calculate the percent of countries that have slopes above and below 0.
all_slopes <- all_slopes %>%
  mutate(
    pct_below = 100 - pct_above,
    label_majority = if_else(
      pct_above >= 50,
      sprintf("%.0f%%", pct_above),
      sprintf("%.0f%%", pct_below)
    ),
    label_x = if_else(
      pct_above >= 50,
      0.002,   # right side, close to center
      -0.002   # left side, close to center
    ),
    label_hjust = if_else(pct_above >= 50, 0, 1),
    color = case_when(pct_above >= 50 ~ "#2166ac",
                      TRUE ~ "#b2182b")
  )

# Plot data
(plot <- ggplot(all_slopes, aes(
  x = estimate,
  y = fct_reorder(group_label, pct_above)
)) +
  geom_density_ridges_gradient(
    aes(fill = after_stat(x > 0)),
    scale = 1.4,
    rel_min_height = 0.01,
    color = "black"
  ) +
  scale_fill_manual(
    values = c(`TRUE` = "#2166ac", `FALSE` = "#b2182b")
  ) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_text(
    aes(x = label_x, label = label_majority, hjust = label_hjust),
    color = all_slopes$color,
    vjust = -2,
    size = 2.5
  ) +
  guides(fill = "none") +
  labs(x = expression(bold(Delta[2010-2020] ~ "Aquatic animal protein reliance %")), y = "Production source") +
  # xlim(c(-0.0105, 0.0105)) +

  # --- Matching theme from aa_reliance_multiple_trends ---
  theme_minimal(base_size = 8) +
  theme(
    # Grid and background
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(color = scales::alpha("black", 0.25), linewidth = 0.3),
    panel.grid.minor.y = element_blank(),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),

    # Axis styling
    axis.text = element_text(size = 6, color = "black"),
    axis.title.x = element_text(size = 7, face = "bold", margin = margin(t = 5)),
    axis.title.y = element_text(size = 7, face = "bold"),

    # Ridge spacing
    axis.text.y = element_text(size = 6),

    # No legend
    legend.position = "none"
  ))



# ggsave("../images/pct_ichange_reliance.jpg", plot = plot, device = "jpg", units = "in", width = 5, height = 4)
```

```{r old color palette for bayesian posterior distribution plot}
# Old color palette
# 1. Base colors per model
# 1. Base colors per model
# model_colors <- c(
#   "Terrestrial animal consumption consumption per capita" = "#b15928",
#   "Aquatic animal consumption consumption per capita"     = "#1f78b4",
#   "Aquatic reliance"                                      = "#e31a1c",
#   "Capture consumption per capita"                        = "#ff7f00",
#   "Aquaculture consumption per capita"                    = "#6a3d9a",
#   "Domestic consumption per capita"                       = "#33a02c",
#   "Foreign consumption per capita"                        = "#a6cee3",
#   "Marine consumption per capita"                         = "#b2df8a",
#   "Inland consumption per capita"                         = "#fb9a99"
# )
```


```{r all 6 continents visualization}
# Set plot title size
axis.title.size <- 8

# Fit linear models to estimate the average annual change in disaggregated
# aquatic animal reliance (2010-2020) by production source (habitat x method)
# for each country
disagg <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "aquatic") %>%
  group_by(year, iso3c, habitat, method) %>%
  summarize(prop_animal_protein = sum(prop_animal_protein)) %>%
  group_by(iso3c, habitat, method) %>%
  nest() %>%
  mutate(
    model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
    tidy  = map(model, broom::tidy)
  ) %>%
  unnest(tidy) %>%
  filter(term == "year") %>%
  mutate(
    consource_habitat_method = paste(habitat, method),
    group_label = consource_habitat_method,
    group_label = case_when(
      group_label == "unknown unknown" ~ "unknown",
      TRUE ~ group_label
    )
  ) %>%
  drop_na(estimate) %>%
  group_by(group_label) %>%
  mutate(pct_above = mean(estimate > 0) * 100) %>%
  ungroup()

# Fit linear models to estimate the average annual change in aggregate
# aquatic animal reliance (2010-2020)
agg <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, year) %>%
  summarise(prop_animal_protein = sum(prop_animal_protein), .groups = "drop") %>%
  group_by(iso3c) %>%
  nest() %>%
  mutate(
    model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
    tidy  = map(model, broom::tidy)
  ) %>%
  unnest(tidy) %>%
  filter(term == "year") %>%
  mutate(group_label = "aggregate") %>%
  drop_na(estimate) %>%
  group_by(group_label) %>%
  mutate(pct_above = mean(estimate > 0) * 100) %>%
  ungroup()

# Clean aggregate estimate data
agg_estimate <- bind_rows(agg, disagg) %>%
  group_by(iso3c) %>%
  mutate(agg_est = estimate[group_label == "aggregate"][1]) %>%
  add_region("iso3c", "region") %>%
  ungroup() %>%
  mutate(
    iso3c  = fct_reorder(iso3c, -agg_est),
    is_agg = group_label == "aggregate"
  ) %>%
  filter(group_label == "aggregate") %>%
  select(iso3c, agg_estimate = estimate)

# Top row of plots
top_row <- disagg %>%
  group_by(iso3c) %>%
  add_region("iso3c", "region") %>%
    filter(!region %in% c("Oceania", "South America")) %>%
  ungroup() %>%
    left_join(agg_estimate, by = "iso3c") %>%
  mutate(
    iso3c  = fct_reorder(iso3c, -agg_estimate),
  ) %>%
  ggplot(aes(x = estimate, y = iso3c, fill = group_label)) +

  geom_col() +
    geom_point(aes(x = agg_estimate, y = iso3c), shape = 16) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.4) +

  scale_alpha_identity() +
  scale_size_identity() +
  scale_fill_manual(
    values = c("#D8C84A","#E68A2E", "#7A5FA3", "#3A77B5", "gray")
  ) +
  labs(
    x = "Average annual change in aquatic animal reliance (2010–2020)",
    y     = "Country",
    fill = "Production source",
  ) +

  facet_wrap(~ region, scales = "free", nrow = 1) +

  theme_minimal(base_size = 8) +
  theme(
    legend.position    = "bottom",
    panel.grid.minor   = element_blank(),
    axis.text = element_text(size = 4.75),
    legend.key.size    = unit(0.3, "cm"),
    legend.box         = "vertical",
    legend.spacing     = unit(0.1, "cm"),
    legend.text        = element_text(size = 5),
    legend.margin      = margin(t = 0, b = 0),
    axis.title.y = element_blank(),
    panel.grid.major.y = element_line(color = scales::alpha("black", 0.05), linewidth = 0.3),
    panel.grid.major.x = element_line(color = scales::alpha("black", 0.25)),
      axis.title.x = element_blank()
  ) +
    guides(fill = "none")


# Bottom row of plots
bottom_row <- disagg %>%
  group_by(iso3c) %>%
  add_region("iso3c", "region") %>%
    filter(region %in% c("Oceania", "South America")) %>%
  ungroup() %>%
    left_join(agg_estimate, by = "iso3c") %>%
  mutate(
    iso3c  = fct_reorder(iso3c, -agg_estimate),
  ) %>%
  ggplot(aes(x = estimate, y = iso3c, fill = group_label)) +

  geom_col() +
    geom_point(aes(x = agg_estimate, y = iso3c), shape = 16) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.4) +

  scale_alpha_identity() +
  scale_size_identity() +
  scale_fill_manual(
    values = c("#D8C84A","#E68A2E", "#7A5FA3", "#3A77B5", "gray")
  ) +
  labs(
    x = "Average annual change in aquatic animal reliance (2010–2020)",
    y     = "Country",
    fill = "Production source",
  ) +

  facet_wrap(~ region, scales = "free") +

  theme_minimal(base_size = 8) +
  theme(
    axis.title.x = element_text(face = "bold"),
    legend.position    = "bottom",
    panel.grid.minor   = element_blank(),
    axis.text        = element_text(size = 5.35),
    legend.key.size    = unit(0.3, "cm"),
    legend.box         = "vertical",
    legend.spacing     = unit(0.1, "cm"),
    legend.text        = element_text(size = 5),
    legend.margin      = margin(t = 0, b = 0),
    axis.title.y = element_blank(),
    panel.grid.major.y = element_line(color = scales::alpha("black", 0.05), linewidth = 0.3),
    panel.grid.major.x = element_line(color = scales::alpha("black", 0.25))
  )

bottom_row_narrowed <- cowplot::plot_grid(
  NULL, bottom_row, NULL,
  ncol = 3,
  rel_widths = c(0.1, 0.3, 0.1)  # adjust 0.8 to make it narrower/wider
)

(bottom_panel <- cowplot::plot_grid(top_row, bottom_row_narrowed, nrow = 2, rel_heights = c(2,1)))

top_panel <- agg %>%
  count(num_countries = estimate > 0) %>%
  ggplot(aes(x = 1, y = n, fill = num_countries)) +
  coord_flip() +
  geom_col(position = "stack") +
  labs(y = "Number of countries", fill = "Direction of change") +
  scale_fill_manual(values = c("#b2182b", "#2166ac"),
                    labels  = c(`FALSE` = "Declining reliance",
              `TRUE` = "Increasing reliance")) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text = element_text(size = 4.75),
        axis.title.y = element_blank(),
        axis.title.x = element_text(size = axis.title.size-2, face = "bold"),
    panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_line(color = scales::alpha("black", 0.25), linewidth = 0.3),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),
    legend.position = "bottom",
    legend.title = element_text(size = 7, face = "bold"),
    legend.text = element_text(size = 6),
    legend.key.size = unit(0.25, "cm"),
    legend.box = "vertical",
    legend.spacing = unit(.05, "cm"),
    legend.margin = margin(t = 0, b = 0),
    legend.box.margin = margin(t = -2, b = -2),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    aspect.ratio = 0.05)
  
(final_plot <- cowplot::plot_grid(top_panel, bottom_panel, nrow = 2, rel_heights = c(0.35,2),
                                  labels = c("A", "B"),label_fontface = "bold", label_size = 9.5))


ggsave("../images/pct_reliance_contribution_four_sources.jpg", plot = final_plot, device = "jpg", units = "in", width = 8, height = 6, dpi = 500)
```

```{r Region counts overlayed on quadrant plot}
# Fit linear models to estimate the average annual change in aquatic protein
# consumption (2010-2020)
aquatic_protein_quant <- aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, year) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap, na.rm = TRUE)) %>%
  distinct() %>%
  split(.$iso3c) %>%
  map(~lm(protein_consumed_kg_percap~year, data = .x)) %>%
  map_df(broom::tidy, .id = 'iso3c') %>%
  filter(term == 'year') %>%
  select(c("iso3c", "aquatic_estimate" = "estimate"))


# Fit linear models to estimate the average annual change in terrestrial protein
# consumption (2010-2020)
terrestrial_protein_quant <- aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         data_source == "new FBS", food_group == "terrestrial") %>%
  group_by(iso3c, year) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap, na.rm = TRUE)) %>%
  distinct() %>%
  split(.$iso3c) %>%
  map(~lm(protein_consumed_kg_percap~year, data = .x)) %>%
  map_df(broom::tidy, .id = 'iso3c') %>%
  filter(term == 'year') %>%
  select(c("iso3c", "terrestrial_estimate" = "estimate"))

# Fit linear models to estimate the average annual change in aggregate
# global aquatic animal reliance (2010-2020) 
aquatic_protein_reliance <- aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, year) %>%
  summarize(prop_animal_protein = sum(prop_animal_protein, na.rm = TRUE)) %>%
  distinct() %>%
  split(.$iso3c) %>%
  map(~lm(prop_animal_protein~year, data = .x)) %>%
  map_df(broom::tidy, .id = 'iso3c') %>%
  filter(term == 'year') %>%
  select(c("iso3c", "aquatic_reliance_estimate" = "estimate"))
  
#join data 
plot_df<- aquatic_protein_quant %>%
  full_join(terrestrial_protein_quant, by="iso3c") %>%
  full_join(aquatic_protein_reliance, by="iso3c") %>%
  add_region(col = "iso3c", region.col.name = "region")


# Create color scale for regions
region_colors <- c(
  "Africa"        = "#b2182b",  # deep red
  "Asia"          = "#2166ac",  # deep blue
  "Europe"        = "#6baed6",  # light blue
  "North America" = "#ef8a62",  # light red/salmon
  "Oceania"       = "#67a9cf",  # mid blue
  "South America" = "#d6604d"   # mid red
)

# Get region legend from bar plot
p_bar <- plot_df %>%
  mutate(quadrant = case_when(
    aquatic_estimate > 0 & terrestrial_estimate > 0 ~ "Q1",
    aquatic_estimate < 0 & terrestrial_estimate > 0 ~ "Q2",
    aquatic_estimate < 0 & terrestrial_estimate < 0 ~ "Q3",
    aquatic_estimate > 0 & terrestrial_estimate < 0 ~ "Q4"
  )) %>%
  group_by(quadrant, region) %>%
  summarize(n = n_distinct(iso3c), .groups = "drop") %>%
  mutate(quadrant = factor(quadrant, levels = c("Q2","Q1","Q3","Q4"))) %>%
  ggplot(aes(x = region, y = n, fill = region)) +
  geom_col(width = 0.75) +
  scale_fill_manual(values = region_colors) +
  labs(x = "", y = "Number of countries", fill = "Region") +
  theme_minimal(base_size = 7) +
  theme(
    axis.text.x = element_blank(),
    legend.position = "bottom",
    legend.title = element_text(size = 6, face = "bold"),
    legend.text  = element_text(size = 5),
    legend.key.size = unit(0.6, "lines"),
    axis.title = element_text(size = 5)
  ) +
  facet_wrap(~ quadrant)

legend_regions <- get_plot_component(p_bar, "guide-box-bottom", return_all = TRUE)

# Scatterplot with legend
p_scatter_base <- plot_df %>%
  filter(iso3c != "MDV") %>%
  ggplot(aes(
    x = aquatic_estimate,
    y = terrestrial_estimate,
    fill = aquatic_reliance_estimate,
    label = iso3c
  )) +
  geom_point(shape = 21, size = 1.8, color = "black", stroke = 0.2) +
  lims(x = c(-0.25,0.25),
       y = c(-4.5,4.5)) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  geom_text_repel(size = 2.4) +
  scale_fill_gradient2(
  low = "#ca0020",     # muted red
  mid = "#f7f7f7",     # very light grey
  high = "#0571b0",    # muted blue
  midpoint = 0
) +
  labs(
    x = "Average annual change in daily aquatic animal protein consumption (kg/capita/day) 2010–2020",
    y = "Average annual change in daily\nterrestrial animal protein consumption (kg/capita/day) 2010–2020",
    fill = "Annual change in daily aquatic\nanimal protein reliance (%) 2010-2020"
    
    
  ) +
  theme_minimal(base_size = 8) +
  theme(
    panel.grid.minor = element_blank(),
    axis.text = element_text(size = 6, color = "black"),
    axis.title = element_text(size = 5, face = "bold"),
    legend.position = "bottom",
    legend.title = element_text(size = 6, face = "bold"),
    legend.text = element_text(size = 5, angle = 35),
    legend.key.size = unit(.8, "lines")
  )

# Extract aquatic reliance legend
legend_reliance <- get_plot_component(p_scatter_base, "guide-box-bottom", return_all = TRUE)

# Remove legend from scatter for main panel
p_scatter <- p_scatter_base + theme(legend.position = "none")

x_max <- max(abs(plot_df$aquatic_estimate), na.rm = TRUE)
y_max <- max(abs(plot_df$terrestrial_estimate), na.rm = TRUE)

# p_scatter <- p_scatter +
#   coord_cartesian(
#     xlim = c(-0.5,0.5),
#     ylim = c(-y_max, y_max)
#   )

# Quadrant counts + insets
quad_counts <- plot_df %>%
  mutate(quadrant = case_when(
    aquatic_estimate > 0 & terrestrial_estimate > 0 ~ "Q1",
    aquatic_estimate < 0 & terrestrial_estimate > 0 ~ "Q2",
    aquatic_estimate < 0 & terrestrial_estimate < 0 ~ "Q3",
    aquatic_estimate > 0 & terrestrial_estimate < 0 ~ "Q4"
  )) %>%
  group_by(quadrant, region) %>%
  summarise(n = n_distinct(iso3c), .groups = "drop")

max_count <- max(quad_counts$n)

make_inset <- function(q) {
  show_y <- q %in% c("Q1", "Q4")
  
  quad_counts %>%
    filter(quadrant == q) %>%
    ggplot(aes(x = region, y = n, fill = region)) +
    geom_col(width = 0.75) +
    scale_fill_manual(values = region_colors) +
    scale_y_continuous(limits = c(0, max_count),
                       breaks = c(0, max_count),
                       position = "right") +
    labs(y = if (show_y) "No. countries" else NULL) +  # label only on right insets
    theme_minimal(base_size = 6) +
    theme(
      axis.title.x  = element_blank(),
      axis.title.y  = if (show_y) element_text(size = 5, color = "black", angle = 90) else element_blank(),
      axis.text.x   = element_blank(),
      axis.text.y   = if (show_y) element_text(size = 5, color = "black") else element_blank(),
      axis.ticks.y  = if (show_y) element_line(linewidth = 0.3) else element_blank(),
      panel.grid    = element_blank(),
      legend.position = "none",
      plot.margin   = margin(0, 0, 0, 0)
    )
}

p_q1 <- make_inset("Q1")
p_q2 <- make_inset("Q2")
p_q3 <- make_inset("Q3")
p_q4 <- make_inset("Q4")

# Add insets inside scatterplot

shrink_x <- 0.18
shrink_y <- 0.18

p_with_insets <- p_scatter +
  inset_element(p_q1, left = 1 - shrink_x, right = 1, bottom = 1 - shrink_y, top = 1) +
  inset_element(p_q2, left = 0, right = shrink_x, bottom = 1 - shrink_y, top = 1) +
  inset_element(p_q3, left = 0, right = shrink_x, bottom = 0, top = shrink_y) +
  inset_element(p_q4, left = 1 - shrink_x, right = 1, bottom = 0, top = shrink_y)

scatter_with_axis <- cowplot::plot_grid(
  p_with_insets,
  ncol = 2,
  rel_widths = c(1, 0.10)
)

# Add reliance and region legends side by side

legend_row <- cowplot::plot_grid(
  legend_reliance,
  legend_regions,
  ncol = 2,
  rel_widths = c(1, 1)
)

# Compile plot

(final_plot <- cowplot::plot_grid(
  scatter_with_axis,
  legend_row,
  ncol = 1,
  rel_heights = c(1, 0.18)
))


# Test for whether the median value of terrestrial protein consumption is greater than the aquatic animal protein consumption
wilcox.test(plot_df$terrestrial_estimate, plot_df$aquatic_estimate, paired = FALSE, alternative = "greater")

# Calculate CV for the terrestrial consumption roc
cv_terr <- sd(plot_df$terrestrial_estimate) / mean(plot_df$terrestrial_estimate)
 
 # Calculate CV for the aquatic animal consumption roc
cv_aquatic <- sd(plot_df$aquatic_estimate) / mean(plot_df$aquatic_reliance_estimate)

# Calcualte CV difference
cv_diff <- abs(cv_aquatic / cv_terr) %>% round(2)


ggsave("../images/scatter_bar_overlay.jpg", plot = final_plot, device = "jpg", units = "in", width = 5, height = 3.5)
```


```{r hierarcical global variable posteriors}
model_type = "realm"

# Create a data frame of posteriors of partial slopes across Bayesian models
slopes_df <- data.frame()

for (i in seq_along(response_vars)) {
  
  # Extract MCMC samples
  x <- rstan::extract(readRDS(file.path(base_path, paste0("fit_", response_vars[i], ".rds"))), permuted = T)
  beta_global <- x$beta       # iter × K (slopes)
  
  X <- readRDS(file.path(base_path, paste0("X_", response_vars[i], ".rds")))
  
  n_iter <- nrow(beta_global)
  K <- ncol(beta_global)
  vars <- colnames(X)

# Extract global slopes
post_long_slopes <- expand_grid(
  iter = 1:n_iter,
  k = 1:K
) %>%
  mutate(
    value = map2_dbl(iter, k, ~beta_global[.x, .y]),
    var = vars[k]
  )

# Calculate credibility intervals
ci_df <- post_long_slopes %>%
  group_by(var) %>%
  summarise(
    lo95 = quantile(value, 0.025),
    lo80 = quantile(value,0.1),
    lo50 = quantile(value, 0.25),
    med = quantile(value, 0.5),
    hi50 = quantile(value, 0.75),
    hi80 = quantile(value,0.9),
    hi95 = quantile(value, 0.975),
    .groups = "drop"
  ) %>%
  mutate(model =  response_vars[i])

slopes_df <- slopes_df %>%
  bind_rows(ci_df)

}

# Clean variable names
var_labels <- c(
  coastline_km = "Coastline length (km)",
  # eez_total = "EEZ area",
  gdp_trade = "GDP from trade (%)",
  inundation_area_log1000km2 = "Inundated area (log 1000 km²)",
  log_gdp = "GDP (log scale)",
  pop_10k_pct_coast = "Population within 10 km of coast (%)",
  supermarkets = "Supermarkets per 100,000 people"
)

# Set model colors for visualization
model_colors <- c(
  "Terrestrial animal consumption per capita" = "#8B4513",
  "Aquatic animal consumption per capita"     = "#3A77B5",
  "Aquatic reliance"                          = "#2166ac",
  "Capture consumption per capita"            = "#E68A2E",
  "Aquaculture consumption per capita"        = "#D8C84A",
  "Domestic consumption per capita"           = "#6BAED6",
  "Foreign consumption per capita"            = "#EF8A62",
  "Marine consumption per capita"             = "#7A5FA3",
  "Inland consumption per capita"             = "#D8A43A"
)

# Get colors for 90, 80, and 50% credibility intervals
col_95 <- lighten(model_colors, 0.65)
col_80 <- lighten(model_colors, 0.4)
col_50 <- lighten(model_colors, 0.10)

# Load in script to clean model names from underscores to be easily readable
source("../R/clean_model_name.R")

# 2. Add clean_name once
slopes_df <- slopes_df %>%
  mutate(clean_name = clean_model_name(model))

# 3. Build plotting data
df_plot <- slopes_df %>%
  mutate(
    clean_name = clean_model_name(model),
    nonzero = (lo95 < 0 & hi95 < 0) | (lo95 > 0 & hi95 > 0),
    suffix = sub(".*_", "", model),
    suffix_num = suppressWarnings(as.numeric(suffix))
  )  %>%
  mutate(
    nonzero    = (lo95 < 0 & hi95 < 0) | (lo95 > 0 & hi95 > 0),
    suffix     = sub(".*_", "", model),
    suffix_num = suppressWarnings(as.numeric(suffix))
  ) %>%
  filter(
    str_detect(clean_name, "capita") |
    str_detect(clean_name, "Aquatic reliance"),
    !model == "animal_cons_cap"
  ) %>%
  
  arrange(suffix_num, suffix) %>%
  mutate(
    clean_name = factor(clean_name, levels = unique(clean_name)),
    col95 = col_95[as.character(clean_name)],
    col80 = col_80[as.character(clean_name)],
    col50 = col_50[as.character(clean_name)],
    panel_id = as.numeric(factor(clean_name)),
    left_col = panel_id %% 3 == 1
  )

panel_tags <- letters[1:9] 

# Function to draw a single panel
make_panel <- function(df, show_yaxis = FALSE, tag = "") {
  ggplot(df, aes(y = var)) +
    geom_segment(aes(x = lo95, xend = hi95, color = col95),
                 linewidth = 2, lineend = "square") +
    geom_segment(aes(x = lo80, xend = hi80, color = col80),
                 linewidth = 2, lineend = "square") +
    geom_segment(aes(x = lo50, xend = hi50, color = col50),
                 linewidth = 2, lineend = "square") +
    geom_point(aes(x = med), size = 1.8, color = "black") +
    geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.4) +
    scale_color_identity() +
    scale_y_discrete(labels = var_labels) +

    # ← add tag in upper-left corner
    annotate("text", x = -Inf, y = Inf, label = tag,
             hjust = -0.3, vjust = 1.3, size = 3.2, fontface = "bold") +

    theme_minimal(base_size = 8) +
    theme(
      plot.title = element_blank(),
      panel.grid.major.y = element_blank(),
      panel.grid.minor   = element_blank(),
      strip.text         = element_text(size = 5, face = "bold"),
      axis.text.x        = element_text(size = 6, color = "black"),
      axis.title         = element_blank(),
      panel.spacing      = unit(1.2, "lines"),
      plot.margin        = margin(5, 5, 5, 5),
      axis.text.y        = if (show_yaxis) element_text(size = 6, color = "black") else element_blank(),
      axis.ticks.y       = if (show_yaxis) element_line() else element_blank()
    )
}

# Split df_plot into 9 panels
panel_names <- levels(df_plot$clean_name)

panels <- lapply(seq_along(panel_names), function(i) {
  name <- panel_names[i]
  df_sub <- df_plot %>% filter(clean_name == name)

  show_y <- i %in% c(1, 4, 7)
  wrapped_name <- stringr::str_wrap(name, width = 20)

  make_panel(df_sub, show_yaxis = show_y, tag = panel_tags[i]) +
    ggtitle(wrapped_name)
})

# Build the x-axis label as its own plot
xlab_plot <- ggplot() +
  theme_void() +
  annotate("text", x = 0.5, y = 0.5,
           label = "Posterior slope estimate",
           size = 2, fontface = "bold")

# Empty plot for left and right columns
empty <- ggplot() + theme_void()

# Assemble 3×3 grid + bottom row with centered x-label
final_plot_with_xlab <-
  (
    (panels[[1]] | panels[[2]] | panels[[3]]) /
    (panels[[4]] | panels[[5]] | panels[[6]]) /
    (panels[[7]] | panels[[8]] | panels[[9]]) /
    (empty | xlab_plot | empty)
  ) +
  plot_layout(heights = c(1, 1, 1, 0.12))

final_plot_with_xlab


ggsave(paste0("../images/posteriors_partial_slopes.jpg"), plot = final_plot_with_xlab, device = "jpg", width = 5, height = 3, units = "in", dpi = 300)
```


```{r individual continent production contribution}
# Fit linear models to estimate the average annual change in disaggregated
# aquatic animal reliance (2010-2020) by production source (habitat x method)
# for each country
disagg <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "aquatic") %>%
  group_by(year, iso3c, habitat, method) %>%
  summarize(prop_animal_protein = sum(prop_animal_protein)) %>%
  group_by(iso3c, habitat, method) %>%
  nest() %>%
  mutate(
    model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
    tidy  = map(model, broom::tidy)
  ) %>%
  unnest(tidy) %>%
  filter(term == "year") %>%
  mutate(
    consource_habitat_method = paste(habitat, method),
    group_label = consource_habitat_method,
    group_label = case_when(
      group_label == "unknown unknown" ~ "unknown",
      TRUE ~ group_label
    )
  ) %>%
  drop_na(estimate) %>%
  group_by(group_label) %>%
  mutate(pct_above = mean(estimate > 0) * 100) %>%
  ungroup()

# Fit linear models to estimate the average annual change in aggregate
# aquatic animal reliance (2010-2020)

agg <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, year) %>%
  summarise(prop_animal_protein = sum(prop_animal_protein), .groups = "drop") %>%
  group_by(iso3c) %>%
  nest() %>%
  mutate(
    model = map(data, ~ lm(prop_animal_protein ~ year, data = .x)),
    tidy  = map(model, broom::tidy)
  ) %>%
  unnest(tidy) %>%
  filter(term == "year") %>%
  mutate(group_label = "aggregate") %>%
  drop_na(estimate) %>%
  group_by(group_label) %>%
  mutate(pct_above = mean(estimate > 0) * 100) %>%
  ungroup()

# Clean aggregate estimate data
agg_estimate <- bind_rows(agg, disagg) %>%
  group_by(iso3c) %>%
  mutate(agg_est = estimate[group_label == "aggregate"][1]) %>%
  add_region("iso3c", "region") %>%
  ungroup() %>%
  mutate(
    iso3c  = fct_reorder(iso3c, -agg_est),
    is_agg = group_label == "aggregate"
  ) %>%
  filter(group_label == "aggregate") %>%
  select(iso3c, agg_estimate = estimate)

# Set plot title size
axis.title.size <- 8

# Iterate through each region (i.e., i = region) and produce a plot to be saved for each region
for (i in c("Asia", "Africa", "Europe", "North America", "South America", "Oceania")) {

# Plot disaggregated aquatic animal reliance by production source (2010-2020)
# for each country in Asia, ordered by aggregate rate of change 
plot <- disagg %>%
  group_by(iso3c) %>%
  add_region("iso3c", "region") %>%
  ungroup() %>%
  left_join(agg_estimate, by = "iso3c") %>%
  mutate(iso3c = fct_reorder(iso3c, -agg_estimate)) %>%
  filter(region == i) %>%
  ggplot() +
  geom_col(aes(x = estimate, y = iso3c, fill = group_label)) +
  geom_point(aes(x = agg_estimate, y = iso3c), shape = 16) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.4) +
  scale_alpha_identity() +
  scale_size_identity() +
  scale_fill_manual(values = c("#D8C84A","#E68A2E", "#7A5FA3", "#3A77B5", "gray")) +
  guides(fill = guide_legend(ncol = 2)) +
  theme(axis.text.y = element_text(size = 4.75),
        axis.title.x = element_text(size = axis.title.size, face = "bold"),
    panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(color = scales::alpha("black", 0.05), linewidth = 0.3),
    panel.grid.major.x = element_line(color = scales::alpha("black", 0.25), linewidth = 0.3),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),
    axis.text = element_text(size = 6, color = "black"),
    axis.title.y = element_blank(),
    legend.position = "bottom",
    legend.title = element_text(size = 7, face = "bold"),
    legend.text = element_text(size = 6),
    legend.key.size = unit(0.25, "cm"),
    legend.box = "vertical",
    legend.spacing = unit(.05, "cm"),
    legend.margin = margin(t = 0, b = 0),
    legend.box.margin = margin(t = -2, b = -2),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA)) +
      labs(fill = "Production\nsource",
           x = "Average annual change in aquatic animal reliance (2010–2020)")

# Lana's analysis calcuating the # of countries with increasing reliance and decreasing relaince from 2010-2020.

# Extract rate of change in aggregate aquatic animal reliance from 2010-2020 
roc_aquatic_total <- aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, year) %>%
  summarize(prop_animal_protein = sum(prop_animal_protein, na.rm = TRUE), .groups = "drop") %>%
  distinct() %>%
  split(list(.$iso3c), drop = TRUE) %>%
  map(~lm(prop_animal_protein~year, data = .x)) %>%
  map_df(broom::tidy, .id = 'combined_id') %>%
  filter(term == 'year') %>%
  select(c("iso3c"="combined_id", "roc" = "estimate"))%>%
  mutate(direction = case_when(roc > 0 ~ "inc", roc < 0 ~ "dec"))%>%
  mutate(method_habitat = "total")%>%
  add_region(col = "iso3c", region.col.name = "continent")

# Get the counts by continent
Counts_by_continent<-roc_aquatic_total%>%
  group_by(continent, method_habitat)%>%
  count(direction)%>%
  pivot_wider(names_from = direction, values_from = n) %>%
  rename(decreasing = dec, increasing = inc) %>%
  pivot_longer(cols = c(decreasing, increasing), names_to = "direction",values_to = "count") %>%
  group_by(continent) %>%
  mutate(prop = count / sum(count))

# Plot counts by target continent
region_counts <- Counts_by_continent %>%
  filter(method_habitat == "total", continent == i) %>%
  ggplot(aes(y = continent, x = count, fill = direction)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("#b2182b", "#2166ac"),
                    labels  = c("decreasing" = "Declining reliance",
              "increasing" = "Increasing reliance")) +
  guides(fill = guide_legend(nrow = 2)) +
  labs(x = "Number of countries with increasing\nor decreasing reliance (aggregate)", fill = "Direction of change") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text = element_text(size = 4.75),
        axis.title.y = element_blank(),
        axis.title.x = element_text(size = axis.title.size-2.5, face = "bold"),
    panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_line(color = scales::alpha("black", 0.25), linewidth = 0.3),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),
    legend.position = "bottom",
    legend.title = element_text(size = 7, face = "bold"),
    legend.text = element_text(size = 6),
    legend.key.size = unit(0.25, "cm"),
    legend.box = "vertical",
    legend.spacing = unit(.05, "cm"),
    legend.margin = margin(t = 0, b = 0),
    legend.box.margin = margin(t = -2, b = -2),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    aspect.ratio = 0.15)

num_regional_countries <- disagg %>%
  add_region("iso3c", "region") %>%
  filter(region == i) %>%
  distinct(iso3c) %>% 
  nrow()

# Proportion of countries with increasing reliance by production source for each region
(disagg_props <- disagg %>%
  add_region("iso3c", "region") %>%
  filter(region == i) %>%
  group_by(region, group_label) %>%
    count(count = estimate > 0) %>%
    group_by(group_label) %>%
  ggplot(aes(y = group_label, x = n, fill = count)) +
  geom_col() +
  scale_fill_manual(values = c("#b2182b", "#2166ac")) +
  guides(fill = "none")  +
  labs(x = paste0("Number of countries with increasing\nor decreasing reliance (disaggregate) (n = ", num_regional_countries, ")")) +
  theme(axis.title.y = element_blank(),
      axis.text = element_text(size = 4.75),
      axis.title.x = element_text(size = axis.title.size-2.5, face = "bold"),
      panel.grid.minor.y = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.y = element_blank(),
      panel.grid.major.x = element_line(color = scales::alpha("black", 0.25), linewidth = 0.3),
      panel.background = element_rect(fill = "transparent", color = NA),
      plot.background = element_rect(fill = "transparent", color = NA)))

# Plot Aggregate aquatic animal reliance for target continent's region,
# ordered by the rate of change in aggregate aquatic animal reliance 
# from 2010-2020
regional_reliance <- aquatic_reliance %>%
    filter(year == 2020, food_group == "aquatic") %>%
    group_by(iso3c) %>%
    summarize(prop_animal_protein = sum(prop_animal_protein)) %>%
    left_join(agg_estimate, by = "iso3c") %>%
    add_region("iso3c", "region") %>%
    filter(region == i) %>%
    ggplot(aes(
      x = fct_reorder(iso3c, agg_estimate), y = prop_animal_protein * 100
    )) +
    geom_col(fill = "#555555") +
    theme(
      axis.text.x = element_text(size = 4.5, angle = 45),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 4.5),
      axis.title.y = element_text(size = axis.title.size, face = "bold"),
      panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(color = scales::alpha("black", 0.25), linewidth = 0.3),
      panel.background = element_rect(fill = "transparent", color = NA),
      plot.background = element_rect(fill = "transparent", color = NA),
      axis.text = element_text(size = 6, color = "black")
    ) +
    labs(y = "Aquatic animal\nreliance, 2020 (%)")


# Function to extract legends
extract_legend <- function(p) {
  legs <- cowplot::get_plot_component(p, "guide-box", return_all = TRUE)
  # Filter out empty (zeroGrob) legends
  good <- Filter(function(x) !inherits(x, "zeroGrob"), legs)
  if (length(good) == 0) return(NULL)
  good[[1]]
}

# Extract legends
legend_plot          <- extract_legend(plot)
legend_region_counts <- extract_legend(region_counts)

# Remove legends from panels
plot_nolegend          <- plot          + theme(legend.position = "none")
region_counts_nolegend <- region_counts + theme(legend.position = "none")

# Combinelegends
legend_row <- cowplot::plot_grid(
  legend_region_counts,
  legend_plot,
  ncol       = 2,
  rel_widths = c(1, 1.5)
)

title <- ggplot() +
  labs(title = i) +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

divider <- ggplot() +
  theme_void() +
  theme(panel.border = element_blank()) +
  geom_hline(yintercept = 0.5, linewidth = 0.5, color = "black")

# Row for global counts and disaggregated counts
top_plot_panel <- cowplot::plot_grid(region_counts_nolegend, disagg_props)
    

# Final plot with legends at the bottom
(final_plot <- cowplot::plot_grid(
  title,
  divider,                    # <-- line between title and panels
  top_plot_panel,
  regional_reliance,
  plot_nolegend,
  legend_row,
  ncol        = 1,
  rel_heights = c(0.15, 0.05, 0.65, 0.85, 2.5, 0.35)  # very small height for divider
))

# Save plot
ggsave(paste0("../images/SI/multi_metrics_",i,"_SI.jpg"), plot = final_plot, device = "jpg", units = "in", width = 5, height = 7, dpi = 500)
}
```

```{r read in and visualize posterior distributional data}
# Variable list with coastline_km area instead of EEZ area
predictors <- c(
  "coastline_km",
  "log_gdp",
  "gdp_trade",
  "inundation_area_log1000km2",
  "supermarkets",
  "pop_10k_pct_coast"
)

# Variable list with EEZ total area instead of coastline_km
# predictors <- c(
#   "eez_total",
#   "log_gdp",
#   "gdp_trade",
#   "inundation_area_log1000km2",
#   "supermarkets",
#   "pop_10k_pct_coast"
# )

# Get which random effect is used
model_type = "realm"

# Get which variable is selected (for unique file path)
var_used = predictors[1]

# Create file path to save files to
  base_path <- file.path(
    "..", "output", "bayesian_output",
    paste0("random_effect_", model_type),
    paste0("var_", var_used)
  )

# Create data frame to hold all posteriors across variables
slopes_df <- data.frame()

for (i in seq_along(response_vars)) {
  
  # Extract MCMC samples
  x <- rstan::extract(readRDS(file.path(base_path, paste0("fit_", response_vars[i], ".rds"))), permuted = T)
  beta_global <- x$beta       # iter × K (slopes)
  
  X <- readRDS(file.path(base_path, paste0("X_", response_vars[i], ".rds")))
  
  n_iter <- nrow(beta_global)
  K <- ncol(beta_global)
  vars <- colnames(X)

# Extract global slopes
post_long_slopes <- expand_grid(
  iter = 1:n_iter,
  k = 1:K
) %>%
  mutate(
    value = map2_dbl(iter, k, ~beta_global[.x, .y]),
    var = vars[k]
  )

# Calculate credibility intervals
ci_df <- post_long_slopes %>%
  group_by(var) %>%
  summarise(
    lo95 = quantile(value, 0.025),
    lo80 = quantile(value,0.1),
    lo50 = quantile(value, 0.25),
    med = quantile(value, 0.5),
    hi50 = quantile(value, 0.75),
    hi80 = quantile(value,0.9),
    hi95 = quantile(value, 0.975),
    .groups = "drop"
  ) %>%
  mutate(model =  response_vars[i])

slopes_df <- slopes_df %>%
  bind_rows(ci_df)

}

# Clean model names
clean_model_name <- function(x) {

  # Suffix dictionary
  suffix_map <- c(
    cap = "consumption per capita",
    rel = "reliance",
    roc = "rate of change",
    prop_of_aquatic_cons = "share of aquatic foods"
  )

  # Domain dictionary
  domain_map <- c(
    terr_animal = "Terrestrial animal",
    terrestrial = "Terrestrial",
    animal = "Animal",
    aquatic_animal = "Aquatic animal",
    aquatic = "Aquatic",
    capture = "Capture",
    aquaculture = "Aquaculture",
    domestic = "Domestic",
    foreign = "Foreign",
    marine = "Marine",
    inland = "Inland"
  )

  # Metric dictionary
  metric_map <- c(
    cons = "consumption",
    protein = "protein"
  )

  # Split name
  parts <- strsplit(x, "_")[[1]]

  # Detect suffix safely
  if (endsWith(x, "prop_of_aquatic_cons")) {
    suffix <- "prop_of_aquatic_cons"
    core <- parts[1:(length(parts)-4)]
  } else {
    suffix <- parts[length(parts)]
    core <- parts[1:(length(parts)-1)]
  }

  suffix_clean <- suffix_map[[suffix]]

  # Detect domain (first one or two parts)
  if (length(core) >= 2 && paste(core[1:2], collapse = "_") %in% names(domain_map)) {
    domain <- paste(core[1:2], collapse = "_")
    rest <- core[-(1:2)]
  } else {
    domain <- core[1]
    rest <- core[-1]
  }

  domain_clean <- domain_map[[domain]]

  # Detect metric (optional)
  if (length(rest) >= 1 && rest[1] %in% names(metric_map)) {
    metric_clean <- metric_map[[rest[1]]]
  } else {
    metric_clean <- NULL
  }

  # Build label
  if (!is.null(metric_clean) && suffix != "prop_of_aquatic_cons") {
    paste(domain_clean, metric_clean, suffix_clean)
  } else {
    paste(domain_clean, suffix_clean)
  }
}

# Clean input variable names to publication friendly text
var_labels <- c(
  coastline_km = "Coastline length (km)",
  gdp_trade = "GDP from trade (%)",
  inundation_area_log1000km2 = "Inundated area (log 1000 km²)",
  log_gdp = "GDP (log scale)",
  prop_pop_coast = "Population within 10 km of coast (%)",
  supermarkets = "Supermarkets per 100,000 people"
)

# Base colors per model
model_colors <- c(
  "Aquatic reliance" = "#1f78b4",
  "Aquaculture reliance"                                      = "#e31a1c",
  "Capture reliance"                    = "#6a3d9a",
  "Domestic reliance"                       = "#b15928",
  "Foreign reliance"                        = "#a6cee3",
  "Marine reliance"                         = "#b2df8a",
  "Inland reliance"                         = "#fb9a99"
)

col_95 <- lighten(model_colors, 0.65)
col_80 <- lighten(model_colors, 0.4)
col_50 <- lighten(model_colors, 0.10)

# Add clean_name once
slopes_df <- slopes_df %>%
  mutate(clean_name = sapply(model, clean_model_name))

# Build plotting data
df_plot <- slopes_df %>%
  mutate(
    clean_name = sapply(model, clean_model_name),
    nonzero = (lo95 < 0 & hi95 < 0) | (lo95 > 0 & hi95 > 0),
    suffix = sub(".*_", "", model),
    suffix_num = suppressWarnings(as.numeric(suffix))
  )  %>%
  filter(model != "animal_cons_cap") %>%
  mutate(
    nonzero    = (lo95 < 0 & hi95 < 0) | (lo95 > 0 & hi95 > 0),
    suffix     = sub(".*_", "", model),
    suffix_num = suppressWarnings(as.numeric(suffix))
  ) %>%
  filter(
    str_detect(clean_name, "reliance")
  ) %>%
  arrange(suffix_num, suffix) %>%
  mutate(
    clean_name = factor(clean_name, levels = unique(clean_name)),
    col95 = col_95[as.character(clean_name)],
    col80 = col_80[as.character(clean_name)],
    col50 = col_50[as.character(clean_name)],
    panel_id = as.numeric(factor(clean_name)),
    left_col = panel_id %% 3 == 1
  )

library(dplyr)
library(ggplot2)
library(patchwork)

panel_tags <- letters[1:9] 


# Function to draw a single panel
make_panel <- function(df, show_yaxis = FALSE, tag = "") {
  ggplot(df, aes(y = var)) +
    geom_segment(aes(x = lo95, xend = hi95, color = col95),
                 linewidth = 2, lineend = "square") +
    geom_segment(aes(x = lo80, xend = hi80, color = col80),
                 linewidth = 2, lineend = "square") +
    geom_segment(aes(x = lo50, xend = hi50, color = col50),
                 linewidth = 2, lineend = "square") +
    geom_point(aes(x = med), size = 1.8, color = "black") +
    geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.4) +
    scale_color_identity() +
    scale_y_discrete(labels = var_labels) +

    # ← add tag in upper-left corner
    annotate("text", x = -Inf, y = Inf, label = tag,
             hjust = -0.3, vjust = 1.3, size = 3.2, fontface = "bold") +

    theme_minimal(base_size = 8) +
    theme(
      plot.title = element_blank(),
      panel.grid.major.y = element_blank(),
      panel.grid.minor   = element_blank(),
      strip.text         = element_text(size = 5, face = "bold"),
      axis.text.x        = element_text(size = 6, color = "black"),
      axis.title         = element_blank(),
      panel.spacing      = unit(1.2, "lines"),
      plot.margin        = margin(5, 5, 5, 5),
      axis.text.y        = if (show_yaxis) element_text(size = 6, color = "black") else element_blank(),
      axis.ticks.y       = if (show_yaxis) element_line() else element_blank()
    )
}

# --- 2. Split df_plot into 9 panels ---
panel_names <- levels(df_plot$clean_name)

panels <- lapply(seq_along(panel_names), function(i) {
  name <- panel_names[i]
  df_sub <- df_plot %>% filter(clean_name == name)

  show_y <- i %in% c(1, 4, 7)
  wrapped_name <- stringr::str_wrap(name, width = 20)

  make_panel(df_sub, show_yaxis = show_y, tag = panel_tags[i]) +
    ggtitle(wrapped_name)
})

library(ggplot2)
library(patchwork)

# x-axis label plot
xlab_plot <- ggplot() +
  theme_void() +
  annotate("text", x = 0.5, y = 0.5,
           label = "Posterior slope estimate",
           size = 2, fontface = "bold")

# empty spacer
empty <- ggplot() + theme_void()

# --- assemble ---
row1 <- (panels[[1]] | panels[[2]] | panels[[3]])
row2 <- (panels[[4]] | panels[[5]] | panels[[6]])
row3 <- (empty | panels[[7]] | empty) + plot_layout(widths = c(0.135, 1, 1))
row4 <- (empty | xlab_plot | empty) + plot_layout(widths = c(1, 1, 1))

(final_plot_with_xlab <- wrap_plots(row1, row2, row3, row4, ncol = 1) +
  plot_layout(heights = c(1, 1, 1, 0.12)))

ggsave(
  filename = "../images/SI/posteriors_partial_slopes_SI.pdf",
  plot     = final_plot_with_xlab,
  device   = cairo_pdf,
  width    = 5,
  height   = 3,
  units    = "in"
)
```


```{r proportion of countries in each quadrant by region}
# Calculate rate of change in aquatic protein quantity
aquatic_protein_quant <- aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, year) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap, na.rm = TRUE)) %>%
  distinct() %>%
  split(.$iso3c) %>%
  map(~lm(protein_consumed_kg_percap ~ year, data = .x)) %>%
  map_df(broom::tidy, .id = 'iso3c') %>%
  filter(term == 'year') %>%
  select(iso3c, aquatic_estimate = estimate)

# Calculate rate of change in terrestrial protein quantity
terrestrial_protein_quant <- aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         data_source == "new FBS", food_group == "terrestrial") %>%
  group_by(iso3c, year) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap, na.rm = TRUE)) %>%
  distinct() %>%
  split(.$iso3c) %>%
  map(~lm(protein_consumed_kg_percap ~ year, data = .x)) %>%
  map_df(broom::tidy, .id = 'iso3c') %>%
  filter(term == 'year') %>%
  select(iso3c, terrestrial_estimate = estimate)

# Calculate rate of change in aquatic reliance
aquatic_protein_reliance <- aquatic_reliance %>%
  filter(!(year %in% c(2010, 2011, 2012, 2013) & data_source == "old FBS"),
         data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, year) %>%
  summarize(prop_animal_protein = sum(prop_animal_protein, na.rm = TRUE)) %>%
  distinct() %>%
  split(.$iso3c) %>%
  map(~lm(prop_animal_protein ~ year, data = .x)) %>%
  map_df(broom::tidy, .id = 'iso3c') %>%
  filter(term == 'year') %>%
  select(iso3c, aquatic_reliance_estimate = estimate)

# Join and compute quadrant proportions per region
quad_df <- aquatic_protein_quant %>%
  full_join(terrestrial_protein_quant, by = "iso3c") %>%
  full_join(aquatic_protein_reliance, by = "iso3c") %>%
  add_region(col = "iso3c", region.col.name = "region") %>%
  mutate(
    quad_raw = case_when(
      aquatic_estimate > 0 & terrestrial_estimate > 0 ~ "Q1",
      aquatic_estimate < 0 & terrestrial_estimate > 0 ~ "Q2",
      aquatic_estimate < 0 & terrestrial_estimate < 0 ~ "Q3",
      aquatic_estimate > 0 & terrestrial_estimate < 0 ~ "Q4"
    )
  ) %>%
  group_by(region, quad_raw) %>%
  summarise(n = n_distinct(iso3c), .groups = "drop") %>%
  group_by(region) %>%
  mutate(
    prop = n / sum(n),
    quadrant = factor(
      quad_raw,
      levels = c("Q1", "Q2", "Q3", "Q4"),
      labels = c("Aquatic + / Terrestrial +",
                 "Aquatic - / Terrestrial +",
                 "Aquatic - / Terrestrial -",
                 "Aquatic + / Terrestrial -")
    )
  ) %>%
  ungroup()

# Plot parameters
max_val   <- 0.75
half_width <- 0.2

# Build spike polygons
# r = prop * max_val * sqrt(2) ensures that at prop=1, the spike tip's
# x and y components each equal max_val (the axis limit)
spike_df <- quad_df %>%
  mutate(
    center_angle = case_when(
      quad_raw == "Q1" ~ pi / 4,
      quad_raw == "Q2" ~ 3 * pi / 4,
      quad_raw == "Q3" ~ 5 * pi / 4,
      quad_raw == "Q4" ~ 7 * pi / 4
    ),
    r       = prop * max_val * sqrt(2),
    x_left  = r * cos(center_angle - half_width),
    y_left  = r * sin(center_angle - half_width),
    x_right = r * cos(center_angle + half_width),
    y_right = r * sin(center_angle + half_width)
  ) %>%
  rowwise() %>%
  reframe(
    region   = region,
    quadrant = quadrant,
    prop     = prop,
    x        = c(0, x_left, x_right),
    y        = c(0, y_left, y_right),
    group    = paste(region, quadrant)
  )

# Set color scale for plot
quad_colors <- c(
  "Aquatic + / Terrestrial +" = "#2a9d8f",
  "Aquatic - / Terrestrial +" = "#e76f51",
  "Aquatic - / Terrestrial -" = "#6d3b47",
  "Aquatic + / Terrestrial -" = "#457b9d" 
)

# Create plot
(plot <- ggplot(spike_df) +
  geom_polygon(
    aes(x = x, y = y, fill = quadrant, group = group),
    alpha = 0.85
  ) +
  geom_hline(yintercept = 0, linewidth = 0.4, colour = "black") +
  geom_vline(xintercept = 0, linewidth = 0.4, colour = "black") +
  facet_wrap(~ region) +
  coord_equal(xlim = c(-max_val, max_val), ylim = c(-max_val, max_val)) +
  scale_x_continuous(
    breaks = c(-max_val, max_val),
    labels = c("75%", "75%"),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    breaks = c(-max_val, max_val),
    labels = c("75%", "75%"),
    expand = c(0, 0)
  ) +
  scale_fill_manual(values = quad_colors) +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE)) +
  labs(
    x    = "Percentage of countries",
    y    = "Percentage of countries",
    fill = "Quadrant"
  ) +
  theme_minimal(base_size = 8) +
  theme(
    panel.grid            = element_blank(),
    panel.background      = element_rect(fill = "transparent", color = NA),
    plot.background       = element_rect(fill = "transparent", color = NA),
    strip.background      = element_blank(),
    strip.text            = element_text(size = 7, face = "bold"),
    axis.text             = element_text(size = 6, color = "black"),
    axis.title            = element_text(size = 7, face = "bold"),
    legend.position       = "bottom",
    legend.box            = "vertical",
    legend.title          = element_text(size = 7, face = "bold"),
    legend.text           = element_text(size = 6),
    legend.key.size       = unit(0.35, "cm"),
    legend.spacing        = unit(0.1, "cm"),
    legend.background     = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    panel.spacing         = unit(1, "lines")
  ))

# Save plot
ggsave("../images/SI/quadrant_bar_region_SI.jpg", plot = plot, device = "jpg", units = "in", width = 5, height = 3.5)
```

```{r collinearity plot}
# Significance stars — flag |r| > 0.7 as potentially collinear
upper_melt <- upper_melt %>%
  mutate(
    label     = sprintf("%.2f", value),
    flag_high = abs(value) >= 0.7
  )

# Bold asterisk on values where |r| >= 0.7 to flag high collinearity
  geom_text(
    data   = upper_melt %>% filter(flag_high),
    aes(x = Var2, y = Var1),
    label  = "",
    size   = 3.5,
    color  = "grey20",
    nudge_x = 0.28,
    nudge_y = 0.22
  ) 
```

```{r extra code for posterior map}


# theme_realm_panel <- function() {
#   theme_minimal(base_size = 8) +
#     theme(
#       panel.grid.major.x = element_blank(),
#       panel.grid.minor.x = element_blank(),
#       panel.grid.major.y = element_line(color = scales::alpha("black", 0.25), linewidth = 0.3),
#       panel.grid.minor.y = element_blank(),
#       panel.background = element_rect(fill = "transparent", color = NA),
#       plot.background = element_rect(fill = "transparent", color = NA),
#       axis.text = element_text(size = 6, color = "black"),
#       axis.title = element_blank(),
#       legend.position = "bottom",
#       legend.title = element_text(size = 7, face = "bold", hjust = 0.5),
#       legend.text = element_text(size = 6),
#       legend.key.size = unit(0.5, "cm"),
#       legend.box = "vertical",
#       legend.spacing = unit(0.1, "cm"),
#       legend.margin = margin(t = 0, b = 0),
#       legend.box.margin = margin(t = -2, b = -2),
#       legend.background = element_rect(fill = "transparent", color = NA),
#       legend.box.background = element_rect(fill = "transparent", color = NA)
#     )
# }

# make_realm_map <- function(model_name, df = intecrept_df) {
# 
#   # Filter model + significance flag
#   df_model <- df %>%
#     filter(model == model_name) %>%
#     mutate(significant = (lo95 > 0 & hi95 > 0) | (lo95 < 0 & hi95 < 0))
# 
#   # Join model values to realm geometry
#   geom_joined <- realm_geom_robin %>%
#     left_join(df_model, by = "realm") %>%
#     filter(!is.na(st_is_empty(geometry)) & !st_is_empty(geometry))
# 
#   ggplot() +
#     # Basemap (always visible)
#     geom_sf(
#       data = continent_basemap,
#       fill = "darkgray",
#       color = "grey70",
#       size = 0.2
#     ) +
# 
#     # Realm shading (outline only for non-NA)
#     geom_sf(
#       data = geom_joined,
#       aes(fill = med, color = !is.na(med)),
#       size = 0.2
#     ) +
#     scale_color_manual(
#       values = c("TRUE" = "grey30", "FALSE" = NA),
#       guide = "none"
#     ) +
# 
#     # Significance outlines
#     geom_sf(
#       data = geom_joined %>% filter(significant),
#       fill = NA,
#       color = "#1F1F1F",
#       linewidth = 0.5
#     ) +
# 
#     # Global symmetric fill scale (computed outside this function)
#     scale_fill_gradient2(
#       low = "#b2182b",
#       mid = "grey85",
#       high = "#2166ac",
#       midpoint = 0,
#       limits = c(-global_limit, global_limit),
#       oob = scales::squish
#     ) +
# 
#     # Hide legend inside each subplot
#     theme_realm_panel() +
#     theme(legend.position = "none") +
# 
#     labs(
#       title = paste(model_name),
#       fill = "Posterior\nmedian\nintercept"
#     )
# }
# 
# 
# # Produce maps for all models
# all_models <- unique(intecrept_df$model)
# realm_maps <- map(all_models, ~ make_realm_map(.x))
# names(realm_maps) <- all_models
# 
# # Create individual panels
# p1 <- realm_maps[["animal_cons_cap"]] +
#   labs(title = "Daily animal protein consumption per capita") +
#   theme_realm_panel() +
#   theme(plot.title = element_text(hjust = 0.5, size = 8, face = "bold"))
# 
# p2 <- realm_maps[["aquatic_rel"]] +
#   labs(title = "Aquatic animal reliance per capita") +
#   theme_realm_panel() +
#   theme(plot.title = element_text(hjust = 0.5, size = 8, face = "bold"))
# 
# # Create combined panel plot
# (plot <- (p1 | p2) +
#   plot_layout(guides = "collect") &
#   theme(
#     legend.position = "bottom",
#     aspect.ratio = 0.45,   # ← try 0.40–0.55 depending on taste
#     plot.margin = margin(0, 0, 0, 0) 
#   ))

# Save plot
# ggsave(paste0("../images/", model_type, "/intercept_map.jpg"), plot = plot, device = "jpg", width = 5, height = 4, units = "in", dpi = 300)
```

```{r create ppd plot}
# Read in posterior predictive distrbituion data
ppd_all_models <- read_parquet(
  paste0("../output/bayesian_output/", model_type, "/posterior_predictive_all_models.parquet")
)

# Get letters for panel numbers
panel_tags <- letters[1:15] 

# Function to draw a single panel ---
make_ppd_panel <- function(df, title, tag = "") {
  ggplot(df, aes(x = ppd, group = draw)) +
    geom_density(alpha = 0.08, color = "grey60") +
    geom_density(aes(x = y_obs), color = "black", linewidth = 1.1) +
    labs(y = "Density") +
    annotate(
      "text",
      x = -Inf, y = Inf,
      label = tag,
      hjust = -0.3, vjust = 1.3,
      size = 3, fontface = "bold"
    ) +
    theme_minimal(base_size = 8) +
    theme(
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.y = element_line(color = scales::alpha("black", 0.25), linewidth = 0.3),
      panel.grid.minor.y = element_blank(),
      panel.background = element_rect(fill = "transparent", color = NA),
      plot.background = element_rect(fill = "transparent", color = NA),
      strip.background = element_blank(),
      strip.text = element_blank(),
      axis.text = element_text(size = 6, color = "black"),
      axis.title.y = element_text(size = 7, margin = margin(r = 5), face = "bold"),
      axis.text.x = element_text(size = 5.5, angle = 30, hjust = 1),
      axis.title.x = element_blank()
    )
}

#  Split df_plot into 9 panels
xlab_plot <- ggplot() +
  theme_void() +
  annotate("text", x = 0.5, y = 0.5,
           label = "Response value",
           size = 2.5, fontface = "bold")

draw_sample <- sample(unique(ppd_all_models$draw), size = 30)

df_plot <- ppd_all_models %>%
  filter(
    model != "animal_cons_cap",
    str_detect(clean_name, "capita") |
      str_detect(clean_name, "Aquatic reliance") |
      str_detect(clean_name, "reliance")
  ) %>%
  filter(draw %in% draw_sample) %>%
  mutate(clean_name = factor(clean_name))

panel_names <- levels(df_plot$clean_name)

layout_order <- c(15, 3, 5,
                  1, 7, 9,
                  13, 11, 4,
                  2, 6, 8,
                  10, 14, 12)

panel_tags <- letters[1:15]

# assign tags in layout order
tags_by_panel_index <- rep("", length(panel_names))
tags_by_panel_index[layout_order] <- panel_tags

# Create panel plots
panels <- lapply(seq_along(panel_names), function(i) {
  nm <- panel_names[i]
  df_sub <- df_plot %>% filter(clean_name == nm)
  make_ppd_panel(df_sub, nm, tag = tags_by_panel_index[i])
})


# Create empty plots for bottom row in the combined plot (so final plot that sits alone at the bottom can be centered)
empty <- ggplot() + theme_void()

# Create panels that will display in the ppd plot
row1 <- panels[[15]]  | panels[[3]]  | panels[[5]]
row2 <- panels[[1]]  | panels[[7]]  | panels[[9]]
row3 <- panels[[13]]  | panels[[11]]  | panels[[4]]
row4 <- panels[[2]] | panels[[6]] | panels[[8]]
row5 <- panels[[10]] | panels[[14]] | panels[[12]]
row6 <- empty | xlab_plot | empty

(final_plot <- wrap_plots(
  row1, row2, row3, row4, row5, row6,
  ncol = 1
) + plot_layout(heights = c(1,1,1,1,1,0.15)))

ggsave(paste0("../images/SI/ppd_SI.jpg"), plot = final_plot, device = "jpg", width = 4, height = 5, units = "in", dpi = 300)
```

```{r}
```{r ppd distributions}
# Variable list with coastline_km area instead of EEZ area
predictors <- c(
  "coastline_km",
  "log_gdp",
  "gdp_trade",
  "inundation_area_log1000km2",
  "supermarkets",
  "pop_10k_pct_coast"
)

# Variable list with EEZ total area instead of coastline_km
# predictors <- c(
#   "eez_total",
#   "log_gdp",
#   "gdp_trade",
#   "inundation_area_log1000km2",
#   "supermarkets",
#   "pop_10k_pct_coast"
# )

# Get which random effect is used
model_type = "realm"

# Get which variable is selected (for unique file path)
var_used = predictors[1]

# Create file path to save files to
base_path <- file.path(
  "..", "output", "bayesian_output",
  paste0("random_effect_", model_type),
  paste0("var_", var_used)
)

if ( !file.exists(paste0("../output/bayesian_output/", model_type, "/posterior_predictive_all_models.parquet")) ) {

# Load in function to clean the model output names from snake case to a publication-friendly read
source("../R/clean_model_name.R")

# Function to extract posterior predictive draws for a model.
extract_ppd <- function(model_name, model_type) {

  # Load fitted Stan model
  fit <- readRDS(file.path(base_path, paste0("fit_", response_vars[i], ".rds")))

  x <- rstan::extract(fit, permuted = TRUE)

  # Posterior predictive draws (S × N)
  y_pred <- x$y_pred

  # Load X and y to get N
  X <- readRDS(file.path(base_path, paste0("X_", response_vars[i], ".rds")))
  y <- readRDS(file.path(base_path, paste0("y_", response_vars[i], ".rds")))

  N <- length(y)

  # Load realm names (vector of length N)
  realm_levels <- readRDS(file.path(base_path, paste0("realms_", response_vars[i], ".rds")))

  # Safety check
  if (length(realm_levels) != N) {
    stop(paste0("realm_levels length mismatch for model ", model_name))
  }

  # Format long
  ppd_long <- as.data.frame(y_pred) %>%
    mutate(draw = row_number()) %>%
    pivot_longer(-draw, names_to = "obs", values_to = "ppd") %>%
    mutate(
      obs = as.integer(gsub("V", "", obs)),
      region = realm_levels[obs],
      y_obs = y[obs],
      model = model_name
    )

  return(ppd_long)
}


# Loop over all models and bind results
ppd_all_models <- map_dfr(response_vars, ~ extract_ppd(.x, model_type)) %>%
  mutate(clean_name = clean_model_name(model)) # Add column for clean model names


# Calculate posterior predictive means
ppd_means <- ppd_all_models %>%
  group_by(model, obs, region) %>%
  summarise(
    ppd_mean = mean(ppd),
    y_obs    = first(y_obs),
    .groups  = "drop"
  )

}
```

```{r significance outlines for bayesian posterior plots}


    # Non-significant rows — reduced opacity, drawn behind
    # geom_segment(data = df_nsig,
    #              aes(x = lo95, xend = hi95, color = col95),
    #              linewidth = 2.2, lineend = "butt", alpha = 0.3) +
    # geom_segment(data = df_nsig,
    #              aes(x = lo80, xend = hi80, color = col80),
    #              linewidth = 2.2, lineend = "butt", alpha = 0.3) +
    # geom_segment(data = df_nsig,
    #              aes(x = lo50, xend = hi50, color = col50),
    #              linewidth = 2.2, lineend = "butt", alpha = 0.3) +
    # geom_point(data = df_nsig,
    #            aes(x = med), shape = 21, size = 1.6,
    #            fill = "white", color = "grey60", stroke = 0.3, alpha = 0.5) +

    # Significant rows — full opacity, drawn on top
    # geom_segment(data = df_sig,
    #              aes(x = lo95, xend = hi95, color = col95),
    #              linewidth = 2.2, lineend = "butt", alpha = 1) +
    # geom_segment(data = df_sig,
    #              aes(x = lo80, xend = hi80, color = col80),
    #              linewidth = 2.2, lineend = "butt", alpha = 1) +
    # geom_segment(data = df_sig,
    #              aes(x = lo50, xend = hi50, color = col50),
    #              linewidth = 2.2, lineend = "butt", alpha = 1) +
    # geom_point(data = df_sig,
    #            aes(x = med), shape = 21, size = 1.6,
    #            fill = "white", color = "grey20", stroke = 0.4) +
```

```{r residuals}
# Load objects
fit <- readRDS(file.path(base_path, paste0("fit_", response_vars[i], ".rds")))
x   <- rstan::extract(fit, permuted = TRUE)

X <- readRDS(file.path(base_path, paste0("X_", response_vars[i], ".rds")))
y <- readRDS(file.path(base_path, paste0("y_", response_vars[i], ".rds")))
realm <- readRDS(file.path(base_path, paste0("realms_", response_vars[i], ".rds")))

province <- as.integer(factor(realm))

# Dimensions
n_iter <- length(x$beta_0)
N <- length(y)

# Storage
residuals <- matrix(NA, nrow = n_iter, ncol = N)

# Compute residuals
for (s in 1:n_iter) {
  mu_s <- x$beta_0i[s, province] + X %*% x$beta[s, ]
  residuals[s, ] <- y - mu_s
}


resid_mean <- colMeans(residuals)

hist(resid_mean)
qqnorm(resid_mean); qqline(resid_mean)
```

```{r extra model names}
if ( !file.exists("../output/response_vars.rds") ) {
  
  # Calculate animal consumption
terr_animal_cons_cap <- aquatic_reliance %>%
  filter(year == 2019) %>%
  group_by(iso3c) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))

animal_cons_cap <- aquatic_reliance %>%
  filter(year == 2019) %>%
  group_by(iso3c) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))

# Calculate aquatic animal consumption
aquatic_animal_cons_cap <- aquatic_reliance %>%
  filter(year == 2019, food_group == "aquatic") %>%
  group_by(iso3c) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))

# Calculate animal reliance consumption
aquatic_rel <- aquatic_animal_cons_cap %>%
  left_join(animal_cons_cap, by = "iso3c") %>%
  group_by(iso3c) %>%
  summarize(prop_animal_protein = protein_consumed_kg_percap.x / protein_consumed_kg_percap.y)

# Add items not created in the loop to the list
response_vars <- c("terr_animal_cons_cap","animal_cons_cap", "aquatic_animal_cons_cap", "aquatic_rel", "aquatic_protein_roc", "terrestrial_protein_roc")

for (i in c("capture", "aquaculture", "domestic", "foreign", "marine", "inland")) {
  
  # Calculate consumption per capita by method
  if (i %in% c("capture", "aquaculture")) {
    assign(
      paste0(i, "_cap"),
      aquatic_reliance %>%
        filter(year == 2019, food_group == "aquatic", method == i) %>%
        group_by(iso3c) %>%
        summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))
    )
  }
  
  # Calculate consumption per capita by consumption source
  if (i %in% c("domestic", "foreign")) {
    assign(
      paste0(i, "_cap"),
      aquatic_reliance %>%
        filter(year == 2019, food_group == "aquatic", consumption_source == i) %>%
        group_by(iso3c) %>%
        summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))
    )
  }
  
  # Calculate consumption per capita by habitat
  if (i %in% c("marine", "inland")) {
    assign(
      paste0(i, "_cap"),
      aquatic_reliance %>%
        filter(year == 2019, food_group == "aquatic", habitat == i) %>%
        group_by(iso3c) %>%
        summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap))
    )
  }
  
  response_vars <- c(response_vars, paste0(i, "_cap"))
  
  # Calculate reliance by method, production source, or habitat
  assign(
    paste0(i, "_rel"),
    get(paste0(i, "_cap")) %>%
      left_join(animal_cons_cap, by = "iso3c") %>%
      group_by(iso3c) %>%
      summarize(prop_animal_protein = protein_consumed_kg_percap.x / protein_consumed_kg_percap.y)
  )
  
  response_vars <- c(response_vars, paste0(i, "_rel"))

}

#calculate rate of change in aquatic protein quantity
aquatic_protein_roc <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "aquatic") %>%
  group_by(iso3c, year, method) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap, na.rm = TRUE)) %>%
  distinct() %>%
  filter(method == "capture") %>%
  split(.$iso3c) %>%
  map(~lm(protein_consumed_kg_percap~year, data = .x)) %>%
  map_df(broom::tidy, .id = 'iso3c') %>%
  filter(term == 'year') %>%
  select(c("iso3c", "aquatic_estimate" = "estimate"))

#calculate rate of change in terrestrial protein quantity
terrestrial_protein_roc <- aquatic_reliance %>%
  filter(data_source == "new FBS", food_group == "terrestrial") %>%
  group_by(iso3c, year) %>%
  summarize(protein_consumed_kg_percap = sum(protein_consumed_kg_percap, na.rm = TRUE)) %>%
  distinct() %>%
  split(.$iso3c) %>%
  map(~lm(protein_consumed_kg_percap~year, data = .x)) %>%
  map_df(broom::tidy, .id = 'iso3c') %>%
  filter(term == 'year') %>%
  select(c("iso3c", "terrestrial_estimate" = "estimate"))

# Save response_vars to RDS format
saveRDS(response_vars, "../data/response_vars.rds")
}
```

### SI Figure: R^2

```{r R^2}
# Calculate hierarchical R² across models, decomposed into fixed, random,
# and combined (conditional) components
R2_df <- data.frame(
  fixed_r2    = numeric(0),
  random_r2   = numeric(0),
  combined_r2 = numeric(0),
  model       = character(0)
)

for (i in seq_along(response_vars)) {

  fit <- rstan::extract(
    readRDS(file.path(base_path, paste0("fit_", response_vars[i], ".rds"))),
    permuted = TRUE
  )

  b_s       <- fit$beta_0i[1:3000, ]
  realm_ids <- readRDS(file.path(base_path, paste0("realms_", response_vars[i], ".rds")))
  X         <- readRDS(file.path(base_path, paste0("X_",      response_vars[i], ".rds")))
  Z         <- model.matrix(~ 0 + factor(realm_ids))

  # Linear predictors: iterations × N
  mu_random <- b_s %*% t(Z)
  beta_s    <- fit$beta[1:3000, ]
  mu_fixed  <- beta_s %*% t(X)

  # Variance partitioning across posterior iterations
  var_fixed_s  <- apply(mu_fixed,  1, var)
  var_random_s <- apply(mu_random, 1, var)
  var_resid_s  <- fit$sigma^2
  mu_total     <- mu_fixed + mu_random
  var_total_s  <- apply(mu_total, 1, var)

  # Marginal R² (fixed), random R², and conditional R² (combined)
  R2_fixed_s  <- var_fixed_s  / (var_total_s + var_resid_s)
  R2_random_s <- var_random_s / (var_total_s + var_resid_s)
  R2_cond_s   <- var_total_s  / (var_total_s + var_resid_s)

  R2_df <- bind_rows(
    R2_df,
    data.frame(
      fixed_r2    = R2_fixed_s,
      random_r2   = R2_random_s,
      combined_r2 = R2_cond_s,
      model       = response_vars[i]
    )
  )
}

# Colours for effect type - drawn from project palette
# Fixed = blue (predictors), random = teal (realm structure), combined = grey
r2_colors <- c(
  "Fixed"    = "#2B6AA8",
  "Random"   = "#1A7A6E",
  "Combined" = "#888780"
)

# Build plot data - filter to capita and reliance models, clean names
r2_plot_df <- R2_df %>%
  filter(
    str_detect(model, "_cap") | str_detect(model, "_rel"),
    model != "animal_cons_cap"
  ) %>%
  pivot_longer(
    cols      = c(fixed_r2, random_r2, combined_r2),
    names_to  = "effect",
    values_to = "r2"
  ) %>%
  mutate(
    clean_name = clean_model_name(model),
    effect     = factor(
      case_when(
        effect == "fixed_r2"    ~ "Fixed",
        effect == "random_r2"   ~ "Random",
        effect == "combined_r2" ~ "Combined"
      ),
      levels = c("Fixed", "Random", "Combined")
    )
  )

# Compute posterior medians and 95% CI per model × effect for reference lines
r2_summary <- r2_plot_df %>%
  group_by(clean_name, effect) %>%
  summarise(
    med   = median(r2),
    lo95  = quantile(r2, 0.025),
    hi95  = quantile(r2, 0.975),
    .groups = "drop"
  )

# Shared theme applied to every individual panel
theme_r2 <- function(show_y = FALSE, show_x = FALSE) {
  base <- theme_minimal(base_size = 8) +
    theme(
      text                  = element_text(family = "sans", color = "black"),
      plot.background       = element_rect(fill = "transparent", color = NA),
      panel.background      = element_rect(fill = "transparent", color = NA),
      panel.grid.minor      = element_blank(),
      panel.grid.major.x    = element_line(
        color = scales::alpha("black", 0.12), linewidth = 0.3
      ),
      panel.grid.major.y    = element_blank(),
      strip.background      = element_blank(),
      plot.title            = element_text(size = 5.5, face = "bold",
                                            hjust = 0.5, margin = margin(b = 2)),
      plot.tag              = element_text(size = 7, face = "bold"),
      plot.tag.position     = "topleft",
      axis.text.x           = element_text(size = 5.5, color = "black"),
      axis.text.y           = element_text(size = 5.5, color = "black"),
      axis.title.x          = element_blank(),
      axis.title.y          = element_blank(),
      legend.position       = "none",
      plot.margin           = margin(3, 4, 3, 4)
    )

  if (show_y) base <- base + theme(
    axis.title.y = element_text(size = 6.5, face = "bold", margin = margin(r = 3))
  )
  if (show_x) base <- base + theme(
    axis.title.x = element_text(size = 6.5, face = "bold", margin = margin(t = 3))
  )
  base
}

# Build one density panel per model
model_names <- unique(r2_plot_df$clean_name)

panels <- lapply(seq_along(model_names), function(i) {
  nm      <- model_names[i]
  df_sub  <- r2_plot_df %>% filter(clean_name == nm)
  sum_sub <- r2_summary  %>% filter(clean_name == nm)

  # Left column panels: 1, 5, 9, 13 - show y-axis title
  # Bottom row panels: 13, 14, 15 - show x-axis title
  show_y <- i %in% c(1, 5, 9, 13)
  show_x <- i %in% c(13, 14, 15)

  ggplot(df_sub, aes(x = r2, colour = effect, fill = effect)) +
    geom_density(alpha = 0.18, linewidth = 0.55) +
    geom_vline(
      data      = sum_sub,
      aes(xintercept = med, colour = effect),
      linewidth = 0.4,
      linetype  = "dashed"
    ) +
    scale_colour_manual(values = r2_colors) +
    scale_fill_manual(  values = r2_colors) +
    scale_x_continuous(
      limits = c(0, 1),
      breaks = c(0, 0.5, 1),
      labels = c("0", "0.5", "1")
    ) +
    labs(
      title = nm,
      tag   = LETTERS[i],
      y     = if (show_y) "Density" else NULL,
      x     = if (show_x) bquote("Posterior" ~ R^2) else NULL
    ) +
    theme_r2(show_y = show_y, show_x = show_x)
})

# Extract a shared legend from a single fully-specified panel
legend_source <- ggplot(r2_plot_df, aes(x = r2, colour = effect, fill = effect)) +
  geom_density(alpha = 0.18, linewidth = 0.55) +
  scale_colour_manual(values = r2_colors, name = "Effect type") +
  scale_fill_manual(  values = r2_colors, name = "Effect type") +
  guides(
    colour = guide_legend(title.position = "top", title.hjust = 0.5, nrow = 1),
    fill   = guide_legend(title.position = "top", title.hjust = 0.5, nrow = 1)
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

legend_grob <- cowplot::get_plot_component(
  legend_source, "guide-box", return_all = TRUE
)
legend_grob <- Filter(function(x) !inherits(x, "zeroGrob"), legend_grob)[[1]]

# Assemble rows manually - bottom row (3 panels) centered with empty spacers
# widths c(0.5, 1, 1, 1, 0.5) flanks the three panels with half-column margins
empty <- ggplot() + theme_void()

row1 <- panels[[1]]  | panels[[2]]  | panels[[3]]  | panels[[4]]
row2 <- panels[[5]]  | panels[[6]]  | panels[[7]]  | panels[[8]]
row3 <- panels[[9]]  | panels[[10]] | panels[[11]] | panels[[12]]
row4 <- (empty | panels[[13]] | panels[[14]] | panels[[15]] | empty) +
  plot_layout(widths = c(0.5, 1, 1, 1, 0.5))

grid_plot <- wrap_plots(row1, row2, row3, row4, ncol = 1) +
  plot_layout(heights = c(1, 1, 1, 1)) &
  theme(plot.margin = margin(2, 3, 2, 3))

# Stack grid above shared legend
plot <- cowplot::plot_grid(
  grid_plot,
  legend_grob,
  ncol        = 1,
  rel_heights = c(1, 0.11)
)

ggsave(
  filename = paste0("../images/SI/r2.pdf"),
  plot     = plot,
  device   = cairo_pdf,
  width    = 9,
  height   = 5,
  units    = "in"
)
```
